### 2018-04-20 working-paper/paper.tex
added working-paper

\documentclass[review]{elsarticle}

\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{}%
 \let\@evenfoot\@oddfoot}
\makeatother

\usepackage{etoolbox}
\patchcmd{\MaketitleBox}{\footnotesize\itshape\elsaddress\par\vskip36pt}{\footnotesize\itshape\elsaddress\par\parbox[b][36pt]{\linewidth}{\vfill\hfill\textnormal{\today}\hfill\null\vfill}}{}{}%
\patchcmd{\pprintMaketitle}{\footnotesize\itshape\elsaddress\par\vskip36pt}{\footnotesize\itshape\elsaddress\par\parbox[b][36pt]{\linewidth}{\vfill\hfill\textnormal{Spring, 2018}\hfill\null\vfill}}{}{}%

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

% For citations
%\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{hyperref}

\newcommand{\theHalgorithm}{\arabic{algorithm}}

%Arash
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{siunitx}

\bibliographystyle{model5-names}

\begin{document}

\begin{frontmatter}

\title{Modeling trend in temperature volatility using generalized LASSO}

\author{Arash Khodadadi}
\address{Department of Statistics, Indiana University, Bloomington, IN, United States}
\emailauthor{arash.khodadadi@gmail.com}{Arash Khodadadi}

\date{\today}

\begin{keyword}
generalized Lasso, linearized ADMM, consensus optimization, trend, volatility, convex optimization.
\end{keyword}

\end{frontmatter}

\section{Introduction}

There is a considerable interest in determining if there is an increasing trend in the climate variability \cite{hansen_perception_2012,huntingford_no_2013}. An increase in the temperature variability will increase the probability of extreme hot outliers. It might be harder for the society to adapt to these extremes than to the gradual increase in the mean temperature \cite{huntingford_no_2013}.

In this project, we consider the problem of detecting the trend in the temperature volatility. All analyses are performed on a sub-set of the European Centre for Medium-Range Weather Forecasts (ECMWF) ERA-40 dataset \cite{uppala_era-40_2005}. This dataset include the temperature measurements over a grid over the earth from 1957 to 2002.

Research on analyzing the trend in the volatility of spatio-temporal data is scarce. \cite{hansen_perception_2012} studied the change in the standard deviation (SD) of the surface temperature in the NASA Goddard Institute for Space Studies gridded temperature data set. In their analysis, for each geographical position, the mean of the temperature computed for the period 1951-1980 (called the base-period) at that position, is subtracted from the corresponding time series. Each time series is then divided by the standard deviation computed at each position and during the same time period. The distribution of the resulting data is then plotted for different periods. These distributions represent the deviation of the temperature for a specific period, from the mean in the base period, in units of the standard deviation in that period. The results showed that these distributions are widen for the resent time periods compared to 1951-1980.  \cite{huntingford_no_2013} took a similar approach in analysing the ERA-40 data set. However, in addition to the aforementioned method, they computed the distribution of the SDs in an alternative way: for each position and each time period, the deviation of the time-series at that position from the mean in that time period at that position was computed, and then divided by the SD of that position in the period before 1981. The results showed that there still is an increase in the SDs from 1958-1970 to 1991-2001, but this is much less than what is obtained from the method used in \cite{hansen_perception_2012}. The authors also computed the time-evolving global SD from the de-trended time-series at each position. The resulting curve suggested that the global SD has been stable.

These previous work (and other related research, e.g., \cite{rhines_frequent_2013}) have several shortcomings. First, no statistical analysis has been performed to examine if the change in the SD is statistically significant. Second, the methodologies for computing the SDs are rather arbitrary. The deviation of each time-series in a given period, is computed from either the mean of a base-period (as in \cite{hansen_perception_2012}), or from the given period (as in \cite{rhines_frequent_2013,huntingford_no_2013}). These deviations are then normalized using the SD of the base-period or the given period. No justification is provided for these choices. Third, the correlation between the observations is ignored. The observations in subsequent days and close geographical positions could be highly correlated. Without considering these correlations, any conclusion based on the averaged data could be flawed.

The main contribution of this work is to develop a new methodology for detecting the trend in the volatility of spatio-temporal data. In this methodology, the variance at each position and time, is considered as a hidden (un-observed) variable. The value of these hidden variables are then estimated by maximizing the likelihood of the observed data. We show that this formulation per se, is not appropriate for detecting the trend. To overcome this issue, we penalize the differences between the estimated variances of the observations which are temporally and/or spatially close to each other. This will result in an optimization problem called the \textit{generalized LASSO problem} \cite{tibshirani_solution_2011}. As we will see, the dimension of this optimization problem is very high and so the standard methods for solving the generalized LASSO cannot be applied directly. We investigate two methods for solving this optimization problem. In the first method, we adopt an optimization technique called alternative direction method of multipliers (ADMM) \cite{boyd_distributed_2011}, to divide the total problem into several sub-problems of much lower dimension and show how the total problem can be solved by iteratively solving these sub-problems. The second method, called the \textit{linearized ADMM algorithm} \cite{parikh_proximal_2014} solves the main problem by iteratively solving a linearized version of it. We will compare the benefits of each method.

\section{Data Exploration}
This section is devoted to exploring some of the properties of the ERA-40 surface temperature data set. The goal here is to demonstrate some of the difficulties in modeling the trend in the temperature volatility and motivate our methodology for doing so.

Figure \ref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil). The time-series of Bloomington and San Diego show clear cyclic behavior. However, while it seems (qualitatively) that these cycles can be modeled by a sinusoidal function for Bloomington, the same is not true for San Diego. Also, the amplitude of the cycles changes from some years to others. The time-series of Manaus does not show any regular cyclic behavior. This demonstrates the first difficulty in analyzing the variance of this data: to analyze the variance, we first need to remove the cyclic terms from all time-series. However, there is a lot of variations in the cyclic behavior of the time-series of different locations. In addition, some of these cycles cannot be easily modeled by a parametric function \footnote{One might try to model the cycles by the summation of sinusoidal terms with different frequencies. However, for some time-series this may need many terms to be included in the summation to achieve a reasonable level of accuracy. In addition, this model cannot capture the non-stationarity in the cycles.}. To overcome these issues, we will use a non-parametric approach to remove the cyclic terms from the time-series and de-trend them. This approach, called \textit{$\ell_1$-trend filtering} is explained in the next section.

\begin{figure}[ht]
	\vskip 0.2in
	\begin{center}
		\centerline{\includegraphics[width=\columnwidth]{Figures/cities_ts}}
		\caption{Time-series of the temperature (in Kelvin) of three cities.}
		\label{fig:cities_ts}
	\end{center}
	\vskip -0.2in
\end{figure}

The right panel of Figure \ref{fig:bloom_detrended} shows the time-series of the temperature of Bloomington, after removing the cyclic terms and de-trending using the method explained in the next section. The goal is to investigate the trend in the variance of this signal. This figure, reveals another issue toward this goal: the variance of this signal, shows cyclic behavior. Also, the cycles are not regular and their amplitude and frequency change. Even if one can describe the behavior of the variance of all the time-series using a single parametric model (for example a variant of the GARCH models \cite{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework. These observations motivate the need to develop a non-parametric framework for the problem at hand.

\begin{figure}[ht]
	\vskip 0.2in
	\begin{center}
		\centerline{\includegraphics[width=\columnwidth]{Figures/bloom_trend_resid}}
		\caption{Left: Time-series of the temperature of Bloomington (blue) and the estimated trend and cycles obtained from the $\ell_1$-trend filtering (red). Right: the same time-series after removing the cyclic terms and de-trending using $\ell_1$-trend filtering.}
		\label{fig:bloom_detrended}
	\end{center}
	\vskip -0.2in
\end{figure}

\section{$\ell_1$-trend filtering for de-trending}

$\ell_1$-trend filtering was proposed by \cite{kim_$ell_1$_2009} as a solution to the following problem: a time-series $y_t$, $t=1,...,T$ is observed and we believe that it consists of a smooth trend and a stochastic component and the goal is to estimate the trend. The estimated trend is desired to be smooth, but at the same time, we seek small estimated residuals (stochastic component). These two objectives cannot be achieved simultaneously and so a trade-off between them should be made. In $\ell_1$-trend filtering, this trade-off is formulated as the following convex optimization problem:

\begin{equation}
\min_{\beta} \frac{1}{2} \sum_{t=1}^{T} (y_t-\beta_t)^2+\lambda \sum_{t=1}^{T-2} |\beta_t-2\beta_{t+1}+\beta_{t+2}|
\end{equation}

or equivalently:

\begin{equation}
\min_{\beta} \frac{1}{2} \lVert y-\beta \lVert_2^2+\lambda \lVert Dh \lVert_1
\label{eq:l1tf}
\end{equation}

\noindent where $||u||_1$ is the $\ell_1$ norm of the vector $u$. Here, $\beta$ is the vector of the free parameters of the model and there is one parameter per observation. The \textit{penalty matrix} is:

\begin{equation}
D=\begin{bmatrix}
1 & -2 & 1 &  &  &  &\\ 
& 1 & -2 & 1 &  &  &\\ 
&  & \ddots & \ddots & \ddots  &  &\\ 
&  & & 1 & -2 & 1 &  \\ 
&  &  &   & 1 & -2 & 1 
\end{bmatrix}
\label{eq:d_matrix}
\end{equation}

The first term in Equation \ref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$. Specifically, with the penalty matrix \ref{eq:d_matrix}, the estimated $\beta$ will be piecewise linear \cite{kim_$ell_1$_2009}. The trade-off between the two aforementioned objectives is determined by the constant $\lambda$. \cite{kim_$ell_1$_2009} proposed a specialized primal-dual interior point (PDIP) method for solving the optimization problem \ref{eq:l1tf} with the penalty matrix \ref{eq:d_matrix}.

The results of applying $\ell_1$-trend filtering on the Bloomington time-series are shown in Figure \ref{fig:bloom_detrended}. The left panel shows the original time-series (blue) together with the estimated $\beta$ (red) and the right panel shows the residuals $y_t-\beta_t$ (blue). As it can be seen, the estimated $\beta$s are piecewise linear. The number of linear segments is determined by the value of $\lambda$. Here, we chose $\lambda=500$. The choice of $\lambda$ is a model selection problem and we will return to it in the next section ???.

\section{$\ell_1$-trend filtering for estimating variance of a time-series}
\label{sec:l1tf_var}

%TODO: talk about the motivation for using non-parametric based on the figure of the residuals.

In this section, inspired by the $\ell_1$-trend filtering algorithm, we propose a non-parametric model for estimating the variance of a time-series. To this end, we assume that at each time step $t$, there is a hidden variable $h_t$ such that conditioned on $h_t$ the observations $y_t$ are independent normal variables with zero mean and variance $\exp(h_t)$. The conditional log-likelihood of the observed data in this model is $l(y|h)=c-\sum_{t=1}^Th_t-y_t^2e^{-h_t}$, where $c$ is a constant that does not depend on $y$ and $h$. Crucially, we assume that the hidden variables $h_t$ vary smoothly. To impose this assumption, we estimate $h_t$ by solving the penalized conditional log-likelihood:

\begin{equation}
\min_h -l(y|h)+\lambda \lVert Dh \lVert_1
\label{eq:l1tf_var}
\end{equation}

\noindent where $D$ is the same as in \ref{eq:d_matrix}.

%TODO: explain the objective more. give the AR(1) example. Explain what you loses by this assumption (ACF,forecasting). Also explain that the covariace matrix is diagonal so it cannot capture the covariance structure. But incontrast to spatial stat literature, it does not make any assumption on estimated variances. Compare to Hallac et al and Lingren et al.

We solve this optimization problem using the PDIP method \footnote{We use PDIP implemented in the cvxopt Python package \cite{andersen_cvxopt:_2013}.}. First, we note that this is a generalized LASSO problem \cite{tibshirani_solution_2011}. The dual of a generalized LASSO with the objective $f(x)+\lambda \lVert Dx \lVert_1$ is:

\begin{equation}
\begin{aligned}
\min_\nu f^*(-D^T\mathsf{\nu}) \\
s.t. \,\,\, \lVert \mathsf{\nu} \lVert_\infty \le \lambda
\end{aligned}
\end{equation}

\noindent where $D^t$ is the transpose of $D$, and $f^*(\cdot)$ is the \textit{conjugate} of $f$ defined by: $f^*(u)=\max_x u^tx-f(x)$. It is simple to show that the conjugate of $f(h)=-l(y|h)$ is:

\begin{equation}
f^*(\mathsf{u})=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1
\label{eq:conj}
\end{equation}

The red curve in the left panel of Figure \ref{fig:bloom_estimatedSD} shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of the time-series of Bloomington. To reduce the number of time-steps in this figure and in the remainder of the paper we work on the weekly averaged of the data.

The curve of the estimated SD captures the periodic variations in the SD of the signal. Just by looking at this curve, it is hard to say if the SD is decreasing or increasing. Therefore, we compute the average of the estimated SD for each year. The estimated SD together with this annual average is shown in the middle panel of Figure \ref{fig:bloom_estimatedSD}. As it can be seen, the annual trend is not smooth. This is because in the optimization problem \ref{eq:l1tf_var}, the smoothness of the annual trend is not encouraged. To remedy this, we add the following long horizon penalty to \ref{eq:l1tf_var}:

\begin{equation}
\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
\label{eq:lh_penalty}
\end{equation}

\noindent where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$. Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$). Since we are working on the weekly averaged data, each year corresponds to 52 observations. In the matrix form, the penalty \ref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$. The estimated SDs using this penalty matrix is shown in the right panel of Figure \ref{fig:bloom_estimatedSD}. The annual average of the estimated SDs shows a linear trend with a positive slope.

\begin{figure}[ht]
	\vskip 0.2in
	\begin{center}
		\centerline{\includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}}
		\caption{Left: The residuals of the time-series of Bloomington (averaged weekly) and the estimated SD obtained from the method of Section \ref{sec:l1tf_var} (red). Middle: the estimated SDs (red) and their annual average (black) without imposing the long horizon penalty. Right: the same as middle panel but here the long horizon penalty is imposed. See the text for more details.}
		\label{fig:bloom_estimatedSD}
	\end{center}
	\vskip -0.2in
\end{figure}

%TODO: how to choose lambda? plot BIC as a function of lambda.

\section{Extension to spatio-temporal data}
\label{sec:exten}

The method in the previous section can be used to estimate the variance of a single time-series. In this section, we extend this method to the estimation of the variance of spatio-temporal data.

At a specific time $t$, the data is measured on a grid of points with $n_r$ rows and $n_c$ columns. Let $y_{ijt}$ denote the value of the observation at time $t$ on the $i^\text{th}$ row and $j^\text{th}$ column of the grid, and $h_{ijt}$ denote the corresponding hidden variable. We seek to impose both temporal and spatial smoothness constraints on the hidden variables. Specifically, we seek a solution for $h$ which is piecewise linear in time and piecewise constant in space \footnote{The assumption that the variance is spatially piecewise constant simplifies the computations. The justification for this assumption is that we are interested in examining the trend in the variance over time and not over space.}. This can be achieved by solving the following optimization problem:

\begin{equation}
\begin{aligned}
\min_h &\sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}}\\
&+\lambda_t \sum_{i,j} \sum_{t=1}^{T-2} \big|h_{ijt}-2h_{ij(t+1)}+h_{ij(t+2)}\big|\\
&+\lambda_s \sum_{t,j} \sum_{i=1}^{n_r-1} \big|h_{ijt}-h_{(i+1)jt}\big|\\
&+\lambda_s \sum_{t,i} \sum_{j=1}^{n_c-1} \big|h_{ijt}-h_{i(j+1)t}\big|
\end{aligned}
\label{eq:l1tf_var_st}
\end{equation}

The first term in the objective is the negative log-likelihood (minus a constant term). The second term is the temporal penalty for the time-series at each location $(i,j)$. The third and fourth terms, penalize the difference between the estimated variance of two vertically and horizontally adjacent points, respectively. This penalty is a special case of the penalty used in the \textit{trend filtering on graphs }\cite{wang_trend_2016} (where the difference between the estimated values of the signal at each two nodes connected with an edge is penalized.). The optimization problem \ref{eq:l1tf_var_st} can be written in the matrix form. Let $h$ be a vector whose first $T$ entries are $h_{11t}$ for $t=1,...,T$, its next $T$ entries are $h_{21t}$ and so on. Then the optimization problem in the matrix form is as follows:

\begin{equation}
\min_h -l(y|h)+ \Lambda^t \lVert D_{total}h \lVert_1
\label{eq:l1tf_var_st_mat}
\end{equation}

Let, $\textbf{e}_n$ denote a vector of size $n$ with all entries being equal to 1. We have: $\Lambda^t=(\lambda_t\textbf{e}_{n_t}^t|\lambda_s\textbf{e}_{n_s}^t)$, $D_{total}^t=[D_{t}^t|D_{s}^t]$, where $n_t$ and $n_s$ are the number of rows of $D_{t}$ and $D_{s}$, respectively (see below). The matrix $D_t$ is the following block-diagonal matrix and corresponds to the temporal penalty:

\begin{equation}
D_t=\begin{bmatrix}
D &  & \\ 
 & \ddots & \\
 &  & D
\end{bmatrix}
\label{eq:d_t_matrix}
\end{equation}

\noindent where $D$ was defined in \ref{eq:d_matrix}. The number of the diagonal blocks is equal to the grid size $n_r \times n_c$. Each row of the matrix $D_s$ corresponds to one spatial constraint in \ref{eq:l1tf_var_st}. For example, the first $T$ rows correspond to $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to $|h_{11t}-h_{12t}|$, and so on.

The dual of this problem is:

\begin{equation}
\begin{aligned}
\min_\nu f^*(-D_{total}^t\mathsf{\nu}) \\
s.t. \,\,\, | \nu_i | \le \Lambda_i
\end{aligned}
\label{eq:dual}
\end{equation}

\noindent where $f^*$ is given in \ref{eq:conj}.

\section{Optimization}
For a spatial grid of size $n_r \times n_c$ and for $T$ time steps, we have $n_t=3n_rn_c-Tn_c-2n_rn_c$ and $n_s=n_rn_cT$. For a grid over the united states and for weekly averaged data of 10 years we have $n_r=32$, $n_c=68$, $T=521$ and so $n_t\approx 3.5\times10^6$ and $n_s \approx 1.0\times10^6$. Therefore, the size of the optimization problem \ref{eq:dual} is about $n_t\approx 4.5\times10^6$. In each step of the PDIP algorithm, we need to solve a linear system of equations in the form $z=Ax$ where $A$ is a square matrix of size $2(n_t+n_s)$ (see \cite{boyd_convex_2004} equation 11.54). Therefore, applying the PDIP directly for solving the optimization problem \ref{eq:dual} is infeasible.

In the next section, we develop an ADMM algorithm for solving this problem. The idea is to cast the problem \ref{eq:dual} as a so-called \textit{consensus optimization problem} \cite{boyd_distributed_2011} and solve it by breaking the problem into smaller sub-problems using an ADMM-based algorithm. Next, we first give a brief overview of the consensus optimization problem and then explain how the problem \ref{eq:l1tf_var_st_mat} can be solved in this framework.

\subsection{Consensus optimization}
\label{sec:consOpt}

A general form consensus optimization problem is in the form $\min_z \sum_i f_i(z)$, $z\in\mathbb{R}^n$. We can define a set of \textit{local variables} $x_i\in \mathbb{R}^{n_i}$ such that $\sum_i f_i(z)=\sum_i f_i(x_i)$. We follow the notation of \cite{boyd_distributed_2011} closely. Let $k=\mathscr{G}(i,j)$ which means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or $(x_i)_j=z_k$) and define $\tilde{z}_i \in \mathbb{R}^{n_i}$ by $(\tilde{z}_i)_j=(x_i)_j$. Then the original unconstrained optimization problem is equivalent to the following constrained optimization problem:

\begin{equation}
\begin{aligned}
\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)\\
 s.t. \,\,\, \tilde{z}_i=x_i \,\,\,\,\,\,
\end{aligned}
\end{equation}

Now, we can apply ADMM to the \textit{augmented Lagrangian} of this problem. This results in the following ADMM updating steps at each iteration $m$:

\begin{equation}
\begin{aligned}
& x_i^{m+1}:=\argmin_{x_i} \bigg( f_i(x_i) + (u_i^m)^t x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2\bigg)\\
& z_k^{m+1}:=(1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i^{m+1})_j\\
& u_i^{m+1}:=u_i^m + \rho (x_i^{m+1}-\tilde{z}_i^{m+1})
\end{aligned}
\label{eq:ADMM_steps}
\end{equation}

Here, $S_k$ is the number of local variable entries that correspond to $z_k$, and $u_i$ are the Lagrange multipliers.

To solve the optimization problem \ref{eq:l1tf_var_st_mat} or \ref{eq:l1tf_var_st} using the method explained in the previous section we need to two address two questions: first, how to choose the local variables $x_i$, and second, how to solve the optimization problem for updating these variables (the first line of \ref{eq:ADMM_steps}).

Obviously, the global variable in the problem \ref{eq:l1tf_var_st} is $z=h$. The global variable can be index by the spatial coordinate and the temporal step. In Figure \ref{fig:data_cube}, $z$ is represented as a cube. We can decompose $z$ into sub-cubes. An example of this decomposition is shown in the figure by white lines. It is easy to see that with this definition of $x_i$, the objective \ref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where $f_i(x_i)=-l(y_i|x_i)+\lVert \Lambda_i^t D_ix_i \lVert_1$.

%TODO: re-write this part. it is very confusing.

\begin{figure}[ht]
	\vskip 0.2in
	\begin{center}
		\centerline{\includegraphics[width=.5\columnwidth]{Figures/data_cube}}
		\caption{The cube represents the global variable $z$ in space and time. The sub-cubes specified by the white lines are $x_i$.}
		\label{fig:data_cube}
	\end{center}
	\vskip -0.2in
\end{figure}

By this definition of $x_i$, the update step for $x_i$ is the following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) + (u_i^m)^t x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.

We solve this using the PDIP method. To this end we first need to compute the dual of the problem. It can be shown that the dual of this problem is:

%TODO: write the appendix

\begin{equation}
\begin{aligned}
\min_\nu f_i^*(-D_i^t\nu_i) \\
s.t. \,\,\, | (\nu_i)_j | \le (\Lambda_i)_j
\end{aligned}
\label{eq:dual_x_update}
\end{equation}

\noindent where:

\begin{equation}
\begin{aligned}
f^*_i(q)&=\sum_j (q_i)_j (w_i)_j -(w_i)_j-(y_i)_j^2 e^{-(w_i)_j} - \\
& (\rho/2)((w_i)_j-(\alpha_i^m)_j)^2\\
&\alpha_i^m =\hat{z}_i^m-u_i^m\\
(w_i)_j&=\mathscr{W}\bigg(\frac{(y_i)_j^2}{\rho}\exp\bigg[\frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}\bigg]\bigg)-\\
& \frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}
\end{aligned}
\label{eq:conj_x_update}
\end{equation}

In these equations, $\mathscr{W}$ is the \textit{Lambert function} \cite{corless_lambertw_1996}. To use the PDIP method, we also need to compute the gradient and Hessian of $f_i^*(-D_i^t\nu_i)$. This involves computing the derivatives of the Lambert functions. The primal solution $x_i$ can be obtained from the dual solution $\nu_i$ in \ref{eq:dual_x_update} by setting $x_i=w_i$ where $w_i$ is defined in the last equation of \ref{eq:conj_x_update}.

The complete ADMM algorithm for estimating the variances is represented in \ref{alg:ADMM}. All the computations in the three updating steps \ref{eq:ADMM_steps} can be performed in parallel. The number of rows and columns of the sub-cubes should be chosen so that the updating of $x_i$ could be performed in one processor. We choose $3\times3\times521$ sub-cubes.

\begin{algorithm}[tb]
	\caption{ADMM for sparse estimation of variance of spatio-temporal data}
	\label{alg:ADMM}
	\begin{algorithmic}
		\STATE {\bfseries Input:} data $y$, mapping $\mathscr{G}(i,j)$, $\rho,\lambda_t,\lambda_s$
		\STATE Initialization: $x_i^0=z^0=u_i^0=\textbf{0}$.
		\FOR{$m=1,2,...$ }
			\FOR{$i=1$ {\bfseries to} $N_{sub-cubes}$}
				\STATE compute $\nu_i$ from \ref{eq:dual_x_update}
				\STATE compute $w_i$ from \ref{eq:conj_x_update}
				\STATE set $x_i^m:=w_i$
			\ENDFOR
		\STATE Compute $z^m$ from  \ref{eq:ADMM_steps}
		\STATE Compute $u_i^m$ from  \ref{eq:ADMM_steps}
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

This algorithm divides the main optimization problem into several sub-problems and in each iteration solves these sub-problems and then performs a consensus step to make the solutions of these sub-problems agree with each other. These sub-problems can be solved independently which makes this algorithm appealing for parallelization: in each iteration, each of these sub-problems can be sent into a separate computer and then all the solutions can be sent into a single computer which performs the consensus step. Therefore, in each iteration, the computation time will be equal to the time of solving each sub-problem plus the time of communicating the solutions to the master computer and then performing the consensus step. Since each sub-problem is small, with parallelization, the computation time in each iteration will be small. In addition, our experiments with several values of $\lambda_t$ and $\lambda_s$ showed that the algorithm converges in few hundreds iterations. Solving each sub-problem on a machine with four 3.20GHz Intel i5-3470 cores takes less than 3 seconds on average, and so for example if we assume that communication time is 10 seconds and the algorithm converges in 300 iterations, with parallelization on $N_{sub-cubes}$ machines, the algorithm will converge in about 1 hour. Assuming that we use $N_{sub-cubes}$ machines and that the convergence rate of the algorithm is independent of the grid size, this time will be independent of the grid size.

If we perform these computations on a single machine, the computation time grows linearly with $N_{sub-cubes}$. For example, for the data in a grid over the united states and using $3\times3\times521$ sub-cubes each iteration of the algorithm will take about 20 minutes on a single machine and so with 300 iterations it will take several days to converge. Given that we need to compute the solution for several values of the parameters $\lambda_t$ and $\lambda_s$, this computation time is not feasible.

Therefore, this algorithm is only useful if we can parallelize the computation over several machines. In the next section, we describe another algorithm which makes the computation feasible on a single machine.

\subsection{Linearized ADMM}
\label{sec:linADMM}

In this section, we describe \textit{Linearized ADMM algorithm} \cite{parikh_proximal_2014} which, as we will see, makes the computation on a single machine feasible.

Consider the following optimization problem:

\begin{equation}
\min_x f(x)+g(Dx)
\label{eq:linADMM_opt}
\end{equation}

\noindent where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each iteration of the linearized ADMM algorithm for solving this problem has the following form:

\begin{eqnarray}
\begin{aligned}
x^{k+1} & := \textbf{prox}_{\mu f} \big(x^k - (\mu/\rho)D^T (D x^k − z^k + u^k )\big)\\
z^{k+1} & := \textbf{prox}_{\rho g} \big(D x^k + u^k\big)\\
u^{k+1} & := u^k + D x^{k+1} - z^{k+1}
\end{aligned}
\label{eq:linADMM_steps}
\end{eqnarray}

\noindent where $z,u\in \mathbb{R}^m$ and the \textit{proximal operator} is defined as follows:

\begin{equation}
 \textbf{prox}_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \lVert x-u \rVert^2
\label{eq:linADMM_prox}
\end{equation}

This algorithm belongs to the general class of \textit{proximal algorithms} for solving convex optimization problems. For more details about these algorithms see \cite{parikh_proximal_2014}.

The optimization problem \ref{eq:l1tf_var_st} can be put into the form \ref{eq:linADMM_opt} as follows:

\begin{equation}
\begin{aligned}
& f(x):= \sum_{k} f_k(x_k) := \sum_{k}x_k+y_{k}^2e^{-x_{k}}\\
& g(z):= \sum_{l} g_l(z_l) := \sum_{l} \lambda_l |z_l|\\
& z= Dx
\end{aligned}
\label{eq:linADMM_fg}
\end{equation}

\noindent where $y_k$ is the $k^{th}$ entry of the vector whose entries are $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector $\Lambda^t=(\lambda_t\textbf{e}_{n_t}^t|\lambda_s\textbf{e}_{n_s}^t)$ (see section \ref{sec:exten}).

To perform the steps in \ref{eq:linADMM_steps}, we need to evaluate $\textbf{prox}_{\mu f}$ and $\textbf{prox}_{\rho g}$. The proximal algorithms are feasible only if these proximal operators can be evaluated efficiently which, as we show next, is the case for our problem.

Let $\big(\textbf{prox}_{\mu f}(u)\big)_k$ be the $k^{th}$ entry of $\textbf{prox}_{\mu f}(u)$. From the \textit{separable sum} property of the proximal operators we have (see \cite{parikh_proximal_2014}, section 2.1):

\begin{equation}
\bigg(\textbf{prox}_{\mu f}(u)\bigg)_k = \textbf{prox}_{\mu f_k}(u_k)\\
\label{eq:linADMM_sepSum}
\end{equation}

Similarly, evaluating $\textbf{prox}_{\rho g}$ reduces to evaluating the proximal operators of scalar functions. We have:

\begin{equation}
\textbf{prox}_{\mu f_k}(u_k)=\min_{x_k} \,\, \mu\big(x_k+y_{k}^2e^{-x_{k}}\big)+\frac{1}{2}  (x_k-u_k)^2\\
\label{eq:linADMM_proxf}
\end{equation}

By setting the derivative with respect to $x_k$ of the function to zero we obtain:

\begin{equation}
\textbf{prox}_{\mu f_k}(u_k)=\mathscr{W}\bigg(\frac{y_k^2}{\mu} \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu u_k}{\mu}\\
\label{eq:linADMM_proxf_sol}
\end{equation}

\noindent where $\mathscr{W}$ is the Lambert function.

Next we compute $\textbf{prox}_{\mu g_l}(u_l)$. The function $\rho \lambda_l |z_l|+1/2(z_l-u_l)^2$ is not differentiable. However, at the optimal solution we have: $\rho \cdot \lambda_l \cdot \partial \big(|z_l| \big)=u_l-z_l$, where $\partial \big(|z_l| \big)$ is the sub-gradient of $|z_l|$. This results in the following solution:

\begin{equation}
\textbf{prox}_{\rho g_l}(u_l)=\left\{\begin{matrix}
 u_l-\rho \lambda_l  & if &  u_l>\rho \lambda_l\\ 
 0  \,\,\,\,\,  & if &  |u_l|\le \rho \lambda_l\\ 
 u_l+\rho \lambda_l  &  if &  u_l<-\rho \lambda_l\\

\end{matrix}\right.
\label{eq:linADMM_proxg_sol}
\end{equation}

Therefore, both proximal operators in \ref{eq:linADMM_steps} can be evaluated easily and so we can use the linearized ADMM algorithm to solve the optimization problem \ref{eq:l1tf_var_st}.

\section{Results}
As it was mentioned before, the algorithm proposed in Section \ref{sec:consOpt} is appropriate only if we parallelize it over multiple machines and so we do not pursue it further here. All the results reported in this section are obtained using the linearize ADMM algorithm of Section \ref{sec:linADMM}. We applied this algorithm on a subset of the ERA-40 dataset. The data is the 2 meter temperature measured daily at 12 p.m from August 31 of 1992 to 2002. To reduce the noise we first computed the weekly average of this data. To further reduce the size of the data, we will only analyze the data of the locations inside a rectangle extended from (\ang{58},\ang{226}) to (\ang{22},\ang{302}). This rectangle covers the united states. All the computations were performed on a machine with four 3.20GHz Intel i5-3470 cores.

\subsection{Convergence}

Figure \ref{fig:loss_vs_iter} shows the value of the loss function \ref{eq:l1tf_var_st_mat} as a function of iteration for $\lambda_t=1$ and $\lambda_s=0$. We used the following rule to determine when to stop the optimization: the optimization was stopped if the value of the loss did not improve by at least 0.1\% in 1000 trials. As we can see, the algorithm converged in about 2000 iterations. This took about 11 minutes. Our experiments showed that the convergence speed depends on the value of $\lambda_t$ and $\lambda_s$. Also, if we use the solution obtained for smaller values of these parameters as the initial value for the larger values (\textit{warm start}), the converges speed improves.

\begin{figure}[ht]
	\vskip 0.2in
	\begin{center}
		\centerline{\includegraphics[width=1\columnwidth]{Figures/loss_vs_iter}}
		\caption{Loss as a function of iteration for $\lambda_t=1$ and $\lambda_s=0$.}
		\label{fig:loss_vs_iter}
	\end{center}
	\vskip -0.2in
\end{figure}

\subsection{Model selection}
One common method for choosing the penalty parameters in the Lasso problems is to find the solution for a range of the values of these parameters and then choose the values which minimize a model selection criterion. However, such analyses needs the computation of the degrees of freedom (df). Several previous work have investigated the df in generalized lasso problems \cite{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}. However, all these studies have considered the linear regression problem and, to the best of our knowledge, the problem of computing the df for generalized lasso with general objective function has not been considered yet.

Another approach is to choose the set of values which minimize an estimate of the expected prediction error obtained by k-fold cross-validation \cite{tibshirani_regression_1996} . Although this method is applicable for our problem, it needs k times more computation.

In this paper, we use a heuristic method for choosing $\lambda_t$ and $\lambda_s$: we compute the optimal solution for a range of values of these parameters and choose the values which minimize $\mathscr{L}(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D_{total}h \lVert$. This objective is a compromise between the negative log likelihood ($-l(y|h)$) and the complexity of the solution ($\sum \lVert D_{total}h \lVert$). For smoother solutions the value of $\sum \lVert D_{total}h \lVert$ will be smaller but with the cost of larger $-l(y|h)$.

We computed the optimal solution for all the combinations of the following sets of values: $\lambda_t \in \{1,5,10,20\} \, \, , \lambda_s \in \{0,.1,1,5,10\}$. The value of $\mathscr{L}(\lambda_t,\lambda_s)$ is shown in Figure \ref{fig:modelSelect}. This function is minimized at $\lambda_t=5$ and $\lambda_s=1$. All the analyses in the next section are performed on the solution for these values.

\begin{figure}[ht]
	\vskip 0.2in
	\begin{center}
		\centerline{\includegraphics[width=1\columnwidth]{Figures/modelSelect}}
		\caption{Value of $\mathscr{L}(\lambda_t,\lambda_s)$ for several values of $\lambda_t$ and $\lambda_s$. }
		\label{fig:modelSelect}
	\end{center}
	\vskip -0.2in
\end{figure}

\subsection{Analysis of trend of temperature volatility}
Figure \ref{fig:estimatedSD} shows the detrended data, the estimated standard deviation and the yearly average of these estimates for two cities in the united states: Bloomington (left) and San Diego (right). The estimated SD captures the periodic behavior in the variance of the time-series. In addition, the number of linear segments changes adaptively in each time window depending on how fast the variance is changing.

\begin{figure}[ht]
	\vskip 0.2in
	\begin{center}
		\centerline{\includegraphics[width=1\columnwidth]{Figures/estimatedSD}}
		\caption{Detrended data and the estimated SD for Bloomington (left) and San Diego (right).}
		\label{fig:estimatedSD}
	\end{center}
	\vskip -0.2in
\end{figure}

The yearly average of the estimated SD captures the trend in the temperature volatility. For example, we can see that in Bloomington, there is a small positive trend. To determine how the volatility has changed in each location, we subtract the average of the estimated variance in 1992 from the average in the following years and compute their sum. The value of this change in the variance in each location is depicted in the right panel of Figure \ref{fig:avg_change_estimatedSD}. The left panel of this figure, shows the average estimated variance in each location.

\begin{figure}[ht]
	\begin{subfigure}
	  \centering
	  \includegraphics[width=.5\linewidth]{Figures/avg_estimatedVar}
	\end{subfigure}%
	\begin{subfigure}
	  \centering
	  \includegraphics[width=.5\linewidth]{Figures/avg_change_estimatedVar}
	\end{subfigure}%
	\caption{Left: the average of the estimated variance in all locations in the grid over the united states. Right: the change in the variance from 1992 to 2002.}
	\label{fig:avg_change_estimatedSD}
\end{figure}

It is interesting to note that the trend in volatility is almost zero over the oceans. The most positive trend can be observed in the south-east and the most negative trend has happened in the north-east.

Finally, Figure \ref{fig:hist_avg_change} shows the histogram of the changes shown in the right panel of Figure \ref{fig:avg_change_estimatedSD}. As we can see, there are two peaks in this histogram: one for locations with no trend (which includes the points on the oceans), and one peak for locations with positive trend. The locations with negative trend form the long tail of the histogram.

\begin{figure}[ht]
	\vskip 0.2in
	\begin{center}
		\centerline{\includegraphics[width=1\columnwidth]{Figures/hist_avg_change}}
		\caption{Histogram of the change in the estimated variances.}
		\label{fig:hist_avg_change}
	\end{center}
	\vskip -0.2in
\end{figure}

\section{Discussion}
In this paper, we proposed a new method for estimating the variance of spatio-temporal data. The main idea is to cast this problem as a constrained optimization problem where the constraints enforce smooth changes in the variance for neighboring points in time and space. In particular, the solution is piecewise linear in time and piecewise constant in space. The resulting optimization is in the form of a generalized LASSO problem with high-dimension, and so applying the PDIP method directly is infeasible. We therefore developed two ADMM-based algorithms to solve this problem: the consensus ADMM and linearized ADMM.

The consensus ADMM algorithm converges in few hundreds of iterations but each iteration takes much longer than the linearized ADMM algorithm. The appealing feature of the consensus ADMM algorithm is that if it is parallelized on enough number of machines the computation time per iteration remains constant as the problem size increases. The linearized ADMM algorithm, on the other hand converges in few thousands of iterations but each iteration is performed in split second. However, since the algorithm converges in many iterations it is not very appropriate for parallelization. The reason is that after each iteration the solution computed in each machine should be broadcast to the master machine and this operation takes some time which depends on the speed of the network connecting the slave machines to the master. A direction for future research would be to combine these two algorithms in the following way: the problem should be split into the sub-problems (as in the consensus ADMM) but each sub-problem can be solved using linearized ADMM.

We applied the linearized ADMM algorithm to the surface temperature data on a grid over the united states, for years 1992-2002. The results showed that in many locations the variance of the temperature has increased about 1 unit in 10 years.

The goal of this paper, however, is not to make any conclusions about the trend in the variance because we solved the problem only for a grid over the united states and for 10 years of the data. A thorough analysis, needs the full solution over the globe and for a longer time period. The goal of the paper, was to propose the idea of estimating the trend in variance of spatio-temporal signals using generalized lasso and to investigate the algorithms for solving the resulting optimization problem.

\section*{References}
\bibliography{references}

\end{document}

### 2018-04-21 code/visualization.py
modified visulization

@@ -20,29 +20,40 @@ elif dataset=='north_hemisphere':
             'globe/Fitted_linADMM'
     saveFigDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/globe/Figures'
 
+#===load data===
 print('\014')
+mu,lam_t,lam_s=(float(mu),float(lam_t),float(lam_s))
+fn='mu_{}_lam_t_{}_lam_s_{}'.format(mu,lam_t,lam_s)
 la = LinearizedADMM()#construct linearizedADMM object
 la.loadData(data_fn,metadata_fn)#load data
+#===load data===
 
-mu,lam_t,lam_s=(float(mu),float(lam_t),float(lam_s))
-fn='mu_'+str(mu)+'_lam_t_'+str(lam_t)+'_lam_s_'+str(lam_s)
+#===analyse fitted value and compute var and ave. var===
 la.analyseFittedValues(join(destDir,'X_'+fn))
+#===analyse fitted value and compute var and ave. var===
 
+#===plot time series of temerature at different locations===
 saveFigPath=join(saveFigDir,'ts_'+fn)
 la.plot_ts_of_locations([39.17,32.7],[-86.52,-117.6],saveFigPath,
                         figureLayout=(1,2),figsize=(12,4))
-la.plotAvgChangeInVariance(saveFigDir)
+#===plot time series of temerature at different locations===
+
+#===plot average change in variance===
+la.plotAvgChangeInVariance(saveFigDir,suffix=fn)
+#===plot average change in variance===
 
 #===plot histogram of avg. change===
 fig=plt.figure()
 plt.hist(la.changeInVar.flatten(),bins=50)
 plt.xlabel('change in variance');plt.ylabel('frequency');
-fig.savefig(join(saveFigDir,'hist_avg_change.pdf'),dpi=300,format='pdf')
+fig.savefig(join(saveFigDir,'hist_avg_change_{}.pdf'.format(fn)),
+            dpi=300,format='pdf')
 #===plot histogram of avg. change===
 
+#===plot solution for a given data===
 la.plot_solution_for_timeStamp('1974-01-01',figsize=(18,6),
-                               saveDir=saveFigDir)
-
+                               saveDir=saveFigDir,suffix=fn)
+#===plot solution for a given data===
 
 
 

### 2018-04-21 code/volatilitytrend/algorithms/base.py

@@ -53,7 +53,7 @@ class BaseAlgorithmClass():
         #===metadata===
         
         #===load fitted values and compute fitted variance===        
-        X=np.exp(np.fromfile(filepath).reshape((gridsize,T)))
+        X=np.fromfile(filepath).reshape((gridsize,T))#???work on log(var)?
         self.fittedVar=X.copy()
         #===load fitted values and compute fitted variance===
         
@@ -69,7 +69,7 @@ class BaseAlgorithmClass():
         del X;gc.collect()                                    
         #===compute the sum of the change in variance at each location===
     
-    def plot_solution_for_timeStamp(self,date,figsize,saveDir):
+    def plot_solution_for_timeStamp(self,date,figsize,saveDir,suffix):
         #===metadata===
         lats,lons,n_rows,n_cols,dates=(self.metadata['lats'],
                                        self.metadata['lons'],
@@ -106,7 +106,7 @@ class BaseAlgorithmClass():
         map.pcolor(lons,lats,data=X,latlon=True)
         plt.colorbar(fraction=.013)   
         
-        fn=join(saveDir,'estimatedVar_in_{}.pdf'.format(date))
+        fn=join(saveDir,'estimatedVar_in_{}_{}.pdf'.format(date,suffix))
         fig.savefig(fn,dpi=300,format='pdf')
         
         
@@ -137,8 +137,9 @@ class BaseAlgorithmClass():
             
             if overlayFittedSD:
                 years=self.metadata['years']
-                plt.plot(dates,np.sqrt(self.fittedVar[idx,:]));
-                plt.plot(years,np.sqrt(self.fittedVar_yearly_avg[idx,:]));
+                plt.plot(dates,2*np.sqrt(np.exp(self.fittedVar[idx,:])/2));
+                plt.plot(years,2*np.sqrt(np.exp(self.\
+                                              fittedVar_yearly_avg[idx,:])/2));
                 
         if overlayFittedSD:
             plt.legend(['data','estimated SD',
@@ -148,7 +149,7 @@ class BaseAlgorithmClass():
                  rotation='vertical')
         fig.savefig(savepath,dpi=300,format='pdf')
         
-    def plotAvgChangeInVariance(self,saveDir):
+    def plotAvgChangeInVariance(self,saveDir,suffix):
         
         #===metadata===
         lats,lons,n_rows,n_cols=(self.metadata['lats'],
[…]

### 2018-04-22 code/fit_using_linearizedADMM.py
added figures

@@ -15,14 +15,14 @@ elif dataset=='north_hemisphere':
     data_fn='1960-01-01_to_2010-12-31_data_avg_sub_north_detrended'
     metadata_fn='1960-01-01_to_2010-12-31_metadata_avg_sub_north_detrended'
     destDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/'+\
-            'globe/Fitted_linADMM/adaptMu'
+            'globe/Fitted_linADMM'
 
 print('\014')
 la = LinearizedADMM()#construct linearizedADMM object
 la.loadData(data_fn,metadata_fn)#load data
-la.fit(destDir,lam_t_vec,lam_s_vec,maxIter=5000,freq=100,
-       ifWarmStart=False,earlyStopping=False,mu=.1,
-       ifAdaptMu=True,mu_adapt_rate=.99,mu_adapt_freq=500)
+la.fit(destDir,lam_t_vec,lam_s_vec,maxIter=5000,freq=500,
+       ifWarmStart=True,earlyStopping=True,mu=.01,
+       ifAdaptMu=False,mu_adapt_rate=.999,mu_adapt_freq=1)
 
 
 

### 2018-04-22 code/visualization.py

@@ -2,7 +2,7 @@ from volatilitytrend.algorithms.base import LinearizedADMM
 from os.path import join
 import matplotlib.pyplot as plt
 
-lam_t=2;lam_s=5;mu=.01
+lam_t=2;lam_s=0;mu=.01
 dataset='north_hemisphere'
 #dataset='us'
 
@@ -51,7 +51,7 @@ fig.savefig(join(saveFigDir,'hist_avg_change_{}.pdf'.format(fn)),
 #===plot histogram of avg. change===
 
 #===plot solution for a given data===
-la.plot_solution_for_timeStamp('1974-01-01',figsize=(18,6),
+la.plot_solution_for_timeStamp('2009-01-01',figsize=(18,6),
                                saveDir=saveFigDir,suffix=fn)
 #===plot solution for a given data===
 

### 2018-04-22 code/volatilitytrend/algorithms/linearized_admm.py

@@ -125,7 +125,7 @@ def linearizedADMM_fit(dataMat,destDir,metadata,
             if (ifWarmStart==0) or (ifWarmStartFlag==0):
                 X=initializeWithMeanVar(o2.reshape((n_cols*n_rows,T)))
                 Z=np.zeros((r_D,1));U=np.zeros((r_D,1));    
-        
+                  
             Dx=D*X        
             totalLoss=[]
             totalLoss.append(computeLoss(X,o2,D,lam_vec))
@@ -142,7 +142,6 @@ def linearizedADMM_fit(dataMat,destDir,metadata,
                     mu*=mu_adapt_rate
                     lam=1.*D_norm*mu
                     scaled_lam=lam*lam_vec 
-                    print('mu={}'.format(mu))
                 #adapt mu
                 
                 X_= X-(mu/lam)*D.transpose()*(Dx-Z+U)

### 2018-05-10 nips-2018/paper.tex
Porting to NIPS style

--- 2018-04-20 working-paper/paper.tex
@@
-\documentclass[review]{elsarticle}
+\documentclass{article}
+\PassOptionsToPackage{sort,numbers,square}{natbib}
 
-\makeatletter
-\def\ps@pprintTitle{%
- \let\@oddhead\@empty
- \let\@evenhead\@empty
- \def\@oddfoot{}%
- \let\@evenfoot\@oddfoot}
-\makeatother
-
-\usepackage{etoolbox}
-\patchcmd{\MaketitleBox}{\footnotesize\itshape\elsaddress\par\vskip36pt}{\footnotesize\itshape\elsaddress\par\parbox[b][36pt]{\linewidth}{\vfill\hfill\textnormal{\today}\hfill\null\vfill}}{}{}%
-\patchcmd{\pprintMaketitle}{\footnotesize\itshape\elsaddress\par\vskip36pt}{\footnotesize\itshape\elsaddress\par\parbox[b][36pt]{\linewidth}{\vfill\hfill\textnormal{Spring, 2018}\hfill\null\vfill}}{}{}%
-
-% use Times
-\usepackage{times}
-% For figures
-\usepackage{graphicx} % more modern
-%\usepackage{epsfig} % less modern
-\usepackage{subfigure}
+\usepackage{nips_2018}
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
 
 % For citations
-%\usepackage{natbib}
+\usepackage{natbib}
 
@@
 
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
+%\newcommand{\theHalgorithm}{\arabic{algorithm}}
+
+\usepackage{graphicx}
+\usepackage{subfigure} 
+\graphicspath{{../working-paper/}}
 
@@
 
-\bibliographystyle{model5-names}
+\usepackage[showonlyrefs]{mathtools}
+
+\renewcommand*{\figureautorefname}{Figure}%
+\renewcommand*{\tableautorefname}{Table}%
+\renewcommand*{\sectionautorefname}{Section}%
+\renewcommand*{\subsectionautorefname}{Section}%
+\renewcommand*{\subsubsectionautorefname}{Section}%
+
+\title{Modeling trend in temperature volatility using generalized LASSO}
+
+\author{Arash Khodadadi\\
+ Department of Statistics\\
+ Indiana University\\
+ Bloomington, IN 47408 \\
+ \texttt{arakhoda@indiana.edu}
+\And
+  Daniel J.\ McDonald \\
+ Department of Statistics\\
+ Indiana University\\
+ Bloomington, IN 47408 \\
+ \texttt{dajmcdon@indiana.cmu.edu}}
 
@@
 
-\begin{frontmatter}
+\maketitle
 
-\title{Modeling trend in temperature volatility using generalized LASSO}
+\begin{abstract}
+words, words,
 
-\author{Arash Khodadadi}
-\address{Department of Statistics, Indiana University, Bloomington, IN, United States}
-\emailauthor{arash.khodadadi@gmail.com}{Arash Khodadadi}
-
-\date{\today}
-
-\begin{keyword}
-generalized Lasso, linearized ADMM, consensus optimization, trend, volatility, convex optimization.
-\end{keyword}
-
-\end{frontmatter}
+\end{abstract}
 
@@
 
-There is a considerable interest in determining if there is an increasing trend in the climate variability \cite{hansen_perception_2012,huntingford_no_2013}.
+There is a considerable interest in determining if there is an increasing trend in the climate variability \citep{hansen_perception_2012,huntingford_no_2013}.
 An increase in the temperature variability will increase the probability of extreme hot outliers.
-It might be harder for the society to adapt to these extremes than to the gradual increase in the mean temperature \cite{huntingford_no_2013}.
+It might be harder for the society to adapt to these extremes than to the gradual increase in the mean temperature \citep{huntingford_no_2013}.
 
 In this project, we consider the problem of detecting the trend in the temperature volatility.
-All analyses are performed on a sub-set of the European Centre for Medium-Range Weather Forecasts (ECMWF) ERA-40 dataset \cite{uppala_era-40_2005}.
+All analyses are performed on a sub-set of the European Centre for Medium-Range Weather Forecasts (ECMWF) ERA-40 dataset \citep{uppala_era-40_2005}.
 This dataset include the temperature measurements over a grid over the earth from 1957 to 2002.
@@
 Research on analyzing the trend in the volatility of spatio-temporal data is scarce.
-\cite{hansen_perception_2012} studied the change in the standard deviation (SD) of the surface temperature in the NASA Goddard Institute for Space Studies gridded temperature data set.
+\citep{hansen_perception_2012} studied the change in the standard deviation (SD) of the surface temperature in the NASA Goddard Institute for Space Studies gridded temperature data set.
 In their analysis, for each geographical position, the mean of the temperature computed for the period 1951-1980 (called the base-period) at that position, is subtracted from the corresponding time series.
@@
 The results showed that these distributions are widen for the resent time periods compared to 1951-1980.
-\cite{huntingford_no_2013} took a similar approach in analysing the ERA-40 data set.
+\citep{huntingford_no_2013} took a similar approach in analysing the ERA-40 data set.
 However, in addition to the aforementioned method, they computed the distribution of the SDs in an alternative way: for each position and each time period, the deviation of the time-series at that position from the mean in that time period at that position was computed, and then divided by the SD of that position in the period before 1981.
-The results showed that there still is an increase in the SDs from 1958-1970 to 1991-2001, but this is much less than what is obtained from the method used in \cite{hansen_perception_2012}.
+The results showed that there still is an increase in the SDs from 1958-1970 to 1991-2001, but this is much less than what is obtained from the method used in \citep{hansen_perception_2012}.
 The authors also computed the time-evolving global SD from the de-trended time-series at each position.
@@
 
-These previous work (and other related research, e.g., \cite{rhines_frequent_2013}) have several shortcomings.
+These previous work (and other related research, e.g., \citep{rhines_frequent_2013}) have several shortcomings.
 First, no statistical analysis has been performed to examine if the change in the SD is statistically significant.
 Second, the methodologies for computing the SDs are rather arbitrary.
-The deviation of each time-series in a given period, is computed from either the mean of a base-period (as in \cite{hansen_perception_2012}), or from the given period (as in \cite{rhines_frequent_2013,huntingford_no_2013}).
+The deviation of each time-series in a given period, is computed from either the mean of a base-period (as in \citep{hansen_perception_2012}), or from the given period (as in \citep{rhines_frequent_2013,huntingford_no_2013}).
 These deviations are then normalized using the SD of the base-period or the given period.
@@
 To overcome this issue, we penalize the differences between the estimated variances of the observations which are temporally and/or spatially close to each other.
-This will result in an optimization problem called the \textit{generalized LASSO problem} \cite{tibshirani_solution_2011}.
+This will result in an optimization problem called the \textit{generalized LASSO problem} \citep{tibshirani_solution_2011}.
 As we will see, the dimension of this optimization problem is very high and so the standard methods for solving the generalized LASSO cannot be applied directly.
 We investigate two methods for solving this optimization problem.
-In the first method, we adopt an optimization technique called alternative direction method of multipliers (ADMM) \cite{boyd_distributed_2011}, to divide the total problem into several sub-problems of much lower dimension and show how the total problem can be solved by iteratively solving these sub-problems.
-The second method, called the \textit{linearized ADMM algorithm} \cite{parikh_proximal_2014} solves the main problem by iteratively solving a linearized version of it.
+In the first method, we adopt an optimization technique called alternative direction method of multipliers (ADMM) \citep{boyd_distributed_2011}, to divide the total problem into several sub-problems of much lower dimension and show how the total problem can be solved by iteratively solving these sub-problems.
+The second method, called the \textit{linearized ADMM algorithm} \citep{parikh_proximal_2014} solves the main problem by iteratively solving a linearized version of it.
 We will compare the benefits of each method.
@@
 
-Figure \ref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil).
+\autoref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil).
 The time-series of Bloomington and San Diego show clear cyclic behavior.
@@
 
-The right panel of Figure \ref{fig:bloom_detrended} shows the time-series of the temperature of Bloomington, after removing the cyclic terms and de-trending using the method explained in the next section.
+The right panel of \autoref{fig:bloom_detrended} shows the time-series of the temperature of Bloomington, after removing the cyclic terms and de-trending using the method explained in the next section.
 The goal is to investigate the trend in the variance of this signal.
@@
 Also, the cycles are not regular and their amplitude and frequency change.
-Even if one can describe the behavior of the variance of all the time-series using a single parametric model (for example a variant of the GARCH models \cite{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework.
+Even if one can describe the behavior of the variance of all the time-series using a single parametric model (for example a variant of the GARCH models \citep{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework.
 These observations motivate the need to develop a non-parametric framework for the problem at hand.
@@
 
-$\ell_1$-trend filtering was proposed by \cite{kim_$ell_1$_2009} as a solution to the following problem: a time-series $y_t$, $t=1,...,T$ is observed and we believe that it consists of a smooth trend and a stochastic component and the goal is to estimate the trend.
+$\ell_1$-trend filtering was proposed by \citep{kim_$ell_1$_2009} as a solution to the following problem: a time-series $y_t$, $t=1,...,T$ is observed and we believe that it consists of a smooth trend and a stochastic component and the goal is to estimate the trend.
 The estimated trend is desired to be smooth, but at the same time, we seek small estimated residuals (stochastic component).
@@
 
-\noindent where $||u||_1$ is the $\ell_1$ norm of the vector $u$.
-Here, $\beta$ is the vector of the free parameters of the model and there is one parameter per observation.
-The \textit{penalty matrix} is:
+ where $||u||_1$ is the $\ell_1$ norm of the vector $u$. Here, $\beta$ is the vector of the free parameters of the model and there is one parameter per observation. The \textit{penalty matrix} is:
 
@@
 
-The first term in Equation \ref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
-Specifically, with the penalty matrix \ref{eq:d_matrix}, the estimated $\beta$ will be piecewise linear \cite{kim_$ell_1$_2009}.
+The first term in \eqref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
+Specifically, with the penalty matrix \eqref{eq:d_matrix}, the estimated $\beta$ will be piecewise linear \citep{kim_$ell_1$_2009}.
 The trade-off between the two aforementioned objectives is determined by the constant $\lambda$.
-\cite{kim_$ell_1$_2009} proposed a specialized primal-dual interior point (PDIP) method for solving the optimization problem \ref{eq:l1tf} with the penalty matrix \ref{eq:d_matrix}.
+\citep{kim_$ell_1$_2009} proposed a specialized primal-dual interior point (PDIP) method for solving the optimization problem \eqref{eq:l1tf} with the penalty matrix \eqref{eq:d_matrix}.
 
-The results of applying $\ell_1$-trend filtering on the Bloomington time-series are shown in Figure \ref{fig:bloom_detrended}.
+The results of applying $\ell_1$-trend filtering on the Bloomington time-series are shown in \autoref{fig:bloom_detrended}.
 The left panel shows the original time-series (blue) together with the estimated $\beta$ (red) and the right panel shows the residuals $y_t-\beta_t$ (blue).
@@
 
-\noindent where $D$ is the same as in \ref{eq:d_matrix}.
+ where $D$ is the same as in \eqref{eq:d_matrix}.
 
@@
 
-We solve this optimization problem using the PDIP method \footnote{We use PDIP implemented in the cvxopt Python package \cite{andersen_cvxopt:_2013}.}.
-First, we note that this is a generalized LASSO problem \cite{tibshirani_solution_2011}.
+We solve this optimization problem using the PDIP method \footnote{We use PDIP implemented in the cvxopt Python package \citep{andersen_cvxopt:_2013}.}.
+First, we note that this is a generalized LASSO problem \citep{tibshirani_solution_2011}.
 The dual of a generalized LASSO with the objective $f(x)+\lambda \lVert Dx \lVert_1$ is:
@@
 
-\noindent where $D^t$ is the transpose of $D$, and $f^*(\cdot)$ is the \textit{conjugate} of $f$ defined by: $f^*(u)=\max_x u^tx-f(x)$. It is simple to show that the conjugate of $f(h)=-l(y|h)$ is:
+ where $D^t$ is the transpose of $D$, and $f^*(\cdot)$ is the \textit{conjugate} of $f$ defined by: $f^*(u)=\max_x u^tx-f(x)$. It is simple to show that the conjugate of $f(h)=-l(y|h)$ is:
 
@@
 
-The red curve in the left panel of Figure \ref{fig:bloom_estimatedSD} shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of the time-series of Bloomington.
+The red curve in the left panel of \autoref{fig:bloom_estimatedSD} shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of the time-series of Bloomington.
 To reduce the number of time-steps in this figure and in the remainder of the paper we work on the weekly averaged of the data.
@@
 Therefore, we compute the average of the estimated SD for each year.
-The estimated SD together with this annual average is shown in the middle panel of Figure \ref{fig:bloom_estimatedSD}.
+The estimated SD together with this annual average is shown in the middle panel of \autoref{fig:bloom_estimatedSD}.
 As it can be seen, the annual trend is not smooth.
-This is because in the optimization problem \ref{eq:l1tf_var}, the smoothness of the annual trend is not encouraged.
-To remedy this, we add the following long horizon penalty to \ref{eq:l1tf_var}:
+This is because in the optimization problem \eqref{eq:l1tf_var}, the smoothness of the annual trend is not encouraged.
+To remedy this, we add the following long horizon penalty to \eqref{eq:l1tf_var}:
 
@@
 
-\noindent where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
+ where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
 Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
 Since we are working on the weekly averaged data, each year corresponds to 52 observations.
-In the matrix form, the penalty \ref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
-The estimated SDs using this penalty matrix is shown in the right panel of Figure \ref{fig:bloom_estimatedSD}.
+In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
+The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
 The annual average of the estimated SDs shows a linear trend with a positive slope.
@@
 		\centerline{\includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}}
-		\caption{Left: The residuals of the time-series of Bloomington (averaged weekly) and the estimated SD obtained from the method of Section \ref{sec:l1tf_var} (red).
+		\caption{Left: The residuals of the time-series of Bloomington (averaged weekly) and the estimated SD obtained from the method of \autoref{sec:l1tf_var} (red).
 Middle: the estimated SDs (red) and their annual average (black) without imposing the long horizon penalty.
@@
 The third and fourth terms, penalize the difference between the estimated variance of two vertically and horizontally adjacent points, respectively.
-This penalty is a special case of the penalty used in the \textit{trend filtering on graphs }\cite{wang_trend_2016} (where the difference between the estimated values of the signal at each two nodes connected with an edge is penalized.).
-The optimization problem \ref{eq:l1tf_var_st} can be written in the matrix form.
+This penalty is a special case of the penalty used in the \textit{trend filtering on graphs }\citep{wang_trend_2016} (where the difference between the estimated values of the signal at each two nodes connected with an edge is penalized.).
+The optimization problem \eqref{eq:l1tf_var_st} can be written in the matrix form.
 Let $h$ be a vector whose first $T$ entries are $h_{11t}$ for $t=1,...,T$, its next $T$ entries are $h_{21t}$ and so on.
@@
 
-\noindent where $D$ was defined in \ref{eq:d_matrix}.
+ where $D$ was defined in \eqref{eq:d_matrix}.
 The number of the diagonal blocks is equal to the grid size $n_r \times n_c$.
-Each row of the matrix $D_s$ corresponds to one spatial constraint in \ref{eq:l1tf_var_st}.
+Each row of the matrix $D_s$ corresponds to one spatial constraint in \eqref{eq:l1tf_var_st}.
 For example, the first $T$ rows correspond to $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to $|h_{11t}-h_{12t}|$, and so on.
@@
 
-\noindent where $f^*$ is given in \ref{eq:conj}.
+ where $f^*$ is given in \eqref{eq:conj}.
 
@@
 For a grid over the united states and for weekly averaged data of 10 years we have $n_r=32$, $n_c=68$, $T=521$ and so $n_t\approx 3.5\times10^6$ and $n_s \approx 1.0\times10^6$.
-Therefore, the size of the optimization problem \ref{eq:dual} is about $n_t\approx 4.5\times10^6$.
-In each step of the PDIP algorithm, we need to solve a linear system of equations in the form $z=Ax$ where $A$ is a square matrix of size $2(n_t+n_s)$ (see \cite{boyd_convex_2004} equation 11.54).
-Therefore, applying the PDIP directly for solving the optimization problem \ref{eq:dual} is infeasible.
+Therefore, the size of the optimization problem \eqref{eq:dual} is about $n_t\approx 4.5\times10^6$.
+In each step of the PDIP algorithm, we need to solve a linear system of equations in the form $z=Ax$ where $A$ is a square matrix of size $2(n_t+n_s)$ (see \citep{boyd_convex_2004} equation 11.54).
+Therefore, applying the PDIP directly for solving the optimization problem \eqref{eq:dual} is infeasible.
 
 In the next section, we develop an ADMM algorithm for solving this problem.
-The idea is to cast the problem \ref{eq:dual} as a so-called \textit{consensus optimization problem} \cite{boyd_distributed_2011} and solve it by breaking the problem into smaller sub-problems using an ADMM-based algorithm.
-Next, we first give a brief overview of the consensus optimization problem and then explain how the problem \ref{eq:l1tf_var_st_mat} can be solved in this framework.
+The idea is to cast the problem \eqref{eq:dual} as a so-called \textit{consensus optimization problem} \citep{boyd_distributed_2011} and solve it by breaking the problem into smaller sub-problems using an ADMM-based algorithm.
+Next, we first give a brief overview of the consensus optimization problem and then explain how the problem \eqref{eq:l1tf_var_st_mat} can be solved in this framework.
 
@@
 We can define a set of \textit{local variables} $x_i\in \mathbb{R}^{n_i}$ such that $\sum_i f_i(z)=\sum_i f_i(x_i)$.
-We follow the notation of \cite{boyd_distributed_2011} closely.
+We follow the notation of \citep{boyd_distributed_2011} closely.
 Let $k=\mathscr{G}(i,j)$ which means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or $(x_i)_j=z_k$) and define $\tilde{z}_i \in \mathbb{R}^{n_i}$ by $(\tilde{z}_i)_j=(x_i)_j$.
@@
 
-To solve the optimization problem \ref{eq:l1tf_var_st_mat} or \ref{eq:l1tf_var_st} using the method explained in the previous section we need to two address two questions: first, how to choose the local variables $x_i$, and second, how to solve the optimization problem for updating these variables (the first line of \ref{eq:ADMM_steps}).
+To solve the optimization problem \eqref{eq:l1tf_var_st_mat} or \eqref{eq:l1tf_var_st} using the method explained in the previous section we need to two address two questions: first, how to choose the local variables $x_i$, and second, how to solve the optimization problem for updating these variables (the first line of \eqref{eq:ADMM_steps}).
 
-Obviously, the global variable in the problem \ref{eq:l1tf_var_st} is $z=h$.
+Obviously, the global variable in the problem \eqref{eq:l1tf_var_st} is $z=h$.
 The global variable can be index by the spatial coordinate and the temporal step.
-In Figure \ref{fig:data_cube}, $z$ is represented as a cube.
+In \autoref{fig:data_cube}, $z$ is represented as a cube.
 We can decompose $z$ into sub-cubes.
 An example of this decomposition is shown in the figure by white lines.
-It is easy to see that with this definition of $x_i$, the objective \ref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where $f_i(x_i)=-l(y_i|x_i)+\lVert \Lambda_i^t D_ix_i \lVert_1$.
+It is easy to see that with this definition of $x_i$, the objective \eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where $f_i(x_i)=-l(y_i|x_i)+\lVert \Lambda_i^t D_ix_i \lVert_1$.
 
@@
 
-\noindent where:
+ where:
 
@@
 
-In these equations, $\mathscr{W}$ is the \textit{Lambert function} \cite{corless_lambertw_1996}.
+In these equations, $\mathscr{W}$ is the \textit{Lambert function} \citep{corless_lambertw_1996}.
 To use the PDIP method, we also need to compute the gradient and Hessian of $f_i^*(-D_i^t\nu_i)$.
 This involves computing the derivatives of the Lambert functions.
-The primal solution $x_i$ can be obtained from the dual solution $\nu_i$ in \ref{eq:dual_x_update} by setting $x_i=w_i$ where $w_i$ is defined in the last equation of \ref{eq:conj_x_update}.
+The primal solution $x_i$ can be obtained from the dual solution $\nu_i$ in \eqref{eq:dual_x_update} by setting $x_i=w_i$ where $w_i$ is defined in the last equation of \eqref{eq:conj_x_update}.
 
-The complete ADMM algorithm for estimating the variances is represented in \ref{alg:ADMM}.
-All the computations in the three updating steps \ref{eq:ADMM_steps} can be performed in parallel.
+The complete ADMM algorithm for estimating the variances is represented in \autoref{alg:ADMM}.
+All the computations in the three updating steps \eqref{eq:ADMM_steps} can be performed in parallel.
 The number of rows and columns of the sub-cubes should be chosen so that the updating of $x_i$ could be performed in one processor.
@@
 			\FOR{$i=1$ {\bfseries to} $N_{sub-cubes}$}
-				\STATE compute $\nu_i$ from \ref{eq:dual_x_update}
-				\STATE compute $w_i$ from \ref{eq:conj_x_update}
+				\STATE compute $\nu_i$ from \eqref{eq:dual_x_update}
+				\STATE compute $w_i$ from \eqref{eq:conj_x_update}
 				\STATE set $x_i^m:=w_i$
 			\ENDFOR
-		\STATE Compute $z^m$ from  \ref{eq:ADMM_steps}
-		\STATE Compute $u_i^m$ from  \ref{eq:ADMM_steps}
+		\STATE Compute $z^m$ from  \eqref{eq:ADMM_steps}
+		\STATE Compute $u_i^m$ from  \eqref{eq:ADMM_steps}
 		\ENDFOR
@@
 
-In this section, we describe \textit{Linearized ADMM algorithm} \cite{parikh_proximal_2014} which, as we will see, makes the computation on a single machine feasible.
+In this section, we describe \textit{Linearized ADMM algorithm} \citep{parikh_proximal_2014} which, as we will see, makes the computation on a single machine feasible.
 
@@
 
-\noindent where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each iteration of the linearized ADMM algorithm for solving this problem has the following form:
+ where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each iteration of the linearized ADMM algorithm for solving this problem has the following form:
 
@@
 \begin{aligned}
-x^{k+1} & := \textbf{prox}_{\mu f} \big(x^k - (\mu/\rho)D^T (D x^k − z^k + u^k )\big)\\
+x^{k+1} & := \textbf{prox}_{\mu f} \big(x^k - (\mu/\rho)D^T (D x^k - z^k + u^k )\big)\\
 z^{k+1} & := \textbf{prox}_{\rho g} \big(D x^k + u^k\big)\\
@@
 
-\noindent where $z,u\in \mathbb{R}^m$ and the \textit{proximal operator} is defined as follows:
+ where $z,u\in \mathbb{R}^m$ and the \textit{proximal operator} is defined as follows:
 
@@
 
-This algorithm belongs to the general class of \textit{proximal algorithms} for solving convex optimization problems. For more details about these algorithms see \cite{parikh_proximal_2014}.
+This algorithm belongs to the general class of \textit{proximal algorithms} for solving convex optimization problems. For more details about these algorithms see \citep{parikh_proximal_2014}.
 
-The optimization problem \ref{eq:l1tf_var_st} can be put into the form \ref{eq:linADMM_opt} as follows:
+The optimization problem \eqref{eq:l1tf_var_st} can be put into the form \eqref{eq:linADMM_opt} as follows:
 
@@
 
-\noindent where $y_k$ is the $k^{th}$ entry of the vector whose entries are $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector $\Lambda^t=(\lambda_t\textbf{e}_{n_t}^t|\lambda_s\textbf{e}_{n_s}^t)$ (see section \ref{sec:exten}).
+ where $y_k$ is the $k^{th}$ entry of the vector whose entries are $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector $\Lambda^t=(\lambda_t\textbf{e}_{n_t}^t|\lambda_s\textbf{e}_{n_s}^t)$ (see \autoref{sec:exten}).
 
-To perform the steps in \ref{eq:linADMM_steps}, we need to evaluate $\textbf{prox}_{\mu f}$ and $\textbf{prox}_{\rho g}$.
+To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate $\textbf{prox}_{\mu f}$ and $\textbf{prox}_{\rho g}$.
 The proximal algorithms are feasible only if these proximal operators can be evaluated efficiently which, as we show next, is the case for our problem.
@@
 Let $\big(\textbf{prox}_{\mu f}(u)\big)_k$ be the $k^{th}$ entry of $\textbf{prox}_{\mu f}(u)$.
-From the \textit{separable sum} property of the proximal operators we have (see \cite{parikh_proximal_2014}, section 2.1):
+From the \textit{separable sum} property of the proximal operators we have (see \citep{parikh_proximal_2014}, section 2.1):
 
@@
 
-\noindent where $\mathscr{W}$ is the Lambert function.
+where $\mathscr{W}$ is the Lambert function.
 
@@
 
-Therefore, both proximal operators in \ref{eq:linADMM_steps} can be evaluated easily and so we can use the linearized ADMM algorithm to solve the optimization problem \ref{eq:l1tf_var_st}.
+Therefore, both proximal operators in \eqref{eq:linADMM_steps} can be evaluated easily and so we can use the linearized ADMM algorithm to solve the optimization problem \eqref{eq:l1tf_var_st}.
 
 \section{Results}
-As it was mentioned before, the algorithm proposed in Section \ref{sec:consOpt} is appropriate only if we parallelize it over multiple machines and so we do not pursue it further here.
-All the results reported in this section are obtained using the linearize ADMM algorithm of Section \ref{sec:linADMM}.
+As it was mentioned before, the algorithm proposed in \autoref{sec:consOpt} is appropriate only if we parallelize it over multiple machines and so we do not pursue it further here.
+All the results reported in this section are obtained using the linearize ADMM algorithm of \autoref{sec:linADMM}.
 We applied this algorithm on a subset of the ERA-40 dataset.
@@
 
-Figure \ref{fig:loss_vs_iter} shows the value of the loss function \ref{eq:l1tf_var_st_mat} as a function of iteration for $\lambda_t=1$ and $\lambda_s=0$.
+\autoref{fig:loss_vs_iter} shows the value of the loss function \eqref{eq:l1tf_var_st_mat} as a function of iteration for $\lambda_t=1$ and $\lambda_s=0$.
 We used the following rule to determine when to stop the optimization: the optimization was stopped if the value of the loss did not improve by at least 0.1\% in 1000 trials.
@@
 However, such analyses needs the computation of the degrees of freedom (df).
-Several previous work have investigated the df in generalized lasso problems \cite{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}.
+Several previous work have investigated the df in generalized lasso problems \citep{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}.
 However, all these studies have considered the linear regression problem and, to the best of our knowledge, the problem of computing the df for generalized lasso with general objective function has not been considered yet.
 
-Another approach is to choose the set of values which minimize an estimate of the expected prediction error obtained by k-fold cross-validation \cite{tibshirani_regression_1996} .
+Another approach is to choose the set of values which minimize an estimate of the expected prediction error obtained by k-fold cross-validation \citep{tibshirani_regression_1996} .
 Although this method is applicable for our problem, it needs k times more computation.
@@
 We computed the optimal solution for all the combinations of the following sets of values: $\lambda_t \in \{1,5,10,20\} \, \, , \lambda_s \in \{0,.1,1,5,10\}$.
-The value of $\mathscr{L}(\lambda_t,\lambda_s)$ is shown in Figure \ref{fig:modelSelect}.
+The value of $\mathscr{L}(\lambda_t,\lambda_s)$ is shown in \autoref{fig:modelSelect}.
 This function is minimized at $\lambda_t=5$ and $\lambda_s=1$.
@@
 \subsection{Analysis of trend of temperature volatility}
-Figure \ref{fig:estimatedSD} shows the detrended data, the estimated standard deviation and the yearly average of these estimates for two cities in the united states: Bloomington (left) and San Diego (right).
+\autoref{fig:estimatedSD} shows the detrended data, the estimated standard deviation and the yearly average of these estimates for two cities in the united states: Bloomington (left) and San Diego (right).
 The estimated SD captures the periodic behavior in the variance of the time-series.
@@
 To determine how the volatility has changed in each location, we subtract the average of the estimated variance in 1992 from the average in the following years and compute their sum.
-The value of this change in the variance in each location is depicted in the right panel of Figure \ref{fig:avg_change_estimatedSD}.
+The value of this change in the variance in each location is depicted in the right panel of \autoref{fig:avg_change_estimatedSD}.
 The left panel of this figure, shows the average estimated variance in each location.
@@
 
-Finally, Figure \ref{fig:hist_avg_change} shows the histogram of the changes shown in the right panel of Figure \ref{fig:avg_change_estimatedSD}.
+Finally, \autoref{fig:hist_avg_change} shows the histogram of the changes shown in the right panel of \autoref{fig:avg_change_estimatedSD}.
 As we can see, there are two peaks in this histogram: one for locations with no trend (which includes the points on the oceans), and one peak for locations with positive trend.
@@
 
-\section*{References}
+\bibliographystyle{abbrvnat}
+\small
 \bibliography{references}

### 2018-05-10 code/AnalyseSyntheticData.py
added synthetic data analysis

@@ -1,0 +1,107 @@
+import numpy as np
+from volatilitytrend.data_utils.simulate_data import simulateSpatioTemporalData
+from os.path import join
+from volatilitytrend.algorithms.base import LinearizedADMM
+import matplotlib.pyplot as plt
+
+bl=[False,True]
+ifSimulateData=bl[0]
+ifFitModel=bl[1]
+ifVisualize=bl[0]
+ifModelSelection=bl[0]
+
+dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/Simulated data'
+
+n_rows=5;n_cols=7;
+T=1000#time-series length
+gridsize=n_rows*n_cols
+
+###################
+###simulate data###
+###################
+if ifSimulateData:
+    
+    #===specify sources properties===
+    var_sources_centers=np.array([[0,0],[0,5],[3,0],[3,5]])
+    var_sources_sigma=[5.]*4
+    #===specify sources properties===
+    
+    #===compute the weight of each source at wach time===
+    times=np.arange(0,T)
+    var_sources_weight1=np.exp(np.sin(10*np.pi*times/float(T)))
+    var_sources_weight2=np.exp(np.sin(10*np.pi*times/float(T)))
+    var_sources_weight3=np.exp(np.cos(10*np.pi*times/float(T)))
+    var_sources_weight4=np.exp(np.cos(10*np.pi*times/float(T)))
+    var_sources_weight_mat=np.concatenate((var_sources_weight1,
+                                           var_sources_weight2,
+                                           var_sources_weight3,
+                                           var_sources_weight4)).\
+                                           reshape((4,-1))
+    #===compute the weight of each source at wach time===
+    
+    simulateSpatioTemporalData(dstDir,n_rows,n_cols,T,var_sources_centers,
+                                   var_sources_weight_mat,var_sources_sigma)
+
+
+
+###################
+###Fit model###
+###################
+lam_t_vec=[0,1,5,10,50,100];lam_s_vec=[0.0,.05,.1,.2,.3];mu=.01
+#lam_t_vec=[10];lam_s_vec=[.05];mu=.01
+
+la = LinearizedADMM()#construct linearizedADMM object
+data_fn=join(dstDir,'simulated_data')
[…]

### 2018-05-10 code/fit_using_linearizedADMM.py

@@ -8,19 +8,19 @@ lam_s_vec=[0.5]
 if dataset=='us':
     data_fn='1992-08-31_to_2002-08-31_data_avg_us_detrended'
     metadata_fn='1992-08-31_to_2002-08-31_metadata_avg_us_detrended'
-    destDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/'+\
+    dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/'+\
     '19920831to20020831/Fitted_LinADMM_test'
     
 elif dataset=='north_hemisphere':
     data_fn='1960-01-01_to_2010-12-31_data_avg_sub_north_detrended'
     metadata_fn='1960-01-01_to_2010-12-31_metadata_avg_sub_north_detrended'
-    destDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/'+\
+    dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/'+\
             'globe/Fitted_linADMM'
 
 print('\014')
 la = LinearizedADMM()#construct linearizedADMM object
 la.loadData(data_fn,metadata_fn)#load data
-la.fit(destDir,lam_t_vec,lam_s_vec,maxIter=10000,freq=500,
+la.fit(dstDir,lam_t_vec,lam_s_vec,maxIter=10000,freq=500,
        ifWarmStart=True,earlyStopping=True,mu=.01,
        ifAdaptMu=False,mu_adapt_rate=.999,mu_adapt_freq=1)
 

### 2018-05-10 code/volatilitytrend/algorithms/base.py

@@ -37,8 +37,8 @@ class BaseAlgorithmClass():
         #===load metadata===
 
         #===load data===
-        self.dataMat=np.fromfile(join(self.dataDir,data_fn),dtype='float32').\
-                            reshape((n_rows*n_cols,T))
+        self.dataMat=np.fromfile(join(self.dataDir,data_fn),
+                                 dtype='float32').reshape((n_rows*n_cols,T))
         #===load data===
         
     def analyseFittedValues(self,filepath):
@@ -207,7 +207,7 @@ class LinearizedADMM(BaseAlgorithmClass):
 
     def fit(self,destDir,lam_t_vec,lam_s_vec,
             mu=.01,maxIter=40000,freq=100,
-            ifWarmStart=True,lh_trend=True,
+            ifWarmStart=True,lh_trend=True,wrapAround=True,
             earlyStopping=True,patience=2,tol=.1,
             ifAdaptMu=False,mu_adapt_rate=.95,mu_adapt_freq=100):
         
@@ -216,6 +216,7 @@ class LinearizedADMM(BaseAlgorithmClass):
                            lam_t_vec,lam_s_vec,mu=mu,
                            maxIter=maxIter,freq=freq,
                            ifWarmStart=ifWarmStart,lh_trend=lh_trend,
+                           wrapAround=wrapAround,
                            earlyStopping=earlyStopping,
                            patience=patience,tol=tol,
                            ifAdaptMu=ifAdaptMu,mu_adapt_rate=mu_adapt_rate,

### 2018-05-10 code/volatilitytrend/algorithms/linearized_admm.py

@@ -78,7 +78,7 @@ def initializeWithMeanVar(o2_mat):
 def linearizedADMM_fit(dataMat,destDir,metadata,
                        lam_t_vec,lam_s_vec,mu=.01,
                        maxIter=40000,freq=100,
-                       ifWarmStart=True,lh_trend=True,
+                       ifWarmStart=True,lh_trend=True,wrapAround=True,
                        earlyStopping=True,patience=2,tol=.1,
                        ifAdaptMu=False,mu_adapt_rate=.95,mu_adapt_freq=100):
     
@@ -97,21 +97,23 @@ def linearizedADMM_fit(dataMat,destDir,metadata,
     lam_s_vec=[float(lam_s) for lam_s in lam_s_vec]
     mu=float(mu)
     #===make sure the parameters are float===   
-    
-    #===compute D and lam_vec===
-    D,lam_vec=compute_D_and_lam(n_rows,n_cols,T,lam_t,lam_s,lh_trend=lh_trend)
-    r_D,c_D=D.shape
-    D_norm=norm(D)    
-    lam=1.*D_norm*mu
-    scaled_lam=lam*lam_vec    
-    #===compute D and lam_vec===
 
     #===compute solution for all lam_t and lam_s=== 
     for i_t,lam_t in enumerate(lam_t_vec):
-        for i_s,lam_s in enumerate(lam_s_vec):
+        for i_s,lam_s in enumerate(lam_s_vec):            
             print('\014')
             print('\n'+ctime()+'...fitting model with lam_t=%.2f, lam_s=%.2f'%\
                   (lam_t,lam_s))
+
+            #===compute D and lam_vec===
+            D,lam_vec=compute_D_and_lam(n_rows,n_cols,T,lam_t,lam_s,
+                                        lh_trend=lh_trend,
+                                        wrapAround=wrapAround)
+            r_D,c_D=D.shape
+            D_norm=norm(D)    
+            lam=1.*D_norm*mu
+            scaled_lam=lam*lam_vec    
+            #===compute D and lam_vec===
             
             #results filename 
             result_fn = 'mu_'+str(mu)+'_lam_t_'+str(lam_t)+\
@@ -122,7 +124,7 @@ def linearizedADMM_fit(dataMat,destDir,metadata,
             if ifWarmStart:
                 X,Z,U,ifWarmStartFlag=warmStart(i_t,i_s,lam_t_vec,lam_s_vec,
[…]

### 2018-05-10 code/volatilitytrend/data_utils/simulate_data.py

@@ -1,0 +1,93 @@
+#from .. import config
+from os.path import join
+import numpy as np
+from scipy import stats
+from cPickle import dump
+import pandas as pd
+
+def computeCovarianceMatrix(n_rows,n_cols,
+                            var_sources_centers,var_sources_weight,
+                            var_sources_sigma):
+    '''
+    This function computes the covariance matrix at each time.
+    The covariance matrix is computed as a weighted sum of several
+    Guassian functions at different locations on the grid and with 
+    different variances. We call each Gaussian a "variance source".
+    
+    Parameters
+    ---------    
+    n_rows: float
+        Number of rows in the grid.
+        
+    n_cols: float
+        Number of columns in the grid.
+
+    var_sources_centers: numpy.array
+        This is a n_sources x 2 array where each row is the center 
+        of a source (the mean of the Gaussian). The first column is 
+        the row number and the second column is the column number
+        of the centers.
+    
+    var_sources_weight: array_like
+        Each element is the weight of a Gaussian source.
+    
+    var_sources_sigma: array_like
+        Each element is the SD of a Gaussian source.
+    
+        
+    Returns
+    -------
+    covarMat: numpu.array
+        The covariance matrix at a time. This will have the same size as
+        the grid.
+        
+    '''
+    
+    n_sources=var_sources_centers.shape[0]
+    covarMat=np.zeros((n_rows,n_cols))
+    
+    Y,X=np.meshgrid(np.arange(n_cols),np.arange(n_rows))
+
+    for s in range(n_sources):
+        center=var_sources_centers[s,];        
+        weight=var_sources_weight[s]
+        sigma=var_sources_sigma[s]
+        
+        dist = (X-center[0])**2 + (Y-center[1])**2
+        covarMat=covarMat + weight*sigma*np.sqrt(2*np.pi)*\
+                            stats.norm.pdf(dist,0,sigma)   
+                
+    return covarMat
+
+
+def simulateSpatioTemporalData(dstDir,n_rows,n_cols,T,var_sources_centers,
+                               var_sources_weight_mat,var_sources_sigma):
+    
+    times=np.arange(0,T)
+    simulated_data=np.zeros((n_rows*n_cols,T))
+    covarMat=np.zeros((n_rows*n_cols,T))
+    for t in times:
+        covarMat[:,[t]]=computeCovarianceMatrix(n_rows,n_cols,
+                                                var_sources_centers,
[…]

### 2018-05-11 code/AnalyseSyntheticData.py
updates on synthetic

@@ -6,15 +6,27 @@ import matplotlib.pyplot as plt
 
 bl=[False,True]
 ifSimulateData=bl[0]
-ifFitModel=bl[1]
-ifVisualize=bl[0]
+ifFitModel=bl[0]
+ifVisualize=bl[1]
 ifModelSelection=bl[0]
 
 dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/Simulated data'
 
 n_rows=5;n_cols=7;
-T=1000#time-series length
+n_yeras=15
+T=n_yeras*52#time-series length
 gridsize=n_rows*n_cols
+
+def computeChangeInVariance(X,n_rows,n_cols,T):
+    '''
+    this is also implemented in 'analyseFittedValues' but there 
+    it is computed for log(var) instead of variance itself.
+    '''    
+    n_years=T/52;gridsize=n_rows*n_cols 
+    X=X[:,0:n_years*52].reshape((n_years*gridsize,52))
+    X=np.diff(np.mean(X,axis=1).reshape((gridsize,n_years)),axis=1)
+    changeInVar=np.sum(X,axis=1).reshape((n_rows,n_cols),order='F')
+    return changeInVar
 
 ###################
 ###simulate data###
@@ -28,10 +40,15 @@ if ifSimulateData:
     
     #===compute the weight of each source at wach time===
     times=np.arange(0,T)
-    var_sources_weight1=np.exp(np.sin(10*np.pi*times/float(T)))
-    var_sources_weight2=np.exp(np.sin(10*np.pi*times/float(T)))
-    var_sources_weight3=np.exp(np.cos(10*np.pi*times/float(T)))
-    var_sources_weight4=np.exp(np.cos(10*np.pi*times/float(T)))
+    trend1=.5*times/float(T)
+    trend2=.1*times/float(T)
+    trend3=-.5*times/float(T)
+    trend4=-.1*times/float(T)
+    
+    var_sources_weight1=np.exp(np.sin(n_yeras*2*np.pi*times/float(T)))+trend1
+    var_sources_weight2=np.exp(np.sin(n_yeras*2*np.pi*times/float(T)))+trend2
+    var_sources_weight3=np.exp(np.cos(n_yeras*2*np.pi*times/float(T))+trend3)
+    var_sources_weight4=np.exp(np.cos(n_yeras*2*np.pi*times/float(T))+trend4)
     var_sources_weight_mat=np.concatenate((var_sources_weight1,
[…]

### 2018-05-11 code/volatilitytrend/algorithms/base.py

@@ -64,8 +64,8 @@ class BaseAlgorithmClass():
         #===compute yearly average of fitted variance===
         
         #===compute the sum of the change in variance at each location===
-        self.changeInVar=np.mean(X-np.tile(X[:,[0]],(1,n_years)),axis=1).\
-                            reshape((n_rows,n_cols),order='F')
+        self.changeInVar=np.sum(np.diff(X,axis=1),axis=1).\
+                        reshape((n_rows,n_cols),order='F')
         del X;gc.collect()                                    
         #===compute the sum of the change in variance at each location===
     

### 2018-05-11 code/AnalyseSyntheticData.py
added consensusADMM

@@ -1,18 +1,18 @@
 import numpy as np
 from volatilitytrend.data_utils.simulate_data import simulateSpatioTemporalData
 from os.path import join
-from volatilitytrend.algorithms.base import LinearizedADMM
+from volatilitytrend.algorithms.base import LinearizedADMM,ConsensusADMM
 import matplotlib.pyplot as plt
 
 bl=[False,True]
-ifSimulateData=bl[0]
+ifSimulateData=bl[1]
 ifFitModel=bl[0]
-ifVisualize=bl[1]
+ifVisualize=bl[0]
 ifModelSelection=bl[0]
 
 dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/Simulated data'
 
-n_rows=5;n_cols=7;
+n_rows=6;n_cols=8;
 n_yeras=15
 T=n_yeras*52#time-series length
 gridsize=n_rows*n_cols

### 2018-05-11 code/test.py

@@ -1,0 +1,44 @@
+import numpy as np
+from os.path import join
+from volatilitytrend.algorithms.base import ConsensusADMM
+import matplotlib.pyplot as plt
+
+bl=[False,True]
+ifSimulateData=bl[0]
+ifFitModel=bl[1]
+ifVisualize=bl[1]
+ifModelSelection=bl[0]
+
+dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/Simulated data'
+
+
+lam_t_vec=[5.];
+lam_s_vec=[0.1];
+rho=.015
+
+ca = ConsensusADMM()#construct linearizedADMM object
+data_fn=join(dstDir,'simulated_data')
+metadata_fn=join(dstDir,'metadata')
+ca.loadData(data_fn,metadata_fn)#load data
+
+if ifFitModel: 
+    ca.fit(dstDir,lam_t_vec,lam_s_vec,rho=rho,maxIter=100,freq=10,
+           n_r_b=3,n_c_b=3,lh_trend=True,wrapAround=False,
+           earlyStopping=False,patience=2,tol=.1)
+    
+if ifVisualize:
+    lam_t=5.;lam_s=.1;
+    
+    rhp,lam_t,lam_s=(float(rho),float(lam_t),float(lam_s))
+    fn='rho_{}_lam_t_{}_lam_s_{}'.format(rho,lam_t,lam_s)
+    
+    #===load fitted values===
+    n_rows=ca.metadata['n_rows'];n_cols=ca.metadata['n_cols']
+    gridsize=n_rows*n_cols
+    
+    covMat = np.fromfile(join(dstDir,'covMat')).reshape((gridsize,-1))
+    filepath=join(dstDir,'Z_'+fn)
+    ca.analyseFittedValues(filepath)
+    plt.figure()
+    i=0;plt.plot(np.exp(ca.fittedVar[i,]/2));plt.plot(covMat[i,:])
+    #===load fitted values===

### 2018-05-11 code/volatilitytrend/algorithms/base.py

@@ -8,6 +8,7 @@ import pandas as pd
 
 from .. import config
 from .linearized_admm import linearizedADMM_fit
+from .consensus_admm import consensusADMM_fit
 from ..utils import latlon_to_rowcol
 
 __metaclass__=type
@@ -223,7 +224,23 @@ class LinearizedADMM(BaseAlgorithmClass):
                            mu_adapt_freq=mu_adapt_freq)
         
     
-    
+class ConsensusADMM(BaseAlgorithmClass):
+
+    def fit(self,destDir,
+            lam_t_vec,lam_s_vec,rho=.1,
+            n_r_b=2,n_c_b=2,
+            maxIter=1000,freq=100,
+            lh_trend=True,wrapAround=True,
+            earlyStopping=True,patience=2,tol=.1):
+        
+        
+        consensusADMM_fit(self.dataMat,destDir,self.metadata,
+                          lam_t_vec,lam_s_vec,rho=rho,
+                          n_r_b=n_r_b,n_c_b=n_c_b,
+                          maxIter=maxIter,freq=freq,
+                          lh_trend=lh_trend,wrapAround=wrapAround,
+                          earlyStopping=earlyStopping,
+                          patience=patience,tol=tol)
     
     
     

### 2018-05-11 code/volatilitytrend/algorithms/consensus_admm.py

@@ -1,0 +1,531 @@
+from scipy.special import lambertw
+from cvxopt import solvers,matrix,spmatrix,sparse,spdiag,div,mul,exp
+
+import numpy as np
+import gc
+from time import ctime
+from os.path import join
+
+
+def compute_cp_matrices(n_rows,n_cols,T,lam_t,lam_s,lh_trend=True,
+                        ifCompute_Gh=False,wrapAround=True):
+    '''
+    This computes the matrices used in cp optimization.
+    '''
+    
+    grid_size=n_rows*n_cols
+    if grid_size>200:
+        print(ctime()+'...computing optimization matrices...')
+
+    if wrapAround:#no. of spatial constraints
+        r_s=T*(2*n_rows*n_cols-n_rows);
+    else:
+        r_s=T*(2*n_rows*n_cols-n_rows-n_cols);
+    r_t=n_rows*n_cols*(T-2);#no. of temporal constraints
+    
+    #===form matrix D===
+    #---spatial penalty---
+    I_s=[];J_s=[];x_s=[];
+
+    for c in range(n_cols):
+        for r in range(n_rows):
+            
+            #---determine the neighbors of the current point---
+            if ((r<(n_rows-1)) & (c<(n_cols-1))):
+                r_n=[r+1,r];c_n=[c,c+1]
+            elif ((r==(n_rows-1)) & (c<(n_cols-1))):
+                r_n=[r];c_n=[c+1]
+            elif (c==(n_cols-1)):
+                if wrapAround:
+                    r_n=[r+1,r];c_n=[c,0]
+                else:
+                    r_n=[r+1];c_n=[c]
+                if (r==(n_rows-1)):
+                    continue
+                    
+            idx_n=np.ravel_multi_index((r_n,c_n),
+                                       dims=(n_rows,n_cols),order='F')
+
+            idx=np.ravel_multi_index((r,c),
+                                       dims=(n_rows,n_cols),order='F')               
+            #---determine the neighbors of the current point---
+            
+            #---add indices corresponding to current point and its neighbors---
+            for i in idx_n:
[…]

### 2018-05-11 code/volatilitytrend/utils.py

@@ -12,8 +12,11 @@ def compute_D_and_lam(n_rows,n_cols,T,lam_t,lam_s,lh_trend=True,
     if grid_size>200:
         print(ctime()+'...computing optimization matrices...')
 
-        
-    r_s=T*(2*n_rows*n_cols-n_rows-n_cols);#no. of spatial constraints
+    if wrapAround:#no. of spatial constraints
+        r_s=T*(2*n_rows*n_cols-n_rows);    
+    else:
+        r_s=T*(2*n_rows*n_cols-n_rows-n_cols);        
+    
     r_t=n_rows*n_cols*(T-2);#no. of temporal constraints
     
     #===form matrix D===

### 2018-05-12 code/AnalyseSyntheticData_conADMM.py
updates on synthetic data analysis

@@ -5,16 +5,18 @@ import matplotlib.pyplot as plt
 
 bl=[False,True]
 ifSimulateData=bl[0]
-ifFitModel=bl[1]
+ifFitModel=bl[0]
 ifVisualize=bl[1]
 ifModelSelection=bl[0]
 
 dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/Simulated data'
-
+savedResultsDir='/home/arash/datasets/ecmwf/Fits_to_simulated_data/'+\
+'LinADMM_withTrend'
+saveFigDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Reports/051218/Figures'
 
 lam_t_vec=[5.];
 lam_s_vec=[0.1];
-rho=.015
+rho=.013
 
 ca = ConsensusADMM()#construct linearizedADMM object
 data_fn=join(dstDir,'simulated_data')
@@ -22,7 +24,7 @@ metadata_fn=join(dstDir,'metadata')
 ca.loadData(data_fn,metadata_fn)#load data
 
 if ifFitModel: 
-    ca.fit(dstDir,lam_t_vec,lam_s_vec,rho=rho,maxIter=100,freq=10,
+    ca.fit(dstDir,lam_t_vec,lam_s_vec,rho=rho,maxIter=400,freq=40,
            n_r_b=3,n_c_b=3,lh_trend=True,wrapAround=False,
            earlyStopping=False,patience=2,tol=.1)
     
@@ -41,4 +43,27 @@ if ifVisualize:
     ca.analyseFittedValues(filepath)
     plt.figure()
     i=0;plt.plot(np.exp(ca.fittedVar[i,]/2));plt.plot(covMat[i,:])
+    plt.xlabel('time');plt.ylabel('standard deviation')
+    plt.legend(['estimated','true'])
     #===load fitted values===
+    
+    #===compare loss-iteration for lin and cons ADMM===
+    savedResultsDir='/home/arash/datasets/ecmwf/Fits_to_simulated_data/'+\
+    'LinADMM_withTrend'
+    filepath=join(savedResultsDir,'loss_'+fn)
+    loss_cons = np.fromfile(filepath).reshape((2,-1))
+    
+    fn='mu_{}_lam_t_{}_lam_s_{}'.format(.01,lam_t,lam_s)
+    filepath=join(savedResultsDir,'loss_'+fn)
+    loss_lin = np.fromfile(filepath).reshape((2,-1))
[…]

### 2018-05-12 code/AnalyseSyntheticData_linADMM.py

@@ -5,14 +5,17 @@ from volatilitytrend.algorithms.base import LinearizedADMM,ConsensusADMM
 import matplotlib.pyplot as plt
 
 bl=[False,True]
-ifSimulateData=bl[1]
+ifSimulateData=bl[0]
 ifFitModel=bl[0]
-ifVisualize=bl[0]
+ifVisualize=bl[1]
 ifModelSelection=bl[0]
 
 dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/Simulated data'
+savedResultsDir='/home/arash/datasets/ecmwf/Fits_to_simulated_data/'+\
+'LinADMM_withTrend'
+saveFigDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Reports/051218/Figures'
 
-n_rows=6;n_cols=8;
+n_rows=5;n_cols=7;
 n_yeras=15
 T=n_yeras*52#time-series length
 gridsize=n_rows*n_cols
@@ -64,9 +67,9 @@ if ifSimulateData:
 ###################
 ###Fit model###
 ###################
-lam_t_vec=[100.];
+lam_t_vec=[5.];
 lam_s_vec=[0.0,.05,.1,.2,.3];
-#lam_s_vec=[0.1];
+lam_s_vec=[0.1];
 mu=.01
 #lam_t_vec=[10];lam_s_vec=[.05];mu=.01
 
@@ -76,9 +79,9 @@ metadata_fn=join(dstDir,'metadata')
 la.loadData(data_fn,metadata_fn)#load data
 
 if ifFitModel:   
-    la.fit(dstDir,lam_t_vec,lam_s_vec,maxIter=15000,freq=500,
+    la.fit(dstDir,lam_t_vec,lam_s_vec,maxIter=40000,freq=500,
            lh_trend=True,wrapAround=False,patience=2,
-           ifWarmStart=False,earlyStopping=True,mu=mu,
+           ifWarmStart=False,earlyStopping=False,mu=mu,
            ifAdaptMu=False,mu_adapt_rate=.999,mu_adapt_freq=1)
 
 
@@ -87,8 +90,9 @@ if ifFitModel:
 ###################
 
 if ifVisualize:
-    lam_t=lam_t_vec[-1];lam_s=lam_s_vec[2]
-    lam_t=5.;lam_s=.2;
[…]

### 2018-05-12 code/fit_using_linearizedADMM.py

@@ -2,8 +2,8 @@ from volatilitytrend.algorithms.base import LinearizedADMM
 
 dataset='north_hemisphere'
 #dataset='us'
-lam_t_vec=[2,4,8,10,15]
-lam_s_vec=[0.5]
+lam_t_vec=[4]
+lam_s_vec=[0,.1,.5,2,5,10]
 
 if dataset=='us':
     data_fn='1992-08-31_to_2002-08-31_data_avg_us_detrended'
@@ -20,7 +20,7 @@ elif dataset=='north_hemisphere':
 print('\014')
 la = LinearizedADMM()#construct linearizedADMM object
 la.loadData(data_fn,metadata_fn)#load data
-la.fit(dstDir,lam_t_vec,lam_s_vec,maxIter=10000,freq=500,
+la.fit(dstDir,lam_t_vec,lam_s_vec,maxIter=10000,freq=500,wrapAround=True,
        ifWarmStart=True,earlyStopping=True,mu=.01,
        ifAdaptMu=False,mu_adapt_rate=.999,mu_adapt_freq=1)
 

### 2018-05-12 code/volatilitytrend/algorithms/consensus_admm.py

@@ -341,7 +341,7 @@ def x_update(Y,D,G,h,Kinv,alpha,rho):
         
         return f, Df, H 
 
-    solvers.options['maxiters']=250;solvers.options['show_progress']=False
+    solvers.options['maxiters']=500;solvers.options['show_progress']=False
     sol=solvers.cp(F=F,G=G,h=h)
     v=sol['x'];#dual solution
     x=compute_primal(Y,v,D,matrix(Kinv),alpha,rho)#primal solution
@@ -502,7 +502,7 @@ def consensusADMM_fit(dataMat,destDir,metadata,
                 W=W+X-Z_hat
                 #---w-update---
                 
-                if  (it+1)%freq==0:
+                if ((it==0) or ((it+1)%freq==0)):
                     totalLoss.append(computeLoss(Z,Y,D,lam_t,lam_s,
                                                  T,n_rows,n_cols,lh_trend))
                     print(ctime()+'          %5.0f          %.1f'\

### 2018-05-12 working-paper/synthetic data/report_AKH_071017.tex
added report on synthetic data analysis

\documentclass[12pt]{article}

\pdfoutput=1
\usepackage{hyperref}
\usepackage{epstopdf}
%\usepackage{fullpage}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{color}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage{boondox-cal}%for calligraphic font
%\usepackage{kbordermatrix}
%\usepackage{algpseudocode}
%\usepackage{algorithm}

%\usepackage{apacite}

\title{Synthetic data}
\author{Arash Khodadadi}

\begin{document}
\maketitle

\section{Introduction}

To analyze the proposed optimization methods, we first fit the model on some synthetic spatio-temporal data. The observations at all time steps and all locations were generated from independent Gaussian random variables with zero mean. However, the variance of these random variables changes smoothly in time and space. Specifically, the observation at time step $t$ and at location $(r,c)$ on the grid is generated from a Gaussian $N(0,\sigma^2(t,r,c))$, where:

\begin{equation}
\begin{aligned}
\sigma(t,r,c) & =\sum_{s=1}^{S} W_s(t) \cdot \exp\bigg( \frac{(r-r_s)^2+(c-c_s)^2}{2\sigma_s^2} \bigg) \\
W_s(t) & =\alpha_s \cdot t + \exp(\sin(2\pi\omega_s t+\phi_s)) 
\label{eq:sourceVar}
\end{aligned}
\end{equation}

In words, the variance at each time and location is computed as the weighted sum of $S$ bell-shaped functions where the weights are time-varying functions. Specifically, the weights consist of a linear trend $\alpha_s \cdot t$ and a periodic term $\beta_s \cdot \sin(2\pi\omega_s t+\phi_s)$. The bell-shaped functions impose the spatial smoothness, and the linear trend and the periodic terms enforce the temporal smoothness. We simulated the data on a 5 by 7 grid and for 780 time steps. We also used $S=4$. The parameters of the variance function are shown in Table \ref{tab:sim_params}. The value of the variance function for all locations on the grid and at $t=25$ and $t=45$ is shown in Figure \ref{fig:true_var_spatial}. Also, the variance for all time steps and at the location (0,0) on the grid is shown in Figure \ref{fig:true_fitted_var}. The linear trend and the period term can be seen clearly in this figure.

\begin{table}[t]
	\caption{Parameters used to simulate data}
	\label{tab:sim_params}
	\vskip 0.15in
	\begin{center}
		\begin{small}
			\begin{sc}
				\begin{tabular}{ccccccc}
					\hline

					$s$ & $r_s$ & $c_s$ & $\sigma_s$ &$\alpha_s$ & $\omega_s$ & $\phi_s$\\
					\hline
					1 & 0 & 0 & 5 & 0.5 & 0.121 & 0 \\
					2 & 0 & 5 & 5 & 0.1 & 0.121 & 0 \\
					3 & 3 & 0 & 5 & -0.5 & 0.121 & $\pi/2$ \\
					4 & 3 & 5 & 5 & -0.1 & 0.121 & $\pi/2$ \\
					\hline
				\end{tabular}
			\end{sc}
		\end{small}
	\end{center}
	\vskip -0.1in
\end{table}

\begin{figure}[!h]
	\centering	
	\includegraphics[width=.7\linewidth]{Figures/true_var_spatial}

	\caption{{\bf Variance function at $t=25$ (left) and $t=45$ (right)}} \label{fig:true_var_spatial}
\end{figure}

\section{Convergence}
Figure \ref{fig:convergence} shows the convergence of the two algorithms. Each iteration of the linearized algorithm takes about 0.01 seconds on average while each iteration of the consensus ADMM takes about 20 seconds.

\begin{figure}[!h]
	\centering	
	\includegraphics[width=.7\linewidth]{Figures/convergence}

	\caption{{\bf Convergence of linearized and consensus ADMM.}} \label{fig:convergence}
\end{figure}

\section{Model selection}
We fitted the linearized ADMM for all combinations of values of $\lambda_t$ and $\lambda_s$ from the sets $\lambda_t \in \{0,1,5,10,50,100\}$ and $\lambda_s \in \{0,0.05,0.1,0.2,0.3\}$. For each pair, we then compute the mean absolute error (MAE) between the estimated variance and the true variance at all locations and all time steps. For $\lambda_t=5$ and $\lambda_s=0.1$ MAE was minimized. Figure \ref{fig:true_fitted_var} shows the true and the estimated standard deviation at location (0,0) using $\lambda_s=0.1$ and $\lambda_t=5$ (blue) and $\lambda_t=100$ (green). As we can see, larger than optimal value of $\lambda_t$ leads to estimated values which are ``too smooth".

\begin{figure}[!h]
	\centering	
	\includegraphics[width=.7\linewidth]{Figures/true_fitted_var}

	\caption{{\bf The true (orange) and estimated standard deviation function at the location (0,0). The estimated values are obtained using linearized ADMM with $\lambda_s=0.1$ and two values of $\lambda_t$: $\lambda_t=5$ (blue) and $\lambda_t=100$ (green).}} \label{fig:true_fitted_var}
\end{figure}

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}

### 2018-05-13 nips-2018/paper.tex
reorganization and adding unique bib

@@
 
-\usepackage[showonlyrefs]{mathtools}
+\usepackage{mathtools}
+\mathtoolsset{showonlyrefs}
 
@@
 \renewcommand*{\subsubsectionautorefname}{Section}%
+
+\usepackage{xcolor}
+\newcommand{\alert}[1]{\textcolor{red}{TODO: #1}}
 
@@
 
+Nonparametric variance estimation for spatio-temporal data.
+
+\subsection{Motivating applications}
+
 There is a considerable interest in determining if there is an increasing trend in the climate variability \citep{hansen_perception_2012,huntingford_no_2013}.
@@
 All analyses are performed on a sub-set of the European Centre for Medium-Range Weather Forecasts (ECMWF) ERA-40 dataset \citep{uppala_era-40_2005}.
-This dataset include the temperature measurements over a grid over the earth from 1957 to 2002.
+This dataset include the temperature measurements over a grid over the earth from 1957 to 2002.~\citep{VasseurDeLong2014,TrenberthZhang2014,StatenKahn2016,Screen2014,FischerBeyerle2013}
 
@@
 
-\section{Data Exploration}
-This section is devoted to exploring some of the properties of the ERA-40 surface temperature data set.
-The goal here is to demonstrate some of the difficulties in modeling the trend in the temperature volatility and motivate our methodology for doing so.
+Also neuroscience.
 
-\autoref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil).
-The time-series of Bloomington and San Diego show clear cyclic behavior.
-However, while it seems (qualitatively) that these cycles can be modeled by a sinusoidal function for Bloomington, the same is not true for San Diego.
-Also, the amplitude of the cycles changes from some years to others.
-The time-series of Manaus does not show any regular cyclic behavior.
-This demonstrates the first difficulty in analyzing the variance of this data: to analyze the variance, we first need to remove the cyclic terms from all time-series.
-However, there is a lot of variations in the cyclic behavior of the time-series of different locations.
-In addition, some of these cycles cannot be easily modeled by a parametric function \footnote{One might try to model the cycles by the summation of sinusoidal terms with different frequencies.
-However, for some time-series this may need many terms to be included in the summation to achieve a reasonable level of accuracy.
-In addition, this model cannot capture the non-stationarity in the cycles.}.
-To overcome these issues, we will use a non-parametric approach to remove the cyclic terms from the time-series and de-trend them.
-This approach, called \textit{$\ell_1$-trend filtering} is explained in the next section.
+\subsection{Related work}
 
-\begin{figure}[ht]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=\columnwidth]{Figures/cities_ts}}
-		\caption{Time-series of the temperature (in Kelvin) of three cities.}
-		\label{fig:cities_ts}
-	\end{center}
-	\vskip -0.2in
-\end{figure}
+Mention ~\citep{KimKoh2009,HallacPark2017}. Also, ~\citep{Tibshirani2014,TibshiraniTaylor2011}. Find Sharpnack Tibshirani. ARCH/GARCH.
 
-The right panel of \autoref{fig:bloom_detrended} shows the time-series of the temperature of Bloomington, after removing the cyclic terms and de-trending using the method explained in the next section.
-The goal is to investigate the trend in the variance of this signal.
-This figure, reveals another issue toward this goal: the variance of this signal, shows cyclic behavior.
-Also, the cycles are not regular and their amplitude and frequency change.
-Even if one can describe the behavior of the variance of all the time-series using a single parametric model (for example a variant of the GARCH models \citep{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework.
-These observations motivate the need to develop a non-parametric framework for the problem at hand.
+\citep{SadhanalaWang2017,LinSharpnack2017,WangSharpnack2016}~\citep{RamdasTibshirani2016}
 
-\begin{figure}[ht]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=\columnwidth]{Figures/bloom_trend_resid}}
-		\caption{Left: Time-series of the temperature of Bloomington (blue) and the estimated trend and cycles obtained from the $\ell_1$-trend filtering (red).
-Right: the same time-series after removing the cyclic terms and de-trending using $\ell_1$-trend filtering.}
-		\label{fig:bloom_detrended}
-	\end{center}
-	\vskip -0.2in
-\end{figure}
+\subsection{Main contributions}
 
-\section{$\ell_1$-trend filtering for de-trending}
+\section{$\ell_1$-trend filtering for estimating variance of a time-series}
 
-$\ell_1$-trend filtering was proposed by \citep{kim_$ell_1$_2009} as a solution to the following problem: a time-series $y_t$, $t=1,...,T$ is observed and we believe that it consists of a smooth trend and a stochastic component and the goal is to estimate the trend.
+$\ell_1$-trend filtering was proposed by \citep{kim_$ell_1$_2009} as a method for estimating a smooth, time-varying trend. It is formulated as the optimization problem
+% : a time-series $y_t$, $t=1,...,T$ is observed and we believe that it consists of a smooth trend and a stochastic component and the goal is to estimate the trend.
 The estimated trend is desired to be smooth, but at the same time, we seek small estimated residuals (stochastic component).
@@
 In $\ell_1$-trend filtering, this trade-off is formulated as the following convex optimization problem:
-
 \begin{equation}
@@
 \end{equation}
-
 or equivalently:
-
 \begin{equation}
-\min_{\beta} \frac{1}{2} \lVert y-\beta \lVert_2^2+\lambda \lVert Dh \lVert_1
+\min_{\beta} \frac{1}{2} \lVert y-\beta \lVert_2^2+\lambda \lVert D\beta\lVert_1
 \label{eq:l1tf}
 \end{equation}
+ where $y_t$ is an observed time-series, $\beta$ is the smooth trend, $D$ is a $T-2\times T$ matrix, and $\lambda$ is a tuning parameter which balances fidelity to the data (small errors in the first term) with a desire for smoothness.
+% The \textit{penalty matrix} is:
+% \begin{equation}
+% D=\begin{bmatrix}
+% 1 & -2 & 1 &  &  &  &\\ 
+% & 1 & -2 & 1 &  &  &\\ 
+% &  & \ddots & \ddots & \ddots  &  &\\ 
+% &  & & 1 & -2 & 1 &  \\ 
+% &  &  &   & 1 & -2 & 1 
+% \end{bmatrix}
+% \label{eq:d_matrix}
+% \end{equation}
+%The first term in \eqref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
+With the penalty matrix $D$, the estimated $\beta$ will be piecewise linear.
+\citep{kim_$ell_1$_2009} proposed a specialized primal-dual interior point (PDIP) algorithm for solving \eqref{eq:l1tf}.
+From a statistical perspective, \eqref{eq:l1tf} is a constrained maximum likelihood problem with independent observations from a normal distribution with common variance, $y_t \sim \mbox{N}(\beta_t, \sigma^2)$, subject to a piecewise linear constraint on $\beta$.
 
- where $||u||_1$ is the $\ell_1$ norm of the vector $u$. Here, $\beta$ is the vector of the free parameters of the model and there is one parameter per observation. The \textit{penalty matrix} is:
-
-\begin{equation}
-D=\begin{bmatrix}
-1 & -2 & 1 &  &  &  &\\ 
-& 1 & -2 & 1 &  &  &\\ 
-&  & \ddots & \ddots & \ddots  &  &\\ 
-&  & & 1 & -2 & 1 &  \\ 
-&  &  &   & 1 & -2 & 1 
-\end{bmatrix}
-\label{eq:d_matrix}
-\end{equation}
-
-The first term in \eqref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
-Specifically, with the penalty matrix \eqref{eq:d_matrix}, the estimated $\beta$ will be piecewise linear \citep{kim_$ell_1$_2009}.
-The trade-off between the two aforementioned objectives is determined by the constant $\lambda$.
-\citep{kim_$ell_1$_2009} proposed a specialized primal-dual interior point (PDIP) method for solving the optimization problem \eqref{eq:l1tf} with the penalty matrix \eqref{eq:d_matrix}.
-
-The results of applying $\ell_1$-trend filtering on the Bloomington time-series are shown in \autoref{fig:bloom_detrended}.
+% The results of applying $\ell_1$-trend filtering on the Bloomington time-series are shown in \autoref{fig:bloom_detrended}.
 The left panel shows the original time-series (blue) together with the estimated $\beta$ (red) and the right panel shows the residuals $y_t-\beta_t$ (blue).
@@
 
-\section{$\ell_1$-trend filtering for estimating variance of a time-series}
+\subsection{Extension to variance estimation}
 \label{sec:l1tf_var}
 
-%TODO: talk about the motivation for using non-parametric based on the figure of the residuals.
-
-In this section, inspired by the $\ell_1$-trend filtering algorithm, we propose a non-parametric model for estimating the variance of a time-series.
+Inspired by the $\ell_1$-trend filtering algorithm, we propose a non-parametric model for estimating the variance of a time-series.
 To this end, we assume that at each time step $t$, there is a hidden variable $h_t$ such that conditioned on $h_t$ the observations $y_t$ are independent normal variables with zero mean and variance $\exp(h_t)$.
-The conditional log-likelihood of the observed data in this model is $l(y|h)=c-\sum_{t=1}^Th_t-y_t^2e^{-h_t}$, where $c$ is a constant that does not depend on $y$ and $h$.
+The negative log-likelihood of the observed data in this model is $l(y|h) \propto -\sum_{t=1}^Th_t - y_t^2e^{-h_t}$.
 Crucially, we assume that the hidden variables $h_t$ vary smoothly.
-To impose this assumption, we estimate $h_t$ by solving the penalized conditional log-likelihood:
-
+To impose this assumption, we estimate $h_t$ by solving the penalized, negative log-likelihood:
 \begin{equation}
@@
 \end{equation}
+ where $D$ has the same structure as above.
 
- where $D$ is the same as in \eqref{eq:d_matrix}.
-
-%TODO: explain the objective more.
+\alert{explain the objective more.
 give the AR(1) example.
@@
 But incontrast to spatial stat literature, it does not make any assumption on estimated variances.
-Compare to Hallac et al and Lingren et al.
+Compare to Hallac et al and Lingren et al.}
 
-We solve this optimization problem using the PDIP method \footnote{We use PDIP implemented in the cvxopt Python package \citep{andersen_cvxopt:_2013}.}.
+As with~\eqref{eq:l1tf}, one can solve~\eqref{eq:l1tf_var} using the PDIP algorithm (as in, e.g., \texttt{cvxopt}~\citep{andersen_cvxopt:_2013}).
 First, we note that this is a generalized LASSO problem \citep{tibshirani_solution_2011}.
 The dual of a generalized LASSO with the objective $f(x)+\lambda \lVert Dx \lVert_1$ is:
-
+\begin{align}
+\min_\nu&\quad f^*(-D^\top\nu) \\
+\mbox{s.t.}&\quad \lVert \nu \lVert_\infty \le \lambda
+\end{align}
+ where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^tx-f(x)$. It is simple to show that 
 \begin{equation}
-\begin{aligned}
-\min_\nu f^*(-D^T\mathsf{\nu}) \\
-s.t. \,\,\, \lVert \mathsf{\nu} \lVert_\infty \le \lambda
-\end{aligned}
-\end{equation}
-
- where $D^t$ is the transpose of $D$, and $f^*(\cdot)$ is the \textit{conjugate} of $f$ defined by: $f^*(u)=\max_x u^tx-f(x)$. It is simple to show that the conjugate of $f(h)=-l(y|h)$ is:
-
-\begin{equation}
-f^*(\mathsf{u})=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1
+f^*(\mathsf{u})=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
 \label{eq:conj}
@@
 
-The red curve in the left panel of \autoref{fig:bloom_estimatedSD} shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of the time-series of Bloomington.
-To reduce the number of time-steps in this figure and in the remainder of the paper we work on the weekly averaged of the data.
-
-The curve of the estimated SD captures the periodic variations in the SD of the signal.
-Just by looking at this curve, it is hard to say if the SD is decreasing or increasing.
-Therefore, we compute the average of the estimated SD for each year.
-The estimated SD together with this annual average is shown in the middle panel of \autoref{fig:bloom_estimatedSD}.
-As it can be seen, the annual trend is not smooth.
-This is because in the optimization problem \eqref{eq:l1tf_var}, the smoothness of the annual trend is not encouraged.
-To remedy this, we add the following long horizon penalty to \eqref{eq:l1tf_var}:
-
-\begin{equation}
-\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
-\label{eq:lh_penalty}
-\end{equation}
-
- where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
-Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
-Since we are working on the weekly averaged data, each year corresponds to 52 observations.
-In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
-The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
-The annual average of the estimated SDs shows a linear trend with a positive slope.
-
-\begin{figure}[ht]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}}
-		\caption{Left: The residuals of the time-series of Bloomington (averaged weekly) and the estimated SD obtained from the method of \autoref{sec:l1tf_var} (red).
-Middle: the estimated SDs (red) and their annual average (black) without imposing the long horizon penalty.
-Right: the same as middle panel but here the long horizon penalty is imposed.
-See the text for more details.}
-		\label{fig:bloom_estimatedSD}
-	\end{center}
-	\vskip -0.2in
-\end{figure}
-
-%TODO: how to choose lambda? plot BIC as a function of lambda.
-
-\section{Extension to spatio-temporal data}
+\subsection{Extension to spatio-temporal data}
 \label{sec:exten}
@@
 
-\begin{figure}[ht]
+\begin{figure}[tb]
 	\vskip 0.2in
@@
 
-\section{Results}
+\section{Empirical evaluation}
+
+\subsection{Simulations}
+\label{sec:simulations}
+
+To analyze the proposed optimization methods, we first fit the model on some synthetic spatio-temporal data.
+The observations at all time steps and all locations were generated from independent Gaussian random variables with zero mean.
+However, the variance of these random variables changes smoothly in time and space.
+Specifically, the observation at time step $t$ and at location $(r,c)$ on the grid is generated from a Gaussian $N(0,\sigma^2(t,r,c))$, where:
+
+\begin{equation}
+\begin{aligned}
+\sigma(t,r,c) & =\sum_{s=1}^{S} W_s(t) \cdot \exp\bigg( \frac{(r-r_s)^2+(c-c_s)^2}{2\sigma_s^2} \bigg) \\
+W_s(t) & =\alpha_s \cdot t + \exp(\sin(2\pi\omega_s t+\phi_s)) 
+\label{eq:sourceVar}
+\end{aligned}
+\end{equation}
+
+In words, the variance at each time and location is computed as the weighted sum of $S$ bell-shaped functions where the weights are time-varying functions.
+Specifically, the weights consist of a linear trend $\alpha_s \cdot t$ and a periodic term $\beta_s \cdot \sin(2\pi\omega_s t+\phi_s)$.
+The bell-shaped functions impose the spatial smoothness, and the linear trend and the periodic terms enforce the temporal smoothness.
+We simulated the data on a 5 by 7 grid and for 780 time steps.
+We also used $S=4$.
+The parameters of the variance function are shown in  \autoref{tab:sim_params}.
+The value of the variance function for all locations on the grid and at $t=25$ and $t=45$ is shown in \autoref{fig:true_var_spatial}.
+Also, the variance for all time steps and at the location (0,0) on the grid is shown in \autoref{fig:true_fitted_var}.
+The linear trend and the period term can be seen clearly in this figure.
+
+\begin{table}[tb]
+	\caption{Parameters used to simulate data}
+	\label{tab:sim_params}
+	\vskip 0.15in
+	\begin{center}
+		\begin{small}
+			\begin{sc}
+				\begin{tabular}{ccccccc}
+					\hline
+
+					$s$ & $r_s$ & $c_s$ & $\sigma_s$ &$\alpha_s$ & $\omega_s$ & $\phi_s$\\
+					\hline
+					1 & 0 & 0 & 5 & 0.5 & 0.121 & 0 \\
+					2 & 0 & 5 & 5 & 0.1 & 0.121 & 0 \\
+					3 & 3 & 0 & 5 & -0.5 & 0.121 & $\pi/2$ \\
+					4 & 3 & 5 & 5 & -0.1 & 0.121 & $\pi/2$ \\
+					\hline
+				\end{tabular}
+			\end{sc}
+		\end{small}
+	\end{center}
+	\vskip -0.1in
+\end{table}
+
+\begin{figure}[tb]
+	\centering	
+	\includegraphics[width=.7\linewidth]{Figures/true_var_spatial}
+
+	\caption{Variance function at $t=25$ (left) and $t=45$ (right)} \label{fig:true_var_spatial}
+\end{figure}
+
+The left panel of \autoref{fig:true_fitted_var} shows the convergence of the two algorithms.
+Each iteration of the linearized algorithm takes about 0.01 seconds on average while each iteration of the consensus ADMM takes about 20 seconds.
+
+We estimated the linearized ADMM for all combinations of values of $\lambda_t$ and $\lambda_s$ from the sets $\lambda_t \in \{0,1,5,10,50,100\}$ and $\lambda_s \in \{0,0.05,0.1,0.2,0.3\}$.
+For each pair, we then compute the mean absolute error (MAE) between the estimated variance and the true variance at all locations and all time steps.
+For $\lambda_t=5$ and $\lambda_s=0.1$ MAE was minimized.
+The right panel of \autoref{fig:true_fitted_var} shows the true and the estimated standard deviation at location (0,0) using $\lambda_s=0.1$ and $\lambda_t=5$ (blue) and $\lambda_t=100$ (green).
+As we can see, larger than optimal value of $\lambda_t$ leads to estimated values which are ``too smooth".
+
+\begin{figure}[!h]
+	\centering	
+	\includegraphics[width=.4\linewidth]{Figures/convergence}			
+	\includegraphics[width=.4\linewidth]{Figures/true_fitted_var}		
+	\caption{Left: Convergence speed of linearized and consensus ADMM.
+Right: The true (orange) and estimated standard deviation function at the location (0,0).
+The estimated values are obtained using linearized ADMM with $\lambda_s=0.1$ and two values of $\lambda_t$: $\lambda_t=5$ (blue) and $\lambda_t=100$ (green).} \label{fig:true_fitted_var}
+\end{figure}
+
+\subsection{Data analysis}
 As it was mentioned before, the algorithm proposed in \autoref{sec:consOpt} is appropriate only if we parallelize it over multiple machines and so we do not pursue it further here.
@@
 
-\subsection{Convergence}
+\paragraph{Data Exploration}
+The red curve in the left panel of \autoref{fig:bloom_estimatedSD} shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of the time-series of Bloomington.
+To reduce the number of time-steps in this figure and in the remainder of the paper we work on the weekly averaged of the data.
 
-\autoref{fig:loss_vs_iter} shows the value of the loss function \eqref{eq:l1tf_var_st_mat} as a function of iteration for $\lambda_t=1$ and $\lambda_s=0$.
+The curve of the estimated SD captures the periodic variations in the SD of the signal.
+Just by looking at this curve, it is hard to say if the SD is decreasing or increasing.
+Therefore, we compute the average of the estimated SD for each year.
+The estimated SD together with this annual average is shown in the middle panel of \autoref{fig:bloom_estimatedSD}.
+As it can be seen, the annual trend is not smooth.
+This is because in the optimization problem \eqref{eq:l1tf_var}, the smoothness of the annual trend is not encouraged.
+To remedy this, we add the following long horizon penalty to \eqref{eq:l1tf_var}:
+
+\begin{equation}
+\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
+\label{eq:lh_penalty}
+\end{equation}
+
+ where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
+Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
+Since we are working on the weekly averaged data, each year corresponds to 52 observations.
+In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
+The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
+The annual average of the estimated SDs shows a linear trend with a positive slope.
+
+\begin{figure}[tb]
+	\vskip 0.2in
+	\begin{center}
+		\centerline{\includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}}
+		\caption{Left: The residuals of the time-series of Bloomington (averaged weekly) and the estimated SD obtained from the method of \autoref{sec:l1tf_var} (red).
+Middle: the estimated SDs (red) and their annual average (black) without imposing the long horizon penalty.
+Right: the same as middle panel but here the long horizon penalty is imposed.
+See the text for more details.}
+		\label{fig:bloom_estimatedSD}
+	\end{center}
+	\vskip -0.2in
+\end{figure}
+
+%TODO: how to choose lambda? plot BIC as a function of lambda.
+
+This section is devoted to exploring some of the properties of the ERA-40 surface temperature data set.
+The goal here is to demonstrate some of the difficulties in modeling the trend in the temperature volatility and motivate our methodology for doing so.
+
+% \autoref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil).
+The time-series of Bloomington and San Diego show clear cyclic behavior.
+However, while it seems (qualitatively) that these cycles can be modeled by a sinusoidal function for Bloomington, the same is not true for San Diego.
+Also, the amplitude of the cycles changes from some years to others.
+The time-series of Manaus does not show any regular cyclic behavior.
+This demonstrates the first difficulty in analyzing the variance of this data: to analyze the variance, we first need to remove the cyclic terms from all time-series.
+However, there is a lot of variations in the cyclic behavior of the time-series of different locations.
+In addition, some of these cycles cannot be easily modeled by a parametric function \footnote{One might try to model the cycles by the summation of sinusoidal terms with different frequencies.
+However, for some time-series this may need many terms to be included in the summation to achieve a reasonable level of accuracy.
+In addition, this model cannot capture the non-stationarity in the cycles.}.
+To overcome these issues, we will use a non-parametric approach to remove the cyclic terms from the time-series and de-trend them.
+This approach, called \textit{$\ell_1$-trend filtering} is explained in the next section.
+
+% \begin{figure}[tb]
+% 	\vskip 0.2in
+% 	\begin{center}
+% 		\centerline{\includegraphics[width=\columnwidth]{Figures/cities_ts}}
+% 		\caption{Time-series of the temperature (in Kelvin) of three cities.}
+% 		\label{fig:cities_ts}
+% 	\end{center}
+% 	\vskip -0.2in
+% \end{figure}
+
+The right panel of \autoref{fig:bloom_detrended} shows the time-series of the temperature of Bloomington, after removing the cyclic terms and de-trending using the method explained in the next section.
+The goal is to investigate the trend in the variance of this signal.
+This figure, reveals another issue toward this goal: the variance of this signal, shows cyclic behavior.
+Also, the cycles are not regular and their amplitude and frequency change.
+Even if one can describe the behavior of the variance of all the time-series using a single parametric model (for example a variant of the GARCH models \citep{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework.
+These observations motivate the need to develop a non-parametric framework for the problem at hand.
+
+\begin{figure}[tb]
+	\vskip 0.2in
+	\begin{center}
+		\centerline{\includegraphics[width=\columnwidth]{Figures/bloom_trend_resid}}
+		\caption{Left: Time-series of the temperature of Bloomington (blue) and the estimated trend and cycles obtained from the $\ell_1$-trend filtering (red).
+Right: the same time-series after removing the cyclic terms and de-trending using $\ell_1$-trend filtering.}
+		\label{fig:bloom_detrended}
+	\end{center}
+	\vskip -0.2in
+\end{figure}
+
+\paragraph{Convergence}
+
 We used the following rule to determine when to stop the optimization: the optimization was stopped if the value of the loss did not improve by at least 0.1\% in 1000 trials.
@@
 
-\begin{figure}[ht]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=1\columnwidth]{Figures/loss_vs_iter}}
-		\caption{Loss as a function of iteration for $\lambda_t=1$ and $\lambda_s=0$.}
-		\label{fig:loss_vs_iter}
-	\end{center}
-	\vskip -0.2in
-\end{figure}
-
-\subsection{Model selection}
+\paragraph{Model selection}
 One common method for choosing the penalty parameters in the Lasso problems is to find the solution for a range of the values of these parameters and then choose the values which minimize a model selection criterion.
@@
 We computed the optimal solution for all the combinations of the following sets of values: $\lambda_t \in \{1,5,10,20\} \, \, , \lambda_s \in \{0,.1,1,5,10\}$.
-The value of $\mathscr{L}(\lambda_t,\lambda_s)$ is shown in \autoref{fig:modelSelect}.
-This function is minimized at $\lambda_t=5$ and $\lambda_s=1$.
+The best combination based on a held out set was $\lambda_t=5$ and $\lambda_s=1$.
 All the analyses in the next section are performed on the solution for these values.
 
-\begin{figure}[ht]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=1\columnwidth]{Figures/modelSelect}}
-		\caption{Value of $\mathscr{L}(\lambda_t,\lambda_s)$ for several values of $\lambda_t$ and $\lambda_s$. }
-		\label{fig:modelSelect}
-	\end{center}
-	\vskip -0.2in
-\end{figure}
+\paragraph{Analysis of trend of temperature volatility}
 
-\subsection{Analysis of trend of temperature volatility}
-\autoref{fig:estimatedSD} shows the detrended data, the estimated standard deviation and the yearly average of these estimates for two cities in the united states: Bloomington (left) and San Diego (right).
+The top row of \autoref{fig:avg_change_estimatedSD} shows the detrended data, the estimated standard deviation and the yearly average of these estimates for two cities in the US.
 The estimated SD captures the periodic behavior in the variance of the time-series.
 In addition, the number of linear segments changes adaptively in each time window depending on how fast the variance is changing.
-
-\begin{figure}[ht]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=1\columnwidth]{Figures/estimatedSD}}
-		\caption{Detrended data and the estimated SD for Bloomington (left) and San Diego (right).}
-		\label{fig:estimatedSD}
-	\end{center}
-	\vskip -0.2in
-\end{figure}
-
 The yearly average of the estimated SD captures the trend in the temperature volatility.
@@
 
-\begin{figure}[ht]
-	\begin{subfigure}
+\begin{figure}[tb]
 	  \centering
-	  \includegraphics[width=.5\linewidth]{Figures/avg_estimatedVar}
-	\end{subfigure}%
-	\begin{subfigure}
-	  \centering
-	  \includegraphics[width=.5\linewidth]{Figures/avg_change_estimatedVar}
-	\end{subfigure}%
-	\caption{Left: the average of the estimated variance in all locations in the grid over the united states. Right: the change in the variance from 1992 to 2002.}
+          \includegraphics[width=.9\columnwidth]{Figures/estimatedSD}\\
+	  \includegraphics[width=.45\linewidth]{Figures/avg_estimatedVar}
+	  \includegraphics[width=.45\linewidth]{Figures/avg_change_estimatedVar}
+	\caption{Top row: Detrended data and the estimated SD for Bloomington (left) and San Diego (right).
+Bottom: the average of the estimated variance over the US (left) and the change in the variance from 1992 to 2002 (right)}
 	\label{fig:avg_change_estimatedSD}
 \end{figure} 
-
 It is interesting to note that the trend in volatility is almost zero over the oceans.
 The most positive trend can be observed in the south-east and the most negative trend has happened in the north-east.
-
-Finally, \autoref{fig:hist_avg_change} shows the histogram of the changes shown in the right panel of \autoref{fig:avg_change_estimatedSD}.
-As we can see, there are two peaks in this histogram: one for locations with no trend (which includes the points on the oceans), and one peak for locations with positive trend.
-The locations with negative trend form the long tail of the histogram.
-
-\begin{figure}[ht]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=1\columnwidth]{Figures/hist_avg_change}}
-		\caption{Histogram of the change in the estimated variances.}
-		\label{fig:hist_avg_change}
-	\end{center}
-	\vskip -0.2in
-\end{figure}
 
@@
 \small
-\bibliography{references}
+\bibliography{nips-references}
 

### 2018-05-13 nips-2018/paper.tex
add gitignore, PDIP algorithm

@@
 \usepackage{xcolor}
-\newcommand{\alert}[1]{\textcolor{red}{TODO: #1}}
+\newcommand{\attn}[1]{\textcolor{red}{TODO: #1}}
+\newcommand{\one}{\mathbf{1}}
+\renewcommand{\algorithmiccomment}[1]{\hfill $\rhd$ #1}
 
@@
 
-\alert{explain the objective more.
+\attn{explain the objective more.
 give the AR(1) example.
@@
 \end{equation}
+Writing
+\[
+  r_w(v,\mu_1,\mu_2):=
+  \begin{bmatrix}
+    \nabla f^*(-D^\top v) + D(v-\lambda \one)^\top \mu_1 -
+    D(v+\lambda \one)^\top \mu_2\\
+    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
+  \end{bmatrix}
+  =
+  \begin{bmatrix}
+  0\\0\end{bmatrix}
+\]
+for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint in the dual problem.
+As $w\rightarrow\infty$, the solution to this nonlinear system reduces to the KKT conditions.
+Therefore, the PDIP method takes Newton steps to solve the system for a series of increasing values of $w$.
+\autoref{alg:pdip} provides the details.
+\begin{algorithm}
+  \caption{PDIP for $\ell_1$ variance estimation}
+  \label{alg:pdip}
+  \begin{algorithmic}
+    \REQUIRE $\lambda>0$, $w>0$, $\nu\leftarrow 0$, $\mu_1\leftarrow 0$, $\mu_2\leftarrow 0$, $J\in\mathbb{Z}^+$, $\{w_k\}$ \COMMENT{Initialization}
+    \FOR[Central path]{$k=1,2,\ldots$}
+    \FOR[Newton updates]{$j=1,\ldots,J$}
+    \STATE Solve $A [\Delta\nu\; \Delta\mu_1\; \Delta\mu_2]^\top=r^j_{w_k}$ to find the search direction.
+    \STATE $\quad\quad A$ is the Jacobian of $r_w$.
+    \STATE Update $r^{j+1} \leftarrow r^j + [\Delta\nu\; \Delta\mu_1\; \Delta\mu_2]^\top$.
+    \ENDFOR
+   \ENDFOR
+   \RETURN $h=-D^\top\nu$ \attn{Is this right??}
+  \end{algorithmic}
+\end{algorithm}
 
@@
 For example, the first $T$ rows correspond to $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to $|h_{11t}-h_{12t}|$, and so on.
+
+\attn{Detailed discussion of the PDIP to here}
 
@@
 By this definition of $x_i$, the update step for $x_i$ is the following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) + (u_i^m)^t x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
+
+\attn{Redo the below and move to the section on PDIP above}
 
@@
 From the \textit{separable sum} property of the proximal operators we have (see \citep{parikh_proximal_2014}, section 2.1):
+
+\attn{Make this a theorem}
 

### 2018-05-15 nips-2018/paper.tex
Rearrangement done. Comments throughout to be addressed

--- 2018-05-13 nips-2018/paper.tex
@@
 %Arash
-\usepackage{amsmath}
+\usepackage{amsmath,amsthm}
 \usepackage{amssymb}
@@
 
+\newtheorem{theorem}{Theorem}%[section]
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{lemma}[theorem]{Lemma}
+
+\usepackage{gensymb}
 \usepackage{xcolor}
-\newcommand{\alert}[1]{\textcolor{red}{TODO: #1}}
+\newcommand{\attn}[1]{\textcolor{red}{TODO: #1}}
+\newcommand{\one}{\mathbf{1}}
+\renewcommand{\algorithmiccomment}[1]{\hfill $\rhd$ #1}
+\newcommand{\given}{\;\vert\;}
+\newcommand*{\algorithmautorefname}{Algorithm}%
+\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
+\DeclareMathOperator*{\prox}{\bf prox}
+\DeclareMathOperator*{\sign}{sign}
 
@@
 
-There is a considerable interest in determining if there is an increasing trend in the climate variability \citep{hansen_perception_2012,huntingford_no_2013}.
-An increase in the temperature variability will increase the probability of extreme hot outliers.
-It might be harder for the society to adapt to these extremes than to the gradual increase in the mean temperature \citep{huntingford_no_2013}.
+\attn{cut this down}
 
-In this project, we consider the problem of detecting the trend in the temperature volatility.
-All analyses are performed on a sub-set of the European Centre for Medium-Range Weather Forecasts (ECMWF) ERA-40 dataset \citep{uppala_era-40_2005}.
-This dataset include the temperature measurements over a grid over the earth from 1957 to 2002.~\citep{VasseurDeLong2014,TrenberthZhang2014,StatenKahn2016,Screen2014,FischerBeyerle2013}
+There is a considerable interest in determining if there is an
+increasing trend in the climate variability
+\citep{hansen_perception_2012,huntingford_no_2013}. An increase in the
+temperature variability will increase the probability of extreme hot
+outliers. It might be harder for the society to adapt to these
+extremes than to the gradual increase in the mean temperature
+\citep{huntingford_no_2013}.
 
-Research on analyzing the trend in the volatility of spatio-temporal data is scarce.
-\citep{hansen_perception_2012} studied the change in the standard deviation (SD) of the surface temperature in the NASA Goddard Institute for Space Studies gridded temperature data set.
-In their analysis, for each geographical position, the mean of the temperature computed for the period 1951-1980 (called the base-period) at that position, is subtracted from the corresponding time series.
-Each time series is then divided by the standard deviation computed at each position and during the same time period.
-The distribution of the resulting data is then plotted for different periods.
-These distributions represent the deviation of the temperature for a specific period, from the mean in the base period, in units of the standard deviation in that period.
-The results showed that these distributions are widen for the resent time periods compared to 1951-1980.
-\citep{huntingford_no_2013} took a similar approach in analysing the ERA-40 data set.
-However, in addition to the aforementioned method, they computed the distribution of the SDs in an alternative way: for each position and each time period, the deviation of the time-series at that position from the mean in that time period at that position was computed, and then divided by the SD of that position in the period before 1981.
-The results showed that there still is an increase in the SDs from 1958-1970 to 1991-2001, but this is much less than what is obtained from the method used in \citep{hansen_perception_2012}.
-The authors also computed the time-evolving global SD from the de-trended time-series at each position.
-The resulting curve suggested that the global SD has been stable.
+In this project, we consider the problem of detecting the trend in the
+temperature volatility. All analyses are performed on a sub-set of the
+European Centre for Medium-Range Weather Forecasts (ECMWF) ERA-40
+dataset \citep{uppala_era-40_2005}. This dataset include the
+temperature measurements over a grid over the earth from 1957 to
+2002.~\citep{VasseurDeLong2014,TrenberthZhang2014,StatenKahn2016,Screen2014,FischerBeyerle2013}
 
-These previous work (and other related research, e.g., \citep{rhines_frequent_2013}) have several shortcomings.
-First, no statistical analysis has been performed to examine if the change in the SD is statistically significant.
-Second, the methodologies for computing the SDs are rather arbitrary.
-The deviation of each time-series in a given period, is computed from either the mean of a base-period (as in \citep{hansen_perception_2012}), or from the given period (as in \citep{rhines_frequent_2013,huntingford_no_2013}).
-These deviations are then normalized using the SD of the base-period or the given period.
-No justification is provided for these choices.
-Third, the correlation between the observations is ignored.
-The observations in subsequent days and close geographical positions could be highly correlated.
-Without considering these correlations, any conclusion based on the averaged data could be flawed.
+Research on analyzing the trend in the volatility of spatio-temporal
+data is scarce. \citep{hansen_perception_2012} studied the change in
+the standard deviation (SD) of the surface temperature in the NASA
+Goddard Institute for Space Studies gridded temperature data set. In
+their analysis, for each geographical position, the mean of the
+temperature computed for the period 1951-1980 (called the base-period)
+at that position, is subtracted from the corresponding time
+series. Each time series is then divided by the standard deviation
+computed at each position and during the same time period. The
+distribution of the resulting data is then plotted for different
+periods. These distributions represent the deviation of the
+temperature for a specific period, from the mean in the base period,
+in units of the standard deviation in that period. The results showed
+that these distributions are widen for the resent time periods
+compared to 1951-1980.  \citep{huntingford_no_2013} took a similar
+approach in analysing the ERA-40 data set. However, in addition to the
+aforementioned method, they computed the distribution of the SDs in an
+alternative way: for each position and each time period, the deviation
+of the time-series at that position from the mean in that time period
+at that position was computed, and then divided by the SD of that
+position in the period before 1981. The results showed that there
+still is an increase in the SDs from 1958-1970 to 1991-2001, but this
+is much less than what is obtained from the method used in
+\citep{hansen_perception_2012}. The authors also computed the
+time-evolving global SD from the de-trended time-series at each
+position. The resulting curve suggested that the global SD has been
+stable.
 
-The main contribution of this work is to develop a new methodology for detecting the trend in the volatility of spatio-temporal data.
-In this methodology, the variance at each position and time, is considered as a hidden (un-observed) variable.
-The value of these hidden variables are then estimated by maximizing the likelihood of the observed data.
-We show that this formulation per se, is not appropriate for detecting the trend.
-To overcome this issue, we penalize the differences between the estimated variances of the observations which are temporally and/or spatially close to each other.
-This will result in an optimization problem called the \textit{generalized LASSO problem} \citep{tibshirani_solution_2011}.
-As we will see, the dimension of this optimization problem is very high and so the standard methods for solving the generalized LASSO cannot be applied directly.
-We investigate two methods for solving this optimization problem.
-In the first method, we adopt an optimization technique called alternative direction method of multipliers (ADMM) \citep{boyd_distributed_2011}, to divide the total problem into several sub-problems of much lower dimension and show how the total problem can be solved by iteratively solving these sub-problems.
-The second method, called the \textit{linearized ADMM algorithm} \citep{parikh_proximal_2014} solves the main problem by iteratively solving a linearized version of it.
-We will compare the benefits of each method.
+These previous work (and other related research, e.g.,
+\citep{rhines_frequent_2013}) have several shortcomings. First, no
+statistical analysis has been performed to examine if the change in
+the SD is statistically significant. Second, the methodologies for
+computing the SDs are rather arbitrary. The deviation of each
+time-series in a given period, is computed from either the mean of a
+base-period (as in \citep{hansen_perception_2012}), or from the given
+period (as in \citep{rhines_frequent_2013,huntingford_no_2013}). These
+deviations are then normalized using the SD of the base-period or the
+given period. No justification is provided for these choices. Third,
+the correlation between the observations is ignored. The observations
+in subsequent days and close geographical positions could be highly
+correlated. Without considering these correlations, any conclusion
+based on the averaged data could be flawed.
+
+The main contribution of this work is to develop a new methodology for
+detecting the trend in the volatility of spatio-temporal data. In this
+methodology, the variance at each position and time, is considered as
+a hidden (un-observed) variable. The value of these hidden variables
+are then estimated by maximizing the likelihood of the observed
+data. We show that this formulation per se, is not appropriate for
+detecting the trend. To overcome this issue, we penalize the
+differences between the estimated variances of the observations which
+are temporally and/or spatially close to each other. This will result
+in an optimization problem called the \textit{generalized LASSO
+  problem} \citep{tibshirani_solution_2011}. As we will see, the
+dimension of this optimization problem is very high and so the
+standard methods for solving the generalized LASSO cannot be applied
+directly. We investigate two methods for solving this optimization
+problem. In the first method, we adopt an optimization technique
+called alternative direction method of multipliers (ADMM)
+\citep{boyd_distributed_2011}, to divide the total problem into
+several sub-problems of much lower dimension and show how the total
+problem can be solved by iteratively solving these sub-problems. The
+second method, called the \textit{linearized ADMM algorithm}
+\citep{parikh_proximal_2014} solves the main problem by iteratively
+solving a linearized version of it. We will compare the benefits of
+each method.
 
@@
 
-Mention ~\citep{KimKoh2009,HallacPark2017}. Also, ~\citep{Tibshirani2014,TibshiraniTaylor2011}. Find Sharpnack Tibshirani. ARCH/GARCH.
+Mention ~\citep{KimKoh2009,HallacPark2017}. Also,
+~\citep{Tibshirani2014,TibshiraniTaylor2011}. ARCH/GARCH.
 
@@
 
-\section{$\ell_1$-trend filtering for estimating variance of a time-series}
+\begin{itemize}
+\item We propose a model for non-parametric variance estimation for a
+  spatio-temporal process (~\autoref{sec:ell_1-trend-filt}).
+\item We derive two algorithms to fit our estimator when applied to
+  very large data (~\autoref{sec:prop-optim-meth}).
+\item We illustrate our methods on a large global temperature dataset
+  with the goal of tracking world-wide trends in variance as well as a
+  simulation constructed to mimic these data's features
+  (~\autoref{sec:empirical-evaluation}). 
+\end{itemize}
+
+\section{$\ell_1$-trend filtering for estimating variance of a
+  time-series}
+\label{sec:ell_1-trend-filt}
 
 $\ell_1$-trend filtering was proposed by \citep{kim_$ell_1$_2009} as a method for estimating a smooth, time-varying trend. It is formulated as the optimization problem
-% : a time-series $y_t$, $t=1,...,T$ is observed and we believe that it consists of a smooth trend and a stochastic component and the goal is to estimate the trend.
-The estimated trend is desired to be smooth, but at the same time, we seek small estimated residuals (stochastic component).
-These two objectives cannot be achieved simultaneously and so a trade-off between them should be made.
-In $\ell_1$-trend filtering, this trade-off is formulated as the following convex optimization problem:
+% : a time-series $y_t$, $t=1,...,T$ is observed and we believe that
+% it consists of a smooth trend and a stochastic component and the
+% goal is to estimate the trend. The estimated trend is desired to be
+% smooth, but at the same time, we seek small estimated residuals
+% (stochastic component). These two objectives cannot be achieved
+% simultaneously and so a trade-off between them should be made. In
+% $\ell_1$-trend filtering, this trade-off is formulated as the
+% following convex optimization problem: 
 \begin{equation}
-\min_{\beta} \frac{1}{2} \sum_{t=1}^{T} (y_t-\beta_t)^2+\lambda \sum_{t=1}^{T-2} |\beta_t-2\beta_{t+1}+\beta_{t+2}|
+\min_{\beta} \frac{1}{2} \sum_{t=1}^{T} (y_t-\beta_t)^2+\lambda
+\sum_{t=1}^{T-2} \left|\beta_t-2\beta_{t+1}+\beta_{t+2}\right| 
 \end{equation}
@@
 \begin{equation}
-\min_{\beta} \frac{1}{2} \lVert y-\beta \lVert_2^2+\lambda \lVert D\beta\lVert_1
+\min_{\beta} \frac{1}{2} \norm{ y-\beta }_2^2+\lambda \norm{ D\beta}_1
 \label{eq:l1tf}
 \end{equation}
- where $y_t$ is an observed time-series, $\beta$ is the smooth trend, $D$ is a $T-2\times T$ matrix, and $\lambda$ is a tuning parameter which balances fidelity to the data (small errors in the first term) with a desire for smoothness.
+ where $y_t$ is an observed time-series, $\beta$ is the smooth trend,
+ $D$ is a $(T-2)\times T$ matrix, and $\lambda$ is a tuning parameter
+ which balances fidelity to the data (small errors in the first term)
+ with a desire for smoothness.  
 % The \textit{penalty matrix} is:
@@
 %The first term in \eqref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
-With the penalty matrix $D$, the estimated $\beta$ will be piecewise linear.
-\citep{kim_$ell_1$_2009} proposed a specialized primal-dual interior point (PDIP) algorithm for solving \eqref{eq:l1tf}.
-From a statistical perspective, \eqref{eq:l1tf} is a constrained maximum likelihood problem with independent observations from a normal distribution with common variance, $y_t \sim \mbox{N}(\beta_t, \sigma^2)$, subject to a piecewise linear constraint on $\beta$.
+With the penalty matrix $D$, the estimated $\beta$ will be piecewise
+linear. \citep{kim_$ell_1$_2009} proposed a specialized primal-dual
+interior point (PDIP) algorithm for solving \eqref{eq:l1tf}. From a
+statistical perspective, \eqref{eq:l1tf} is a constrained maximum
+likelihood problem with independent observations from a normal
+distribution with common variance, $y_t \sim \mbox{N}(\beta_t,
+\sigma^2)$, subject to a piecewise linear constraint on $\beta$.
 
-% The results of applying $\ell_1$-trend filtering on the Bloomington time-series are shown in \autoref{fig:bloom_detrended}.
-The left panel shows the original time-series (blue) together with the estimated $\beta$ (red) and the right panel shows the residuals $y_t-\beta_t$ (blue).
-As it can be seen, the estimated $\beta$s are piecewise linear.
-The number of linear segments is determined by the value of $\lambda$.
-Here, we chose $\lambda=500$.
-The choice of $\lambda$ is a model selection problem and we will return to it in the next section ???.
+% The results of applying $\ell_1$-trend filtering on the Bloomington
+% time-series are shown in \autoref{fig:bloom_detrended}. The left
+% panel shows the original time-series (blue) together with the
+% estimated $\beta$ (red) and the right panel shows the residuals
+% $y_t-\beta_t$ (blue). As it can be seen, the estimated $\beta$s are
+% piecewise linear. The number of linear segments is determined by the
+% value of $\lambda$. Here, we chose $\lambda=500$. The choice of
+% $\lambda$ is a model selection problem and we will return to it in
+% the next section ???.
 
-\subsection{Extension to variance estimation}
+\subsection{Estimating the variance}
 \label{sec:l1tf_var}
 
-Inspired by the $\ell_1$-trend filtering algorithm, we propose a non-parametric model for estimating the variance of a time-series.
-To this end, we assume that at each time step $t$, there is a hidden variable $h_t$ such that conditioned on $h_t$ the observations $y_t$ are independent normal variables with zero mean and variance $\exp(h_t)$.
-The negative log-likelihood of the observed data in this model is $l(y|h) \propto -\sum_{t=1}^Th_t - y_t^2e^{-h_t}$.
-Crucially, we assume that the hidden variables $h_t$ vary smoothly.
-To impose this assumption, we estimate $h_t$ by solving the penalized, negative log-likelihood:
+Inspired by the $\ell_1$-trend filtering algorithm, we propose a
+non-parametric model for estimating the variance of a time-series. To
+this end, we assume that at each time step $t$, there is a hidden
+variable $h_t$ such that conditioned on $h_t$ the observations $y_t$
+are independent normal variables with zero mean and variance
+$\exp(h_t)$. The negative log-likelihood of the observed data in this
+model is $l(y\given h) \propto -\sum_{t=1}^Th_t - y_t^2e^{-h_t}$. Crucially,
+we assume that the hidden variables $h_t$ vary smoothly. To impose
+this assumption, we estimate $h_t$ by solving the penalized, negative
+log-likelihood: 
 \begin{equation}
-\min_h -l(y|h)+\lambda \lVert Dh \lVert_1
+\min_h -l(y\given h)+\lambda \norm{ Dh }_1
 \label{eq:l1tf_var}
@@
 
-\alert{explain the objective more.
-give the AR(1) example.
-Explain what you loses by this assumption (ACF,forecasting).
-Also explain that the covariace matrix is diagonal so it cannot capture the covariance structure.
-But incontrast to spatial stat literature, it does not make any assumption on estimated variances.
-Compare to Hallac et al and Lingren et al.}
+\attn{explain the objective more. give the AR(1) example. Explain what
+  you loses by this assumption (ACF,forecasting). Also explain that
+  the covariace matrix is diagonal so it cannot capture the covariance
+  structure. But incontrast to spatial stat literature, it does not
+  make any assumption on estimated variances. Compare to Hallac et al
+  and Lingren et al.}
 
-As with~\eqref{eq:l1tf}, one can solve~\eqref{eq:l1tf_var} using the PDIP algorithm (as in, e.g., \texttt{cvxopt}~\citep{andersen_cvxopt:_2013}).
-First, we note that this is a generalized LASSO problem \citep{tibshirani_solution_2011}.
-The dual of a generalized LASSO with the objective $f(x)+\lambda \lVert Dx \lVert_1$ is:
+As with~\eqref{eq:l1tf}, one can solve~\eqref{eq:l1tf_var} using the
+PDIP algorithm (as in, e.g.,
+\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}). First, we note that
+this is a generalized LASSO problem
+\citep{tibshirani_solution_2011}. The dual of a generalized LASSO with
+the objective $f(x)+\lambda \norm{ Dx }_1$ is:  
 \begin{align}
-\min_\nu&\quad f^*(-D^\top\nu) \\
-\mbox{s.t.}&\quad \lVert \nu \lVert_\infty \le \lambda
+\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
 \end{align}
- where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^tx-f(x)$. It is simple to show that 
+ where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x
+ u^\top x-f(x)$. It is simple to show that 
 \begin{equation}
-f^*(\mathsf{u})=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
+f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
 \label{eq:conj}
 \end{equation}
+Writing
+\begin{equation}
+  r_w(v,\mu_1,\mu_2):=
+  \begin{bmatrix}
+    \nabla f^*(-D^\top v) + D(v-\lambda \one)^\top \mu_1 -
+    D(v+\lambda \one)^\top \mu_2\\
+    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
+  \end{bmatrix}
+  =
+  \begin{bmatrix}
+  0\\0\end{bmatrix}
+\label{eq:resid}
+\end{equation}
+for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the
+$\ell_\infty$ constraint, as $w\rightarrow\infty$,
+the solution to this nonlinear system reduces to the KKT
+conditions. Therefore, the PDIP method takes Newton steps to solve the
+system for a series of increasing values of $w$. \autoref{alg:pdip}
+provides the details. 
+\begin{algorithm}[tb]
+  \caption{PDIP for $\ell_1$ variance estimation}
+  \label{alg:pdip}
+  \begin{algorithmic}
+    \REQUIRE $\lambda>0$, $w>0$, $\nu\leftarrow 0$, $\mu_1\leftarrow
+    0$, $\mu_2\leftarrow 0$, $J\in\mathbb{Z}^+$, $\{w_k\}$
+    \COMMENT{Initialization} 
+    \FOR[Central path]{$k=1,2,\ldots$}
+    \FOR[Newton updates]{$j=1,\ldots,J$}
+    \STATE Solve $A [\Delta\nu\; \Delta\mu_1\; \Delta\mu_2]^\top=r_{w_k}$ to find the search direction.
+    \STATE $\quad\quad A$ is the Jacobian of $r_w$ in \eqref{eq:resid}.
+    \STATE Update $[\nu^{j+1}\; \mu^{j+1}_1\; \mu^{j+1}_2] \leftarrow
+    [\nu^{j}\; \mu^{j}_1\; \mu^{j}_2] + [\Delta\nu\; \Delta\mu_1\;
+    \Delta\mu_2]$. 
+    \ENDFOR
+   \ENDFOR
+   \RETURN $h=\log\frac{y^2}{1+D^\top\nu}$ \attn{Is this right?
+     Explicit form of $A$}
+  \end{algorithmic}
+\end{algorithm}
 
-\subsection{Extension to spatio-temporal data}
+\subsection{Adding spatial constraints}
 \label{sec:exten}
 
-The method in the previous section can be used to estimate the variance of a single time-series. In this section, we extend this method to the estimation of the variance of spatio-temporal data.
+The method in the previous section can be used to estimate the
+variance of a single time-series. In this section, we extend this
+method to the estimation of the variance of spatio-temporal data.
 
-At a specific time $t$, the data is measured on a grid of points with $n_r$ rows and $n_c$ columns.
-Let $y_{ijt}$ denote the value of the observation at time $t$ on the $i^\text{th}$ row and $j^\text{th}$ column of the grid, and $h_{ijt}$ denote the corresponding hidden variable.
-We seek to impose both temporal and spatial smoothness constraints on the hidden variables.
-Specifically, we seek a solution for $h$ which is piecewise linear in time and piecewise constant in space \footnote{The assumption that the variance is spatially piecewise constant simplifies the computations.
-The justification for this assumption is that we are interested in examining the trend in the variance over time and not over space.}.
-This can be achieved by solving the following optimization problem:
+At a specific time $t$, the data is measured on a grid of points with
+$n_r$ rows and $n_c$ columns for a total of $S=n_r\times n_c$ spatial
+locations. Let $y_{ijt}$ denote the value of the 
+observation at time $t$ on the $i^\text{th}$ row and $j^\text{th}$
+column of the grid, and $h_{ijt}$ denote the corresponding hidden
+variable. We seek to impose both temporal and spatial smoothness
+constraints on the hidden variables. Specifically, we seek a solution
+for $h$ which is piecewise linear in time and piecewise constant in
+space (although higher-order smoothness can be imposed with minimal
+alterations to the methodology). We achieve this goal
+by solving the following optimization problem: 
+\begin{align}
+\min_h &\sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}}
++\lambda_1 \sum_{i,j} \sum_{t=1}^{T-2} \left|h_{ijt}-2h_{ij(t+1)}+h_{ij(t+2)}\right|\\
+&\quad+\lambda_2 \sum_{t,j} \sum_{i=1}^{n_r-1} \left|h_{ijt}-h_{(i+1)jt}\right|
++\lambda_2 \sum_{t,i} \sum_{j=1}^{n_c-1} \left|h_{ijt}-h_{i(j+1)t}\right|
+\label{eq:l1tf_var_st}
+\end{align}
 
+The first term in the objective is proportional to the negative
+log-likelihood, the second is the temporal penalty for the
+time-series at each location $(i,j)$, while the third and fourth,
+penalize the difference between the estimated variance of two
+vertically and horizontally adjacent points, respectively. The spatial
+component of this
+penalty is a special case of trend
+filtering on graphs~\citep{wang_trend_2016} which penalizes the difference
+between the estimated values of the signal on the connected
+nodes. As before, we can write
+\eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times
+S$ vector and $D$ is replaced by $D_{TS}$. Then, as we
+have two different tuning parameters for the temporal and spatial
+components, we write $\lambda =\left[\lambda_1\one_T^\top,\;
+  \lambda_2\one_{S}^\top\right]^\top$ and $\Lambda =\texttt{diag}(\lambda)$ leading to 
+% let $h$ be a
+% vector whose first $T$ entries are $h_{11t}$ for $t=1,...,T$, its next
+% $T$ entries are $h_{21t}$ and so on, then the optimization problem in
+% the matrix form is as
 \begin{equation}
-\begin{aligned}
-\min_h &\sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}}\\
-&+\lambda_t \sum_{i,j} \sum_{t=1}^{T-2} \big|h_{ijt}-2h_{ij(t+1)}+h_{ij(t+2)}\big|\\
-&+\lambda_s \sum_{t,j} \sum_{i=1}^{n_r-1} \big|h_{ijt}-h_{(i+1)jt}\big|\\
-&+\lambda_s \sum_{t,i} \sum_{j=1}^{n_c-1} \big|h_{ijt}-h_{i(j+1)t}\big|
-\end{aligned}
-\label{eq:l1tf_var_st}
-\end{equation}
-
-The first term in the objective is the negative log-likelihood (minus a constant term).
-The second term is the temporal penalty for the time-series at each location $(i,j)$.
-The third and fourth terms, penalize the difference between the estimated variance of two vertically and horizontally adjacent points, respectively.
-This penalty is a special case of the penalty used in the \textit{trend filtering on graphs }\citep{wang_trend_2016} (where the difference between the estimated values of the signal at each two nodes connected with an edge is penalized.).
-The optimization problem \eqref{eq:l1tf_var_st} can be written in the matrix form.
-Let $h$ be a vector whose first $T$ entries are $h_{11t}$ for $t=1,...,T$, its next $T$ entries are $h_{21t}$ and so on.
-Then the optimization problem in the matrix form is as follows:
-
-\begin{equation}
-\min_h -l(y|h)+ \Lambda^t \lVert D_{total}h \lVert_1
+\min_h -l(y\given h)+ \norm{\Lambda D_{TS}h }_1.
 \label{eq:l1tf_var_st_mat}
 \end{equation}
-
-Let, $\textbf{e}_n$ denote a vector of size $n$ with all entries being equal to 1.
-We have: $\Lambda^t=(\lambda_t\textbf{e}_{n_t}^t|\lambda_s\textbf{e}_{n_s}^t)$, $D_{total}^t=[D_{t}^t|D_{s}^t]$, where $n_t$ and $n_s$ are the number of rows of $D_{t}$ and $D_{s}$, respectively (see below).
-The matrix $D_t$ is the following block-diagonal matrix and corresponds to the temporal penalty:
-
-\begin{equation}
-D_t=\begin{bmatrix}
-D &  & \\ 
- & \ddots & \\
- &  & D
-\end{bmatrix}
-\label{eq:d_t_matrix}
-\end{equation}
-
- where $D$ was defined in \eqref{eq:d_matrix}.
-The number of the diagonal blocks is equal to the grid size $n_r \times n_c$.
-Each row of the matrix $D_s$ corresponds to one spatial constraint in \eqref{eq:l1tf_var_st}.
-For example, the first $T$ rows correspond to $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to $|h_{11t}-h_{12t}|$, and so on.
-
-The dual of this problem is:
-
-\begin{equation}
-\begin{aligned}
-\min_\nu f^*(-D_{total}^t\mathsf{\nu}) \\
-s.t. \,\,\, | \nu_i | \le \Lambda_i
-\end{aligned}
-\label{eq:dual}
-\end{equation}
-
+% where $\Lambda = \mathbf{I}_{Tn_rn_c}\left[\lambda_1\one_T^\top,\;
+%   \lambda_2\one_{n_rn_c}^\top\right]$ and 
+% $D_{total}^t=[D_{t}^t|D_{s}^t]$, where $n_t$ and $n_s$ are the number
+% of rows of $D_{t}$ and $D_{s}$, respectively (see below). The matrix
+% $D_t$ is the following block-diagonal matrix and corresponds to the
+% temporal penalty: 
+% \begin{equation}
+% D_t=\begin{bmatrix}
+% D &  & \\ 
+%  & \ddots & \\
+%  &  & D
+% \end{bmatrix}
+% \label{eq:d_t_matrix}
+% \end{equation}
+%  where $D$ was defined in \eqref{eq:d_matrix}. The number of the
+%  diagonal blocks is equal to the grid size $n_r \times n_c$. Each row
+%  of the matrix $D_s$ corresponds to one spatial constraint in
+%  \eqref{eq:l1tf_var_st}. For example, the first $T$ rows correspond to
+%  $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to
+%  $|h_{11t}-h_{12t}|$, and so on. 
+% \attn{Detailed discussion of the PDIP to here}
+The dual of this problem is 
+$\min_\nu f^*(-D_{ST}^\top\nu)$ subject to 
+$| \nu_k | \le \Lambda_{kk},\; k=1,\ldots,ST$,
  where $f^*$ is given in \eqref{eq:conj}.
 
-\section{Optimization}
-For a spatial grid of size $n_r \times n_c$ and for $T$ time steps, we have $n_t=3n_rn_c-Tn_c-2n_rn_c$ and $n_s=n_rn_cT$.
-For a grid over the united states and for weekly averaged data of 10 years we have $n_r=32$, $n_c=68$, $T=521$ and so $n_t\approx 3.5\times10^6$ and $n_s \approx 1.0\times10^6$.
-Therefore, the size of the optimization problem \eqref{eq:dual} is about $n_t\approx 4.5\times10^6$.
-In each step of the PDIP algorithm, we need to solve a linear system of equations in the form $z=Ax$ where $A$ is a square matrix of size $2(n_t+n_s)$ (see \citep{boyd_convex_2004} equation 11.54).
-Therefore, applying the PDIP directly for solving the optimization problem \eqref{eq:dual} is infeasible.
+\attn{Add back the full form of $D$ above? It was too cumbersome
+  before, but possibly necessary}
 
-In the next section, we develop an ADMM algorithm for solving this problem.
-The idea is to cast the problem \eqref{eq:dual} as a so-called \textit{consensus optimization problem} \citep{boyd_distributed_2011} and solve it by breaking the problem into smaller sub-problems using an ADMM-based algorithm.
-Next, we first give a brief overview of the consensus optimization problem and then explain how the problem \eqref{eq:l1tf_var_st_mat} can be solved in this framework.
+\section{Proposed optimization methods}
+\label{sec:prop-optim-meth}
+
+For a spatial grid of size $S$ and $T$ time steps, $D_{ST}$ will have
+$3n_rn_c+Tn_c+2Tn_r$ rows and $ST$ columns. For a $1^\circ\times
+1^\circ$ grid over the
+entire northern hemisphere and daily data over 10 years, we have
+$n_r=90$, $n_c=180$, $T=3650$ and so $D_{ST}$ has approximately
+$10^8$ columns and $10^8$
+rows. In each
+step of \autoref{alg:pdip}, we need to solve a linear system of
+equations in $A$ which depends on $D_{ST}^\top D_{ST}$ (see \citep{boyd_convex_2004}
+equation 11.54). Therefore, 
+applying the PDIP directly is infeasible for our data.\footnote{We note that this
+  is a highly structured and sparse matrix, but, unlike trend
+  filtering alone, it is not banded. We are unaware of general
+  linear algebra techniques for inverting such matrix, despite our
+  best efforts.}
+
+In the next section, we develop two ADMM algorithms for solving this
+problem efficiently. The first casts the problem as a
+so-called consensus optimization problem~\citep{boyd_distributed_2011}
+which solves smaller sub-problems using PDIP and then recombines the results. The
+second uses proximal methods to avoid matrix inversions. 
+\attn{We note that stochastic gradient descent could be used but...}
 
@@
 
-A general form consensus optimization problem is in the form $\min_z \sum_i f_i(z)$, $z\in\mathbb{R}^n$.
-We can define a set of \textit{local variables} $x_i\in \mathbb{R}^{n_i}$ such that $\sum_i f_i(z)=\sum_i f_i(x_i)$.
-We follow the notation of \citep{boyd_distributed_2011} closely.
-Let $k=\mathscr{G}(i,j)$ which means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or $(x_i)_j=z_k$) and define $\tilde{z}_i \in \mathbb{R}^{n_i}$ by $(\tilde{z}_i)_j=(x_i)_j$.
-Then the original unconstrained optimization problem is equivalent to the following constrained optimization problem:
+General (unconstrained) optimization problems $\min_z
+\sum_i f_i(z)$, $z\in\mathbb{R}^n$ can be rewritten equivalently using a
+collection of local variables, $x_i\in \mathbb{R}^{n_i}$, and a
+constraint. Following the notation of 
+\citep{boyd_distributed_2011}, let $k=\mathscr{G}(i,j)$ which
+means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or
+$(x_i)_j=z_k$) and define $\tilde{z}_i \in \mathbb{R}^{n_i}$ by
+$(\tilde{z}_i)_j=(x_i)_j$. Then the original unconstrained
+optimization problem is equivalent to the following constrained
+problem
+$\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)$ subject to $\tilde{z}_i=x_i$.
+Now, we can apply ADMM to the augmented Lagrangian of this
+problem: 
+\begin{align}
+x_i&\leftarrow \argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
+     (\rho/2) \norm{ x_i-\tilde{z}_i }_2^2\\
+\label{eq:ADMM_steps}
+z_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j\\
+u_i&\leftarrow u_i + \rho (x_i-\tilde{z}_i)
+\end{align}
+Here, $S_k$ is the number of local variable entries that correspond to
+$z_k$, and $u_i$ are the Lagrange multipliers.  The $x$-update is
+often a simple modification as the original optimization, though over only a small
+partition, the $z$-update simply averages over collections of
+analogous local variables while the $u$ update enforces sharing. Both
+the $x$ and $u$ updates are parallelizable over the partition of local
+variables.
+\attn{I don't much like this explanation or notation. Seems unclear}
 
-\begin{equation}
-\begin{aligned}
-\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)\\
- s.t. \,\,\, \tilde{z}_i=x_i \,\,\,\,\,\,
-\end{aligned}
-\end{equation}
+To solve the optimization problem \eqref{eq:l1tf_var_st_mat} or
+\eqref{eq:l1tf_var_st} using this method, we need to two address two
+questions: first, how to choose the 
+local variables $x_i$, and second, how to solve the optimization
+problem for updating these variables (the first line of
+\eqref{eq:ADMM_steps}).
 
-Now, we can apply ADMM to the \textit{augmented Lagrangian} of this problem. This results in the following ADMM updating steps at each iteration $m$:
-
-\begin{equation}
-\begin{aligned}
-& x_i^{m+1}:=\argmin_{x_i} \bigg( f_i(x_i) + (u_i^m)^t x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2\bigg)\\
-& z_k^{m+1}:=(1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i^{m+1})_j\\
-& u_i^{m+1}:=u_i^m + \rho (x_i^{m+1}-\tilde{z}_i^{m+1})
-\end{aligned}
-\label{eq:ADMM_steps}
-\end{equation}
-
-Here, $S_k$ is the number of local variable entries that correspond to $z_k$, and $u_i$ are the Lagrange multipliers.
-
-To solve the optimization problem \eqref{eq:l1tf_var_st_mat} or \eqref{eq:l1tf_var_st} using the method explained in the previous section we need to two address two questions: first, how to choose the local variables $x_i$, and second, how to solve the optimization problem for updating these variables (the first line of \eqref{eq:ADMM_steps}).
-
-Obviously, the global variable in the problem \eqref{eq:l1tf_var_st} is $z=h$.
-The global variable can be index by the spatial coordinate and the temporal step.
-In \autoref{fig:data_cube}, $z$ is represented as a cube.
-We can decompose $z$ into sub-cubes.
-An example of this decomposition is shown in the figure by white lines.
-It is easy to see that with this definition of $x_i$, the objective \eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where $f_i(x_i)=-l(y_i|x_i)+\lVert \Lambda_i^t D_ix_i \lVert_1$.
-
-%TODO: re-write this part. it is very confusing.
+In \autoref{fig:data_cube}, the global variable $z$ is represented
+as a cube (using the subset of the US as an example). We decompose $z$
+into sub-cubes as shown in the figure by white lines. It is easy to see
+that with this definition of $x_i$, the objective
+\eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where
+$f_i(x_i)=-l(y_i\given x_i)+\norm{\Lambda_{(i)} D_{(i)}x_i}$, and
+$\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
+penalties corresponding to $x_i$ only.
+\attn{I would like to be able to write all this in terms of $h$
+  instead of $x$}
 
 \begin{figure}[tb]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=.5\columnwidth]{Figures/data_cube}}
-		\caption{The cube represents the global variable $z$ in space and time. The sub-cubes specified by the white lines are $x_i$.}
+  \centering
+  \includegraphics[height=.2\textheight]{Figures/data_cube}
+    \caption{The cube represents the global variable $z$ in space and
+      time. The sub-cubes specified by the white lines are
+      $x_i$. \attn{Can we use $h_i$ here instead of $x_i$?}}
     \label{fig:data_cube}
-	\end{center}
-	\vskip -0.2in
 \end{figure}
 
-By this definition of $x_i$, the update step for $x_i$ is the following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) + (u_i^m)^t x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
+By this definition of $x_i$, the update step for $x_i$ is the
+following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
+(u_i^m)^t x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
 
-We solve this using the PDIP method. To this end we first need to compute the dual of the problem. It can be shown that the dual of this problem is:
+We solve this using the PDIP method. To this end we first need to
+compute the dual of the problem. It can be shown that the dual of this
+problem is: 
+%TODO: write the appendix
+\begin{align}
+\min_\nu &f_i^*(-D_i^t\nu_i) & s.t. &| (\nu_i)_j | \le (\Lambda_i)_j
+\label{eq:dual_x_update}
+\end{align}
+ where:
+\begin{align}
+f^*_i(q)&=\sum_j (q_i)_j (w_i)_j -(w_i)_j-(y_i)_j^2 e^{-(w_i)_j} - 
+(\rho/2)((w_i)_j-(\alpha_i^m)_j)^2\\
+\alpha_i^m &=\hat{z}_i^m-u_i^m \label{eq:conj_x_update}\\
+(w_i)_j&=\mathscr{W}\bigg(\frac{(y_i)_j^2}{\rho}\exp
+         \bigg[\frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}\bigg]\bigg)- 
+\frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}
+\end{align}
+\attn{I'm not quite sure what's going on here. These need to match the
+  generic discussion in \eqref{eq:ADMM_steps}. Or are these just
+  defining some other variables for convenience? This needs to be
+  clarified and incorporated into \autoref{alg:ADMM}.}
 
-%TODO: write the appendix
+In these equations, $\mathscr{W}$ is the \textit{Lambert function}
+\citep{corless_lambertw_1996}. To use the PDIP method, we also need to
+compute the gradient and Hessian of $f_i^*(-D_i^t\nu_i)$. This
+involves computing the derivatives of the Lambert functions. The
+primal solution $x_i$ can be obtained from the dual solution $\nu_i$
+in \eqref{eq:dual_x_update} by setting $x_i=w_i$ where $w_i$ is
+defined in the last equation of \eqref{eq:conj_x_update}.
 
-\begin{equation}
-\begin{aligned}
-\min_\nu f_i^*(-D_i^t\nu_i) \\
-s.t. \,\,\, | (\nu_i)_j | \le (\Lambda_i)_j
-\end{aligned}
-\label{eq:dual_x_update}
-\end{equation}
-
- where:
-
-\begin{equation}
-\begin{aligned}
-f^*_i(q)&=\sum_j (q_i)_j (w_i)_j -(w_i)_j-(y_i)_j^2 e^{-(w_i)_j} - \\
-& (\rho/2)((w_i)_j-(\alpha_i^m)_j)^2\\
-&\alpha_i^m =\hat{z}_i^m-u_i^m\\
-(w_i)_j&=\mathscr{W}\bigg(\frac{(y_i)_j^2}{\rho}\exp\bigg[\frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}\bigg]\bigg)-\\
-& \frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}
-\end{aligned}
-\label{eq:conj_x_update}
-\end{equation}
-
-In these equations, $\mathscr{W}$ is the \textit{Lambert function} \citep{corless_lambertw_1996}.
-To use the PDIP method, we also need to compute the gradient and Hessian of $f_i^*(-D_i^t\nu_i)$.
-This involves computing the derivatives of the Lambert functions.
-The primal solution $x_i$ can be obtained from the dual solution $\nu_i$ in \eqref{eq:dual_x_update} by setting $x_i=w_i$ where $w_i$ is defined in the last equation of \eqref{eq:conj_x_update}.
-
-The complete ADMM algorithm for estimating the variances is represented in \autoref{alg:ADMM}.
-All the computations in the three updating steps \eqref{eq:ADMM_steps} can be performed in parallel.
-The number of rows and columns of the sub-cubes should be chosen so that the updating of $x_i$ could be performed in one processor.
-We choose $3\times3\times521$ sub-cubes.
+The complete ADMM algorithm for estimating the variances is
+represented in \autoref{alg:ADMM}. All the computations in the three
+updating steps \eqref{eq:ADMM_steps} can be performed in parallel. The
+number of rows and columns of the sub-cubes should be chosen so that
+the updating of $x_i$ could be performed in one processor. We choose
+$3\times3\times521$ sub-cubes.
 
 \begin{algorithm}[tb]
-	\caption{ADMM for sparse estimation of variance of spatio-temporal data}
+  \caption{ADMM for sparse estimation of variance of spatio-temporal
+    data \attn{Fix this}}
   \label{alg:ADMM}
   \begin{algorithmic}
-		\STATE {\bfseries Input:} data $y$, mapping $\mathscr{G}(i,j)$, $\rho,\lambda_t,\lambda_s$
+    \STATE {\bfseries Input:} data $y$, mapping $\mathscr{G}(i,j)$,
+    $\rho,\lambda_t,\lambda_s$ 
     \STATE Initialization: $x_i^0=z^0=u_i^0=\textbf{0}$.
@@
 
-This algorithm divides the main optimization problem into several sub-problems and in each iteration solves these sub-problems and then performs a consensus step to make the solutions of these sub-problems agree with each other.
-These sub-problems can be solved independently which makes this algorithm appealing for parallelization: in each iteration, each of these sub-problems can be sent into a separate computer and then all the solutions can be sent into a single computer which performs the consensus step.
-Therefore, in each iteration, the computation time will be equal to the time of solving each sub-problem plus the time of communicating the solutions to the master computer and then performing the consensus step.
-Since each sub-problem is small, with parallelization, the computation time in each iteration will be small.
-In addition, our experiments with several values of $\lambda_t$ and $\lambda_s$ showed that the algorithm converges in few hundreds iterations.
-Solving each sub-problem on a machine with four 3.20GHz Intel i5-3470 cores takes less than 3 seconds on average, and so for example if we assume that communication time is 10 seconds and the algorithm converges in 300 iterations, with parallelization on $N_{sub-cubes}$ machines, the algorithm will converge in about 1 hour.
-Assuming that we use $N_{sub-cubes}$ machines and that the convergence rate of the algorithm is independent of the grid size, this time will be independent of the grid size.
+Because \autoref{alg:ADMM} breaks the large optimization into
+sub-problems that can be solved independently, it is amenable to a
+split-gather parallelization strategy via, e.g., the map reduce framework.
+In each iteration, the
+computation time will be equal to the time to solve each sub-problem
+plus the time to communicate the solutions on the master processor
+and perform the consensus step. Since each sub-problem is
+small, with parallelization, the computation time in each iteration
+will be small. In addition, our experiments with several values of
+$\lambda_t$ and $\lambda_s$ showed that the algorithm converges in few
+hundreds iterations. 
+\attn{Need to redo this:}Solving each sub-problem on a machine with four
+3.20GHz Intel i5-3470 cores takes less than 3 seconds on average, and
+so for example if we assume that communication time is 10 seconds and
+the algorithm converges in 300 iterations, with parallelization on
+$N_{sub-cubes}$ machines, the algorithm will converge in about 1
+hour. Assuming that we use $N_{sub-cubes}$ machines and that the
+convergence rate of the algorithm is independent of the grid size,
+this time will be independent of the grid size.
 
-If we perform these computations on a single machine, the computation time grows linearly with $N_{sub-cubes}$.
-For example, for the data in a grid over the united states and using $3\times3\times521$ sub-cubes each iteration of the algorithm will take about 20 minutes on a single machine and so with 300 iterations it will take several days to converge.
-Given that we need to compute the solution for several values of the parameters $\lambda_t$ and $\lambda_s$, this computation time is not feasible.
+If we perform these computations on a single machine, the computation
+time grows linearly with $N_{sub-cubes}$. For example, for the data in
+a grid over the united states and using $3\times3\times521$ sub-cubes
+each iteration of the algorithm will take about 20 minutes on a single
+machine and so with 300 iterations it will take several days to
+converge. Given that we need to compute the solution for several
+values of the parameters $\lambda_t$ and $\lambda_s$, this computation
+time is not feasible.
 
-Therefore, this algorithm is only useful if we can parallelize the computation over several machines.
-In the next section, we describe another algorithm which makes the computation feasible on a single machine.
+Therefore, this algorithm is only useful if we can parallelize the
+computation over several machines. In the next section, we describe
+another algorithm which makes the computation feasible on a single
+machine.
 
@@
 
-In this section, we describe \textit{Linearized ADMM algorithm} \citep{parikh_proximal_2014} which, as we will see, makes the computation on a single machine feasible.
+% In this section, we describe \textit{Linearized ADMM algorithm} \citep{parikh_proximal_2014} which, as we will see, makes the computation on a single machine feasible.
 
-Consider the following optimization problem:
-
+\attn{Need a good dummy variable here, it can't be $u$.} Consider the generic optimization problem
+$\min_x f(x)+g(Dx)$
+where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each
+iteration of the linearized ADMM
+algorithm~\citep{parikh_proximal_2014} for solving this problem 
+has the form
+\begin{align}
+x & \leftarrow \prox_{\mu f} \left(x - (\mu/\rho)D^T (D x - z + u )\right)\\
+z & \leftarrow \prox_{\rho g} \left(z + u\right) \label{eq:linADMM_steps}\\
+u & \leftarrow u + D x - z
+\end{align}
+ where $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
 \begin{equation}
-\min_x f(x)+g(Dx)
-\label{eq:linADMM_opt}
-\end{equation}
-
- where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each iteration of the linearized ADMM algorithm for solving this problem has the following form:
-
-\begin{eqnarray}
-\begin{aligned}
-x^{k+1} & := \textbf{prox}_{\mu f} \big(x^k - (\mu/\rho)D^T (D x^k - z^k + u^k )\big)\\
-z^{k+1} & := \textbf{prox}_{\rho g} \big(D x^k + u^k\big)\\
-u^{k+1} & := u^k + D x^{k+1} - z^{k+1}
-\end{aligned}
-\label{eq:linADMM_steps}
-\end{eqnarray}
-
- where $z,u\in \mathbb{R}^m$ and the \textit{proximal operator} is defined as follows:
-
-\begin{equation}
- \textbf{prox}_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \lVert x-u \rVert^2
+ \prox_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \norm{ x-u}_2^2.
 \label{eq:linADMM_prox}
 \end{equation}
+Clearly,
+\eqref{eq:l1tf_var_st} has this form.\attn{What are $\mu$ and $\rho$?}
+% The optimization problem 
+% form \eqref{eq:linADMM_opt} by defining
+% \begin{align}
+% & f(x):= \sum_{k} f_k(x_k) := \sum_{k}x_k+y_{k}^2e^{-x_{k}}\\
+% & g(z):= \sum_{l} g_l(z_l) := \sum_{l} \lambda_l |z_l|\label{eq:linADMM_fg}\\
+% & z = Dx
+% \end{align}
+% where $y_k$ is the $k^{th}$ entry of the vector whose entries are
+% $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector
+% $\Lambda^t=(\lambda_t\textbf{e}_{n_t}^t|\lambda_s\textbf{e}_{n_s}^t)$
+% (see \autoref{sec:exten}). 
+To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate
+$\prox_{\mu f}$ and $\prox_{\rho g}$. Proximal
+algorithms are feasible only if these proximal operators can be
+evaluated efficiently which, as we show next, is the case for our
+problem.
 
-This algorithm belongs to the general class of \textit{proximal algorithms} for solving convex optimization problems. For more details about these algorithms see \citep{parikh_proximal_2014}.
+\begin{theorem}
+  Let $f(h) = \sum_{k} h_{k} + y_{k}^2e^{-h_{k}}$ and $g(x) =
+  \norm{x}_1$. Then, 
+  \begin{align}
+    [\prox_{\mu f}(u)]_k &= \mathscr{W}\bigg(\frac{y_k^2}{\mu}
+    \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu u_k}{\mu},\\
+    \prox_{\rho g}(u) &= S_{\rho\lambda}(u)
+  \end{align}
+where $[S_{\alpha}(u)]_k = \sign(u_k)(|u_k| -\alpha_k)_+$ and
+$(v)_+=v\vee 0$.
+\end{theorem}
+\begin{proof}
+  If $f(x)=\sum_k f_k(x_k)$ then $[\prox_{\mu f}(x)]_k =
+  \prox_{\mu f_k}(u_k)$. So 
+  $[\prox_{\mu f}(u)]_k=\min_{x_k} \,\,
+  \mu\big(x_k+y_{k}^2e^{-x_{k}}\big)+\frac{1}{2}  (x_k-u_k)^2.$
+  Setting the derivative to 0 and solving for $u_k$ gives the
+  result. Similarly, $[\prox_{\rho g}(u)]_\ell=\rho
+  \lambda_\ell |z_\ell|+1/2(z_\ell-u_\ell)^2$. This is not differentiable,
+  but the solution must satisfy $\rho \cdot \lambda_\ell \cdot \partial
+  \big(|z_\ell| \big)=u_\ell-z_\ell$ where $\partial \big(|z_\ell| \big)$ is the
+  sub-differential of $|z_\ell|$. The solution is the soft-thresholding
+  operator $S_{\rho\lambda_\ell}(u_\ell)$.
+\end{proof}
 
-The optimization problem \eqref{eq:l1tf_var_st} can be put into the form \eqref{eq:linADMM_opt} as follows:
-
-\begin{equation}
-\begin{aligned}
-& f(x):= \sum_{k} f_k(x_k) := \sum_{k}x_k+y_{k}^2e^{-x_{k}}\\
-& g(z):= \sum_{l} g_l(z_l) := \sum_{l} \lambda_l |z_l|\\
-& z= Dx
-\end{aligned}
-\label{eq:linADMM_fg}
-\end{equation}
-
- where $y_k$ is the $k^{th}$ entry of the vector whose entries are $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector $\Lambda^t=(\lambda_t\textbf{e}_{n_t}^t|\lambda_s\textbf{e}_{n_s}^t)$ (see \autoref{sec:exten}).
-
-To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate $\textbf{prox}_{\mu f}$ and $\textbf{prox}_{\rho g}$.
-The proximal algorithms are feasible only if these proximal operators can be evaluated efficiently which, as we show next, is the case for our problem.
-
-Let $\big(\textbf{prox}_{\mu f}(u)\big)_k$ be the $k^{th}$ entry of $\textbf{prox}_{\mu f}(u)$.
-From the \textit{separable sum} property of the proximal operators we have (see \citep{parikh_proximal_2014}, section 2.1):
-
-\begin{equation}
-\bigg(\textbf{prox}_{\mu f}(u)\bigg)_k = \textbf{prox}_{\mu f_k}(u_k)\\
-\label{eq:linADMM_sepSum}
-\end{equation}
-
-Similarly, evaluating $\textbf{prox}_{\rho g}$ reduces to evaluating the proximal operators of scalar functions. We have:
-
-\begin{equation}
-\textbf{prox}_{\mu f_k}(u_k)=\min_{x_k} \,\, \mu\big(x_k+y_{k}^2e^{-x_{k}}\big)+\frac{1}{2}  (x_k-u_k)^2\\
-\label{eq:linADMM_proxf}
-\end{equation}
-
-By setting the derivative with respect to $x_k$ of the function to zero we obtain:
-
-\begin{equation}
-\textbf{prox}_{\mu f_k}(u_k)=\mathscr{W}\bigg(\frac{y_k^2}{\mu} \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu u_k}{\mu}\\
-\label{eq:linADMM_proxf_sol}
-\end{equation}
-
-where $\mathscr{W}$ is the Lambert function.
-
-Next we compute $\textbf{prox}_{\mu g_l}(u_l)$.
-The function $\rho \lambda_l |z_l|+1/2(z_l-u_l)^2$ is not differentiable.
-However, at the optimal solution we have: $\rho \cdot \lambda_l \cdot \partial \big(|z_l| \big)=u_l-z_l$, where $\partial \big(|z_l| \big)$ is the sub-gradient of $|z_l|$.
-This results in the following solution:
-
-\begin{equation}
-\textbf{prox}_{\rho g_l}(u_l)=\left\{\begin{matrix}
- u_l-\rho \lambda_l  & if &  u_l>\rho \lambda_l\\ 
- 0  \,\,\,\,\,  & if &  |u_l|\le \rho \lambda_l\\ 
- u_l+\rho \lambda_l  &  if &  u_l<-\rho \lambda_l\\
-
-\end{matrix}\right.
-\label{eq:linADMM_proxg_sol}
-\end{equation}
-
-Therefore, both proximal operators in \eqref{eq:linADMM_steps} can be evaluated easily and so we can use the linearized ADMM algorithm to solve the optimization problem \eqref{eq:l1tf_var_st}.
+\begin{algorithm}[tb]
+  \caption{Linearized ADMM }
+  \label{alg:linADMM}
+  \begin{algorithmic}
+    \STATE {\bfseries Input:} data $y$, penalty matrix $D$,
+    $\rho,\lambda_t,\lambda_s >0$.
+    \STATE {\bf Set:} $h\leftarrow 0$, $z\leftarrow 0$, $u\leftarrow 0$. 
+    \FOR{$m=1,2,...$ }
+    \STATE $h_k\leftarrow \mathscr{W}\bigg(\frac{y_k^2}{\mu}
+    \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu u_k}{\mu}$,
+    \STATE $z\leftarrow S_{\rho\lambda}(u)$.
+    \STATE $u\leftarrow u + Dh-z$
+    \ENDFOR
+  \end{algorithmic}
+\end{algorithm}
 
 \section{Empirical evaluation}
+\label{sec:empirical-evaluation}
+
+In this section, we examine both simulated and real spatio-temporal
+climate data. All the computations were performed on a Linux machine
+with four 3.20GHz Intel i5-3470 cores.
 
@@
 
-To analyze the proposed optimization methods, we first fit the model on some synthetic spatio-temporal data.
-The observations at all time steps and all locations were generated from independent Gaussian random variables with zero mean.
-However, the variance of these random variables changes smoothly in time and space.
-Specifically, the observation at time step $t$ and at location $(r,c)$ on the grid is generated from a Gaussian $N(0,\sigma^2(t,r,c))$, where:
+\attn{Can we add some type of performance measure? What if we fit 2
+  marginal models, spatial only and temporal only. Plus maybe a
+  marginal GARCH?}
 
-\begin{equation}
-\begin{aligned}
-\sigma(t,r,c) & =\sum_{s=1}^{S} W_s(t) \cdot \exp\bigg( \frac{(r-r_s)^2+(c-c_s)^2}{2\sigma_s^2} \bigg) \\
-W_s(t) & =\alpha_s \cdot t + \exp(\sin(2\pi\omega_s t+\phi_s)) 
+We generate observations at all time steps and all locations from
+independent Gaussian random variables with zero mean. However, the
+variance of these random variables follows a smoothly varying function
+in time and space
+\begin{align}
+\sigma^2(t,r,c) & =\sum_{s=1}^{S} W_s(t) \cdot \exp\bigg( \frac{(r-r_s)^2+(c-c_s)^2}{2\sigma_s^2} \bigg); &
+W_s(t) & =\alpha_s \cdot t + \exp(\sin(2\pi\omega_s t+\phi_s)) .
 \label{eq:sourceVar}
-\end{aligned}
-\end{equation}
-
-In words, the variance at each time and location is computed as the weighted sum of $S$ bell-shaped functions where the weights are time-varying functions.
-Specifically, the weights consist of a linear trend $\alpha_s \cdot t$ and a periodic term $\beta_s \cdot \sin(2\pi\omega_s t+\phi_s)$.
-The bell-shaped functions impose the spatial smoothness, and the linear trend and the periodic terms enforce the temporal smoothness.
-We simulated the data on a 5 by 7 grid and for 780 time steps.
-We also used $S=4$.
-The parameters of the variance function are shown in  \autoref{tab:sim_params}.
-The value of the variance function for all locations on the grid and at $t=25$ and $t=45$ is shown in \autoref{fig:true_var_spatial}.
-Also, the variance for all time steps and at the location (0,0) on the grid is shown in \autoref{fig:true_fitted_var}.
-The linear trend and the period term can be seen clearly in this figure.
+\end{align}
+In words, the variance at each time and location is computed as the
+weighted sum of $S$ bell-shaped functions where the weights are
+time-varying, consist of a linear trend $\alpha_s \cdot t$ and a
+periodic term $\beta_s \cdot \sin(2\pi\omega_s t+\phi_s)$. The
+bell-shaped functions impose the spatial smoothness, and the linear
+trend and the periodic terms enforce the temporal smoothness similar
+to the seasonal component in the real climate data. We
+simulated the data on a 5 by 7 grid and for 780 time steps with
+$S=4$. The parameters of the variance function are shown in
+\autoref{tab:sim_params}. For reference, we plot the variance function for all
+locations at $t=25$ and $t=45$ in
+as well as
+the variance across time at $(0,0)$ in
+\autoref{fig:true_var_spatial}.
 
 \begin{table}[tb]
-	\caption{Parameters used to simulate data}
+  \caption{Parameters used to simulate data. \attn{Any ideas to take
+      up less space with this info?}}
   \label{tab:sim_params}
-	\vskip 0.15in
   \begin{center}
-		\begin{small}
-			\begin{sc}
     \begin{tabular}{ccccccc}
       \hline
-
       $s$ & $r_s$ & $c_s$ & $\sigma_s$ &$\alpha_s$ & $\omega_s$ & $\phi_s$\\
@@
     \end{tabular}
-			\end{sc}
-		\end{small}
   \end{center}
-	\vskip -0.1in
 \end{table}
@@
   \centering	
-	\includegraphics[width=.7\linewidth]{Figures/true_var_spatial}
-
-	\caption{Variance function at $t=25$ (left) and $t=45$ (right)} \label{fig:true_var_spatial}
+  \includegraphics[height=.15\textheight]{Figures/true_var_spatial}
+  \includegraphics[height=.15\textheight]{Figures/true_fitted_var}
+  \caption{Variance function at $t=25$ (left) and $t=45$
+    (center). Right: the true (orange) and estimated standard deviation 
+    function at the location (0,0). The estimated values are
+    obtained using linearized ADMM with $\lambda_s=0.1$ and two
+    values of $\lambda_t$: $\lambda_t=5$ (blue) and
+    $\lambda_t=100$ (green).} \label{fig:true_var_spatial}
 \end{figure}
 
-The left panel of \autoref{fig:true_fitted_var} shows the convergence of the two algorithms.
-Each iteration of the linearized algorithm takes about 0.01 seconds on average while each iteration of the consensus ADMM takes about 20 seconds.
+We estimated the linearized ADMM for all combinations of values of
+$\lambda_t$ and $\lambda_s$ from the sets $\lambda_t \in
+\{0,1,5,10,50,100\}$ and $\lambda_s \in \{0,0.05,0.1,0.2,0.3\}$. For
+each pair, we then compute the mean absolute error (MAE) between the
+estimated variance and the true variance at all locations and all time
+steps. For $\lambda_t=5$ and $\lambda_s=0.1$ MAE was minimized. The
+right panel of \autoref{fig:true_fitted_var} shows the true and the
+estimated standard deviation at location (0,0) using $\lambda_s=0.1$
+and $\lambda_t=5$ (blue) and $\lambda_t=100$ (green). As we can see,
+larger than optimal value of $\lambda_t$ leads to estimated values
+which are ``too smooth".
 
-We estimated the linearized ADMM for all combinations of values of $\lambda_t$ and $\lambda_s$ from the sets $\lambda_t \in \{0,1,5,10,50,100\}$ and $\lambda_s \in \{0,0.05,0.1,0.2,0.3\}$.
-For each pair, we then compute the mean absolute error (MAE) between the estimated variance and the true variance at all locations and all time steps.
-For $\lambda_t=5$ and $\lambda_s=0.1$ MAE was minimized.
-The right panel of \autoref{fig:true_fitted_var} shows the true and the estimated standard deviation at location (0,0) using $\lambda_s=0.1$ and $\lambda_t=5$ (blue) and $\lambda_t=100$ (green).
-As we can see, larger than optimal value of $\lambda_t$ leads to estimated values which are ``too smooth".
-
-\begin{figure}[!h]
+\autoref{fig:true_fitted_var} shows the convergence
+of Algorithms~\ref{alg:ADMM} and~\ref{alg:linADMM}. Each iteration of
+the linearized algorithm 
+takes 0.01 seconds on average while each iteration of the
+consensus ADMM takes about 20 seconds.  \attn{These data are pretty
+  small. Is it possible to do the PDIP?}
+\begin{figure}[tb]
   \centering				
-	\includegraphics[width=.4\linewidth]{Figures/convergence}			
-	\includegraphics[width=.4\linewidth]{Figures/true_fitted_var}		
-	\caption{Left: Convergence speed of linearized and consensus ADMM.
-Right: The true (orange) and estimated standard deviation function at the location (0,0).
-The estimated values are obtained using linearized ADMM with $\lambda_s=0.1$ and two values of $\lambda_t$: $\lambda_t=5$ (blue) and $\lambda_t=100$ (green).} \label{fig:true_fitted_var}
+  \includegraphics[height=.15\textheight]{Figures/convergence}	
+  \caption{Convergence speed of linearized and consensus
+    ADMM.} \label{fig:true_fitted_var}  
 \end{figure}
@@
 \subsection{Data analysis}
-As it was mentioned before, the algorithm proposed in \autoref{sec:consOpt} is appropriate only if we parallelize it over multiple machines and so we do not pursue it further here.
-All the results reported in this section are obtained using the linearize ADMM algorithm of \autoref{sec:linADMM}.
-We applied this algorithm on a subset of the ERA-40 dataset.
-The data is the 2 meter temperature measured daily at 12 p.m from August 31 of 1992 to 2002.
-To reduce the noise we first computed the weekly average of this data.
-To further reduce the size of the data, we will only analyze the data of the locations inside a rectangle extended from (\ang{58},\ang{226}) to (\ang{22},\ang{302}).
-This rectangle covers the united states.
-All the computations were performed on a machine with four 3.20GHz Intel i5-3470 cores.
+
+\autoref{alg:ADMM} is appropriate only if
+we parallelize it over multiple machines, and it is significantly
+slower on our simulated data, so we do not pursue it
+further here. All the results reported in this section are obtained
+using \autoref{alg:linADMM}. We
+applied this algorithm to the northern hemisphere of the ERA-40
+dataset available from the \href{European Center for Medium-Range
+  Weather Forecasts}{https://www.ecmwf.int}. The data are
+the 2 meter temperature measured daily at 12 p.m from August 31 of
+1992 to 2002.
+
+\attn{From here down needs to be cut down and corrected with the right numbers.}
 
 \paragraph{Data Exploration}
-The red curve in the left panel of \autoref{fig:bloom_estimatedSD} shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of the time-series of Bloomington.
-To reduce the number of time-steps in this figure and in the remainder of the paper we work on the weekly averaged of the data.
+The red curve in the left panel of \autoref{fig:bloom_estimatedSD}
+shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
+the time-series of Bloomington. To reduce the number of time-steps in
+this figure and in the remainder of the paper we work on the weekly
+averaged of the data.
 
-The curve of the estimated SD captures the periodic variations in the SD of the signal.
-Just by looking at this curve, it is hard to say if the SD is decreasing or increasing.
-Therefore, we compute the average of the estimated SD for each year.
-The estimated SD together with this annual average is shown in the middle panel of \autoref{fig:bloom_estimatedSD}.
-As it can be seen, the annual trend is not smooth.
-This is because in the optimization problem \eqref{eq:l1tf_var}, the smoothness of the annual trend is not encouraged.
-To remedy this, we add the following long horizon penalty to \eqref{eq:l1tf_var}:
-
+The curve of the estimated SD captures the periodic variations in the
+SD of the signal. Just by looking at this curve, it is hard to say if
+the SD is decreasing or increasing. Therefore, we compute the average
+of the estimated SD for each year. The estimated SD together with this
+annual average is shown in the middle panel of
+\autoref{fig:bloom_estimatedSD}. As it can be seen, the annual trend
+is not smooth. This is because in the optimization problem
+\eqref{eq:l1tf_var}, the smoothness of the annual trend is not
+encouraged. To remedy this, we add the following long horizon penalty
+to \eqref{eq:l1tf_var}: 
 \begin{equation}
@@
 \end{equation}
-
- where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
-Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
-Since we are working on the weekly averaged data, each year corresponds to 52 observations.
-In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
-The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
-The annual average of the estimated SDs shows a linear trend with a positive slope.
+ where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$. Also,
+ $N_{year}$ is the number of years over which we are performing our
+ analysis (here $N_{year}=10$). Since we are working on the weekly
+ averaged data, each year corresponds to 52 observations. In the
+ matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows
+ to the matrix $D$. The estimated SDs using this penalty matrix is
+ shown in the right panel of \autoref{fig:bloom_estimatedSD}. The
+ annual average of the estimated SDs shows a linear trend with a
+ positive slope.
 
 \begin{figure}[tb]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}}
-		\caption{Left: The residuals of the time-series of Bloomington (averaged weekly) and the estimated SD obtained from the method of \autoref{sec:l1tf_var} (red).
-Middle: the estimated SDs (red) and their annual average (black) without imposing the long horizon penalty.
-Right: the same as middle panel but here the long horizon penalty is imposed.
-See the text for more details.}
+  \centering
+  \includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}
+  \caption{Left: The residuals of the time-series of Bloomington
+    (averaged weekly) and the estimated SD obtained from the method of
+    \autoref{sec:l1tf_var} (red). Middle: the estimated SDs (red) and
+    their annual average (black) without imposing the long horizon
+    penalty. Right: the same as middle panel but here the long horizon
+    penalty is imposed. See the text for more details.} 
   \label{fig:bloom_estimatedSD}
-	\end{center}
-	\vskip -0.2in
 \end{figure}
@@
 
-This section is devoted to exploring some of the properties of the ERA-40 surface temperature data set.
-The goal here is to demonstrate some of the difficulties in modeling the trend in the temperature volatility and motivate our methodology for doing so.
+This section is devoted to exploring some of the properties of the
+ERA-40 surface temperature data set. The goal here is to demonstrate
+some of the difficulties in modeling the trend in the temperature
+volatility and motivate our methodology for doing so.
 
@@
 
-The right panel of \autoref{fig:bloom_detrended} shows the time-series of the temperature of Bloomington, after removing the cyclic terms and de-trending using the method explained in the next section.
-The goal is to investigate the trend in the variance of this signal.
-This figure, reveals another issue toward this goal: the variance of this signal, shows cyclic behavior.
-Also, the cycles are not regular and their amplitude and frequency change.
-Even if one can describe the behavior of the variance of all the time-series using a single parametric model (for example a variant of the GARCH models \citep{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework.
-These observations motivate the need to develop a non-parametric framework for the problem at hand.
+The right panel of \autoref{fig:bloom_detrended} shows the time-series
+of the temperature of Bloomington, after removing the cyclic terms and
+de-trending using the method explained in the next section. The goal
+is to investigate the trend in the variance of this signal. This
+figure, reveals another issue toward this goal: the variance of this
+signal, shows cyclic behavior. Also, the cycles are not regular and
+their amplitude and frequency change. Even if one can describe the
+behavior of the variance of all the time-series using a single
+parametric model (for example a variant of the GARCH models
+\citep{bollerslev_generalized_1986}), it is not clear how the trend in
+the variance should be investigated in this framework. These
+observations motivate the need to develop a non-parametric framework
+for the problem at hand.
 
 \begin{figure}[tb]
-	\vskip 0.2in
-	\begin{center}
-		\centerline{\includegraphics[width=\columnwidth]{Figures/bloom_trend_resid}}
-		\caption{Left: Time-series of the temperature of Bloomington (blue) and the estimated trend and cycles obtained from the $\ell_1$-trend filtering (red).
-Right: the same time-series after removing the cyclic terms and de-trending using $\ell_1$-trend filtering.}
+  \centering
+  \includegraphics[width=\columnwidth]{Figures/bloom_trend_resid}
+  \caption{Left: Time-series of the temperature of Bloomington (blue)
+    and the estimated trend and cycles obtained from the
+    $\ell_1$-trend filtering (red). Right: the same time-series after
+    removing the cyclic terms and de-trending using $\ell_1$-trend
+    filtering.} 
   \label{fig:bloom_detrended}
-	\end{center}
-	\vskip -0.2in
 \end{figure}
@@
 
-We used the following rule to determine when to stop the optimization: the optimization was stopped if the value of the loss did not improve by at least 0.1\% in 1000 trials.
-As we can see, the algorithm converged in about 2000 iterations.
-This took about 11 minutes.
-Our experiments showed that the convergence speed depends on the value of $\lambda_t$ and $\lambda_s$.
-Also, if we use the solution obtained for smaller values of these parameters as the initial value for the larger values (\textit{warm start}), the converges speed improves.
+We used the following rule to determine when to stop the optimization:
+the optimization was stopped if the value of the loss did not improve
+by at least 0.1\% in 1000 trials. As we can see, the algorithm
+converged in about 2000 iterations. This took about 11 minutes. Our
+experiments showed that the convergence speed depends on the value of
+$\lambda_t$ and $\lambda_s$. Also, if we use the solution obtained for
+smaller values of these parameters as the initial value for the larger
+values (\textit{warm start}), the converges speed improves.
 
 \paragraph{Model selection}
-One common method for choosing the penalty parameters in the Lasso problems is to find the solution for a range of the values of these parameters and then choose the values which minimize a model selection criterion.
-However, such analyses needs the computation of the degrees of freedom (df).
-Several previous work have investigated the df in generalized lasso problems \citep{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}.
-However, all these studies have considered the linear regression problem and, to the best of our knowledge, the problem of computing the df for generalized lasso with general objective function has not been considered yet.
+One common method for choosing the penalty parameters in the Lasso
+problems is to find the solution for a range of the values of these
+parameters and then choose the values which minimize a model selection
+criterion. However, such analyses needs the computation of the degrees
+of freedom (df). Several previous work have investigated the df in
+generalized lasso problems
+\citep{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}. However,
+all these studies have considered the linear regression problem and,
+to the best of our knowledge, the problem of computing the df for
+generalized lasso with general objective function has not been
+considered yet.
 
-Another approach is to choose the set of values which minimize an estimate of the expected prediction error obtained by k-fold cross-validation \citep{tibshirani_regression_1996} .
-Although this method is applicable for our problem, it needs k times more computation.
+Another approach is to choose the set of values which minimize an
+estimate of the expected prediction error obtained by k-fold
+cross-validation \citep{tibshirani_regression_1996} . Although this
+method is applicable for our problem, it needs k times more
+computation.
 
-In this paper, we use a heuristic method for choosing $\lambda_t$ and $\lambda_s$: we compute the optimal solution for a range of values of these parameters and choose the values which minimize $\mathscr{L}(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D_{total}h \lVert$.
-This objective is a compromise between the negative log likelihood ($-l(y|h)$) and the complexity of the solution ($\sum \lVert D_{total}h \lVert$).
-For smoother solutions the value of $\sum \lVert D_{total}h \lVert$ will be smaller but with the cost of larger $-l(y|h)$.
+In this paper, we use a heuristic method for choosing $\lambda_t$ and
+$\lambda_s$: we compute the optimal solution for a range of values of
+these parameters and choose the values which minimize
+$\mathscr{L}(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D_{total}h
+\lVert$. This objective is a compromise between the negative log
+likelihood ($-l(y|h)$) and the complexity of the solution ($\sum
+\lVert D_{total}h \lVert$). For smoother solutions the value of $\sum
+\lVert D_{total}h \lVert$ will be smaller but with the cost of larger
+$-l(y|h)$.
 
-We computed the optimal solution for all the combinations of the following sets of values: $\lambda_t \in \{1,5,10,20\} \, \, , \lambda_s \in \{0,.1,1,5,10\}$.
-The best combination based on a held out set was $\lambda_t=5$ and $\lambda_s=1$.
-All the analyses in the next section are performed on the solution for these values.
+We computed the optimal solution for all the combinations of the
+following sets of values: $\lambda_t \in \{1,5,10,20\} \, \, ,
+\lambda_s \in \{0,.1,1,5,10\}$. The best combination based on a held
+out set was $\lambda_t=5$ and $\lambda_s=1$. All the analyses in the
+next section are performed on the solution for these values.
 
@@
 
-The top row of \autoref{fig:avg_change_estimatedSD} shows the detrended data, the estimated standard deviation and the yearly average of these estimates for two cities in the US.
-The estimated SD captures the periodic behavior in the variance of the time-series.
-In addition, the number of linear segments changes adaptively in each time window depending on how fast the variance is changing.
-The yearly average of the estimated SD captures the trend in the temperature volatility.
-For example, we can see that in Bloomington, there is a small positive trend.
-To determine how the volatility has changed in each location, we subtract the average of the estimated variance in 1992 from the average in the following years and compute their sum.
-The value of this change in the variance in each location is depicted in the right panel of \autoref{fig:avg_change_estimatedSD}.
-The left panel of this figure, shows the average estimated variance in each location.
+The top row of \autoref{fig:avg_change_estimatedSD} shows the
+detrended data, the estimated standard deviation and the yearly
+average of these estimates for two cities in the US. The estimated SD
+captures the periodic behavior in the variance of the time-series. In
+addition, the number of linear segments changes adaptively in each
+time window depending on how fast the variance is changing.
+
+The yearly average of the estimated SD captures the trend in the
+temperature volatility. For example, we can see that in Bloomington,
+there is a small positive trend. To determine how the volatility has
+changed in each location, we subtract the average of the estimated
+variance in 1992 from the average in the following years and compute
+their sum. The value of this change in the variance in each location
+is depicted in the right panel of
+\autoref{fig:avg_change_estimatedSD}. The left panel of this figure,
+shows the average estimated variance in each location.
 
@@
   \includegraphics[width=.45\linewidth]{Figures/avg_change_estimatedVar}
-	\caption{Top row: Detrended data and the estimated SD for Bloomington (left) and San Diego (right).
-Bottom: the average of the estimated variance over the US (left) and the change in the variance from 1992 to 2002 (right)}
+  \caption{Top row: Detrended data and the estimated SD for
+    Bloomington (left) and San Diego (right). Bottom: the
+    average of the estimated variance over the US (left) and the
+    change in the variance from 1992 to 2002 (right)} 
   \label{fig:avg_change_estimatedSD}
 \end{figure} 
-It is interesting to note that the trend in volatility is almost zero over the oceans.
-The most positive trend can be observed in the south-east and the most negative trend has happened in the north-east.
+It is interesting to note that the trend in volatility is almost zero
+over the oceans. The most positive trend can be observed in the
+south-east and the most negative trend has happened in the north-east.
 
 \section{Discussion}
-In this paper, we proposed a new method for estimating the variance of spatio-temporal data.
-The main idea is to cast this problem as a constrained optimization problem where the constraints enforce smooth changes in the variance for neighboring points in time and space.
-In particular, the solution is piecewise linear in time and piecewise constant in space.
-The resulting optimization is in the form of a generalized LASSO problem with high-dimension, and so applying the PDIP method directly is infeasible.
-We therefore developed two ADMM-based algorithms to solve this problem: the consensus ADMM and linearized ADMM.
+In this paper, we proposed a new method for estimating the variance of
+spatio-temporal data. The main idea is to cast this problem as a
+constrained optimization problem where the constraints enforce smooth
+changes in the variance for neighboring points in time and space. In
+particular, the solution is piecewise linear in time and piecewise
+constant in space. The resulting optimization is in the form of a
+generalized LASSO problem with high-dimension, and so applying the
+PDIP method directly is infeasible. We therefore developed two
+ADMM-based algorithms to solve this problem: the consensus ADMM and
+linearized ADMM.
 
-The consensus ADMM algorithm converges in few hundreds of iterations but each iteration takes much longer than the linearized ADMM algorithm.
-The appealing feature of the consensus ADMM algorithm is that if it is parallelized on enough number of machines the computation time per iteration remains constant as the problem size increases.
-The linearized ADMM algorithm, on the other hand converges in few thousands of iterations but each iteration is performed in split second.
-However, since the algorithm converges in many iterations it is not very appropriate for parallelization.
-The reason is that after each iteration the solution computed in each machine should be broadcast to the master machine and this operation takes some time which depends on the speed of the network connecting the slave machines to the master.
-A direction for future research would be to combine these two algorithms in the following way: the problem should be split into the sub-problems (as in the consensus ADMM) but each sub-problem can be solved using linearized ADMM.
+The consensus ADMM algorithm converges in few hundreds of iterations
+but each iteration takes much longer than the linearized ADMM
+algorithm. The appealing feature of the consensus ADMM algorithm is
+that if it is parallelized on enough number of machines the
+computation time per iteration remains constant as the problem size
+increases. The linearized ADMM algorithm, on the other hand converges
+in few thousands of iterations but each iteration is performed in
+split second. However, since the algorithm converges in many
+iterations it is not very appropriate for parallelization. The reason
+is that after each iteration the solution computed in each machine
+should be broadcast to the master machine and this operation takes
+some time which depends on the speed of the network connecting the
+slave machines to the master. A direction for future research would be
+to combine these two algorithms in the following way: the problem
+should be split into the sub-problems (as in the consensus ADMM) but
+each sub-problem can be solved using linearized ADMM.
 
-We applied the linearized ADMM algorithm to the surface temperature data on a grid over the united states, for years 1992-2002.
-The results showed that in many locations the variance of the temperature has increased about 1 unit in 10 years.
+We applied the linearized ADMM algorithm to the surface temperature
+data on a grid over the united states, for years 1992-2002. The
+results showed that in many locations the variance of the temperature
+has increased about 1 unit in 10 years.
 
-The goal of this paper, however, is not to make any conclusions about the trend in the variance because we solved the problem only for a grid over the united states and for 10 years of the data.
-A thorough analysis, needs the full solution over the globe and for a longer time period.
-The goal of the paper, was to propose the idea of estimating the trend in variance of spatio-temporal signals using generalized lasso and to investigate the algorithms for solving the resulting optimization problem.
+The goal of this paper, however, is not to make any conclusions about
+the trend in the variance because we solved the problem only for a
+grid over the united states and for 10 years of the data. A thorough
+analysis, needs the full solution over the globe and for a longer time
+period. The goal of the paper, was to propose the idea of estimating
+the trend in variance of spatio-temporal signals using generalized
+lasso and to investigate the algorithms for solving the resulting
+optimization problem.
 

### 2018-05-15 code/ModelSelection.py
added model selection

@@ -1,0 +1,44 @@
+from volatilitytrend.algorithms.base import LinearizedADMM
+from os.path import join
+import matplotlib.pyplot as plt
+import numpy as np
+
+lam_t_vec=[0,2,4,8,10,15,200]
+lam_s_vec=[0,.1,.5,2,5,10]
+mu_vec=[.01]*6+[.001]*3
+dataset='north_hemisphere'
+#dataset='us'
+
+
+if dataset=='us':
+    data_fn='1992-08-31_to_2002-08-31_data_avg_us_detrended'
+    metadata_fn='1992-08-31_to_2002-08-31_metadata_avg_us_detrended'
+    fittedModelsDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/'+\
+    '19920831to20020831/Fitted_LinADMM_test'
+    
+    
+elif dataset=='north_hemisphere':
+    data_fn='1960-01-01_to_2010-12-31_data_avg_sub_north_detrended'
+    metadata_fn='1960-01-01_to_2010-12-31_metadata_avg_sub_north_detrended'
+    fittedModelsDir='/home/arash/datasets/ecmwf/Fitted_linADMM'
+    saveFigDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/globe/Figures'
+    
+
+#===load data===
+print('\014')
+la = LinearizedADMM()#construct linearizedADMM object
+la.loadData(data_fn,metadata_fn)#load data
+#===load data===
+
+la.modelSelection(fittedModelsDir,lam_t_vec,lam_s_vec,mu_vec,
+                  lh_trend=True,wrapAround=True)
+
+print('optimal penalty parameters are: lam_t={:.2f} , lam_s={:.2f}'.\
+      format(la.opt_lamt,la.opt_lams))
+  
+
+
+
+
+
+ 

### 2018-05-15 code/visualization.py

@@ -2,7 +2,7 @@ from volatilitytrend.algorithms.base import LinearizedADMM
 from os.path import join
 import matplotlib.pyplot as plt
 
-lam_t=10;lam_s=10;mu=.01
+lam_t=200;lam_s=5;mu=.001
 dataset='north_hemisphere'
 #dataset='us'
 
@@ -10,14 +10,14 @@ dataset='north_hemisphere'
 if dataset=='us':
     data_fn='1992-08-31_to_2002-08-31_data_avg_us_detrended'
     metadata_fn='1992-08-31_to_2002-08-31_metadata_avg_us_detrended'
-    destDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/'+\
+    fittedModelsDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/'+\
     '19920831to20020831/Fitted_LinADMM_test'
+    
     
 elif dataset=='north_hemisphere':
     data_fn='1960-01-01_to_2010-12-31_data_avg_sub_north_detrended'
     metadata_fn='1960-01-01_to_2010-12-31_metadata_avg_sub_north_detrended'
-    destDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/'+\
-            'globe/Fitted_linADMM'
+    fittedModelsDir='/home/arash/datasets/ecmwf/Fitted_linADMM'
     saveFigDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/globe/Figures'
 
 #===load data===
@@ -29,7 +29,7 @@ la.loadData(data_fn,metadata_fn)#load data
 #===load data===
 
 #===analyse fitted value and compute var and ave. var===
-la.analyseFittedValues(join(destDir,'X_'+fn))
+la.analyseFittedValues(join(fittedModelsDir,'X_'+fn))
 #===analyse fitted value and compute var and ave. var===
 
 #===plot time series of temerature at different locations===

### 2018-05-15 code/volatilitytrend/algorithms/base.py

@@ -5,11 +5,12 @@ import gc,datetime
 import matplotlib.pyplot as plt
 from mpl_toolkits.basemap import Basemap
 import pandas as pd
+from time import ctime 
 
 from .. import config
 from .linearized_admm import linearizedADMM_fit
 from .consensus_admm import consensusADMM_fit
-from ..utils import latlon_to_rowcol
+from ..utils import latlon_to_rowcol,compute_D_and_lam,computeLoss
 
 __metaclass__=type
 class BaseAlgorithmClass():
@@ -26,6 +27,7 @@ class BaseAlgorithmClass():
         self.metadata={'n_rows':[],'n_cols':[],'T':[],'dates':[],
                        'lats':[],'lons':[]}
         self.dataMat=[]
+        self.fittedVar=[]
         
     def loadData(self,data_fn,metadata_fn):
         
@@ -202,7 +204,7 @@ class BaseAlgorithmClass():
         plt.colorbar(fraction=.013)
         fn=join(saveDir,'avg_logVar_{}.pdf'.format(suffix))
         fig.savefig(fn,dpi=300,format='pdf')
-        #===plot average variance===
+        #===plot average variance===        
         
 class LinearizedADMM(BaseAlgorithmClass):
 
@@ -222,8 +224,66 @@ class LinearizedADMM(BaseAlgorithmClass):
                            patience=patience,tol=tol,
                            ifAdaptMu=ifAdaptMu,mu_adapt_rate=mu_adapt_rate,
                            mu_adapt_freq=mu_adapt_freq)
+      
+    def modelSelection(self,fittedModelsDir,lam_t_vec,lam_s_vec,mu_vec,
+                       lh_trend=True,wrapAround=True):
+        '''
+        This function computes the value of the loss function 
+        for all combinations of given sets of values of lam_t and 
+        lam_s.
+        '''  
         
+        #===check if the data is loaded===
+        if len(self.dataMat)==0:
+            msg='''Data has not been loaded.
+            Use self.loadData() to load the data first'''
+            raise ValueError(msg)
+        #===check if the data is loaded===
     
+        #===metadata===
+        n_rows,n_cols,T=(self.metadata['n_rows'],self.metadata['n_cols'],
+                         self.metadata['T'])    
+        #===metadata===        
+        
+        #===make sure the parameters are float===
[…]

### 2018-05-15 code/volatilitytrend/algorithms/consensus_admm.py

@@ -400,7 +400,7 @@ def consensusADMM_fit(dataMat,destDir,metadata,
     #===check the compatability of grid size with sub-blocks size===
     if (((n_rows-1)%(n_r_b-1)!=0) or ((n_cols-1)%(n_c_b-1)!=0)):
         msg='In current implementation, n_rows-1 and n_cols-1 should be'+\
-        'divisive to (n_r_b-1) and (n_c_b)-1, respectively.'
+        'divisable by (n_r_b-1) and (n_c_b)-1, respectively.'
         raise ValueError(msg)
     #===check the compatability of grid size with sub-blocks size===
     

### 2018-05-15 nips-2018/paper.tex
updates on paper

@@
 \usepackage{siunitx}
-
+\usepackage{comment}
 \usepackage{mathtools}
@@
 \section{Introduction}
+
+\attn{Some equations do not have numbering and some have.}
 
@@
 
+\attn{Arash: should't the title of this section be "...estimating variance of spatio-temporal data" or something similar?}
+
 $\ell_1$-trend filtering was proposed by \citep{kim_$ell_1$_2009} as a method for estimating a smooth, time-varying trend. It is formulated as the optimization problem
@@
 $\exp(h_t)$. The negative log-likelihood of the observed data in this
-model is $l(y\given h) \propto -\sum_{t=1}^Th_t - y_t^2e^{-h_t}$. Crucially,
+model is $l(y\given h) \propto -\sum_{t=1}^T h_t - y_t^2e^{-h_t}$. Crucially,
 we assume that the hidden variables $h_t$ vary smoothly. To impose
@@
 the objective $f(x)+\lambda \norm{ Dx }_1$ is:
+
 \begin{align}
@@
 \end{equation}
+
 for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the
@@
 \eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times
-S$ vector and $D$ is replaced by $D_{TS}$. Then, as we
+S$ vector and $D$ is replaced by $D_{TS} \in \mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and spatial constraints, respectively\footnote{$N_s$ is obtained by counting the number of unique constraints at each location and at all times.}.
+Then, as we
 have two different tuning parameters for the temporal and spatial
-components, we write $\lambda =\left[\lambda_1\one_T^\top,\;
-  \lambda_2\one_{S}^\top\right]^\top$ and $\Lambda =\texttt{diag}(\lambda)$ leading to 
+components, we write $\Lambda =\left[\lambda_1\one_{N_t}^\top,\;
+  \lambda_2\one_{N_s}^\top\right]^\top$ leading to \footnote{Throughout the paper, we use $|x|$ for both scalars and vectors.
+For vectors we use this to denote a vector obtained by taking the absolute value of each entry of $x$.}:
 % let $h$ be a
@@
 \begin{equation}
-\min_h -l(y\given h)+ \norm{\Lambda D_{TS}h }_1.
+\min_h -y\given h)+ \Lambda^\top | D_{TS}h |
 \label{eq:l1tf_var_st_mat}
@@
 % \attn{Detailed discussion of the PDIP to here}
-The dual of this problem is 
-$\min_\nu f^*(-D_{ST}^\top\nu)$ subject to 
-$| \nu_k | \le \Lambda_{kk},\; k=1,\ldots,ST$,
- where $f^*$ is given in \eqref{eq:conj}.
 
@@
 For a spatial grid of size $S$ and $T$ time steps, $D_{ST}$ will have
-$3n_rn_c+Tn_c+2Tn_r$ rows and $ST$ columns. For a $1^\circ\times
+$3Tn_rn_c-2n_rn_c-Tn_r$ rows and $ST$ columns. For a $1^\circ\times
 1^\circ$ grid over the
@@
 
+\begin{comment}
 General (unconstrained) optimization problems $\min_z
@@
      (\rho/2) \norm{ x_i-\tilde{z}_i }_2^2\\
-\label{eq:ADMM_steps}
+%\label{eq:ADMM_steps}
 z_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j\\
@@
 \end{align}
+\end{comment}
+
+Consider an optimization problem of the form $\min_h f(h)$, where $h\in\mathbb{R}^n$ is the \textit{global variable} and $f(h):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is convex.
+The goal in consensus optimization is to break this problem into several smaller sub-problems which can be solved independently in each iteration of optimization.
+
+Assume that it is possible to define a set of \textit{local variables} $x_i \in \mathbb{R}^{n_i}$ such that $f(h)=\sum_i f_i(x_i)$, where each $x_i$ is a subset of the global variable $h$.
+More specifically, each entry of the local variables corresponds to an entry of the global variable.
+Therefore we can define a mapping $\mathscr{G}(i,j)$ from the local variables indices into the global variable indices: $k=\mathscr{G}(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is $h_k$ (or $(x_i)_j=h_k$).
+For ease of notation, define $\tilde{h}_i \in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{\mathscr{G}(i,j)}$.
+Then, the original optimization problem is equivalent to the following problem:
+
+\begin{equation}
+\begin{aligned}
+\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)\\
+ s.t. \,\,\, \tilde{h}_i=x_i \,\,\,\,\,\,
+\end{aligned}
+\label{eq:consADMM}
+\end{equation}
+
+It is important to note that each entry of the global variable may correspond to several entries of the local variables and so the constraints $\tilde{h}_i=x_i$ enforce the consensus between the local variables corresponding to the same global variable.
+
+The \textit{augmented Lagrangian} corresponding to the problem \ref{eq:consADMM} is $L_\rho(x,h,y)=\sum_i \big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert x_i-\tilde{h}_i \lVert_2^2 \big)$.
+Now, we can apply ADMM to $L_\rho$ which results in the following ADMM updating steps at each iteration $m$:
+
+\begin{equation}
+\begin{aligned}
+& x_i^{m+1}:=\argmin_{x_i} \bigg( f_i(x_i) + (u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{h}_i^m \lVert_2^2\bigg)\\
+& h_k^{m+1}:=(1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i^{m+1})_j\\
+& u_i^{m+1}:=u_i^m + \rho (x_i^{m+1}-\tilde{h}_i^{m+1})
+\end{aligned}
+\label{eq:ADMM_steps}
+\end{equation}
+
 Here, $S_k$ is the number of local variable entries that correspond to
-$z_k$, and $u_i$ are the Lagrange multipliers.  The $x$-update is
+$h_k$, and $u_i$ are the Lagrange multipliers.
+
+\begin{comment}
+The $x$-update is
 often a simple modification as the original optimization, though over only a small
-partition, the $z$-update simply averages over collections of
+partition, the $h$-update simply averages over collections of
 analogous local variables while the $u$ update enforces sharing. Both
@@
 variables.
+\end{comment}
+
 \attn{I don't much like this explanation or notation. Seems unclear}
 
+\attn{Arash: I re-wrote this section.}
+
 To solve the optimization problem \eqref{eq:l1tf_var_st_mat} or
-\eqref{eq:l1tf_var_st} using this method, we need to two address two
+\eqref{eq:l1tf_var_st} using this method, we need to address two
 questions: first, how to choose the 
@@
 problem for updating these variables (the first line of
-\eqref{eq:ADMM_steps}).
-
-In \autoref{fig:data_cube}, the global variable $z$ is represented
-as a cube (using the subset of the US as an example). We decompose $z$
+\eqref{eq:ADMM_steps} which we will refer to it as \textit{x-update step}).
+
+In \autoref{fig:data_cube}, the global variable $h$ is represented
+as a cube (using the subset of the US as an example). We decompose $h$
 into sub-cubes as shown in the figure by white lines. It is easy to see
-that with this definition of $x_i$, the objective
+that by this definition of $x_i$, the objective
 \eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where
-$f_i(x_i)=-l(y_i\given x_i)+\norm{\Lambda_{(i)} D_{(i)}x_i}$, and
+$f_i(x_i)=-l(y_i\given x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|$, and
 $\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
-penalties corresponding to $x_i$ only.
-\attn{I would like to be able to write all this in terms of $h$
-  instead of $x$}
+penalties corresponding to $x_i$ only. The x-update step in \autoref{eq:ADMM_steps} is the following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
+(u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
+
+\attn{I would like to be able to write all this in terms of $h$ instead of $x$}
+
+\attn{Arash: now I'm using $h$ as the global variables. We cannot write the local variables in terms of $h$ since they are different from $h$.}
 
@@
   \includegraphics[height=.2\textheight]{Figures/data_cube}
-    \caption{The cube represents the global variable $z$ in space and
+    \caption{The cube represents the global variable $h$ in space and
       time. The sub-cubes specified by the white lines are
@@
 following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
-(u_i^m)^t x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
-
-We solve this using the PDIP method. To this end we first need to
-compute the dual of the problem. It can be shown that the dual of this
-problem is: 
-%TODO: write the appendix
-\begin{align}
-\min_\nu &f_i^*(-D_i^t\nu_i) & s.t. &| (\nu_i)_j | \le (\Lambda_i)_j
-\label{eq:dual_x_update}
-\end{align}
- where:
-\begin{align}
-f^*_i(q)&=\sum_j (q_i)_j (w_i)_j -(w_i)_j-(y_i)_j^2 e^{-(w_i)_j} - 
-(\rho/2)((w_i)_j-(\alpha_i^m)_j)^2\\
-\alpha_i^m &=\hat{z}_i^m-u_i^m \label{eq:conj_x_update}\\
-(w_i)_j&=\mathscr{W}\bigg(\frac{(y_i)_j^2}{\rho}\exp
-         \bigg[\frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}\bigg]\bigg)- 
-\frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}
-\end{align}
-\attn{I'm not quite sure what's going on here. These need to match the
-  generic discussion in \eqref{eq:ADMM_steps}. Or are these just
-  defining some other variables for convenience? This needs to be
-  clarified and incorporated into \autoref{alg:ADMM}.}
-
-In these equations, $\mathscr{W}$ is the \textit{Lambert function}
-\citep{corless_lambertw_1996}. To use the PDIP method, we also need to
-compute the gradient and Hessian of $f_i^*(-D_i^t\nu_i)$. This
-involves computing the derivatives of the Lambert functions. The
-primal solution $x_i$ can be obtained from the dual solution $\nu_i$
-in \eqref{eq:dual_x_update} by setting $x_i=w_i$ where $w_i$ is
-defined in the last equation of \eqref{eq:conj_x_update}.
+(u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
+We solve this using the PDIP method.
+As it was explained in \autoref{sec:l1tf_var}, to use PDIP we need to compute the dual problem, which in turn needs the computation of the Fenchel conjugate of the loss function.
+In addition, referring to \autoref{alg:pdip}, we need to compute the gradient and Jacobian of the conjugate functions.
+It is important to note that compare to the optimization problem \autoref{eq:l1tf_var}, the loss function in the x-update step in \autoref{eq:ADMM_steps} includes the quadratic term $(\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2$.
+This makes the computation more involved than \autoref{sec:l1tf_var}.
+The details of the computations are explained in Appendix 1.
+
+\attn{I'm not sure if we want to devote some space to write this as a separate algorithm.
+The steps are those mentioned in \ref{eq:ADMM_steps} and the x-update is now explained in the appendix.
+So we may want to remove algorithm \ref{alg:ADMM}.}
 
@@
 \begin{align}
-x & \leftarrow \prox_{\mu f} \left(x - (\mu/\rho)D^T (D x - z + u )\right)\\
+x & \leftarrow \prox_{\mu f} \left(x - (\mu/\rho)D^\top (D x - z + u )\right)\\
 z & \leftarrow \prox_{\rho g} \left(z + u\right) \label{eq:linADMM_steps}\\
@@
 % $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector
-% $\Lambda^t=(\lambda_t\textbf{e}_{n_t}^t|\lambda_s\textbf{e}_{n_s}^t)$
+% $\Lambda^\top=(\lambda_t\textbf{e}_{n_t}^\top|\lambda_s\textbf{e}_{n_s}^\top)$
 % (see \autoref{sec:exten}). 
@@
 to \eqref{eq:l1tf_var}:
+
 \begin{equation}
@@
 \end{equation}
- where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$. Also,
- $N_{year}$ is the number of years over which we are performing our
- analysis (here $N_{year}=10$). Since we are working on the weekly
- averaged data, each year corresponds to 52 observations. In the
- matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows
- to the matrix $D$. The estimated SDs using this penalty matrix is
- shown in the right panel of \autoref{fig:bloom_estimatedSD}. The
- annual average of the estimated SDs shows a linear trend with a
- positive slope.
+
+where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
+Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
+Since we are working on the weekly averaged data, each year corresponds to 52 observations.
+In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
+The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
+The annual average of the estimated SDs shows a linear trend with a positive slope.
 
@@
 
+\section{Appendix}
+
+\attn{put this in nips appendix format}
+
+In this Appendix we give more details on performing the x-update step in \autoref{eq:ADMM_steps}. We need to solve the following optimization problem:
+
+\begin{equation}
+\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
+\label{eq:x_update_opt}
+\end{equation}
+
+\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$.
+Let $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$.
+As it was explained in \autoref{sec:l1tf_var}, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$.
+So to use PDIP we first need to compute the conjugate function $f^*(\cdot)$.
+We have:
+
+\begin{align}
+f^*(\xi) & = \max_x \quad \xi^\top x - f(x)\\
+& =  \max_x \quad \sum_{j=1}^{n_b} (\xi_jx_j - x_j - y_j^2e^{-x_j} - (\rho/2)(x_j-\tilde{z}_j+u_j))
+\label{eq:conjugate}
+\end{align}
+
+Setting the derivative of the terms inside the summation to 0, we obtain:
+
+\begin{equation}
+\xi_j-y_j^2e^{-x_j^*}-\rho x_j^* + \rho (\tilde{z}_j-u_j)=0
+\label{eq:x_start}
+\end{equation}
+
+\noindent where $x^*$ is the maximizer in \ref{eq:conjugate}. Then, it can be shown that $x_j^*$ which satisfies \eqref{eq:x_start} can be obtained as follows:
+
+\begin{align}
+x^*_j & = \mathscr{W}\bigg(\frac{y_j^2}{\rho} e^{\phi_j} \bigg) - \phi_j \\
+\phi_j & =\frac{1-\xi_j-\rho(\tilde{z}_j-u_j)}{\rho}
+\end{align}
+
+In this equation, $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996}.
+Finally, the conjugate function is: $f^*(\xi) = \sum_{j=1}^{n_b} (\xi_jx^*_j - x^*_j - y_j^2e^{-x^*_j} - (\rho/2)(x^*_j-\tilde{z}_j+u_j))$.
+
+To use PDIP, we also need to evaluate $\nabla f^*$ and $\nabla^2 f^*$.
+First note that $\frac{\partial \mathscr{W}(q)}{\partial q} = \frac{\mathscr{W}(q)}{q(1+\mathscr{W}(q))}$ and $\frac{\partial^2 \mathscr{W}(q)}{\partial q^2} = - \frac{\mathscr{W}^2(q)(\mathscr{W}(q)+q)}{q^2(1+\mathscr{W}(q))^3}$.
+Using the chain rule we get:
+
+%\x^*_j + \xi_j + \frac{\partial \x^*_j}{\partial \xi_j}
+
+\begin{equation}
+\frac{\partial f^*(\xi)}{\partial \xi_j}  =  x^*_j  + \frac{\partial x^*_j}{\partial \xi_j} \bigg[ \xi_j -1 + y_j^2 e^{-x_j^*} + \rho (\tilde{z}_j - u_j - x_j^*) \bigg]
+\label{eq:d_f*_start}
+\end{equation}
+
+\noindent where we have:
+
+\begin{equation}
+\frac{\partial x_j^*}{\partial \xi_j} & = \frac{1}{\rho(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))}
+\label{eq:d_x*_start}
+\end{equation}
+
+By some tedious but straightforward computation we can obtain the second derivatives:
+
+\begin{align}
+\frac{\partial^2 f^*(\xi)}{\partial \xi_j^2} & =  \frac{\partial x_j^*}{\partial \xi_j} - \rho \frac{\partial^2 x_j^*}{\partial \xi_j^2} \bigg[ \phi_j +x_j^* - \tilde{z}_j + u_j \bigg]\\
+& \quad + \frac{\partial x_j^*}{\partial \xi_j} \bigg[ 1-y_j^2 \frac{\partial x_j^*}{\partial \xi_j} e^{-x_j^*} -\rho \frac{\partial x_j^*}{\partial \xi_j} \bigg]
+\label{eq:d2_f*_start}
+\end{align}
+
+\begin{equation}
+\frac{\partial^2 x_j^*}{\partial \xi_j^2}  = \frac{\mathscr{W}((y_j^2/\rho) e^{-\phi_j} )}{\rho^2(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))^3}
+\label{eq:d2_x*_start}
+\end{equation}
+
+Having computed the conjugate function and its gradient and Jacobian, now we can use a number of convex optimization software packages which have an implementation of PDIP to perform the x-update step inside the ADMM loop.
+We chose the python API of the \texttt{cvxopt}~\citep{andersen_cvxopt:_2013} package.
+
 \bibliographystyle{abbrvnat}

### 2018-05-15 working-paper/paper.tex
updates on paper

@@
 
-A general form consensus optimization problem is in the form $\min_z \sum_i f_i(z)$, $z\in\mathbb{R}^n$.
-We can define a set of \textit{local variables} $x_i\in \mathbb{R}^{n_i}$ such that $\sum_i f_i(z)=\sum_i f_i(x_i)$.
-We follow the notation of \cite{boyd_distributed_2011} closely.
-Let $k=\mathscr{G}(i,j)$ which means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or $(x_i)_j=z_k$) and define $\tilde{z}_i \in \mathbb{R}^{n_i}$ by $(\tilde{z}_i)_j=(x_i)_j$.
-Then the original unconstrained optimization problem is equivalent to the following constrained optimization problem:
+Consider an optimization problem of the form $\min_z f(z)$, where $z\in\mathbb{R}^n$ is the \textit{global variable} and $f(z):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is convex.
+The goal in consensus optimization is to break this problem into several smaller sub-problems which can be solved independently in each iteration of optimization.
+
+Assume that it is possible to define a set of \textit{local variables} $x_i \in \mathbb{R}^{n_i}$ such that $f(z)=\sum_i f_i(x_i)$, where each $x_i$ is a subset of the global variable $z$.
+More specifically, each entry of the local variables corresponds to an entry of the global variable.
+Therefore we can define a mapping $\mathscr{G}(i,j)$ from the local variables indices into the global variable indices: $k=\mathscr{G}(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or $(x_i)_j=z_k$).
+For ease of notation, define $\tilde{z}_i \in \mathbb{R}^{n_i}$ as $(\tilde{z}_i)_j=z_{\mathscr{G}(i,j)}$.
+Then, the original optimization problem is equivalent to the following problem:
 
@@
 \end{aligned}
+\label{eq:consADMM}
 \end{equation}
 
-Now, we can apply ADMM to the \textit{augmented Lagrangian} of this problem. This results in the following ADMM updating steps at each iteration $m$:
+It is important to note that each entry of the global variable may correspond to several entries of the local variables and so the constraints $\tilde{z}_i=x_i$ enforce the consensus between the local variables corresponding to the same global variable.
+
+The \textit{augmented Lagrangian} corresponding to the problem \ref{eq:consADMM} is $L_\rho(x,z,y)=\sum_i \big(f_i(x_i)+y_i^T(x_i-\tilde{z}_i) + (\rho/2) \lVert x_i-\tilde{z}_i \lVert_2^2 \big)$.
+Now, we can apply ADMM to $L_\rho$ which results in the following ADMM updating steps at each iteration $m$:
 

### 2018-05-16 nips-2018/paper.tex
updates on paper

@@
 has the form
+
 \begin{align}
@@
 \end{align}
- where $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
+
+\noindent where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu < \rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
+
 \begin{equation}
@@
 \end{equation}
-Clearly,
-\eqref{eq:l1tf_var_st} has this form.\attn{What are $\mu$ and $\rho$?}
+
+\attn{What are $\mu$ and $\rho$?}
+
+\attn{Arash: I explained it above. They are algorithm parameters.}
+
+Clearly, \eqref{eq:l1tf_var_st} has this form necessary for using this algorithm.
 % The optimization problem 

### 2018-05-16 code/AnalyseSyntheticData_modelComp.py
model comparison

@@ -1,0 +1,135 @@
+'''
+In this script we compare the performance of four models for
+in estimating the variance. The models are:
+    1. linearizedADMM with optimal penalty parameters
+    2. linearizedADMM with lam_t=0
+    3. linearizedADMM with lam_s=0
+    4. GARCH(1,1)
+
+'''
+
+import numpy as np
+from volatilitytrend.data_utils.simulate_data import simulateSpatioTemporalData
+from os.path import join
+from volatilitytrend.algorithms.base import LinearizedADMM
+from arch import arch_model
+import pandas as pd
+from time import ctime
+
+n_sim=50;#number of simulations
+
+#===models parameters===
+#---linearizedADMM optimal---
+lam_t_opt=5.;lam_s_opt=0.1;mu=.01
+
+#---linearizedADMM temporal---
+lam_t_temp=5.;lam_s_temp=0.;mu=.01
+
+#---linearizedADMM spatial---
+lam_t_sp=0.;lam_s_sp=.1;mu=.01
+#===models parameters===
+
+#===Specify parameters of data simulator===
+n_rows=5;n_cols=7;
+n_yeras=15
+T=n_yeras*52#time-series length
+gridsize=n_rows*n_cols
+
+#---specify sources properties---
+var_sources_centers=np.array([[0,0],[0,5],[3,0],[3,5]])
+var_sources_sigma=[5.]*4
+#---specify sources properties---
+
+#---compute the weight of each source at wach time---
+times=np.arange(0,T)
+trend1=.5*times/float(T)
+trend2=.1*times/float(T)
+trend3=-.5*times/float(T)
+trend4=-.1*times/float(T)
+
+var_sources_weight1=np.exp(np.sin(n_yeras*2*np.pi*times/float(T)))+trend1
+var_sources_weight2=np.exp(np.sin(n_yeras*2*np.pi*times/float(T)))+trend2
+var_sources_weight3=np.exp(np.cos(n_yeras*2*np.pi*times/float(T))+trend3)
+var_sources_weight4=np.exp(np.cos(n_yeras*2*np.pi*times/float(T))+trend4)
+var_sources_weight_mat=np.concatenate((var_sources_weight1,
+                                       var_sources_weight2,
+                                       var_sources_weight3,
+                                       var_sources_weight4)).\
+                                       reshape((4,-1))
[…]

### 2018-05-16 nips-2018/paper.tex
update paper

@@
 $\ell_\infty$ constraint, as $w\rightarrow\infty$,
-the solution to this nonlinear system reduces to the KKT
+the solution to this nonlinear system reduces to the Karush–Kuhn–Tucker (KKT)
 conditions. Therefore, the PDIP method takes Newton steps to solve the
@@
 provides the details.
+
 \begin{algorithm}[tb]
@@
 This makes the computation more involved than \autoref{sec:l1tf_var}.
-The details of the computations are explained in Appendix 1.
+The details of the computations are explained in \autoref{sec:app_consADMM}.\attn{change this to appendix}
 
@@
   \end{align}
-where $[S_{\alpha}(u)]_k = \sign(u_k)(|u_k| -\alpha_k)_+$ and
+where $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996},  $[S_{\alpha}(u)]_k = \sign(u_k)(|u_k| -\alpha_k)_+$ and
 $(v)_+=v\vee 0$.
@@
 
-\section{Appendix}
-
+\section{Appendix A}
+\label{sec:app_l1tf_var}
+
+In this appendix we provide more details on how to solve the optimization problem \ref{eq:l1tf_var}
+
+\section{Appendix B}
+\label{sec:app_consADMM}
 \attn{put this in nips appendix format}

### 2018-05-16 code/AnalyseSyntheticData_modelComp.py

@@ -15,8 +15,9 @@ from volatilitytrend.algorithms.base import LinearizedADMM
 from arch import arch_model
 import pandas as pd
 from time import ctime
+import matplotlib.pyplot as plt
 
-n_sim=50;#number of simulations
+n_sim=20;#number of simulations
 
 #===models parameters===
 #---linearizedADMM optimal---
@@ -37,7 +38,7 @@ gridsize=n_rows*n_cols
 
 #---specify sources properties---
 var_sources_centers=np.array([[0,0],[0,5],[3,0],[3,5]])
-var_sources_sigma=[5.]*4
+var_sources_sigma=[7.]*4
 #---specify sources properties---
 
 #---compute the weight of each source at wach time---
@@ -82,6 +83,7 @@ def computeMAE_ADMM(data_fn,metadata_fn,savedResultsDir,lam_t,lam_s,covMat):
     filepath=join(dstDir,'X_'+fn)
     X=np.fromfile(filepath).reshape((gridsize,T))
     return np.mean(np.abs(np.exp(X/2)-covMat))
+#    return np.mean(np.abs((X/2.)-np.log(covMat)))
 
 def computeMAE_GARCH(data_fn,metadata_fn,covMat):
     la = LinearizedADMM()#construct linearizedADMM object    
@@ -95,6 +97,8 @@ def computeMAE_GARCH(data_fn,metadata_fn,covMat):
         am=arch_model(y)
         res=am.fit()
         err.append(np.mean(np.abs(res.conditional_volatility-covMat[r,:])))
+#        err.append(np.mean(np.abs(np.log(res.conditional_volatility)-\
+#                                  np.log(covMat[r,:]))))
     return np.mean(err)
     
 
@@ -104,6 +108,7 @@ for i in range(n_sim):
     print(ctime()+'...simulation number {}'.format(i))
     
     #simulate data
+    var_sources_sigma=[4+3*np.random.rand()]*4
     simulateSpatioTemporalData(dstDir,n_rows,n_cols,T,var_sources_centers,
                                var_sources_weight_mat,var_sources_sigma)
     covMat = np.fromfile(join(savedResultsDir,'covMat')).\
@@ -127,7 +132,12 @@ for i in range(n_sim):
 
 errors=pd.DataFrame({'admm_opt':MAE_opt,'admm_temp':MAE_temp,'admm_sp':MAE_sp,
            'garch':MAE_ga})
+    
[…]

### 2018-05-16 nips-2018/paper.tex
update paper

@@
 
-\attn{Some equations do not have numbering and some have.}
+\attn{Arash: Some equations do not have numbering and some have.}
+
+\attn{Arash: fix appendicies and the references to them.}
 
@@
 PDIP algorithm (as in, e.g.,
-\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}). First, we note that
-this is a generalized LASSO problem
-\citep{tibshirani_solution_2011}. The dual of a generalized LASSO with
-the objective $f(x)+\lambda \norm{ Dx }_1$ is:
-
-\begin{align}
-\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
-\end{align}
- where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x
- u^\top x-f(x)$. It is simple to show that 
-\begin{equation}
-f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
-\label{eq:conj}
-\end{equation}
-Writing
-\begin{equation}
-  r_w(v,\mu_1,\mu_2):=
-  \begin{bmatrix}
-    \nabla f^*(-D^\top v) + D(v-\lambda \one)^\top \mu_1 -
-    D(v+\lambda \one)^\top \mu_2\\
-    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
-  \end{bmatrix}
-  =
-  \begin{bmatrix}
-  0\\0\end{bmatrix}
-\label{eq:resid}
-\end{equation}
-
-for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the
-$\ell_\infty$ constraint, as $w\rightarrow\infty$,
-the solution to this nonlinear system reduces to the Karush–Kuhn–Tucker (KKT)
-conditions. Therefore, the PDIP method takes Newton steps to solve the
-system for a series of increasing values of $w$. \autoref{alg:pdip}
-provides the details.
-
+\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}).
+In each iteration of PDIP we need to compute a \textit{search direction} by taking a Newton step on a system of nonlinear equations.
+Appendix B provides details on how this computation can be performed for the objective function \ref{eq:l1tf_var}.
+
+\attn{Arash: I think we should remove Algorithm 1.
+There are other steps in PDIP which involve adapting the algorithm parameter $w$ and doing Backtracking line search.
+Also there is only one loop in PDIP.}
+
+\begin{comment}
 \begin{algorithm}[tb]
@@
     \STATE Solve $A [\Delta\nu\; \Delta\mu_1\; \Delta\mu_2]^\top=r_{w_k}$ to find the search direction.
-    \STATE $\quad\quad A$ is the Jacobian of $r_w$ in \eqref{eq:resid}.
+    \STATE $\quad\quad A$ is given in \ref{eq:delta_r}
     \STATE Update $[\nu^{j+1}\; \mu^{j+1}_1\; \mu^{j+1}_2] \leftarrow
@@
    \ENDFOR
-   \RETURN $h=\log\frac{y^2}{1+D^\top\nu}$ \attn{Is this right?
-     Explicit form of $A$}
+   \RETURN $h=\log\frac{y^2}{1+D^\top\nu}$ 
+   \\
+   \attn{Is this right? Explicit form of $A$}\\
+   \attn{Arash: see appendix A}
   \end{algorithmic}
 \end{algorithm}
+\end{comment}
 
@@
 \begin{equation}
-\min_h -y\given h)+ \Lambda^\top | D_{TS}h |
+\min_h -l(y\given h)+ \Lambda^\top | D_{TS}h |
 \label{eq:l1tf_var_st_mat}
@@
 
-For a spatial grid of size $S$ and $T$ time steps, $D_{ST}$ will have
-$3Tn_rn_c-2n_rn_c-Tn_r$ rows and $ST$ columns. For a $1^\circ\times
-1^\circ$ grid over the
-entire northern hemisphere and daily data over 10 years, we have
-$n_r=90$, $n_c=180$, $T=3650$ and so $D_{ST}$ has approximately
-$10^8$ columns and $10^8$
-rows. In each
-step of \autoref{alg:pdip}, we need to solve a linear system of
-equations in $A$ which depends on $D_{ST}^\top D_{ST}$ (see \citep{boyd_convex_2004}
-equation 11.54). Therefore, 
-applying the PDIP directly is infeasible for our data.\footnote{We note that this
-  is a highly structured and sparse matrix, but, unlike trend
-  filtering alone, it is not banded. We are unaware of general
-  linear algebra techniques for inverting such matrix, despite our
-  best efforts.}
+For a spatial grid of size $S$ and $T$ time steps, $D_{ST}$ will have $3Tn_rn_c-2n_rn_c-Tn_r$ rows and $ST$ columns.
+For a $1^\circ\times 1^\circ$ grid over the entire northern hemisphere and daily data over 10 years, we have $n_r=90$, $n_c=180$, $T=3650$ and so $D_{ST}$ has approximately $10^8$ columns and $10^8$ rows.
+In each step of the PDIP algorithm, we need to solve a linear system of equations in $A$ which depends on $D_{ST}^\top D_{ST}$ (see appendix A and B).
+Therefore, applying the PDIP directly is infeasible for our data.\footnote{We note that this is a highly structured and sparse matrix, but, unlike trend filtering alone, it is not banded.
+We are unaware of general linear algebra techniques for inverting such matrix, despite our best efforts.}
 
@@
 second uses proximal methods to avoid matrix inversions.
+
 \attn{We note that stochastic gradient descent could be used but...}
+
+\attn{Arash: The primal objective is not differentiable and so we cannot use SGD.
+The dual has constriants and so the plane SGD cannot be used.
+We can use projected SGD but I think it would converge slowly.
+Since other papers on LASSO have not considered SGD I don't think we need to justify why we did not use it.}
 
@@
 The steps are those mentioned in \ref{eq:ADMM_steps} and the x-update is now explained in the appendix.
-So we may want to remove algorithm \ref{alg:ADMM}.}
+So we may want to remove algorithm \ref{alg:ADMM}.
+}
 
@@
 
-In this appendix we provide more details on how to solve the optimization problem \ref{eq:l1tf_var}
+In this appendix we provide more details on how to solve the optimization problem \ref{eq:l1tf_var} using PDIP.
+The objective function is convex but not differentiable.
+Therefore, to be able to use PDIP we first need to derive the dual of this problem.
+We note that this is a generalized LASSO problem \citep{tibshirani_solution_2011}.
+The dual of a generalized LASSO with the objective $f(x)+\lambda \norm{ Dx }_1$ is:
+
+\begin{align}
+\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
+\end{align}
+
+where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that
+
+\begin{equation}
+f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
+\label{eq:conj}
+\end{equation}
+
+Each iteration of PDIP involves computing a search direction by taking a Newton step for the system of nonlinear equations $r_w(v,\mu_1,\mu_2)=0$, where $w>0$ is a parameter and
+
+\begin{equation}
+  r_w(v,\mu_1,\mu_2):=
+	\begin{bmatrix}
+	r_{dual}\\
+	r_{cent}	
+	\end{bmatrix}=
+  \begin{bmatrix}
+    \nabla f^*(-D^\top v) + \mu_1 - \mu_2\\
+    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
+  \end{bmatrix}
+\label{eq:resid}
+\end{equation}
+
+for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form
+
+\begin{equation}
+r_w(v,\mu_1,\mu_2)+A
+\begin{bmatrix}
+	\nabla v\\
+	\nabla \mu_1\\
+	\nabla \mu_2	
+	\end{bmatrix}= 0
+\label{eq:newton_step}
+\end{equation}
+
+We have:
+
+\begin{equation}
+A=
+\begin{bmatrix}
+	\nabla^2 f^*(-D^\top v) & I & -I\\
+	-\mathbf{diag(\mu_1)}\one & -v+\lambda\one & \mathbf{0}\\
+	\mathbf{diag(\mu_2)}\one & v+\lambda\one & \mathbf{0}
+	\end{bmatrix}
+\label{eq:delta_r}
+\end{equation}
+
+Therefore, to perform the Newton step we need to compute $\nabla f^*(-D^\top v)$ and $\nabla^2 f^*(-D^\top v)$. It is straightforward to show that
+
+\begin{align}
+\nabla f^*(-D^\top v) = -\nabla_u f^*(u) D^\top ,\quad u=-D^\top v, \quad (\nabla_u f^*(u))_j=\log\bigg(\frac{y_j^2}{1-u_j}\bigg) \\
+\nabla^2 f^*(-D^\top v)=D\nabla_u^2 f^*(u)D^\top, \quad (\nabla_u^2 f^*(u))_j=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg)
+\end{align}
 

### 2018-05-17 nips-2018/paper.tex
update paper

@@
 \usepackage{subfigure} 
-\graphicspath{{../working-paper/}}
+\graphicspath{{../nips-2018/}}
 
@@
 
-\renewcommand*{\figureautorefname}{Figure}%
+\renewcommand*{\figureautorefname}{}%
 \renewcommand*{\tableautorefname}{Table}%
@@
 
-In this project, we consider the problem of detecting the trend in the
+In this paper, we consider the problem of detecting the trend in the
 temperature volatility. All analyses are performed on a sub-set of the
-European Centre for Medium-Range Weather Forecasts (ECMWF) ERA-40
+European Centre for Medium-Range Weather Forecasts (ECMWF) ERA-20C
 dataset \citep{uppala_era-40_2005}. This dataset include the
@@
 in time and space
+
 \begin{align}
@@
 \end{align}
-In words, the variance at each time and location is computed as the
-weighted sum of $S$ bell-shaped functions where the weights are
-time-varying, consist of a linear trend $\alpha_s \cdot t$ and a
-periodic term $\beta_s \cdot \sin(2\pi\omega_s t+\phi_s)$. The
-bell-shaped functions impose the spatial smoothness, and the linear
-trend and the periodic terms enforce the temporal smoothness similar
-to the seasonal component in the real climate data. We
-simulated the data on a 5 by 7 grid and for 780 time steps with
-$S=4$. The parameters of the variance function are shown in
-\autoref{tab:sim_params}. For reference, we plot the variance function for all
-locations at $t=25$ and $t=45$ in
-as well as
-the variance across time at $(0,0)$ in
-\autoref{fig:true_var_spatial}.
+
+In words, the variance at each time and location is computed as the weighted sum of $S$ bell-shaped functions where the weights are time-varying, consist of a linear trend $\alpha_s \cdot t$ and a periodic term $\beta_s \cdot \sin(2\pi\omega_s t+\phi_s)$.
+The bell-shaped functions impose the spatial smoothness, and the linear trend and the periodic terms enforce the temporal smoothness similar to the seasonal component in the real climate data.
+We simulated the data on a 5 by 7 grid and for 780 time steps with $S=4$.
+The parameters of the variance function are shown in \autoref{tab:sim_params} in Appendix C.
+For reference, we plot the variance function for all locations at $t=25$ and $t=45$ in as well as the variance across time at $(0,0)$ in \autoref{fig:true_var_spatial} in Appendix C.
+
+We estimated the linearized ADMM for all combinations of values of
+$\lambda_t$ and $\lambda_s$ from the sets $\lambda_t \in
+\{0,1,5,10,50,100\}$ and $\lambda_s \in \{0,0.05,0.1,0.2,0.3\}$. For
+each pair, we then compute the mean absolute error (MAE) between the
+estimated variance and the true variance at all locations and all time
+steps. For $\lambda_t=5$ and $\lambda_s=0.1$ MAE was minimized. The
+left panel of \autoref{fig:true_fitted_var} shows the true and the
+estimated standard deviation at location (0,0) using $\lambda_s=0.1$
+and $\lambda_t=5$ (blue) and $\lambda_t=100$ (green). As we can see,
+larger than optimal value of $\lambda_t$ leads to estimated values
+which are ``too smooth".
+
+\begin{figure}[tb]
+  \centering	
+  \includegraphics[height=.12\textheight]{Figures/true_fitted_var}
+  \includegraphics[height=.12\textheight]{Figures/convergence}  
+  \includegraphics[height=.12\textheight]{Figures/modelComp_MAE}  
+  \caption{Left: The true (orange) and estimated standard deviation 
+    function at the location (0,0). The estimated values are
+    obtained using linearized ADMM with $\lambda_s=0.1$ and two
+    values of $\lambda_t$: $\lambda_t=5$ (blue) and
+    $\lambda_t=100$ (green).
+Middle: Convergence speed of linearized and consensus ADMM.
+Right: MAE for four models: admm\_opt: the proposed model with optimal values of $\lambda_t$ and $\lambda_s$, admm\_temp: no spatial penalty, admm\_sp: no temporal penalty.} \label{fig:true_fitted_var}
+\end{figure}
+
+The middle panel of \autoref{fig:true_fitted_var} shows the convergence of Algorithms~\ref{alg:ADMM} and~\ref{alg:linADMM}.
+Each iteration of the linearized algorithm takes 0.01 seconds on average while each iteration of the consensus ADMM takes about 20 seconds.
+
+To further examine the performance of the proposed model, we next compare it to three models: a model which does not consider the spatial smoothness and so is equivalent to fitting the model in \autoref{sec:l1tf_var} to each time-series separately, a model which does not consider the temporal smoothness, and a GARCH(1,1) model.
+We simulated 100 datasets using the method explained above with $\sigma_s \sim \mathsf{unif}(4,7)$.
+The right panel of \autoref{fig:true_fitted_var} shows the boxplot of MAE for these models.
+Interestingly, the proposed model with optimal parameters outperforms GARCH(1,1) in estimating the true value of the variance.
+
+\attn{These data are pretty small. Is it possible to do the PDIP?}
+
+\attn{Arash: we can, but why? it will be very slow.}
+
+\subsection{Data analysis}
+
+\autoref{alg:ADMM} is appropriate only if
+we parallelize it over multiple machines, and it is significantly
+slower on our simulated data, so we do not pursue it
+further here. All the results reported in this section are obtained
+using \autoref{alg:linADMM}. We
+applied this algorithm to the northern hemisphere of the ERA-20C
+dataset available from the \href{European Center for Medium-Range
+  Weather Forecasts}{https://www.ecmwf.int}. The data are
+the 2 meter temperature measured daily at 12 p.m from January 1, 1960
+to December 24, 2010.
+
+???
+
+To remedy this, we add the following long horizon penalty
+to \eqref{eq:l1tf_var}:
+
+\begin{equation}
+\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
+\label{eq:lh_penalty}
+\end{equation}
+
+where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
+Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
+Since we are working on the weekly averaged data, each year corresponds to 52 observations.
+In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
+The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
+The annual average of the estimated SDs shows a linear trend with a positive slope.
+
+\attn{From here down needs to be cut down and corrected with the right numbers.}
+
+\paragraph{Convergence}
+
+We used the following rule to determine when to stop the optimization:
+the optimization was stopped if the value of the loss did not improve
+by at least 0.1\% in 1000 trials. As we can see, the algorithm
+converged in about 2000 iterations. This took about 11 minutes. Our
+experiments showed that the convergence speed depends on the value of
+$\lambda_t$ and $\lambda_s$. Also, if we use the solution obtained for
+smaller values of these parameters as the initial value for the larger
+values (\textit{warm start}), the converges speed improves.
+
+\paragraph{Model selection}
+One common method for choosing the penalty parameters in the Lasso
+problems is to find the solution for a range of the values of these
+parameters and then choose the values which minimize a model selection
+criterion. However, such analyses needs the computation of the degrees
+of freedom (df). Several previous work have investigated the df in
+generalized lasso problems
+\citep{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}. However,
+all these studies have considered the linear regression problem and,
+to the best of our knowledge, the problem of computing the df for
+generalized lasso with general objective function has not been
+considered yet.
+
+Another approach is to choose the set of values which minimize an
+estimate of the expected prediction error obtained by k-fold
+cross-validation \citep{tibshirani_regression_1996} . Although this
+method is applicable for our problem, it needs k times more
+computation.
+
+In this paper, we use a heuristic method for choosing $\lambda_t$ and
+$\lambda_s$: we compute the optimal solution for a range of values of
+these parameters and choose the values which minimize
+$\mathscr{L}(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D_{total}h
+\lVert$. This objective is a compromise between the negative log
+likelihood ($-l(y|h)$) and the complexity of the solution ($\sum
+\lVert D_{total}h \lVert$). For smoother solutions the value of $\sum
+\lVert D_{total}h \lVert$ will be smaller but with the cost of larger
+$-l(y|h)$.
+
+We computed the optimal solution for all the combinations of the
+following sets of values: $\lambda_t \in \{0,2,4,8,10,15,200,1000\} \, \, ,
+\lambda_s \in \{0,.1,.5,2,5,10\}$. The best combination based on a held
+out set was $\lambda_t=4$ and $\lambda_s=2$. All the analyses in the
+next section are performed on the solution for these values.
+
+\paragraph{Analysis of trend of temperature volatility}
+
+The top row of \autoref{fig:avg_change_estimatedSD} shows the
+detrended data, the estimated standard deviation and the yearly
+average of these estimates for two cities in the US: Bloomington (left) and San Diego (right).
+The estimated SD captures the periodic behavior in the variance of the time-series.
+In addition, the number of linear segments changes adaptively in each time window depending on how fast the variance is changing.
+
+The yearly average of the estimated SD captures the trend in the
+temperature volatility. For example, we can see that in Bloomington,
+there is a small positive trend. To determine how the volatility has
+changed in each location, we subtract the average of the estimated
+variance in 1992 from the average in the following years and compute
+their sum. The value of this change in the variance in each location
+is depicted in the right panel of \autoref{fig:avg_change_estimatedSD}.
+The left panel of this figure, shows the average estimated variance in each location.
+Since the optimal value of the spatial penalty is rather large ($\lambda_s=2$) the estimated variance is spatially very smooth.
+
+It is interesting to note that the trend in volatility is almost zero over the oceans. The most positive trend can be observed in Asia and particularly in south-east Asia.
+
+The bottom panel of \autoref{fig:avg_change_estimatedSD} shows the histogram of change in the estimated SD across the northern hemisphere.
+As we can see, the SD in most locations on the northern hemisphere had a negative trend in this time period.
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=.9 \columnwidth]{Figures/ts_estimatedVar}\\
+  \includegraphics[width=.45\linewidth]{Figures/avg_logVar.pdf}
+  \includegraphics[width=.45\linewidth]{Figures/avg_change_logVar.pdf}\\
+    \includegraphics[width=.3\linewidth]{Figures/hist_avg_change.pdf}
+  \caption{Top row: Detrended data and the estimated SD for
+    Bloomington (left) and San Diego (right). Middle row: the
+    average of the estimated variance over the northern hemisphere (left) and the
+    change in the variance from 1961 to 2011 (right). Bottom: the histogram of change in estimated SD.} 
+  \label{fig:avg_change_estimatedSD}
+\end{figure}
+
+\section{Discussion}
+In this paper, we proposed a new method for estimating the variance of
+spatio-temporal data. The main idea is to cast this problem as a
+constrained optimization problem where the constraints enforce smooth
+changes in the variance for neighboring points in time and space. In
+particular, the solution is piecewise linear in time and piecewise
+constant in space. The resulting optimization is in the form of a
+generalized LASSO problem with high-dimension, and so applying the
+PDIP method directly is infeasible. We therefore developed two
+ADMM-based algorithms to solve this problem: the consensus ADMM and
+linearized ADMM.
+
+The consensus ADMM algorithm converges in few hundreds of iterations
+but each iteration takes much longer than the linearized ADMM
+algorithm. The appealing feature of the consensus ADMM algorithm is
+that if it is parallelized on enough number of machines the
+computation time per iteration remains constant as the problem size
+increases. The linearized ADMM algorithm, on the other hand converges
+in few thousands of iterations but each iteration is performed in
+split second. However, since the algorithm converges in many
+iterations it is not very appropriate for parallelization. The reason
+is that after each iteration the solution computed in each machine
+should be broadcast to the master machine and this operation takes
+some time which depends on the speed of the network connecting the
+slave machines to the master. A direction for future research would be
+to combine these two algorithms in the following way: the problem
+should be split into the sub-problems (as in the consensus ADMM) but
+each sub-problem can be solved using linearized ADMM.
+
+We applied the linearized ADMM algorithm to the surface temperature
+data on a grid over the united states, for years 1992-2002. The
+results showed that in many locations the variance of the temperature
+has increased about 1 unit in 10 years.
+
+The goal of this paper, however, is not to make any conclusions about
+the trend in the variance because we solved the problem only for a
+grid over the united states and for 10 years of the data. A thorough
+analysis, needs the full solution over the globe and for a longer time
+period. The goal of the paper, was to propose the idea of estimating
+the trend in variance of spatio-temporal signals using generalized
+lasso and to investigate the algorithms for solving the resulting
+optimization problem.
+
+\section{Appendix A}
+\label{sec:app_l1tf_var}
+
+In this appendix we provide more details on how to solve the optimization problem \ref{eq:l1tf_var} using PDIP.
+The objective function is convex but not differentiable.
+Therefore, to be able to use PDIP we first need to derive the dual of this problem.
+We note that this is a generalized LASSO problem \citep{tibshirani_solution_2011}.
+The dual of a generalized LASSO with the objective $f(x)+\lambda \norm{ Dx }_1$ is:
+
+\begin{align}
+\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
+\end{align}
+
+where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that
+
+\begin{equation}
+f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
+\label{eq:conj}
+\end{equation}
+
+Each iteration of PDIP involves computing a search direction by taking a Newton step for the system of nonlinear equations $r_w(v,\mu_1,\mu_2)=0$, where $w>0$ is a parameter and
+
+\begin{equation}
+  r_w(v,\mu_1,\mu_2):=
+	\begin{bmatrix}
+	r_{dual}\\
+	r_{cent}	
+	\end{bmatrix}=
+  \begin{bmatrix}
+    \nabla f^*(-D^\top v) + \mu_1 - \mu_2\\
+    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
+  \end{bmatrix}
+\label{eq:resid}
+\end{equation}
+
+for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form
+
+\begin{equation}
+r_w(v,\mu_1,\mu_2)+A
+\begin{bmatrix}
+	\nabla v\\
+	\nabla \mu_1\\
+	\nabla \mu_2	
+	\end{bmatrix}= 0
+\label{eq:newton_step}
+\end{equation}
+
+We have:
+
+\begin{equation}
+A=
+\begin{bmatrix}
+	\nabla^2 f^*(-D^\top v) & I & -I\\
+	-\mathbf{diag(\mu_1)}\one & -v+\lambda\one & \mathbf{0}\\
+	\mathbf{diag(\mu_2)}\one & v+\lambda\one & \mathbf{0}
+	\end{bmatrix}
+\label{eq:delta_r}
+\end{equation}
+
+Therefore, to perform the Newton step we need to compute $\nabla f^*(-D^\top v)$ and $\nabla^2 f^*(-D^\top v)$. It is straightforward to show that
+
+\begin{align}
+\nabla f^*(-D^\top v) = -\nabla_u f^*(u) D^\top ,\quad u=-D^\top v, \quad (\nabla_u f^*(u))_j=\log\bigg(\frac{y_j^2}{1-u_j}\bigg) \\
+\nabla^2 f^*(-D^\top v)=D\nabla_u^2 f^*(u)D^\top, \quad (\nabla_u^2 f^*(u))_j=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg)
+\end{align}
+
+\section{Appendix B}
+\label{sec:app_consADMM}
+\attn{put this in nips appendix format}
+
+In this Appendix we give more details on performing the x-update step in \autoref{eq:ADMM_steps}. We need to solve the following optimization problem:
+
+\begin{equation}
+\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
+\label{eq:x_update_opt}
+\end{equation}
+
+\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$.
+Let $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$.
+As it was explained in \autoref{sec:l1tf_var}, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$.
+So to use PDIP we first need to compute the conjugate function $f^*(\cdot)$.
+We have:
+
+\begin{align}
+f^*(\xi) & = \max_x \quad \xi^\top x - f(x)\\
+& =  \max_x \quad \sum_{j=1}^{n_b} (\xi_jx_j - x_j - y_j^2e^{-x_j} - (\rho/2)(x_j-\tilde{z}_j+u_j))
+\label{eq:conjugate}
+\end{align}
+
+Setting the derivative of the terms inside the summation to 0, we obtain:
+
+\begin{equation}
+\xi_j-y_j^2e^{-x_j^*}-\rho x_j^* + \rho (\tilde{z}_j-u_j)=0
+\label{eq:x_start}
+\end{equation}
+
+\noindent where $x^*$ is the maximizer in \ref{eq:conjugate}. Then, it can be shown that $x_j^*$ which satisfies \eqref{eq:x_start} can be obtained as follows:
+
+\begin{align}
+x^*_j & = \mathscr{W}\bigg(\frac{y_j^2}{\rho} e^{\phi_j} \bigg) - \phi_j \\
+\phi_j & =\frac{1-\xi_j-\rho(\tilde{z}_j-u_j)}{\rho}
+\end{align}
+
+In this equation, $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996}.
+Finally, the conjugate function is: $f^*(\xi) = \sum_{j=1}^{n_b} (\xi_jx^*_j - x^*_j - y_j^2e^{-x^*_j} - (\rho/2)(x^*_j-\tilde{z}_j+u_j))$.
+
+To use PDIP, we also need to evaluate $\nabla f^*$ and $\nabla^2 f^*$.
+First note that $\frac{\partial \mathscr{W}(q)}{\partial q} = \frac{\mathscr{W}(q)}{q(1+\mathscr{W}(q))}$ and $\frac{\partial^2 \mathscr{W}(q)}{\partial q^2} = - \frac{\mathscr{W}^2(q)(\mathscr{W}(q)+q)}{q^2(1+\mathscr{W}(q))^3}$.
+Using the chain rule we get:
+
+%\x^*_j + \xi_j + \frac{\partial \x^*_j}{\partial \xi_j}
+
+\begin{equation}
+\frac{\partial f^*(\xi)}{\partial \xi_j}  =  x^*_j  + \frac{\partial x^*_j}{\partial \xi_j} \bigg[ \xi_j -1 + y_j^2 e^{-x_j^*} + \rho (\tilde{z}_j - u_j - x_j^*) \bigg]
+\label{eq:d_f*_start}
+\end{equation}
+
+\noindent where we have:
+
+\begin{equation}
+\frac{\partial x_j^*}{\partial \xi_j} & = \frac{1}{\rho(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))}
+\label{eq:d_x*_start}
+\end{equation}
+
+By some tedious but straightforward computation we can obtain the second derivatives:
+
+\begin{align}
+\frac{\partial^2 f^*(\xi)}{\partial \xi_j^2} & =  \frac{\partial x_j^*}{\partial \xi_j} - \rho \frac{\partial^2 x_j^*}{\partial \xi_j^2} \bigg[ \phi_j +x_j^* - \tilde{z}_j + u_j \bigg]\\
+& \quad + \frac{\partial x_j^*}{\partial \xi_j} \bigg[ 1-y_j^2 \frac{\partial x_j^*}{\partial \xi_j} e^{-x_j^*} -\rho \frac{\partial x_j^*}{\partial \xi_j} \bigg]
+\label{eq:d2_f*_start}
+\end{align}
+
+\begin{equation}
+\frac{\partial^2 x_j^*}{\partial \xi_j^2}  = \frac{\mathscr{W}((y_j^2/\rho) e^{-\phi_j} )}{\rho^2(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))^3}
+\label{eq:d2_x*_start}
+\end{equation}
+
+Having computed the conjugate function and its gradient and Jacobian, now we can use a number of convex optimization software packages which have an implementation of PDIP to perform the x-update step inside the ADMM loop.
+We chose the python API of the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.
+
+\section{Appendix C}
+Table \autoref{tab:sim_params} lists the parameters used for simulating data in \autoref{sec:simulations}.
+\autoref{fig:true_var_spatial} shows the variance function obtained from there parameters at $t=25$ (left) and $t=45$ (center).
 
@@
 
-We estimated the linearized ADMM for all combinations of values of
-$\lambda_t$ and $\lambda_s$ from the sets $\lambda_t \in
-\{0,1,5,10,50,100\}$ and $\lambda_s \in \{0,0.05,0.1,0.2,0.3\}$. For
-each pair, we then compute the mean absolute error (MAE) between the
-estimated variance and the true variance at all locations and all time
-steps. For $\lambda_t=5$ and $\lambda_s=0.1$ MAE was minimized. The
-right panel of \autoref{fig:true_fitted_var} shows the true and the
-estimated standard deviation at location (0,0) using $\lambda_s=0.1$
-and $\lambda_t=5$ (blue) and $\lambda_t=100$ (green). As we can see,
-larger than optimal value of $\lambda_t$ leads to estimated values
-which are ``too smooth".
-
-\autoref{fig:true_fitted_var} shows the convergence
-of Algorithms~\ref{alg:ADMM} and~\ref{alg:linADMM}. Each iteration of
-the linearized algorithm 
-takes 0.01 seconds on average while each iteration of the
-consensus ADMM takes about 20 seconds.  \attn{These data are pretty
-  small. Is it possible to do the PDIP?}
+\section{Appendix D}
+
+In this appendix we examine some of the properties of the time-series of the temperature in ERA-20C dataset.
+The goal here is to demonstrate some of the difficulties in modeling the trend in the temperature volatility and motivate our methodology.
+
+\autoref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil).
+The time-series of Bloomington and San Diego show clear cyclic behavior.
+However, while it seems (qualitatively) that these cycles can be modeled by a sinusoidal function for Bloomington, the same is not true for San Diego.
+Also, the amplitude of the cycles changes from some years to others.
+The time-series of Manaus does not show any regular cyclic behavior.
+This demonstrates the first difficulty in analyzing the variance of this data: to analyze the variance, we first need to remove the cyclic terms from all time-series.
+However, there is a lot of variations in the cyclic behavior of the time-series of different locations.
+In addition, some of these cycles cannot be easily modeled by a parametric function \footnote{One might try to model the cycles by the summation of sinusoidal terms with different frequencies.
+However, for some time-series this may need many terms to be included in the summation to achieve a reasonable level of accuracy.
+In addition, this model cannot capture the non-stationarity in the cycles.}.
+To overcome these issues, we use a non-parametric approach to remove the cyclic terms from the time-series and de-trend them.
+This approach, called \textit{$\ell_1$-trend filtering} is explained in Section 2 of the text.
+We detrended each time-series separately using this method.
+For each time-series, we found the optimal value of the penalty parameter using \textit{k-fold cross-validation} with $k=5$.
+We used the R package \textbf{genlasso} to perform these computations \cite{arnold_efficient_2016}.
+
 \begin{figure}[tb]
 	\centering
-  \includegraphics[height=.15\textheight]{Figures/convergence}	
-  \caption{Convergence speed of linearized and consensus
-    ADMM.} \label{fig:true_fitted_var}  
+	\includegraphics[width=.2\textheight]{Figures/cities_ts}
+ 	\caption{Time-series of the temperature (in Kelvin) of three cities.}
+ 	\label{fig:cities_ts}
 \end{figure}
 
-\subsection{Data analysis}
-
-\autoref{alg:ADMM} is appropriate only if
-we parallelize it over multiple machines, and it is significantly
-slower on our simulated data, so we do not pursue it
-further here. All the results reported in this section are obtained
-using \autoref{alg:linADMM}. We
-applied this algorithm to the northern hemisphere of the ERA-40
-dataset available from the \href{European Center for Medium-Range
-  Weather Forecasts}{https://www.ecmwf.int}. The data are
-the 2 meter temperature measured daily at 12 p.m from August 31 of
-1992 to 2002.
-
-\attn{From here down needs to be cut down and corrected with the right numbers.}
-
-\paragraph{Data Exploration}
-The red curve in the left panel of \autoref{fig:bloom_estimatedSD}
-shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
-the time-series of Bloomington. To reduce the number of time-steps in
-this figure and in the remainder of the paper we work on the weekly
-averaged of the data.
-
-The curve of the estimated SD captures the periodic variations in the
-SD of the signal. Just by looking at this curve, it is hard to say if
-the SD is decreasing or increasing. Therefore, we compute the average
-of the estimated SD for each year. The estimated SD together with this
-annual average is shown in the middle panel of
-\autoref{fig:bloom_estimatedSD}. As it can be seen, the annual trend
-is not smooth. This is because in the optimization problem
-\eqref{eq:l1tf_var}, the smoothness of the annual trend is not
-encouraged. To remedy this, we add the following long horizon penalty
-to \eqref{eq:l1tf_var}:
-
-\begin{equation}
-\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
-\label{eq:lh_penalty}
-\end{equation}
-
-where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
-Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
-Since we are working on the weekly averaged data, each year corresponds to 52 observations.
-In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
-The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
-The annual average of the estimated SDs shows a linear trend with a positive slope.
+The blue curve in the left panel of \autoref{fig:bloom_estimatedSD} shows the time-series of the temperature of Bloomington after detrending using this method.
+This figure, reveals another difficulty in estimating the trend of volatility in this data: the variance of this signal, shows cyclic behavior.
+Also, the cycles are not regular and their amplitude and frequency change.
+Even if one can describe the behavior of the variance of the time-series at all locations using a single parametric model (for example a variant of the GARCH models \citep{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework.
+These observations motivate the need to develop a non-parametric framework for the problem at hand.
 
@@
 
-%TODO: how to choose lambda? plot BIC as a function of lambda.
-
-This section is devoted to exploring some of the properties of the
-ERA-40 surface temperature data set. The goal here is to demonstrate
-some of the difficulties in modeling the trend in the temperature
-volatility and motivate our methodology for doing so.
-
-% \autoref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil).
-The time-series of Bloomington and San Diego show clear cyclic behavior.
-However, while it seems (qualitatively) that these cycles can be modeled by a sinusoidal function for Bloomington, the same is not true for San Diego.
-Also, the amplitude of the cycles changes from some years to others.
-The time-series of Manaus does not show any regular cyclic behavior.
-This demonstrates the first difficulty in analyzing the variance of this data: to analyze the variance, we first need to remove the cyclic terms from all time-series.
-However, there is a lot of variations in the cyclic behavior of the time-series of different locations.
-In addition, some of these cycles cannot be easily modeled by a parametric function \footnote{One might try to model the cycles by the summation of sinusoidal terms with different frequencies.
-However, for some time-series this may need many terms to be included in the summation to achieve a reasonable level of accuracy.
-In addition, this model cannot capture the non-stationarity in the cycles.}.
-To overcome these issues, we will use a non-parametric approach to remove the cyclic terms from the time-series and de-trend them.
-This approach, called \textit{$\ell_1$-trend filtering} is explained in the next section.
-
-% \begin{figure}[tb]
-% 	\vskip 0.2in
-% 	\begin{center}
-% 		\centerline{\includegraphics[width=\columnwidth]{Figures/cities_ts}}
-% 		\caption{Time-series of the temperature (in Kelvin) of three cities.}
-% 		\label{fig:cities_ts}
-% 	\end{center}
-% 	\vskip -0.2in
-% \end{figure}
-
-The right panel of \autoref{fig:bloom_detrended} shows the time-series
-of the temperature of Bloomington, after removing the cyclic terms and
-de-trending using the method explained in the next section. The goal
-is to investigate the trend in the variance of this signal. This
-figure, reveals another issue toward this goal: the variance of this
-signal, shows cyclic behavior. Also, the cycles are not regular and
-their amplitude and frequency change. Even if one can describe the
-behavior of the variance of all the time-series using a single
-parametric model (for example a variant of the GARCH models
-\citep{bollerslev_generalized_1986}), it is not clear how the trend in
-the variance should be investigated in this framework. These
-observations motivate the need to develop a non-parametric framework
-for the problem at hand.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=\columnwidth]{Figures/bloom_trend_resid}
-  \caption{Left: Time-series of the temperature of Bloomington (blue)
-    and the estimated trend and cycles obtained from the
-    $\ell_1$-trend filtering (red). Right: the same time-series after
-    removing the cyclic terms and de-trending using $\ell_1$-trend
-    filtering.} 
-  \label{fig:bloom_detrended}
-\end{figure}
-
-\paragraph{Convergence}
-
-We used the following rule to determine when to stop the optimization:
-the optimization was stopped if the value of the loss did not improve
-by at least 0.1\% in 1000 trials. As we can see, the algorithm
-converged in about 2000 iterations. This took about 11 minutes. Our
-experiments showed that the convergence speed depends on the value of
-$\lambda_t$ and $\lambda_s$. Also, if we use the solution obtained for
-smaller values of these parameters as the initial value for the larger
-values (\textit{warm start}), the converges speed improves.
-
-\paragraph{Model selection}
-One common method for choosing the penalty parameters in the Lasso
-problems is to find the solution for a range of the values of these
-parameters and then choose the values which minimize a model selection
-criterion. However, such analyses needs the computation of the degrees
-of freedom (df). Several previous work have investigated the df in
-generalized lasso problems
-\citep{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}. However,
-all these studies have considered the linear regression problem and,
-to the best of our knowledge, the problem of computing the df for
-generalized lasso with general objective function has not been
-considered yet.
-
-Another approach is to choose the set of values which minimize an
-estimate of the expected prediction error obtained by k-fold
-cross-validation \citep{tibshirani_regression_1996} . Although this
-method is applicable for our problem, it needs k times more
-computation.
-
-In this paper, we use a heuristic method for choosing $\lambda_t$ and
-$\lambda_s$: we compute the optimal solution for a range of values of
-these parameters and choose the values which minimize
-$\mathscr{L}(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D_{total}h
-\lVert$. This objective is a compromise between the negative log
-likelihood ($-l(y|h)$) and the complexity of the solution ($\sum
-\lVert D_{total}h \lVert$). For smoother solutions the value of $\sum
-\lVert D_{total}h \lVert$ will be smaller but with the cost of larger
-$-l(y|h)$.
-
-We computed the optimal solution for all the combinations of the
-following sets of values: $\lambda_t \in \{1,5,10,20\} \, \, ,
-\lambda_s \in \{0,.1,1,5,10\}$. The best combination based on a held
-out set was $\lambda_t=5$ and $\lambda_s=1$. All the analyses in the
-next section are performed on the solution for these values.
-
-\paragraph{Analysis of trend of temperature volatility}
-
-The top row of \autoref{fig:avg_change_estimatedSD} shows the
-detrended data, the estimated standard deviation and the yearly
-average of these estimates for two cities in the US. The estimated SD
-captures the periodic behavior in the variance of the time-series. In
-addition, the number of linear segments changes adaptively in each
-time window depending on how fast the variance is changing.
-
-The yearly average of the estimated SD captures the trend in the
-temperature volatility. For example, we can see that in Bloomington,
-there is a small positive trend. To determine how the volatility has
-changed in each location, we subtract the average of the estimated
-variance in 1992 from the average in the following years and compute
-their sum. The value of this change in the variance in each location
-is depicted in the right panel of
-\autoref{fig:avg_change_estimatedSD}. The left panel of this figure,
-shows the average estimated variance in each location.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=.9\columnwidth]{Figures/estimatedSD}\\
-  \includegraphics[width=.45\linewidth]{Figures/avg_estimatedVar}
-  \includegraphics[width=.45\linewidth]{Figures/avg_change_estimatedVar}
-  \caption{Top row: Detrended data and the estimated SD for
-    Bloomington (left) and San Diego (right). Bottom: the
-    average of the estimated variance over the US (left) and the
-    change in the variance from 1992 to 2002 (right)} 
-  \label{fig:avg_change_estimatedSD}
-\end{figure} 
-It is interesting to note that the trend in volatility is almost zero
-over the oceans. The most positive trend can be observed in the
-south-east and the most negative trend has happened in the north-east.
-
-\section{Discussion}
-In this paper, we proposed a new method for estimating the variance of
-spatio-temporal data. The main idea is to cast this problem as a
-constrained optimization problem where the constraints enforce smooth
-changes in the variance for neighboring points in time and space. In
-particular, the solution is piecewise linear in time and piecewise
-constant in space. The resulting optimization is in the form of a
-generalized LASSO problem with high-dimension, and so applying the
-PDIP method directly is infeasible. We therefore developed two
-ADMM-based algorithms to solve this problem: the consensus ADMM and
-linearized ADMM.
-
-The consensus ADMM algorithm converges in few hundreds of iterations
-but each iteration takes much longer than the linearized ADMM
-algorithm. The appealing feature of the consensus ADMM algorithm is
-that if it is parallelized on enough number of machines the
-computation time per iteration remains constant as the problem size
-increases. The linearized ADMM algorithm, on the other hand converges
-in few thousands of iterations but each iteration is performed in
-split second. However, since the algorithm converges in many
-iterations it is not very appropriate for parallelization. The reason
-is that after each iteration the solution computed in each machine
-should be broadcast to the master machine and this operation takes
-some time which depends on the speed of the network connecting the
-slave machines to the master. A direction for future research would be
-to combine these two algorithms in the following way: the problem
-should be split into the sub-problems (as in the consensus ADMM) but
-each sub-problem can be solved using linearized ADMM.
-
-We applied the linearized ADMM algorithm to the surface temperature
-data on a grid over the united states, for years 1992-2002. The
-results showed that in many locations the variance of the temperature
-has increased about 1 unit in 10 years.
-
-The goal of this paper, however, is not to make any conclusions about
-the trend in the variance because we solved the problem only for a
-grid over the united states and for 10 years of the data. A thorough
-analysis, needs the full solution over the globe and for a longer time
-period. The goal of the paper, was to propose the idea of estimating
-the trend in variance of spatio-temporal signals using generalized
-lasso and to investigate the algorithms for solving the resulting
-optimization problem.
-
-\section{Appendix A}
-\label{sec:app_l1tf_var}
-
-In this appendix we provide more details on how to solve the optimization problem \ref{eq:l1tf_var} using PDIP.
-The objective function is convex but not differentiable.
-Therefore, to be able to use PDIP we first need to derive the dual of this problem.
-We note that this is a generalized LASSO problem \citep{tibshirani_solution_2011}.
-The dual of a generalized LASSO with the objective $f(x)+\lambda \norm{ Dx }_1$ is:
-
-\begin{align}
-\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
-\end{align}
-
-where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that
-
-\begin{equation}
-f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
-\label{eq:conj}
-\end{equation}
-
-Each iteration of PDIP involves computing a search direction by taking a Newton step for the system of nonlinear equations $r_w(v,\mu_1,\mu_2)=0$, where $w>0$ is a parameter and
-
-\begin{equation}
-  r_w(v,\mu_1,\mu_2):=
-	\begin{bmatrix}
-	r_{dual}\\
-	r_{cent}	
-	\end{bmatrix}=
-  \begin{bmatrix}
-    \nabla f^*(-D^\top v) + \mu_1 - \mu_2\\
-    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
-  \end{bmatrix}
-\label{eq:resid}
-\end{equation}
-
-for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form
-
-\begin{equation}
-r_w(v,\mu_1,\mu_2)+A
-\begin{bmatrix}
-	\nabla v\\
-	\nabla \mu_1\\
-	\nabla \mu_2	
-	\end{bmatrix}= 0
-\label{eq:newton_step}
-\end{equation}
-
-We have:
-
-\begin{equation}
-A=
-\begin{bmatrix}
-	\nabla^2 f^*(-D^\top v) & I & -I\\
-	-\mathbf{diag(\mu_1)}\one & -v+\lambda\one & \mathbf{0}\\
-	\mathbf{diag(\mu_2)}\one & v+\lambda\one & \mathbf{0}
-	\end{bmatrix}
-\label{eq:delta_r}
-\end{equation}
-
-Therefore, to perform the Newton step we need to compute $\nabla f^*(-D^\top v)$ and $\nabla^2 f^*(-D^\top v)$. It is straightforward to show that
-
-\begin{align}
-\nabla f^*(-D^\top v) = -\nabla_u f^*(u) D^\top ,\quad u=-D^\top v, \quad (\nabla_u f^*(u))_j=\log\bigg(\frac{y_j^2}{1-u_j}\bigg) \\
-\nabla^2 f^*(-D^\top v)=D\nabla_u^2 f^*(u)D^\top, \quad (\nabla_u^2 f^*(u))_j=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg)
-\end{align}
-
-\section{Appendix B}
-\label{sec:app_consADMM}
-\attn{put this in nips appendix format}
-
-In this Appendix we give more details on performing the x-update step in \autoref{eq:ADMM_steps}. We need to solve the following optimization problem:
-
-\begin{equation}
-\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
-\label{eq:x_update_opt}
-\end{equation}
-
-\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$.
-Let $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$.
-As it was explained in \autoref{sec:l1tf_var}, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$.
-So to use PDIP we first need to compute the conjugate function $f^*(\cdot)$.
-We have:
-
-\begin{align}
-f^*(\xi) & = \max_x \quad \xi^\top x - f(x)\\
-& =  \max_x \quad \sum_{j=1}^{n_b} (\xi_jx_j - x_j - y_j^2e^{-x_j} - (\rho/2)(x_j-\tilde{z}_j+u_j))
-\label{eq:conjugate}
-\end{align}
-
-Setting the derivative of the terms inside the summation to 0, we obtain:
-
-\begin{equation}
-\xi_j-y_j^2e^{-x_j^*}-\rho x_j^* + \rho (\tilde{z}_j-u_j)=0
-\label{eq:x_start}
-\end{equation}
-
-\noindent where $x^*$ is the maximizer in \ref{eq:conjugate}. Then, it can be shown that $x_j^*$ which satisfies \eqref{eq:x_start} can be obtained as follows:
-
-\begin{align}
-x^*_j & = \mathscr{W}\bigg(\frac{y_j^2}{\rho} e^{\phi_j} \bigg) - \phi_j \\
-\phi_j & =\frac{1-\xi_j-\rho(\tilde{z}_j-u_j)}{\rho}
-\end{align}
-
-In this equation, $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996}.
-Finally, the conjugate function is: $f^*(\xi) = \sum_{j=1}^{n_b} (\xi_jx^*_j - x^*_j - y_j^2e^{-x^*_j} - (\rho/2)(x^*_j-\tilde{z}_j+u_j))$.
-
-To use PDIP, we also need to evaluate $\nabla f^*$ and $\nabla^2 f^*$.
-First note that $\frac{\partial \mathscr{W}(q)}{\partial q} = \frac{\mathscr{W}(q)}{q(1+\mathscr{W}(q))}$ and $\frac{\partial^2 \mathscr{W}(q)}{\partial q^2} = - \frac{\mathscr{W}^2(q)(\mathscr{W}(q)+q)}{q^2(1+\mathscr{W}(q))^3}$.
-Using the chain rule we get:
-
-%\x^*_j + \xi_j + \frac{\partial \x^*_j}{\partial \xi_j}
-
-\begin{equation}
-\frac{\partial f^*(\xi)}{\partial \xi_j}  =  x^*_j  + \frac{\partial x^*_j}{\partial \xi_j} \bigg[ \xi_j -1 + y_j^2 e^{-x_j^*} + \rho (\tilde{z}_j - u_j - x_j^*) \bigg]
-\label{eq:d_f*_start}
-\end{equation}
-
-\noindent where we have:
-
-\begin{equation}
-\frac{\partial x_j^*}{\partial \xi_j} & = \frac{1}{\rho(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))}
-\label{eq:d_x*_start}
-\end{equation}
-
-By some tedious but straightforward computation we can obtain the second derivatives:
-
-\begin{align}
-\frac{\partial^2 f^*(\xi)}{\partial \xi_j^2} & =  \frac{\partial x_j^*}{\partial \xi_j} - \rho \frac{\partial^2 x_j^*}{\partial \xi_j^2} \bigg[ \phi_j +x_j^* - \tilde{z}_j + u_j \bigg]\\
-& \quad + \frac{\partial x_j^*}{\partial \xi_j} \bigg[ 1-y_j^2 \frac{\partial x_j^*}{\partial \xi_j} e^{-x_j^*} -\rho \frac{\partial x_j^*}{\partial \xi_j} \bigg]
-\label{eq:d2_f*_start}
-\end{align}
-
-\begin{equation}
-\frac{\partial^2 x_j^*}{\partial \xi_j^2}  = \frac{\mathscr{W}((y_j^2/\rho) e^{-\phi_j} )}{\rho^2(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))^3}
-\label{eq:d2_x*_start}
-\end{equation}
-
-Having computed the conjugate function and its gradient and Jacobian, now we can use a number of convex optimization software packages which have an implementation of PDIP to perform the x-update step inside the ADMM loop.
-We chose the python API of the \texttt{cvxopt}~\citep{andersen_cvxopt:_2013} package.
+The red curve in the left panel of \autoref{fig:bloom_estimatedSD}
+shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
+the time-series of Bloomington. To reduce the number of time-steps in
+this figure and in the remainder of the paper we work on the weekly
+averaged of the data.
+
+The curve of the estimated SD captures the periodic variations in the
+SD of the signal. Just by looking at this curve, it is hard to say if
+the SD is decreasing or increasing. Therefore, we compute the average
+of the estimated SD for each year. The estimated SD together with this
+annual average is shown in the middle panel of
+\autoref{fig:bloom_estimatedSD}. As it can be seen, the annual trend
+is not smooth. This is because in the optimization problem
+\eqref{eq:l1tf_var}, the smoothness of the annual trend is not
+encouraged.
+To remedy this, we proposed a method to encourage the smoothness of the annual average trend.
+The estimated SDs using this method is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
+The annual average of the estimated SDs shows a linear trend with a positive slope.
 

### 2018-05-17 code/ModelSelection.py

@@ -3,7 +3,7 @@ from os.path import join
 import matplotlib.pyplot as plt
 import numpy as np
 
-lam_t_vec=[0,2,4,8,10,15,200]
+lam_t_vec=[0,2,4,8,10,15,200,1000]
 lam_s_vec=[0,.1,.5,2,5,10]
 mu_vec=[.01]*6+[.001]*3
 dataset='north_hemisphere'

### 2018-05-17 code/visualization.py

@@ -2,7 +2,7 @@ from volatilitytrend.algorithms.base import LinearizedADMM
 from os.path import join
 import matplotlib.pyplot as plt
 
-lam_t=200;lam_s=5;mu=.001
+lam_t=4;lam_s=2;mu=.01
 dataset='north_hemisphere'
 #dataset='us'
 

### 2018-05-17 code/volatilitytrend/algorithms/base.py

@@ -272,7 +272,8 @@ class LinearizedADMM(BaseAlgorithmClass):
                 X=self.fittedVar.reshape((-1,1))
                 
                 NLL=np.sum(X + o2*np.exp(-X))
-                df=np.sum(np.abs(D.dot(X))>.005)#or: np.sum(np.abs(D.dot(X)))
+#                df=np.sum(np.abs(D.dot(X))>.005)
+                df = np.sum(np.abs(D.dot(X)))
                 self.AIC[i,j] =  NLL + 2*df
                 self.BIC[i,j] =  NLL + np.log(X.size)*df
         #===compute loss for all combinations of values===

### 2018-05-17 nips-2018/paper.tex
updates on paper

@@
 \subsection{Data analysis}
+\label{sec:data}
 
@@
 
-???
-
-To remedy this, we add the following long horizon penalty
-to \eqref{eq:l1tf_var}:
-
-\begin{equation}
-\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
-\label{eq:lh_penalty}
-\end{equation}
-
-where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
-Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
-Since we are working on the weekly averaged data, each year corresponds to 52 observations.
-In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
-The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
+In Appendix D, we investigate some properties of the time-series of different locations on earth.
+We also explain how each time-series is detrended using $\ell_1$-trend filtering method explained in \autoref{sec:ell_1-trend-filt}.
+\autoref{fig:bloom_estimatedSD} shows a time-series after detrending using this method.
+The variance of this time-series have a cyclic behavior.
+The cycles are not regular and their amplitude and frequency change.
+In addition, the time-series of other locations show different patterns.
+These observations motivate the need to develop a non-parametric framework for the problem at hand.
+In this figure, the estimate SD obtained from the method of \autoref{sec:l1tf_var} is also shown.
+Sine the estimated SD captures the periodic behavior of volatility, it is hard to see the trend in volatility based on these estimated values.
+Therefore, we compute the annual average of these estimates.
+However, as this figure shows, the annual trend is not smooth.
+This is because in the optimization problem \eqref{eq:l1tf_var}, the smoothness of the annual trend is not encouraged.
+Therefore, in Appendix D we propose a simple method to encourage the smoothness of the annual average trend.
+The estimated SDs using this method is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
 The annual average of the estimated SDs shows a linear trend with a positive slope.
 
-\attn{From here down needs to be cut down and corrected with the right numbers.}
-
 \paragraph{Convergence}
 
-We used the following rule to determine when to stop the optimization:
-the optimization was stopped if the value of the loss did not improve
-by at least 0.1\% in 1000 trials. As we can see, the algorithm
-converged in about 2000 iterations. This took about 11 minutes. Our
-experiments showed that the convergence speed depends on the value of
-$\lambda_t$ and $\lambda_s$. Also, if we use the solution obtained for
-smaller values of these parameters as the initial value for the larger
-values (\textit{warm start}), the converges speed improves.
+We used the following rule to determine when to stop the optimization: the optimization was stopped if the value of the loss did not improve by at least 0.1\% in 1000 trials.
+Our experiments showed that the convergence speed depends on the value of $\lambda_t$ and $\lambda_s$.
+Also, if we use the solution obtained for smaller values of these parameters as the initial value for the larger values (\textit{warm start}), the converges speed improves.
 
@@
 following sets of values: $\lambda_t \in \{0,2,4,8,10,15,200,1000\} \, \, ,
-\lambda_s \in \{0,.1,.5,2,5,10\}$. The best combination based on a held
-out set was $\lambda_t=4$ and $\lambda_s=2$. All the analyses in the
-next section are performed on the solution for these values.
+\lambda_s \in \{0,.1,.5,2,5,10\}$. The best combination was $\lambda_t=4$ and $\lambda_s=2$. All the analyses in the next section are performed on the solution for these values.
 
@@
 shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
-the time-series of Bloomington. To reduce the number of time-steps in
-this figure and in the remainder of the paper we work on the weekly
-averaged of the data.
-
+the time-series of Bloomington obtained from out proposed model.
+To reduce the number of time-steps we work on the weekly averaged of the data.
 The curve of the estimated SD captures the periodic variations in the
@@
 \eqref{eq:l1tf_var}, the smoothness of the annual trend is not
-encouraged.
-To remedy this, we proposed a method to encourage the smoothness of the annual average trend.
-The estimated SDs using this method is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
+encouraged. To remedy this, we add the following long horizon penalty to \eqref{eq:l1tf_var}:
+
+\begin{equation}
+\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
+\label{eq:lh_penalty}
+\end{equation}
+
+where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
+Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
+Since we are working on the weekly averaged data, each year corresponds to 52 observations.
+In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
+The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
 The annual average of the estimated SDs shows a linear trend with a positive slope.

### 2018-05-17 code/AnalyseSyntheticData_modelComp.py

@@ -63,7 +63,7 @@ var_sources_weight_mat=np.concatenate((var_sources_weight1,
 dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/Simulated data'
 savedResultsDir='/home/arash/datasets/ecmwf/Fits_to_simulated_data/'+\
 'LinADMM_withTrend'
-saveFigDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Reports/051218/Figures'
+saveFigDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/globe/Figures'
 
 data_fn=join(dstDir,'simulated_data')
 metadata_fn=join(dstDir,'metadata')

### 2018-05-18 nips-2018/paper.tex
add appendix, cut down

--- 2018-05-17 nips-2018/paper.tex
@@
 \usepackage{hyperref}
+\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}
 
@@
 
-\renewcommand*{\figureautorefname}{}%
+\renewcommand*{\figureautorefname}{Figure}%
 \renewcommand*{\tableautorefname}{Table}%
@@
 \DeclareMathOperator*{\sign}{sign}
+
+\usepackage{xr}
+\externaldocument{paper-supplement}
 
@@
 
-\section{$\ell_1$-trend filtering for estimating variance of a
-  time-series}
+\section{Estimating the variance of spatio-temporal data}
 \label{sec:ell_1-trend-filt}
 
-\attn{Arash: should't the title of this section be "...estimating variance of spatio-temporal data" or something similar?}
-
-$\ell_1$-trend filtering was proposed by \citep{kim_$ell_1$_2009} as a method for estimating a smooth, time-varying trend. It is formulated as the optimization problem
+$\ell_1$-trend filtering was proposed by \citep{KimKoh2009} as a method for estimating a smooth, time-varying trend. It is formulated as the optimization problem
 % : a time-series $y_t$, $t=1,...,T$ is observed and we believe that
@@
 With the penalty matrix $D$, the estimated $\beta$ will be piecewise
-linear. \citep{kim_$ell_1$_2009} proposed a specialized primal-dual
+linear. \citep{KimKoh2009} proposed a specialized primal-dual
 interior point (PDIP) algorithm for solving \eqref{eq:l1tf}. From a
@@
 
-\attn{explain the objective more. give the AR(1) example. Explain what
-  you loses by this assumption (ACF,forecasting). Also explain that
-  the covariace matrix is diagonal so it cannot capture the covariance
-  structure. But incontrast to spatial stat literature, it does not
-  make any assumption on estimated variances. Compare to Hallac et al
-  and Lingren et al.}
-
 As with~\eqref{eq:l1tf}, one can solve~\eqref{eq:l1tf_var} using the
 PDIP algorithm (as in, e.g.,
-\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}).
-In each iteration of PDIP we need to compute a \textit{search direction} by taking a Newton step on a system of nonlinear equations.
-Appendix B provides details on how this computation can be performed for the objective function \ref{eq:l1tf_var}.
-
-\attn{Arash: I think we should remove Algorithm 1.
-There are other steps in PDIP which involve adapting the algorithm parameter $w$ and doing Backtracking line search.
-Also there is only one loop in PDIP.}
+\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}). In each iteration of
+PDIP we need to compute a search direction by taking a Newton
+step on a system of nonlinear equations. Due to space limitations, we
+defer details to Appendix B in the Supplement.
 
@@
     \FOR[Newton updates]{$j=1,\ldots,J$}
-    \STATE Solve $A [\Delta\nu\; \Delta\mu_1\; \Delta\mu_2]^\top=r_{w_k}$ to find the search direction.
+    \STATE Solve $A [\Delta\nu\; \Delta\mu_1\;
+    \Delta\mu_2]^\top=r_{w_k}$ to find the search direction. 
     \STATE $\quad\quad A$ is given in \ref{eq:delta_r}
@@
 penalty is a special case of trend
-filtering on graphs~\citep{wang_trend_2016} which penalizes the difference
+filtering on graphs~\citep{WangSharpnack2016} which penalizes the difference
 between the estimated values of the signal on the connected
@@
 \eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times
-S$ vector and $D$ is replaced by $D_{TS} \in \mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and spatial constraints, respectively\footnote{$N_s$ is obtained by counting the number of unique constraints at each location and at all times.}.
-Then, as we
+S$ vector and $D$ is replaced by $D_{TS} \in \mathbb{R}^{(N_t+N_s)
+  \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot
+(2n_rn_c-n_r)$ are the number of temporal and spatial constraints,
+respectively\footnote{$N_s$ is obtained by counting the number of
+  unique constraints at each location and at all times.}. Then, as we 
 have two different tuning parameters for the temporal and spatial
 components, we write $\Lambda =\left[\lambda_1\one_{N_t}^\top,\;
-  \lambda_2\one_{N_s}^\top\right]^\top$ leading to \footnote{Throughout the paper, we use $|x|$ for both scalars and vectors.
-For vectors we use this to denote a vector obtained by taking the absolute value of each entry of $x$.}:
+  \lambda_2\one_{N_s}^\top\right]^\top$ leading
+to:\footnote{Throughout the paper, we use $|x|$ for both scalars and
+  vectors. For vectors we use this to denote a vector obtained by
+  taking the absolute value of each entry of $x$.} 
 % let $h$ be a
@@
 \begin{equation}
-\min_h -l(y\given h)+ \Lambda^\top | D_{TS}h |
+\min_h -l(y\given h)+ \Lambda^\top | D_{TS}h |.
 \label{eq:l1tf_var_st_mat}
 \end{equation}
+\attn{equations should have proper punctuation as if they were a sentence.}
 % where $\Lambda = \mathbf{I}_{Tn_rn_c}\left[\lambda_1\one_T^\top,\;
@@
 
-\attn{Add back the full form of $D$ above? It was too cumbersome
-  before, but possibly necessary}
+% \attn{Add back the full form of $D$ above? It was too cumbersome
+%   before, but possibly necessary}
 
@@
 
-For a spatial grid of size $S$ and $T$ time steps, $D_{ST}$ will have $3Tn_rn_c-2n_rn_c-Tn_r$ rows and $ST$ columns.
-For a $1^\circ\times 1^\circ$ grid over the entire northern hemisphere and daily data over 10 years, we have $n_r=90$, $n_c=180$, $T=3650$ and so $D_{ST}$ has approximately $10^8$ columns and $10^8$ rows.
-In each step of the PDIP algorithm, we need to solve a linear system of equations in $A$ which depends on $D_{ST}^\top D_{ST}$ (see appendix A and B).
-Therefore, applying the PDIP directly is infeasible for our data.\footnote{We note that this is a highly structured and sparse matrix, but, unlike trend filtering alone, it is not banded.
-We are unaware of general linear algebra techniques for inverting such matrix, despite our best efforts.}
+For a spatial grid of size $S$ and $T$ time steps, $D_{ST}$ will have
+$3Tn_rn_c-2n_rn_c-Tn_r$ rows and $ST$ columns. For a $1^\circ\times
+1^\circ$ grid over the entire northern hemisphere and daily data over
+10 years, we have $n_r=90$, $n_c=180$, $T=3650$ and so $D_{ST}$ has
+approximately $10^8$ columns and $10^8$ rows. In each step of the PDIP
+algorithm, we need to solve a linear system of equations in $A$ which
+depends on $D_{ST}^\top D_{ST}$ (see appendix A and B). Therefore,
+applying the PDIP directly is infeasible for our data.\footnote{We
+  note that this is a highly structured and sparse matrix, but, unlike
+  trend filtering alone, it is not banded. We are unaware of general
+  linear algebra techniques for inverting such matrix, despite our
+  best efforts.}
 
@@
 second uses proximal methods to avoid matrix inversions.
-
-\attn{We note that stochastic gradient descent could be used but...}
-
-\attn{Arash: The primal objective is not differentiable and so we cannot use SGD.
-The dual has constriants and so the plane SGD cannot be used.
-We can use projected SGD but I think it would converge slowly.
-Since other papers on LASSO have not considered SGD I don't think we need to justify why we did not use it.}
 
@@
 
-Consider an optimization problem of the form $\min_h f(h)$, where $h\in\mathbb{R}^n$ is the \textit{global variable} and $f(h):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is convex.
-The goal in consensus optimization is to break this problem into several smaller sub-problems which can be solved independently in each iteration of optimization.
-
-Assume that it is possible to define a set of \textit{local variables} $x_i \in \mathbb{R}^{n_i}$ such that $f(h)=\sum_i f_i(x_i)$, where each $x_i$ is a subset of the global variable $h$.
-More specifically, each entry of the local variables corresponds to an entry of the global variable.
-Therefore we can define a mapping $\mathscr{G}(i,j)$ from the local variables indices into the global variable indices: $k=\mathscr{G}(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is $h_k$ (or $(x_i)_j=h_k$).
-For ease of notation, define $\tilde{h}_i \in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{\mathscr{G}(i,j)}$.
-Then, the original optimization problem is equivalent to the following problem:
-
+Given an optimization problem of the form $\min_h f(h)$, where
+$h\in\mathbb{R}^n$ is the global variable and
+$f(h):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is
+convex. Consensus optimization breaks this problem
+into several smaller sub-problems that can be solved independently in
+each iteration of optimization.
+
+Assume it is possible to define a set of local variables
+$x_i \in \mathbb{R}^{n_i}$ such that $f(h)=\sum_i f_i(x_i)$, where
+each $x_i$ is a subset of the global variable $h$. More specifically,
+each entry of the local variables corresponds to an entry of the
+global variable. Therefore we can define a mapping $\mathscr{G}(i,j)$
+from the local variable indices into the global variable indices:
+$k=\mathscr{G}(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is
+$h_k$ (or $(x_i)_j=h_k$). For ease of notation, define $\tilde{h}_i
+\in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{\mathscr{G}(i,j)}$. Then,
+the original optimization problem is equivalent to the following
+problem:  
 \begin{equation}
@@
 \min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)\\
- s.t. \,\,\, \tilde{h}_i=x_i \,\,\,\,\,\,
+ s.t. \,\,\, \tilde{h}_i=x_i. \,\,\,\,\,\,
 \end{aligned}
@@
 \end{equation}
-
-It is important to note that each entry of the global variable may correspond to several entries of the local variables and so the constraints $\tilde{h}_i=x_i$ enforce the consensus between the local variables corresponding to the same global variable.
-
-The \textit{augmented Lagrangian} corresponding to the problem \ref{eq:consADMM} is $L_\rho(x,h,y)=\sum_i \big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert x_i-\tilde{h}_i \lVert_2^2 \big)$.
-Now, we can apply ADMM to $L_\rho$ which results in the following ADMM updating steps at each iteration $m$:
-
-\begin{equation}
-\begin{aligned}
-& x_i^{m+1}:=\argmin_{x_i} \bigg( f_i(x_i) + (u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{h}_i^m \lVert_2^2\bigg)\\
-& h_k^{m+1}:=(1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i^{m+1})_j\\
-& u_i^{m+1}:=u_i^m + \rho (x_i^{m+1}-\tilde{h}_i^{m+1})
-\end{aligned}
-\label{eq:ADMM_steps}
-\end{equation}
-
+It is important to note that each entry of the global variable may
+correspond to several entries of the local variables and so the
+constraints $\tilde{h}_i=x_i$ enforce the consensus between the local
+variables corresponding to the same global variable.
+
+The augmented Lagrangian corresponding to 
+\eqref{eq:consADMM} is $L_\rho(x,h,y)=\sum_i
+\big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert
+x_i-\tilde{h}_i \lVert_2^2 \big)$. Now, we can apply ADMM to $L_\rho$
+which results in the following ADMM updates:
+\begin{align}
+x_i&\leftarrow\argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
+     (\rho/2)  \norm{ x_i-\tilde{h}_i }_2^2\\
+h_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j \label{eq:ADMM_steps}\\
+u_i&\leftarrow u_i + \rho (x_i-\tilde{h}_i).
+\end{align}
+\attn{If you hate the left arrow, that's fine. But change it
+  everywhere, and use $=$. The $:=$ notation means that ``thing on
+  left is defined to be thing on right'' and isn't really appropriate here.}
 Here, $S_k$ is the number of local variable entries that correspond to
@@
 
-\begin{comment}
-The $x$-update is
-often a simple modification as the original optimization, though over only a small
-partition, the $h$-update simply averages over collections of
-analogous local variables while the $u$ update enforces sharing. Both
-the $x$ and $u$ updates are parallelizable over the partition of local
-variables.
-\end{comment}
-
-\attn{I don't much like this explanation or notation. Seems unclear}
-
-\attn{Arash: I re-wrote this section.}
-
-To solve the optimization problem \eqref{eq:l1tf_var_st_mat} or
-\eqref{eq:l1tf_var_st} using this method, we need to address two
+To solve the optimization problem \eqref{eq:l1tf_var_st_mat} using
+this method, we need to address two 
 questions: first, how to choose the 
-local variables $x_i$, and second, how to solve the optimization
-problem for updating these variables (the first line of
-\eqref{eq:ADMM_steps} which we will refer to it as \textit{x-update step}).
-
-In \autoref{fig:data_cube}, the global variable $h$ is represented
-as a cube (using the subset of the US as an example). We decompose $h$
-into sub-cubes as shown in the figure by white lines. It is easy to see
-that by this definition of $x_i$, the objective
-\eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where
-$f_i(x_i)=-l(y_i\given x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|$, and
-$\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
-penalties corresponding to $x_i$ only. The x-update step in \autoref{eq:ADMM_steps} is the following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
-(u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
-
-\attn{I would like to be able to write all this in terms of $h$ instead of $x$}
-
-\attn{Arash: now I'm using $h$ as the global variables. We cannot write the local variables in terms of $h$ since they are different from $h$.}
+local variables $x_i$, and second, how to the update in the first line of
+\eqref{eq:ADMM_steps}.
 
@@
     time. The sub-cubes specified by the white lines are
-      $x_i$. \attn{Can we use $h_i$ here instead of $x_i$?}}
+    $x_i$.}
   \label{fig:data_cube}
@@
 
-By this definition of $x_i$, the update step for $x_i$ is the
-following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
-(u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
-We solve this using the PDIP method.
-As it was explained in \autoref{sec:l1tf_var}, to use PDIP we need to compute the dual problem, which in turn needs the computation of the Fenchel conjugate of the loss function.
-In addition, referring to \autoref{alg:pdip}, we need to compute the gradient and Jacobian of the conjugate functions.
-It is important to note that compare to the optimization problem \autoref{eq:l1tf_var}, the loss function in the x-update step in \autoref{eq:ADMM_steps} includes the quadratic term $(\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2$.
-This makes the computation more involved than \autoref{sec:l1tf_var}.
-The details of the computations are explained in \autoref{sec:app_consADMM}.\attn{change this to appendix}
-
-\attn{I'm not sure if we want to devote some space to write this as a separate algorithm.
-The steps are those mentioned in \ref{eq:ADMM_steps} and the x-update is now explained in the appendix.
-So we may want to remove algorithm \ref{alg:ADMM}.
-}
-
-The complete ADMM algorithm for estimating the variances is
-represented in \autoref{alg:ADMM}. All the computations in the three
-updating steps \eqref{eq:ADMM_steps} can be performed in parallel. The
-number of rows and columns of the sub-cubes should be chosen so that
-the updating of $x_i$ could be performed in one processor. We choose
-$3\times3\times521$ sub-cubes.
-
-\begin{algorithm}[tb]
-  \caption{ADMM for sparse estimation of variance of spatio-temporal
-    data \attn{Fix this}}
-  \label{alg:ADMM}
-  \begin{algorithmic}
-    \STATE {\bfseries Input:} data $y$, mapping $\mathscr{G}(i,j)$,
-    $\rho,\lambda_t,\lambda_s$ 
-    \STATE Initialization: $x_i^0=z^0=u_i^0=\textbf{0}$.
-    \FOR{$m=1,2,...$ }
-    \FOR{$i=1$ {\bfseries to} $N_{sub-cubes}$}
-    \STATE compute $\nu_i$ from \eqref{eq:dual_x_update}
-    \STATE compute $w_i$ from \eqref{eq:conj_x_update}
-    \STATE set $x_i^m:=w_i$
-    \ENDFOR
-    \STATE Compute $z^m$ from  \eqref{eq:ADMM_steps}
-    \STATE Compute $u_i^m$ from  \eqref{eq:ADMM_steps}
-    \ENDFOR
-  \end{algorithmic}
-\end{algorithm}
-
-Because \autoref{alg:ADMM} breaks the large optimization into
+In \autoref{fig:data_cube}, the global variable $h$ is represented
+as a cube (using the subset of the US as an example). We decompose $h$
+into sub-cubes as shown by white lines. With
+this definition of $x_i$, the objective
+\eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where
+$f_i(x_i)=-l(y_i\given x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|$, and
+$\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
+penalties corresponding to $x_i$ only. 
+% The $x$-update step in
+% \eqref{eq:ADMM_steps} is the following optimization problem:
+% $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) + 
+% (u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
+Thus, with this choice of the local variables $x_i$, we solve the
+$x$-update using the PDIP method. 
+% As it 
+% was explained in \autoref{sec:l1tf_var}, to use PDIP we need to
+% compute the dual problem, which in turn needs the computation of the
+% Fenchel conjugate of the loss function. In addition, referring to
+% \autoref{alg:pdip}, we need to compute the gradient and Jacobian of
+% the conjugate functions. It is important to note that compare to the
+% optimization problem \autoref{eq:l1tf_var}, the loss function in the
+% x-update step in \autoref{eq:ADMM_steps} includes the quadratic term
+% $(\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2$. This makes the
+% computation more involved than \autoref{sec:l1tf_var}. 
+The details of the remaining computations
+the computations are explained in the Supplement.
+
+% The complete ADMM algorithm for estimating the variances is
+% represented in \autoref{alg:ADMM}. All the computations in the three
+% updating steps \eqref{eq:ADMM_steps} can be performed in parallel. The
+% number of rows and columns of the sub-cubes should be chosen so that
+% the updating of $x_i$ could be performed in one processor. We choose
+% $3\times3\times521$ sub-cubes.
+
+% \begin{algorithm}[tb]
+%   \caption{ADMM for sparse estimation of variance of spatio-temporal
+%     data \attn{Fix this}}
+%   \label{alg:ADMM}
+%   \begin{algorithmic}
+%     \STATE {\bfseries Input:} data $y$, mapping $\mathscr{G}(i,j)$,
+%     $\rho,\lambda_t,\lambda_s$ 
+%     \STATE Initialization: $x_i^0=z^0=u_i^0=\textbf{0}$.
+%     \FOR{$m=1,2,...$ }
+%     \FOR{$i=1$ {\bfseries to} $N_{sub-cubes}$}
+%     \STATE compute $\nu_i$ from \eqref{eq:dual_x_update}
+%     \STATE compute $w_i$ from \eqref{eq:conj_x_update}
+%     \STATE set $x_i^m:=w_i$
+%     \ENDFOR
+%     \STATE Compute $z^m$ from  \eqref{eq:ADMM_steps}
+%     \STATE Compute $u_i^m$ from  \eqref{eq:ADMM_steps}
+%     \ENDFOR
+%   \end{algorithmic}
+% \end{algorithm}
+
+Because consensus ADMM breaks the large optimization into
 sub-problems that can be solved independently, it is amenable to a
@@
 hundreds iterations. 
-\attn{Need to redo this:}Solving each sub-problem on a machine with four
-3.20GHz Intel i5-3470 cores takes less than 3 seconds on average, and
-so for example if we assume that communication time is 10 seconds and
-the algorithm converges in 300 iterations, with parallelization on
-$N_{sub-cubes}$ machines, the algorithm will converge in about 1
-hour. Assuming that we use $N_{sub-cubes}$ machines and that the
-convergence rate of the algorithm is independent of the grid size,
-this time will be independent of the grid size.
-
-If we perform these computations on a single machine, the computation
-time grows linearly with $N_{sub-cubes}$. For example, for the data in
-a grid over the united states and using $3\times3\times521$ sub-cubes
-each iteration of the algorithm will take about 20 minutes on a single
-machine and so with 300 iterations it will take several days to
-converge. Given that we need to compute the solution for several
-values of the parameters $\lambda_t$ and $\lambda_s$, this computation
-time is not feasible.
-
-Therefore, this algorithm is only useful if we can parallelize the
+% \attn{Need to redo this:}Solving each sub-problem on a machine with four
+% 3.20GHz Intel i5-3470 cores takes less than 3 seconds on average, and
+% so for example if we assume that communication time is 10 seconds and
+% the algorithm converges in 300 iterations, with parallelization on
+% $N_{sub-cubes}$ machines, the algorithm will converge in about 1
+% hour. Assuming that we use $N_{sub-cubes}$ machines and that the
+% convergence rate of the algorithm is independent of the grid size,
+% this time will be independent of the grid size. 
+% If we perform these computations on a single machine, the computation
+% time grows linearly with $N_{sub-cubes}$. For example, for the data in
+% a grid over the united states and using $3\times3\times521$ sub-cubes
+% each iteration of the algorithm will take about 20 minutes on a single
+% machine and so with 300 iterations it will take several days to
+% converge. Given that we need to compute the solution for several
+% values of the parameters $\lambda_t$ and $\lambda_s$, this computation
+% time is not feasible. 
+However, this algorithm is only useful if we can parallelize the
 computation over several machines. In the next section, we describe
@@
 
-\attn{Need a good dummy variable here, it can't be $u$.} Consider the generic optimization problem
+Consider the generic optimization problem
 $\min_x f(x)+g(Dx)$
@@
 has the form
-
 \begin{align}
@@
 \end{align}
-
-\noindent where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu < \rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
-
+where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu < \rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
 \begin{equation}
@@
 
-\attn{What are $\mu$ and $\rho$?}
-
-\attn{Arash: I explained it above. They are algorithm parameters.}
-
-Clearly, \eqref{eq:l1tf_var_st} has this form necessary for using this algorithm.
+%Clearly, \eqref{eq:l1tf_var_st} has this form necessary for using this algorithm.
 % The optimization problem 
@@
 % (see \autoref{sec:exten}). 
-To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate
-$\prox_{\mu f}$ and $\prox_{\rho g}$. Proximal
-algorithms are feasible only if these proximal operators can be
+% To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate
+% $\prox_{\mu f}$ and $\prox_{\rho g}$. 
+Proximal
+algorithms are feasible when these proximal operators can be
 evaluated efficiently which, as we show next, is the case for our
@@
 
-\begin{theorem}
+\begin{lemma}
   Let $f(h) = \sum_{k} h_{k} + y_{k}^2e^{-h_{k}}$ and $g(x) =
@@
 $(v)_+=v\vee 0$.
-\end{theorem}
+\end{lemma}
 \begin{proof}
@@
 
+Therefore, \autoref{alg:linADMM} gives a different method for solving
+the same problem.
+
 \begin{algorithm}[tb]
@@
     \STATE {\bfseries Input:} data $y$, penalty matrix $D$, 
-    $\rho,\lambda_t,\lambda_s >0$.
-    \STATE {\bf Set:} $h\leftarrow 0$, $z\leftarrow 0$, $u\leftarrow 0$. 
-    \FOR{$m=1,2,...$ }
+    $\epsilon, \rho,\lambda_t,\lambda_s >0$.
+    \STATE {\bf Set:} $h\leftarrow 0$, $z\leftarrow 0$, $u\leftarrow
+    0$. \COMMENT{Initialization} 
+    \REPEAT
     \STATE $h_k\leftarrow \mathscr{W}\bigg(\frac{y_k^2}{\mu} 
-    \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu u_k}{\mu}$,
-    \STATE $z\leftarrow S_{\rho\lambda}(u)$.
-    \STATE $u\leftarrow u + Dh-z$
-    \ENDFOR
+    \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu
+      u_k}{\mu}\quad k=1,\ldots TS$. \COMMENT{Primal update}
+    \STATE $z\leftarrow S_{\rho\lambda}(u)$. \COMMENT{Elementwise soft thresholding}
+    \STATE $u\leftarrow u + Dh-z$. \COMMENT{Dual update}
+    \UNTIL {$\max\{\norm{h-z}_2^2,\; \norm{z^{m+1}-z^m}_2^2\} < epsilon$}
+    \STATE {\bf Return:} $z$.
   \end{algorithmic}
@@
 \label{sec:simulations}
-
-\attn{Can we add some type of performance measure? What if we fit 2
-  marginal models, spatial only and temporal only. Plus maybe a
-  marginal GARCH?}
 
@@
 
-In words, the variance at each time and location is computed as the weighted sum of $S$ bell-shaped functions where the weights are time-varying, consist of a linear trend $\alpha_s \cdot t$ and a periodic term $\beta_s \cdot \sin(2\pi\omega_s t+\phi_s)$.
-The bell-shaped functions impose the spatial smoothness, and the linear trend and the periodic terms enforce the temporal smoothness similar to the seasonal component in the real climate data.
-We simulated the data on a 5 by 7 grid and for 780 time steps with $S=4$.
-The parameters of the variance function are shown in \autoref{tab:sim_params} in Appendix C.
-For reference, we plot the variance function for all locations at $t=25$ and $t=45$ in as well as the variance across time at $(0,0)$ in \autoref{fig:true_var_spatial} in Appendix C.
+In words, the variance at each time and location is computed as the
+weighted sum of $S$ bell-shaped functions where the weights are
+time-varying, consist of a linear trend $\alpha_s \cdot t$ and a
+periodic term $\beta_s \cdot \sin(2\pi\omega_s t+\phi_s)$. The
+bell-shaped functions impose the spatial smoothness, and the linear
+trend and the periodic terms enforce the temporal smoothness similar
+to the seasonal component in the real climate data. We simulated the
+data on a 5 by 7 grid and for 780 time steps with $S=4$. 
+\attn{The
+parameters of the variance function are shown in
+\autoref{tab:sim_params} in the Supplement. For reference, we plot the
+variance function for all locations at $t=25$ and $t=45$ in as well as
+the variance across time at $(0,0)$ in \autoref{fig:true_var_spatial}
+in Appendix C.}
 
@@
 estimated variance and the true variance at all locations and all time
-steps. For $\lambda_t=5$ and $\lambda_s=0.1$ MAE was minimized. The
+steps. For $\lambda_t=5$ and $\lambda_s=0.1$, the MAE was minimized. The
 left panel of \autoref{fig:true_fitted_var} shows the true and the
@@
 
-The middle panel of \autoref{fig:true_fitted_var} shows the convergence of Algorithms~\ref{alg:ADMM} and~\ref{alg:linADMM}.
-Each iteration of the linearized algorithm takes 0.01 seconds on average while each iteration of the consensus ADMM takes about 20 seconds.
-
-To further examine the performance of the proposed model, we next compare it to three models: a model which does not consider the spatial smoothness and so is equivalent to fitting the model in \autoref{sec:l1tf_var} to each time-series separately, a model which does not consider the temporal smoothness, and a GARCH(1,1) model.
-We simulated 100 datasets using the method explained above with $\sigma_s \sim \mathsf{unif}(4,7)$.
-The right panel of \autoref{fig:true_fitted_var} shows the boxplot of MAE for these models.
-Interestingly, the proposed model with optimal parameters outperforms GARCH(1,1) in estimating the true value of the variance.
-
-\attn{These data are pretty small. Is it possible to do the PDIP?}
-
-\attn{Arash: we can, but why? it will be very slow.}
+The middle panel of \autoref{fig:true_fitted_var} shows the
+convergence of both methods. Each iteration of the linearized
+algorithm takes 0.01 seconds on average while each iteration of the
+consensus ADMM takes about 20 seconds.
+
+To further examine the performance of the proposed model, we next
+compare it to three alternatives: a model which does not consider the
+spatial smoothness (equivalent to fitting the model in
+\autoref{sec:l1tf_var} to each time-series separately), a model which
+does not consider imposes only spatial smoothness, and a GARCH(1,1) model. We
+simulated 100 datasets using the method explained above with $\sigma_s
+\sim \mathsf{uniform}(4,7)$. The right panel of
+\autoref{fig:true_fitted_var} shows the boxplot of the MAE for these
+models. Interestingly, the proposed model with optimal parameters
+outperforms GARCH(1,1) in estimating the true value of the variance.
 
 \subsection{Data analysis}
-
-\autoref{alg:ADMM} is appropriate only if
+\label{sec:data}
+
+Consensus ADMM in~\autoref{sec:consOpt} is appropriate only if
 we parallelize it over multiple machines, and it is significantly
@@
 applied this algorithm to the northern hemisphere of the ERA-20C
-dataset available from the \href{European Center for Medium-Range
-  Weather Forecasts}{https://www.ecmwf.int}. The data are
+dataset available from the \href {https://www.ecmwf.int}{European Center for Medium-Range
+  Weather Forecasts}. The data are
 the 2 meter temperature measured daily at 12 p.m from January 1, 1960
@@
 
-???
-
-To remedy this, we add the following long horizon penalty
-to \eqref{eq:l1tf_var}:
-
-\begin{equation}
-\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
-\label{eq:lh_penalty}
-\end{equation}
-
-where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$.
-Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$).
-Since we are working on the weekly averaged data, each year corresponds to 52 observations.
-In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$.
-The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
-The annual average of the estimated SDs shows a linear trend with a positive slope.
-
-\attn{From here down needs to be cut down and corrected with the right numbers.}
+The Supplement explains some preprocessing and investigates some
+properties of the time-series of different locations on
+earth. \autoref{fig:bloom_estimatedSD} a processed time-series for a
+single location.
+The variance of this time-series has an irregular cyclic
+behavior. Additionally, the time-series of other locations show 
+different patterns. These observations motivated the need to develop a
+non-parametric framework for this
+problem. \autoref{fig:bloom_estimatedSD} also shows the
+estimated SD obtained using the method of
+\autoref{sec:l1tf_var}. 
+% Since the estimated SD captures the periodic
+% behavior of 
+% volatility, it is hard to distinguish any long-term trend. Therefore,
+% we compute the annual average of these 
+% estimates. However, as this figure shows, the annual trend is not
+% smooth. This is because in the optimization problem
+% \eqref{eq:l1tf_var}, the smoothness of the annual trend is not
+% encouraged. Therefore, in Appendix D we propose a simple method to
+% encourage the smoothness of the annual average trend. The estimated
+% SDs using this method is shown in the right panel of
+% \autoref{fig:bloom_estimatedSD}. The annual average of the estimated
+% SDs shows a linear trend with a positive slope.
 
@@
 
-We used the following rule to determine when to stop the optimization:
-the optimization was stopped if the value of the loss did not improve
-by at least 0.1\% in 1000 trials. As we can see, the algorithm
-converged in about 2000 iterations. This took about 11 minutes. Our
-experiments showed that the convergence speed depends on the value of
-$\lambda_t$ and $\lambda_s$. Also, if we use the solution obtained for
-smaller values of these parameters as the initial value for the larger
-values (\textit{warm start}), the converges speed improves.
+As shown in \autoref{alg:linADMM}, we evaluated convergence using
+$\epsilon=0.001\%$ of the MSE of the data.
+Our simulation experiments showed that the
+convergence speed depends on the value of $\lambda_t$ and
+$\lambda_s$. Furthermore, using the solution obtained for smaller values
+of these parameters as a warm start for the larger values, the
+converges speed improves.
 
@@
 criterion. However, such analyses needs the computation of the degrees
-of freedom (df). Several previous work have investigated the df in
+of freedom. Several previous work have investigated the df in
 generalized lasso problems
@@
 considered yet.
-
-Another approach is to choose the set of values which minimize an
-estimate of the expected prediction error obtained by k-fold
-cross-validation \citep{tibshirani_regression_1996} . Although this
-method is applicable for our problem, it needs k times more
-computation.
 
@@
 following sets of values: $\lambda_t \in \{0,2,4,8,10,15,200,1000\} \, \, ,
-\lambda_s \in \{0,.1,.5,2,5,10\}$. The best combination based on a held
-out set was $\lambda_t=4$ and $\lambda_s=2$. All the analyses in the
-next section are performed on the solution for these values.
+\lambda_s \in \{0,.1,.5,2,5,10\}$. The best combination was
+$\lambda_t=4$ and $\lambda_s=2$. All the analyses in the next section
+are performed on the solution for these values.
 
 \paragraph{Analysis of trend of temperature volatility}
-
-The top row of \autoref{fig:avg_change_estimatedSD} shows the
-detrended data, the estimated standard deviation and the yearly
-average of these estimates for two cities in the US: Bloomington (left) and San Diego (right).
-The estimated SD captures the periodic behavior in the variance of the time-series.
-In addition, the number of linear segments changes adaptively in each time window depending on how fast the variance is changing.
-
-The yearly average of the estimated SD captures the trend in the
-temperature volatility. For example, we can see that in Bloomington,
-there is a small positive trend. To determine how the volatility has
-changed in each location, we subtract the average of the estimated
-variance in 1992 from the average in the following years and compute
-their sum. The value of this change in the variance in each location
-is depicted in the right panel of \autoref{fig:avg_change_estimatedSD}.
-The left panel of this figure, shows the average estimated variance in each location.
-Since the optimal value of the spatial penalty is rather large ($\lambda_s=2$) the estimated variance is spatially very smooth.
-
-It is interesting to note that the trend in volatility is almost zero over the oceans. The most positive trend can be observed in Asia and particularly in south-east Asia.
-
-The bottom panel of \autoref{fig:avg_change_estimatedSD} shows the histogram of change in the estimated SD across the northern hemisphere.
-As we can see, the SD in most locations on the northern hemisphere had a negative trend in this time period.
 
@@
   \includegraphics[width=.9 \columnwidth]{Figures/ts_estimatedVar}\\
-  \includegraphics[width=.45\linewidth]{Figures/avg_logVar.pdf}
-  \includegraphics[width=.45\linewidth]{Figures/avg_change_logVar.pdf}\\
-    \includegraphics[width=.3\linewidth]{Figures/hist_avg_change.pdf}
-  \caption{Top row: Detrended data and the estimated SD for
-    Bloomington (left) and San Diego (right). Middle row: the
-    average of the estimated variance over the northern hemisphere (left) and the
-    change in the variance from 1961 to 2011 (right). Bottom: the histogram of change in estimated SD.} 
+  \caption{Detrended data and the estimated SD for
+    a small midwestern city (left) and San Diego (right).} 
   \label{fig:avg_change_estimatedSD}
 \end{figure}
+
+The top row of \autoref{fig:avg_change_estimatedSD} shows the
+detrended data, the estimated standard deviation and the yearly
+average of these estimates for two cities in the US: a small
+midwestern city (left) and San Diego (right). The estimated SD
+captures the periodic behavior in the variance of the time-series. In
+addition, the number of linear segments changes adaptively in each
+time window depending on how fast the variance is changing.  
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=.45\linewidth]{Figures/avg_logVar.pdf}
+  \includegraphics[width=.45\linewidth]{Figures/avg_change_logVar.pdf}
+  \caption{The
+    average of the estimated variance over the northern hemisphere (left) and the
+    change in the variance from 1961 to 2011 (right).}
+\label{fig:avg_chg}
+\end{figure}
+
+The yearly average of the estimated SD captures the trend in the
+temperature volatility. For example, we can see that the variance in
+the midwestern city displays a small positive trend. To determine how the volatility has
+changed in each location, we subtract the average of the estimated
+variance in \attn{1992} from the average in the following years and compute
+their sum. The value of this change in the variance in each location
+is depicted in the right panel of
+\autoref{fig:avg_chg}. The left panel of this
+shows the average estimated variance in each location. Since the
+optimal value of the spatial penalty is rather large ($\lambda_s=2$)
+the estimated variance is spatially very smooth.
+
+It is interesting to note that the trend in volatility is almost zero
+over the oceans. The most positive trend can be observed in Asia and
+particularly in south-east Asia.
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=.3\linewidth]{Figures/hist_avg_change.pdf}
+  \caption{The histogram of changes in estimated SD.}
+  \label{fig:hist}
+\end{figure}
+
+\autoref{fig:hist} shows the
+histogram of change in the estimated SD across the northern
+hemisphere. The SD in most locations on the northern
+hemisphere had a negative trend in this time period.
 
@@
 
-The consensus ADMM algorithm converges in few hundreds of iterations
+The consensus ADMM algorithm converges in a few hundreds of iterations
 but each iteration takes much longer than the linearized ADMM
 algorithm. The appealing feature of the consensus ADMM algorithm is
-that if it is parallelized on enough number of machines the
+that if it is parallelized on enough machines the
 computation time per iteration remains constant as the problem size
-increases. The linearized ADMM algorithm, on the other hand converges
-in few thousands of iterations but each iteration is performed in
+increases. The linearized ADMM algorithm on the other hand converges
+in a few thousand iterations but each iteration is performed in a
 split second. However, since the algorithm converges in many
@@
 
+\attn{We did not do this. Change the below.}
 We applied the linearized ADMM algorithm to the surface temperature
@@
 
-\section{Appendix A}
-\label{sec:app_l1tf_var}
-
-In this appendix we provide more details on how to solve the optimization problem \ref{eq:l1tf_var} using PDIP.
-The objective function is convex but not differentiable.
-Therefore, to be able to use PDIP we first need to derive the dual of this problem.
-We note that this is a generalized LASSO problem \citep{tibshirani_solution_2011}.
-The dual of a generalized LASSO with the objective $f(x)+\lambda \norm{ Dx }_1$ is:
-
-\begin{align}
-\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
-\end{align}
-
-where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that
-
-\begin{equation}
-f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
-\label{eq:conj}
-\end{equation}
-
-Each iteration of PDIP involves computing a search direction by taking a Newton step for the system of nonlinear equations $r_w(v,\mu_1,\mu_2)=0$, where $w>0$ is a parameter and
-
-\begin{equation}
-  r_w(v,\mu_1,\mu_2):=
-	\begin{bmatrix}
-	r_{dual}\\
-	r_{cent}	
-	\end{bmatrix}=
-  \begin{bmatrix}
-    \nabla f^*(-D^\top v) + \mu_1 - \mu_2\\
-    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
-  \end{bmatrix}
-\label{eq:resid}
-\end{equation}
-
-for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form
-
-\begin{equation}
-r_w(v,\mu_1,\mu_2)+A
-\begin{bmatrix}
-	\nabla v\\
-	\nabla \mu_1\\
-	\nabla \mu_2	
-	\end{bmatrix}= 0
-\label{eq:newton_step}
-\end{equation}
-
-We have:
-
-\begin{equation}
-A=
-\begin{bmatrix}
-	\nabla^2 f^*(-D^\top v) & I & -I\\
-	-\mathbf{diag(\mu_1)}\one & -v+\lambda\one & \mathbf{0}\\
-	\mathbf{diag(\mu_2)}\one & v+\lambda\one & \mathbf{0}
-	\end{bmatrix}
-\label{eq:delta_r}
-\end{equation}
-
-Therefore, to perform the Newton step we need to compute $\nabla f^*(-D^\top v)$ and $\nabla^2 f^*(-D^\top v)$. It is straightforward to show that
-
-\begin{align}
-\nabla f^*(-D^\top v) = -\nabla_u f^*(u) D^\top ,\quad u=-D^\top v, \quad (\nabla_u f^*(u))_j=\log\bigg(\frac{y_j^2}{1-u_j}\bigg) \\
-\nabla^2 f^*(-D^\top v)=D\nabla_u^2 f^*(u)D^\top, \quad (\nabla_u^2 f^*(u))_j=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg)
-\end{align}
-
-\section{Appendix B}
-\label{sec:app_consADMM}
-\attn{put this in nips appendix format}
-
-In this Appendix we give more details on performing the x-update step in \autoref{eq:ADMM_steps}. We need to solve the following optimization problem:
-
-\begin{equation}
-\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
-\label{eq:x_update_opt}
-\end{equation}
-
-\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$.
-Let $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$.
-As it was explained in \autoref{sec:l1tf_var}, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$.
-So to use PDIP we first need to compute the conjugate function $f^*(\cdot)$.
-We have:
-
-\begin{align}
-f^*(\xi) & = \max_x \quad \xi^\top x - f(x)\\
-& =  \max_x \quad \sum_{j=1}^{n_b} (\xi_jx_j - x_j - y_j^2e^{-x_j} - (\rho/2)(x_j-\tilde{z}_j+u_j))
-\label{eq:conjugate}
-\end{align}
-
-Setting the derivative of the terms inside the summation to 0, we obtain:
-
-\begin{equation}
-\xi_j-y_j^2e^{-x_j^*}-\rho x_j^* + \rho (\tilde{z}_j-u_j)=0
-\label{eq:x_start}
-\end{equation}
-
-\noindent where $x^*$ is the maximizer in \ref{eq:conjugate}. Then, it can be shown that $x_j^*$ which satisfies \eqref{eq:x_start} can be obtained as follows:
-
-\begin{align}
-x^*_j & = \mathscr{W}\bigg(\frac{y_j^2}{\rho} e^{\phi_j} \bigg) - \phi_j \\
-\phi_j & =\frac{1-\xi_j-\rho(\tilde{z}_j-u_j)}{\rho}
-\end{align}
-
-In this equation, $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996}.
-Finally, the conjugate function is: $f^*(\xi) = \sum_{j=1}^{n_b} (\xi_jx^*_j - x^*_j - y_j^2e^{-x^*_j} - (\rho/2)(x^*_j-\tilde{z}_j+u_j))$.
-
-To use PDIP, we also need to evaluate $\nabla f^*$ and $\nabla^2 f^*$.
-First note that $\frac{\partial \mathscr{W}(q)}{\partial q} = \frac{\mathscr{W}(q)}{q(1+\mathscr{W}(q))}$ and $\frac{\partial^2 \mathscr{W}(q)}{\partial q^2} = - \frac{\mathscr{W}^2(q)(\mathscr{W}(q)+q)}{q^2(1+\mathscr{W}(q))^3}$.
-Using the chain rule we get:
-
-%\x^*_j + \xi_j + \frac{\partial \x^*_j}{\partial \xi_j}
-
-\begin{equation}
-\frac{\partial f^*(\xi)}{\partial \xi_j}  =  x^*_j  + \frac{\partial x^*_j}{\partial \xi_j} \bigg[ \xi_j -1 + y_j^2 e^{-x_j^*} + \rho (\tilde{z}_j - u_j - x_j^*) \bigg]
-\label{eq:d_f*_start}
-\end{equation}
-
-\noindent where we have:
-
-\begin{equation}
-\frac{\partial x_j^*}{\partial \xi_j} & = \frac{1}{\rho(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))}
-\label{eq:d_x*_start}
-\end{equation}
-
-By some tedious but straightforward computation we can obtain the second derivatives:
-
-\begin{align}
-\frac{\partial^2 f^*(\xi)}{\partial \xi_j^2} & =  \frac{\partial x_j^*}{\partial \xi_j} - \rho \frac{\partial^2 x_j^*}{\partial \xi_j^2} \bigg[ \phi_j +x_j^* - \tilde{z}_j + u_j \bigg]\\
-& \quad + \frac{\partial x_j^*}{\partial \xi_j} \bigg[ 1-y_j^2 \frac{\partial x_j^*}{\partial \xi_j} e^{-x_j^*} -\rho \frac{\partial x_j^*}{\partial \xi_j} \bigg]
-\label{eq:d2_f*_start}
-\end{align}
-
-\begin{equation}
-\frac{\partial^2 x_j^*}{\partial \xi_j^2}  = \frac{\mathscr{W}((y_j^2/\rho) e^{-\phi_j} )}{\rho^2(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))^3}
-\label{eq:d2_x*_start}
-\end{equation}
-
-Having computed the conjugate function and its gradient and Jacobian, now we can use a number of convex optimization software packages which have an implementation of PDIP to perform the x-update step inside the ADMM loop.
-We chose the python API of the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.
-
-\section{Appendix C}
-Table \autoref{tab:sim_params} lists the parameters used for simulating data in \autoref{sec:simulations}.
-\autoref{fig:true_var_spatial} shows the variance function obtained from there parameters at $t=25$ (left) and $t=45$ (center).
-
-\begin{table}[tb]
-  \caption{Parameters used to simulate data. \attn{Any ideas to take
-      up less space with this info?}}
-  \label{tab:sim_params}
-  \begin{center}
-    \begin{tabular}{ccccccc}
-      \hline
-      $s$ & $r_s$ & $c_s$ & $\sigma_s$ &$\alpha_s$ & $\omega_s$ & $\phi_s$\\
-      \hline
-      1 & 0 & 0 & 5 & 0.5 & 0.121 & 0 \\
-      2 & 0 & 5 & 5 & 0.1 & 0.121 & 0 \\
-      3 & 3 & 0 & 5 & -0.5 & 0.121 & $\pi/2$ \\
-      4 & 3 & 5 & 5 & -0.1 & 0.121 & $\pi/2$ \\
-      \hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-\begin{figure}[tb]
-  \centering	
-  \includegraphics[height=.15\textheight]{Figures/true_var_spatial}
-  \includegraphics[height=.15\textheight]{Figures/true_fitted_var}
-  \caption{Variance function at $t=25$ (left) and $t=45$
-    (center). Right: the true (orange) and estimated standard deviation 
-    function at the location (0,0). The estimated values are
-    obtained using linearized ADMM with $\lambda_s=0.1$ and two
-    values of $\lambda_t$: $\lambda_t=5$ (blue) and
-    $\lambda_t=100$ (green).} \label{fig:true_var_spatial}
-\end{figure}
-
-\section{Appendix D}
-
-In this appendix we examine some of the properties of the time-series of the temperature in ERA-20C dataset.
-The goal here is to demonstrate some of the difficulties in modeling the trend in the temperature volatility and motivate our methodology.
-
-\autoref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil).
-The time-series of Bloomington and San Diego show clear cyclic behavior.
-However, while it seems (qualitatively) that these cycles can be modeled by a sinusoidal function for Bloomington, the same is not true for San Diego.
-Also, the amplitude of the cycles changes from some years to others.
-The time-series of Manaus does not show any regular cyclic behavior.
-This demonstrates the first difficulty in analyzing the variance of this data: to analyze the variance, we first need to remove the cyclic terms from all time-series.
-However, there is a lot of variations in the cyclic behavior of the time-series of different locations.
-In addition, some of these cycles cannot be easily modeled by a parametric function \footnote{One might try to model the cycles by the summation of sinusoidal terms with different frequencies.
-However, for some time-series this may need many terms to be included in the summation to achieve a reasonable level of accuracy.
-In addition, this model cannot capture the non-stationarity in the cycles.}.
-To overcome these issues, we use a non-parametric approach to remove the cyclic terms from the time-series and de-trend them.
-This approach, called \textit{$\ell_1$-trend filtering} is explained in Section 2 of the text.
-We detrended each time-series separately using this method.
-For each time-series, we found the optimal value of the penalty parameter using \textit{k-fold cross-validation} with $k=5$.
-We used the R package \textbf{genlasso} to perform these computations \cite{arnold_efficient_2016}.
-
-\begin{figure}[tb]
-	\centering
-	\includegraphics[width=.2\textheight]{Figures/cities_ts}
- 	\caption{Time-series of the temperature (in Kelvin) of three cities.}
- 	\label{fig:cities_ts}
-\end{figure}
-
-The blue curve in the left panel of \autoref{fig:bloom_estimatedSD} shows the time-series of the temperature of Bloomington after detrending using this method.
-This figure, reveals another difficulty in estimating the trend of volatility in this data: the variance of this signal, shows cyclic behavior.
-Also, the cycles are not regular and their amplitude and frequency change.
-Even if one can describe the behavior of the variance of the time-series at all locations using a single parametric model (for example a variant of the GARCH models \citep{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework.
-These observations motivate the need to develop a non-parametric framework for the problem at hand.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}
-  \caption{Left: The residuals of the time-series of Bloomington
-    (averaged weekly) and the estimated SD obtained from the method of
-    \autoref{sec:l1tf_var} (red). Middle: the estimated SDs (red) and
-    their annual average (black) without imposing the long horizon
-    penalty. Right: the same as middle panel but here the long horizon
-    penalty is imposed. See the text for more details.} 
-  \label{fig:bloom_estimatedSD}
-\end{figure}
-
-The red curve in the left panel of \autoref{fig:bloom_estimatedSD}
-shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
-the time-series of Bloomington. To reduce the number of time-steps in
-this figure and in the remainder of the paper we work on the weekly
-averaged of the data.
-
-The curve of the estimated SD captures the periodic variations in the
-SD of the signal. Just by looking at this curve, it is hard to say if
-the SD is decreasing or increasing. Therefore, we compute the average
-of the estimated SD for each year. The estimated SD together with this
-annual average is shown in the middle panel of
-\autoref{fig:bloom_estimatedSD}. As it can be seen, the annual trend
-is not smooth. This is because in the optimization problem
-\eqref{eq:l1tf_var}, the smoothness of the annual trend is not
-encouraged.
-To remedy this, we proposed a method to encourage the smoothness of the annual average trend.
-The estimated SDs using this method is shown in the right panel of \autoref{fig:bloom_estimatedSD}.
-The annual average of the estimated SDs shows a linear trend with a positive slope.
-
 \bibliographystyle{abbrvnat}

### 2018-05-18 nips-2018/paper-supplement.tex
added supplement

\documentclass{article}
\PassOptionsToPackage{sort,numbers,square}{natbib}

\usepackage{nips_2018}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{hyperref}
\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}

%\newcommand{\theHalgorithm}{\arabic{algorithm}}

\usepackage{graphicx}
\usepackage{subfigure} 
%\graphicspath{{../nips-2018/}}

%Arash
\usepackage{amsmath,amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{siunitx}
\usepackage{comment}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

\renewcommand*{\figureautorefname}{}%
\renewcommand*{\tableautorefname}{Table}%
\renewcommand*{\sectionautorefname}{Section}%
\renewcommand*{\subsectionautorefname}{Section}%
\renewcommand*{\subsubsectionautorefname}{Section}%

\newtheorem{theorem}{Theorem}%[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}

\usepackage{gensymb}
\usepackage{xcolor}
\newcommand{\attn}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\one}{\mathbf{1}}
\renewcommand{\algorithmiccomment}[1]{\hfill $\rhd$ #1}
\newcommand{\given}{\;\vert\;}
\newcommand*{\algorithmautorefname}{Algorithm}%
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\DeclareMathOperator*{\prox}{\bf prox}
\DeclareMathOperator*{\sign}{sign}

\usepackage{xr}
\externaldocument{paper}

\title{Supplement for "Modeling trend in temperature volatility using generalized LASSO"}

\author{Arash Khodadadi\\
 Department of Statistics\\
 Indiana University\\
 Bloomington, IN 47408 \\
 \texttt{arakhoda@indiana.edu}
\And
  Daniel J.\ McDonald \\
 Department of Statistics\\
 Indiana University\\
 Bloomington, IN 47408 \\
 \texttt{dajmcdon@indiana.cmu.edu}}

\begin{document}

\maketitle

\section{Appendix A}
\label{sec:app_l1tf_var}

In this appendix we provide more details on how to solve the optimization problem \ref{eq:l1tf_var} using PDIP. The objective function is convex but not differentiable. Therefore, to be able to use PDIP we first need to derive the dual of this problem. We note that this is a generalized LASSO problem \citep{tibshirani_solution_2011}. The dual of a generalized LASSO with the objective $f(x)+\lambda \norm{ Dx }_1$ is:

\begin{align}
\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
\end{align}

where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that

\begin{equation}
f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
\label{eq:conj}
\end{equation}

Each iteration of PDIP involves computing a search direction by taking a Newton step for the system of nonlinear equations $r_w(v,\mu_1,\mu_2)=0$, where $w>0$ is a parameter and

\begin{equation}
  r_w(v,\mu_1,\mu_2):=
	\begin{bmatrix}
	r_{dual}\\
	r_{cent}	
	\end{bmatrix}=
  \begin{bmatrix}
    \nabla f^*(-D^\top v) + \mu_1 - \mu_2\\
    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
  \end{bmatrix}
\label{eq:resid}
\end{equation}

for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form

\begin{equation}
r_w(v,\mu_1,\mu_2)+A
\begin{bmatrix}
	\nabla v\\
	\nabla \mu_1\\
	\nabla \mu_2	
	\end{bmatrix}= 0
\label{eq:newton_step}
\end{equation}

We have:

\begin{equation}
A=
\begin{bmatrix}
	\nabla^2 f^*(-D^\top v) & I & -I\\
	-\mathbf{diag(\mu_1)}\one & -v+\lambda\one & \mathbf{0}\\
	\mathbf{diag(\mu_2)}\one & v+\lambda\one & \mathbf{0}
	\end{bmatrix}
\label{eq:delta_r}
\end{equation}

Therefore, to perform the Newton step we need to compute $\nabla f^*(-D^\top v)$ and $\nabla^2 f^*(-D^\top v)$. It is straightforward to show that

\begin{align}
\nabla f^*(-D^\top v) = -\nabla_u f^*(u) D^\top ,\quad u=-D^\top v, \quad (\nabla_u f^*(u))_j=\log\bigg(\frac{y_j^2}{1-u_j}\bigg) \\
\nabla^2 f^*(-D^\top v)=D\nabla_u^2 f^*(u)D^\top, \quad (\nabla_u^2 f^*(u))_j=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg)
\end{align}

\section{Appendix B}
\label{sec:app_consADMM}

In this Appendix we give more details on performing the x-update step in \autoref{eq:ADMM_steps}. We need to solve the following optimization problem:

\begin{equation}
\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
\label{eq:x_update_opt}
\end{equation}

\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$. Let $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$. As it was explained in \autoref{sec:l1tf_var}, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$. So to use PDIP we first need to compute the conjugate function $f^*(\cdot)$. We have:

\begin{align}
f^*(\xi) & = \max_x \quad \xi^\top x - f(x)\\
& =  \max_x \quad \sum_{j=1}^{n_b} (\xi_jx_j - x_j - y_j^2e^{-x_j} - (\rho/2)(x_j-\tilde{z}_j+u_j))
\label{eq:conjugate}
\end{align}

Setting the derivative of the terms inside the summation to 0, we obtain:

\begin{equation}
\xi_j-y_j^2e^{-x_j^*}-\rho x_j^* + \rho (\tilde{z}_j-u_j)=0
\label{eq:x_start}
\end{equation}

\noindent where $x^*$ is the maximizer in \ref{eq:conjugate}. Then, it can be shown that $x_j^*$ which satisfies \eqref{eq:x_start} can be obtained as follows:

\begin{align}
x^*_j & = \mathscr{W}\bigg(\frac{y_j^2}{\rho} e^{\phi_j} \bigg) - \phi_j \\
\phi_j & =\frac{1-\xi_j-\rho(\tilde{z}_j-u_j)}{\rho}
\end{align}

In this equation, $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996}. Finally, the conjugate function is: $f^*(\xi) = \sum_{j=1}^{n_b} (\xi_jx^*_j - x^*_j - y_j^2e^{-x^*_j} - (\rho/2)(x^*_j-\tilde{z}_j+u_j))$.

To use PDIP, we also need to evaluate $\nabla f^*$ and $\nabla^2 f^*$. First note that $\frac{\partial \mathscr{W}(q)}{\partial q} = \frac{\mathscr{W}(q)}{q(1+\mathscr{W}(q))}$ and $\frac{\partial^2 \mathscr{W}(q)}{\partial q^2} = - \frac{\mathscr{W}^2(q)(\mathscr{W}(q)+q)}{q^2(1+\mathscr{W}(q))^3}$. Using the chain rule we get:

%\x^*_j + \xi_j + \frac{\partial \x^*_j}{\partial \xi_j}

\begin{equation}
\frac{\partial f^*(\xi)}{\partial \xi_j}  =  x^*_j  + \frac{\partial x^*_j}{\partial \xi_j} \bigg[ \xi_j -1 + y_j^2 e^{-x_j^*} + \rho (\tilde{z}_j - u_j - x_j^*) \bigg]
\label{eq:d_f*_start}
\end{equation}
where we have:

\begin{equation}
\frac{\partial x_j^*}{\partial \xi_j}  = \frac{1}{\rho(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))}
\label{eq:d_x*_start}
\end{equation}

By some tedious but straightforward computation we can obtain the second derivatives:

\begin{align}
\frac{\partial^2 f^*(\xi)}{\partial \xi_j^2} & =  \frac{\partial x_j^*}{\partial \xi_j} - \rho \frac{\partial^2 x_j^*}{\partial \xi_j^2} \bigg[ \phi_j +x_j^* - \tilde{z}_j + u_j \bigg]\\
& \quad + \frac{\partial x_j^*}{\partial \xi_j} \bigg[ 1-y_j^2 \frac{\partial x_j^*}{\partial \xi_j} e^{-x_j^*} -\rho \frac{\partial x_j^*}{\partial \xi_j} \bigg]
\label{eq:d2_f*_start}
\end{align}

\begin{equation}
\frac{\partial^2 x_j^*}{\partial \xi_j^2}  = \frac{\mathscr{W}((y_j^2/\rho) e^{-\phi_j} )}{\rho^2(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))^3}
\label{eq:d2_x*_start}
\end{equation}

Having computed the conjugate function and its gradient and Jacobian, now we can use a number of convex optimization software packages which have an implementation of PDIP to perform the x-update step inside the ADMM loop. We chose the python API of the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.

\section{Appendix C}
\autoref{tab:sim_params} lists the parameters used for simulating data
in \autoref{sec:simulations}. \autoref{fig:true_var_spatial} shows the
variance function obtained from there parameters at $t=25$ (left) and
$t=45$ (center).

\begin{table}[tb]
  \caption{Parameters used to simulate data. \attn{Any ideas to take
      up less space with this info?}}
  \label{tab:sim_params}
  \begin{center}
    \begin{tabular}{ccccccc}
      \hline
      $s$ & $r_s$ & $c_s$ & $\sigma_s$ &$\alpha_s$ & $\omega_s$ & $\phi_s$\\
      \hline
      1 & 0 & 0 & 5 & 0.5 & 0.121 & 0 \\
      2 & 0 & 5 & 5 & 0.1 & 0.121 & 0 \\
      3 & 3 & 0 & 5 & -0.5 & 0.121 & $\pi/2$ \\
      4 & 3 & 5 & 5 & -0.1 & 0.121 & $\pi/2$ \\
      \hline
    \end{tabular}
  \end{center}
\end{table}

\begin{figure}[tb]
  \centering	
  \includegraphics[height=.15\textheight]{Figures/true_var_spatial}
  \includegraphics[height=.15\textheight]{Figures/true_fitted_var}
  \caption{Variance function at $t=25$ (left) and $t=45$
    (center). Right: the true (orange) and estimated standard deviation 
    function at the location (0,0). The estimated values are
    obtained using linearized ADMM with $\lambda_s=0.1$ and two
    values of $\lambda_t$: $\lambda_t=5$ (blue) and
    $\lambda_t=100$ (green).} \label{fig:true_var_spatial}
\end{figure}

\section{Appendix D}

In this appendix we examine some of the properties of the time-series of the temperature in ERA-20C dataset. The goal here is to demonstrate some of the difficulties in modeling the trend in the temperature volatility and motivate our methodology.

\autoref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil). The time-series of Bloomington and San Diego show clear cyclic behavior. However, while it seems (qualitatively) that these cycles can be modeled by a sinusoidal function for Bloomington, the same is not true for San Diego. Also, the amplitude of the cycles changes from some years to others. The time-series of Manaus does not show any regular cyclic behavior. This demonstrates the first difficulty in analyzing the variance of this data: to analyze the variance, we first need to remove the cyclic terms from all time-series. However, there is a lot of variations in the cyclic behavior of the time-series of different locations. In addition, some of these cycles cannot be easily modeled by a parametric function \footnote{One might try to model the cycles by the summation of sinusoidal terms with different frequencies. However, for some time-series this may need many terms to be included in the summation to achieve a reasonable level of accuracy. In addition, this model cannot capture the non-stationarity in the cycles.}. To overcome these issues, we use a non-parametric approach to remove the cyclic terms from the time-series and de-trend them. This approach, called \textit{$\ell_1$-trend filtering} is explained in Section 2 of the text. We detrended each time-series separately using this method. For each time-series, we found the optimal value of the penalty parameter using \textit{k-fold cross-validation} with $k=5$. We used the R package \textbf{genlasso} to perform these computations \cite{arnold_efficient_2016}.

\begin{figure}[tb]
	\centering
	\includegraphics[width=.2\textheight]{Figures/cities_ts}
 	\caption{Time-series of the temperature (in Kelvin) of three cities.}
 	\label{fig:cities_ts}
\end{figure}

The blue curve in the left panel of \autoref{fig:bloom_estimatedSD} shows the time-series of the temperature of Bloomington after detrending using this method. This figure, reveals another difficulty in estimating the trend of volatility in this data: the variance of this signal, shows cyclic behavior. Also, the cycles are not regular and their amplitude and frequency change. Even if one can describe the behavior of the variance of the time-series at all locations using a single parametric model (for example a variant of the GARCH models \citep{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework. These observations motivate the need to develop a non-parametric framework for the problem at hand.

\begin{figure}[tb]
  \centering
  \includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}
  \caption{Left: The residuals of the time-series of Bloomington
    (averaged weekly) and the estimated SD obtained from the method of
    \autoref{sec:l1tf_var} (red). Middle: the estimated SDs (red) and
    their annual average (black) without imposing the long horizon
    penalty. Right: the same as middle panel but here the long horizon
    penalty is imposed. See the text for more details.} 
  \label{fig:bloom_estimatedSD}
\end{figure}

The red curve in the left panel of \autoref{fig:bloom_estimatedSD}
shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
the time-series of Bloomington obtained from out proposed model. To reduce the number of time-steps we work on the weekly averaged of the data. The curve of the estimated SD captures the periodic variations in the
SD of the signal. Just by looking at this curve, it is hard to say if
the SD is decreasing or increasing. Therefore, we compute the average
of the estimated SD for each year. The estimated SD together with this
annual average is shown in the middle panel of
\autoref{fig:bloom_estimatedSD}. As it can be seen, the annual trend
is not smooth. This is because in the optimization problem
\eqref{eq:l1tf_var}, the smoothness of the annual trend is not
encouraged. To remedy this, we add the following long horizon penalty to \eqref{eq:l1tf_var}:

\begin{equation}
\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
\label{eq:lh_penalty}
\end{equation}

where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$. Also,
$N_{year}$ is the number of years over which we are performing our
analysis (here $N_{year}=10$). Since we are working on the weekly
averaged data, each year corresponds to 52 observations. In the matrix
form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the
matrix $D$. The estimated SDs using this penalty matrix is shown in
the right panel of \autoref{fig:bloom_estimatedSD}. The annual average
of the estimated SDs shows a linear trend with a positive slope.

\bibliographystyle{abbrvnat}
\small
\bibliography{nips-references}

\end{document}

### 2018-05-18 nips-2018/paper.tex
page length ok, to arash for final edits

@@
 \begin{abstract}
-words, words,
-
+In this paper, we present methodology for estimating trends in
+spatio-temporal volatility. We give two algorithms for computing our
+estimator which are tailored for dense, gridded observations over both
+space and time, though these can be easily extended to other
+structures (time-varying network flows, neuroimaging). We motivate our
+methodology using by applying it to a massive climate dataset and
+discuss the implications for climate analysis.
 \end{abstract}
@@
 
-\attn{Arash: Some equations do not have numbering and some have.}
+Estimating smooth trends over time for large collections of time
+series is a common problem in economics, finance, meteorology,
+neuroscience and more. Most of this work focuses on analyzing trends
+(or removing trends) in the temporal average, but trends in variance
+can be more relevant, especially for financial data but also for
+climate science.
 
-\attn{Arash: fix appendicies and the references to them.}
+Trends in terrestrial temperature variability are perhaps more relavant for species
+viability than trends in mean
+temperature~\citep{huntingford_no_2013},
+because, an increase in the
+temperature variability will increase the probability of extreme hot
+outliers~\citep{VasseurDeLong2014}. Recent climate literature
+suggests that it is more difficult for society to adapt to these
+extremes than to the gradual increase in the mean temperature
+\citep{hansen_perception_2012,huntingford_no_2013}. Furthermore, the willingness of popular media to
+emphasize the prevalence extreme cold events coupled with a
+fundamental misunderstanding of the relationship between climate (the global
+distribution of weather over the long run) and weather (observed
+short-term, localized behavior) leads to public misunderstanding of climate
+change. In fact, increased frequency of extreme cold events in the
+northern hemisphere is can be partially attributed to increases in
+mean climate but is also due to increases in temperature
+variance~\citep{Screen2014,FischerBeyerle2013,TrenberthZhang2014}.
 
-Nonparametric variance estimation for spatio-temporal data.
-
-\subsection{Motivating applications}
-
-\attn{cut this down}
-
-There is a considerable interest in determining if there is an
-increasing trend in the climate variability
-\citep{hansen_perception_2012,huntingford_no_2013}. An increase in the
-temperature variability will increase the probability of extreme hot
-outliers. It might be harder for the society to adapt to these
-extremes than to the gradual increase in the mean temperature
-\citep{huntingford_no_2013}.
-
-In this paper, we consider the problem of detecting the trend in the
-temperature volatility. All analyses are performed on a sub-set of the
-European Centre for Medium-Range Weather Forecasts (ECMWF) ERA-20C
-dataset \citep{uppala_era-40_2005}. This dataset include the
-temperature measurements over a grid over the earth from 1957 to
-2002.~\citep{VasseurDeLong2014,TrenberthZhang2014,StatenKahn2016,Screen2014,FischerBeyerle2013}
-
-Research on analyzing the trend in the volatility of spatio-temporal
-data is scarce. \citep{hansen_perception_2012} studied the change in
+Nevertheless, research examining trends in the volatility of
+spatio-temporal climate data is scarce. \citep{hansen_perception_2012} studied the change in
 the standard deviation (SD) of the surface temperature in the NASA
-Goddard Institute for Space Studies gridded temperature data set. In
-their analysis, for each geographical position, the mean of the
-temperature computed for the period 1951-1980 (called the base-period)
-at that position, is subtracted from the corresponding time
-series. Each time series is then divided by the standard deviation
-computed at each position and during the same time period. The
-distribution of the resulting data is then plotted for different
-periods. These distributions represent the deviation of the
-temperature for a specific period, from the mean in the base period,
-in units of the standard deviation in that period. The results showed
-that these distributions are widen for the resent time periods
-compared to 1951-1980.  \citep{huntingford_no_2013} took a similar
-approach in analysing the ERA-40 data set. However, in addition to the
-aforementioned method, they computed the distribution of the SDs in an
-alternative way: for each position and each time period, the deviation
-of the time-series at that position from the mean in that time period
-at that position was computed, and then divided by the SD of that
-position in the period before 1981. The results showed that there
+Goddard Institute for Space Studies gridded temperature data set by
+examining the empirical SD at each spatial location relative to that
+location's SD over a base period, and showed
+that these estimates are increasing.
+\citep{huntingford_no_2013} took a similar
+approach in analyzing the ERA-40 data set. Their results showed that there
 still is an increase in the SDs from 1958-1970 to 1991-2001, but this
@@
 time-evolving global SD from the de-trended time-series at each
-position. The resulting curve suggested that the global SD has been
+position, which suggests that the global SD has been
 stable.
 
-These previous work (and other related research, e.g.,
+These and other related research, e.g.,
 \citep{rhines_frequent_2013}) have several shortcomings. First, no
-statistical analysis has been performed to examine if the change in
-the SD is statistically significant. Second, the methodologies for
-computing the SDs are rather arbitrary. The deviation of each
-time-series in a given period, is computed from either the mean of a
-base-period (as in \citep{hansen_perception_2012}), or from the given
-period (as in \citep{rhines_frequent_2013,huntingford_no_2013}). These
-deviations are then normalized using the SD of the base-period or the
-given period. No justification is provided for these choices. Third,
-the correlation between the observations is ignored. The observations
-in subsequent days and close geographical positions could be highly
-correlated. Without considering these correlations, any conclusion
-based on the averaged data could be flawed.
+statistical analysis has been performed to examine if the changes in
+the SD are statistically significant. Second, the methodologies for
+computing the SDs are highly sensitive to the choice of base period. Third,
+and most importantly, temporal and spatial correlations between the
+observations are ignored.
+
+\subsection{Related work}
+
+Variance estimation for financial time series has a lengthy history,
+focused especially on parametric models like the generalized
+autoregressive conditional heteroskedasticity (GARCH) process~\citep{engle2002dynamic} and
+stochastic volatility models~\citep{HarveyRuiz1994}. These models (and
+related AR processes) are specifically for parametric modelling of
+short ``bursts'' of high volatility, behavior typical of financial
+instruments. Parametric models for spatial data go back at least
+to~\citep{besag1974spatial} who proposed a conditional probability
+model on the lattice for examining plant ecology.
+
+More recently, nonparametric models for both spatial and temporal data
+have focused on using $\ell_1$-regularization for trend
+estimation. \citep{KimKoh2009} proposed $\ell_1$-trend filtering for
+univariate time series, which forms the basis of our methods. These
+methods have been generalized to higher order temporal smoothness
+~\citep{Tibshirani2014}, graph dependencies~\citep{WangSharpnack2016},
+and, most recently, small, time-varying
+graphs~\citep{HallacPark2017}. Our methodology is similar in flavor
+to~\citep{HallacPark2017}, though it uses a different likelihood
+function to emphasize variance estimation rather than trends in mean
+signal. Furthermore, the focus is in high-dimensional, regular data
+rather than a small number of changing graphs.
+
+\subsection{Main contributions}
 
@@
 detecting the trend in the volatility of spatio-temporal data. In this
-methodology, the variance at each position and time, is considered as
-a hidden (un-observed) variable. The value of these hidden variables
+methodology, the variance at each position and time is considered as
+a hidden variable. The values of these hidden variables
 are then estimated by maximizing the likelihood of the observed
-data. We show that this formulation per se, is not appropriate for
-detecting the trend. To overcome this issue, we penalize the
-differences between the estimated variances of the observations which
-are temporally and/or spatially close to each other. This will result
-in an optimization problem called the \textit{generalized LASSO
-  problem} \citep{tibshirani_solution_2011}. As we will see, the
-dimension of this optimization problem is very high and so the
-standard methods for solving the generalized LASSO cannot be applied
-directly. We investigate two methods for solving this optimization
-problem. In the first method, we adopt an optimization technique
-called alternative direction method of multipliers (ADMM)
-\citep{boyd_distributed_2011}, to divide the total problem into
-several sub-problems of much lower dimension and show how the total
-problem can be solved by iteratively solving these sub-problems. The
-second method, called the \textit{linearized ADMM algorithm}
-\citep{parikh_proximal_2014} solves the main problem by iteratively
-solving a linearized version of it. We will compare the benefits of
-each method.
+data. We show that this formulation is not appropriate for
+detecting the trend, so, following ~\citep{Tibshirani2014}, we penalize the
+differences between the estimated variances which
+are temporally and spatially ``close'', resulting
+in a generalized LASSO problem. % However, the
+% dimension of this optimization problem is very high and so the
+% standard methods solvers are inadequate. We develop two algorithms
+% which are computationally feasible. In the first method, we adopt an optimization technique
+% called alternative direction method of multipliers (ADMM)
+% \citep{boyd_distributed_2011}, to divide the total problem into
+% several sub-problems of much lower dimension and show how the total
+% problem can be solved by iteratively solving these sub-problems. The
+% second method, called the \textit{linearized ADMM algorithm}
+% \citep{parikh_proximal_2014} solves the main problem by iteratively
+% solving a linearized version of it. We will compare the benefits of
+% each method.
 
-Also neuroscience.
-
-\subsection{Related work}
-
-Mention ~\citep{KimKoh2009,HallacPark2017}. Also,
-~\citep{Tibshirani2014,TibshiraniTaylor2011}. ARCH/GARCH.
-
-\citep{SadhanalaWang2017,LinSharpnack2017,WangSharpnack2016}~\citep{RamdasTibshirani2016}
-
-\subsection{Main contributions}
-
-\begin{itemize}
-\item We propose a model for non-parametric variance estimation for a
-  spatio-temporal process (~\autoref{sec:ell_1-trend-filt}).
-\item We derive two algorithms to fit our estimator when applied to
-  very large data (~\autoref{sec:prop-optim-meth}).
+Our main contributions are as follows:
+\begin{enumerate}
+\item We propose a model for nonparametric variance estimation for a
+  spatio-temporal process (\autoref{sec:ell_1-trend-filt}).
+\item We derive two alternating direction method of multiplier
+  algorithms (ADMM) to fit our estimator when applied to 
+  very large data (\autoref{sec:prop-optim-meth}). We give situations
+  under which each algorithm is most likely to be useful.
 \item We illustrate our methods on a large global temperature dataset
@@
   simulation constructed to mimic these data's features
-  (~\autoref{sec:empirical-evaluation}). 
-\end{itemize}
+  (\autoref{sec:empirical-evaluation}). 
+\end{enumerate}
+
+While we motivate and illustrate our methods on large, gridded climate
+data, we note that our algorithms are also applicable to neuroimaging
+data or large collections of financial instruments.
 
@@
 
-$\ell_1$-trend filtering was proposed by \citep{KimKoh2009} as a method for estimating a smooth, time-varying trend. It is formulated as the optimization problem
+$\ell_1$-trend filtering was proposed by \citep{KimKoh2009} as a
+method for estimating a smooth, time-varying trend. It is formulated
+as the optimization problem 
 % : a time-series $y_t$, $t=1,...,T$ is observed and we believe that
@@
 % following convex optimization problem: 
-\begin{equation}
+$
 \min_{\beta} \frac{1}{2} \sum_{t=1}^{T} (y_t-\beta_t)^2+\lambda
 \sum_{t=1}^{T-2} \left|\beta_t-2\beta_{t+1}+\beta_{t+2}\right| 
-\end{equation}
+$
 or equivalently:
@@
 \eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times
-S$ vector and $D$ is replaced by $D_{TS} \in \mathbb{R}^{(N_t+N_s)
+S$ vector and $D$ is replaced by $D_{ST} \in \mathbb{R}^{(N_t+N_s)
   \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot
@@
 \begin{equation}
-\min_h -l(y\given h)+ \Lambda^\top | D_{TS}h |.
+\min_h -l(y\given h)+ \Lambda^\top | D_{ST}h |.
 \label{eq:l1tf_var_st_mat}
@@
 variables corresponding to the same global variable.  
-
 The augmented Lagrangian corresponding to 
@@
 \big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert
-x_i-\tilde{h}_i \lVert_2^2 \big)$. Now, we can apply ADMM to $L_\rho$
-which results in the following ADMM updates:
-\begin{align}
-x_i&\leftarrow\argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
-     (\rho/2)  \norm{ x_i-\tilde{h}_i }_2^2\\
-h_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j \label{eq:ADMM_steps}\\
-u_i&\leftarrow u_i + \rho (x_i-\tilde{h}_i).
-\end{align}
-\attn{If you hate the left arrow, that's fine. But change it
-  everywhere, and use $=$. The $:=$ notation means that ``thing on
-  left is defined to be thing on right'' and isn't really appropriate here.}
-Here, $S_k$ is the number of local variable entries that correspond to
-$h_k$, and $u_i$ are the Lagrange multipliers.
-
+x_i-\tilde{h}_i \lVert_2^2 \big)$. Now, we can apply ADMM to $L_\rho$.
+% which results in the following ADMM updates:
+% \begin{align}
+% x_i&\leftarrow\argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
+%      (\rho/2)  \norm{ x_i-\tilde{h}_i }_2^2\\
+% h_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j \label{eq:ADMM_steps}\\
+% u_i&\leftarrow u_i + \rho (x_i-\tilde{h}_i).
+% \end{align}
+% \attn{If you hate the left arrow, that's fine. But change it
+%   everywhere, and use $=$. The $:=$ notation means that ``thing on
+%   left is defined to be thing on right'' and isn't really appropriate here.}
+% Here, $S_k$ is the number of local variable entries that correspond to
+% $h_k$, and $u_i$ are the Lagrange multipliers.  
 To solve the optimization problem \eqref{eq:l1tf_var_st_mat} using
@@
 questions: first, how to choose the 
-local variables $x_i$, and second, how to the update in the first line of
-\eqref{eq:ADMM_steps}.
-
+local variables $x_i$, and second, how to the update them.
 \begin{figure}[tb]
@@
 % computation more involved than \autoref{sec:l1tf_var}. 
-The details of the remaining computations
-the computations are explained in the Supplement.
+\autoref{alg:conADMM} gives the general version of this algorithm. A
+more detailed discussion of the is in the Supplement.
 
-% The complete ADMM algorithm for estimating the variances is
-% represented in \autoref{alg:ADMM}. All the computations in the three
-% updating steps \eqref{eq:ADMM_steps} can be performed in parallel. The
-% number of rows and columns of the sub-cubes should be chosen so that
-% the updating of $x_i$ could be performed in one processor. We choose
-% $3\times3\times521$ sub-cubes.
-
-% \begin{algorithm}[tb]
-%   \caption{ADMM for sparse estimation of variance of spatio-temporal
-%     data \attn{Fix this}}
-%   \label{alg:ADMM}
-%   \begin{algorithmic}
-%     \STATE {\bfseries Input:} data $y$, mapping $\mathscr{G}(i,j)$,
-%     $\rho,\lambda_t,\lambda_s$ 
-%     \STATE Initialization: $x_i^0=z^0=u_i^0=\textbf{0}$.
-%     \FOR{$m=1,2,...$ }
-%     \FOR{$i=1$ {\bfseries to} $N_{sub-cubes}$}
-%     \STATE compute $\nu_i$ from \eqref{eq:dual_x_update}
-%     \STATE compute $w_i$ from \eqref{eq:conj_x_update}
-%     \STATE set $x_i^m:=w_i$
-%     \ENDFOR
-%     \STATE Compute $z^m$ from  \eqref{eq:ADMM_steps}
-%     \STATE Compute $u_i^m$ from  \eqref{eq:ADMM_steps}
-%     \ENDFOR
-%   \end{algorithmic}
-% \end{algorithm}
+\begin{algorithm}[tb]
+  \caption{Consensus ADMM }
+  \label{alg:conADMM}
+  \begin{algorithmic}
+    \STATE {\bfseries Input:} data $y$, penalty matrix $D$, 
+    $\epsilon, \rho,\lambda_t,\lambda_s >0$.
+    \STATE {\bf Set:} $h\leftarrow 0$, $z\leftarrow 0$, $u\leftarrow
+    0$. \COMMENT{Initialization} 
+    \REPEAT
+    \STATE $x_i\leftarrow\argmin_{x_i} -l(y_i\given
+    x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|$
+    \STATE $\quad\quad\quad+ (u_i)^\top x_i +
+    (\rho/2)  \lVert x_i-\tilde{h}_i \rVert_2^2$. \COMMENT{Update local
+      vars using PDIP}
+    \STATE $h_k\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j
+    $. \COMMENT{Global update.}
+    \STATE $ u_i\leftarrow u_i + \rho (x_i-\tilde{h}_i)$. \COMMENT{Dual update}
+    \UNTIL {$\norm{h^{m+1}-h^m}_2^2\ < \epsilon$}
+    \STATE {\bf Return:} $h$.
+  \end{algorithmic}
+\end{algorithm}
 
@@
 where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu < \rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
-\begin{equation}
+$
  \prox_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \norm{ x-u}_2^2.
-\label{eq:linADMM_prox}
-\end{equation}
+$
 
@@
     \STATE $u\leftarrow u + Dh-z$. \COMMENT{Dual update}
-    \UNTIL {$\max\{\norm{h-z}_2^2,\; \norm{z^{m+1}-z^m}_2^2\} < epsilon$}
+    \UNTIL {$\max\{\norm{h-z}_2^2,\; \norm{z^{m+1}-z^m}_2^2\} < \epsilon$}
     \STATE {\bf Return:} $z$.
@@
 data on a 5 by 7 grid and for 780 time steps with $S=4$. 
-\attn{The
-parameters of the variance function are shown in
-\autoref{tab:sim_params} in the Supplement. For reference, we plot the
+Specific parameter choices of the variance function are shown in
+\autoref{tab:sim_params} in the Supplement. For illustration, we also plot the
 variance function for all locations at $t=25$ and $t=45$ in as well as
 the variance across time at $(0,0)$ in \autoref{fig:true_var_spatial}
-in Appendix C.}
+in the Supplement.
 
@@
 
-\paragraph{Analysis of trend of temperature volatility}
+\paragraph{Analysis of trend in temperature volatility}
 
@@
 hemisphere. The SD in most locations on the northern
-hemisphere had a negative trend in this time period.
+hemisphere had a negative trend in this time period, though spatially,
+this decreasing pattern is localized mainly toward the extreme
+northern latitudes and over oceans. In many ways, this is consistent
+with climate change predictions: oceans tend to operate as a local
+thermostat, regulating deviations in local temperature, while warming polar
+regions display fewer days of extreme cold.
 
@@
 
-\attn{We did not do this. Change the below.}
-We applied the linearized ADMM algorithm to the surface temperature
-data on a grid over the united states, for years 1992-2002. The
-results showed that in many locations the variance of the temperature
-has increased about 1 unit in 10 years.
+% \attn{We did not do this. Change the below.}
+% We applied the linearized ADMM algorithm to the surface temperature
+% data on a grid over the united states, for years 1992-2002. The
+% results showed that in many locations the variance of the temperature
+% has increased about 1 unit in 10 years.
 
-The goal of this paper, however, is not to make any conclusions about
-the trend in the variance because we solved the problem only for a
-grid over the united states and for 10 years of the data. A thorough
-analysis, needs the full solution over the globe and for a longer time
-period. The goal of the paper, was to propose the idea of estimating
-the trend in variance of spatio-temporal signals using generalized
-lasso and to investigate the algorithms for solving the resulting
-optimization problem.
+% The goal of this paper, however, is not to make any conclusions about
+% the trend in the variance because we solved the problem only for a
+% grid over the united states and for 10 years of the data. A thorough
+% analysis, needs the full solution over the globe and for a longer time
+% period. The goal of the paper, was to propose the idea of estimating
+% the trend in variance of spatio-temporal signals using generalized
+% lasso and to investigate the algorithms for solving the resulting
+% optimization problem.
 

### 2018-05-18 nips-2018/paper-supplement.tex
ready to submit

@@
 
+The matrix $D$ in these equations have the following form:  $D_{total}^t=[D_{t}^t|D_{s}^t]$. The matrix $D_t$ is the following block-diagonal matrix and corresponds to the temporal penalty: 
+ \begin{equation}
+ D_t=\begin{bmatrix}
+ D &  & \\ 
+  & \ddots & \\
+  &  & D
+ \end{bmatrix}
+ \label{eq:d_t_matrix}
+ \end{equation}
+
+where $D$ was defined in Section 2 of the main text.
+The number of the diagonal blocks is equal to the grid size $n_r \times n_c$.
+Each row of the matrix $D_s$ corresponds to one spatial constraint in Equation \eqref{eq:l1tf_var_st} in the text.
+For example, the first $T$ rows correspond to $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to $|h_{11t}-h_{12t}|$, and so on.
+
 \section{Appendix B}
@@
 
-In this Appendix we give more details on performing the x-update step in \autoref{eq:ADMM_steps}. We need to solve the following optimization problem:
+In this Appendix we give more details on performing the x-update step in Algorithm 1. We need to solve the following optimization problem:
 
@@
 \begin{table}[tb]
-  \caption{Parameters used to simulate data. \attn{Any ideas to take
-      up less space with this info?}}
+  \caption{Parameters used to simulate data.}
   \label{tab:sim_params}
@@
 
-\autoref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil).
+Figure \autoref{fig:cities_ts} shows the time-series of the temperature of three cities: Bloomington (USA), San Diego (USA) and Manaus (Brazil).
 The time-series of Bloomington and San Diego show clear cyclic behavior.
@@
 
-The blue curve in the left panel of \autoref{fig:bloom_estimatedSD} shows the time-series of the temperature of Bloomington after detrending using this method.
+The blue curve in the left panel of Figure \autoref{fig:bloom_estimatedSD} shows the time-series of the temperature of Bloomington after detrending using this method.
 This figure, reveals another difficulty in estimating the trend of volatility in this data: the variance of this signal, shows cyclic behavior.
@@
 
-The red curve in the left panel of \autoref{fig:bloom_estimatedSD}
+The red curve in the left panel of Figure \autoref{fig:bloom_estimatedSD}
 shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of

### 2018-05-18 nips-2018/paper.tex
ready to submit

@@
  Bloomington, IN 47408 \\
- \texttt{dajmcdon@indiana.cmu.edu}}
+ \texttt{dajmcdon@indiana.edu}}
 
@@
 structures (time-varying network flows, neuroimaging). We motivate our
-methodology using by applying it to a massive climate dataset and
+methodology by applying it to a massive climate dataset and
 discuss the implications for climate analysis.
@@
 change. In fact, increased frequency of extreme cold events in the
-northern hemisphere is can be partially attributed to increases in
+northern hemisphere can be partially attributed to increases in
 mean climate but is also due to increases in temperature
@@
 These and other related research, e.g.,
-\citep{rhines_frequent_2013}) have several shortcomings. First, no
+\citep{rhines_frequent_2013}, have several shortcomings. First, no
 statistical analysis has been performed to examine if the changes in
@@
 step on a system of nonlinear equations. Due to space limitations, we
-defer details to Appendix B in the Supplement.
+defer details to Appendix A in the Supplement, where we show how to derive the dual of this optimization problem and compute the first and second derivatives of the dual objective function.
 
@@
 +\lambda_2 \sum_{t,i} \sum_{j=1}^{n_c-1} \left|h_{ijt}-h_{i(j+1)t}\right|
-\label{eq:l1tf_var_st}
+\label{eq:l1tf_var_st}.
 \end{align}
@@
 component of this
-penalty is a special case of trend
-filtering on graphs~\citep{WangSharpnack2016} which penalizes the difference
-between the estimated values of the signal on the connected
-nodes. As before, we can write
-\eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times
-S$ vector and $D$ is replaced by $D_{ST} \in \mathbb{R}^{(N_t+N_s)
-  \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot
-(2n_rn_c-n_r)$ are the number of temporal and spatial constraints,
-respectively\footnote{$N_s$ is obtained by counting the number of
-  unique constraints at each location and at all times.}. Then, as we 
-have two different tuning parameters for the temporal and spatial
-components, we write $\Lambda =\left[\lambda_1\one_{N_t}^\top,\;
-  \lambda_2\one_{N_s}^\top\right]^\top$ leading
-to:\footnote{Throughout the paper, we use $|x|$ for both scalars and
-  vectors. For vectors we use this to denote a vector obtained by
-  taking the absolute value of each entry of $x$.} 
+penalty is a special case of trend filtering on graphs~\citep{WangSharpnack2016} which penalizes the difference between the estimated values of the signal on the connected nodes.
+As before, we can write \eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times S$ vector and $D$ is replaced by $D_{ST} \in \mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and spatial constraints, respectively\footnote{$N_s$ is obtained by counting the number of unique constraints at each location and at all times.}.
+The exact form of this matrix in clarified in Appendix A in the Supplement.
+Then, as we have two different tuning parameters for the temporal and spatial components, we write $\Lambda =\left[\lambda_1\one_{N_t}^\top,\;  \lambda_2\one_{N_s}^\top\right]^\top$ leading to:\footnote{Throughout the paper, we use $|x|$ for both scalars and vectors.
+For vectors we use this to denote a vector obtained by taking the absolute value of each entry of $x$.}
 % let $h$ be a
@@
 \end{equation}
-\attn{equations should have proper punctuation as if they were a sentence.}
-% where $\Lambda = \mathbf{I}_{Tn_rn_c}\left[\lambda_1\one_T^\top,\;
-%   \lambda_2\one_{n_rn_c}^\top\right]$ and 
-% $D_{total}^t=[D_{t}^t|D_{s}^t]$, where $n_t$ and $n_s$ are the number
-% of rows of $D_{t}$ and $D_{s}$, respectively (see below). The matrix
-% $D_t$ is the following block-diagonal matrix and corresponds to the
-% temporal penalty: 
-% \begin{equation}
-% D_t=\begin{bmatrix}
-% D &  & \\ 
-%  & \ddots & \\
-%  &  & D
-% \end{bmatrix}
-% \label{eq:d_t_matrix}
-% \end{equation}
-%  where $D$ was defined in \eqref{eq:d_matrix}. The number of the
-%  diagonal blocks is equal to the grid size $n_r \times n_c$. Each row
-%  of the matrix $D_s$ corresponds to one spatial constraint in
-%  \eqref{eq:l1tf_var_st}. For example, the first $T$ rows correspond to
-%  $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to
-%  $|h_{11t}-h_{12t}|$, and so on. 
-% \attn{Detailed discussion of the PDIP to here}
-
-% \attn{Add back the full form of $D$ above? It was too cumbersome
-%   before, but possibly necessary}
 
@@
 1^\circ$ grid over the entire northern hemisphere and daily data over
-10 years, we have $n_r=90$, $n_c=180$, $T=3650$ and so $D_{ST}$ has
+10 years, we have $n_r=90$, $n_c=360$, $T=3650$ and so $D_{ST}$ has
 approximately $10^8$ columns and $10^8$ rows. In each step of the PDIP
-algorithm, we need to solve a linear system of equations in $A$ which
+algorithm, we need to solve a linear system of equations which
 depends on $D_{ST}^\top D_{ST}$ (see appendix A and B). Therefore,
@@
 
-Given an optimization problem of the form $\min_h f(h)$, where
-$h\in\mathbb{R}^n$ is the global variable and
+Consider an optimization problem of the form $\min_h f(h)$, where
+$h\in\mathbb{R}^n$ is the \textit{global variable} and
 $f(h):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is
@@
 variables corresponding to the same global variable.  
-The augmented Lagrangian corresponding to 
+The \textit{augmented Lagrangian} corresponding to 
 \eqref{eq:consADMM} is $L_\rho(x,h,y)=\sum_i
 \big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert
-x_i-\tilde{h}_i \lVert_2^2 \big)$. Now, we can apply ADMM to $L_\rho$.
+x_i-\tilde{h}_i \lVert_2^2 \big)$.
+Now, we can apply ADMM to $L_\rho$.
+This results in solving $N$ independent optimization problems followed by a step to achieve consensus among the solutions in each iteration.
 % which results in the following ADMM updates:
@@
   \centering
-  \includegraphics[height=.2\textheight]{Figures/data_cube}
+  \includegraphics[height=.13\textheight]{Figures/data_cube}
   \caption{The cube represents the global variable $h$ in space and
@@
 In \autoref{fig:data_cube}, the global variable $h$ is represented
-as a cube (using the subset of the US as an example). We decompose $h$
-into sub-cubes as shown by white lines. With
+as a cube. We decompose $h$
+into sub-cubes as shown by white lines.
+Each local variable inside the sub-cubes corresponds to only one global variable.
+The local variables on the border (white lines), however, correspond to more than one global variable.
+With
 this definition of $x_i$, the objective
@@
 $\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
-penalties corresponding to $x_i$ only. 
-% The $x$-update step in
-% \eqref{eq:ADMM_steps} is the following optimization problem:
-% $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) + 
-% (u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
-Thus, with this choice of the local variables $x_i$, we solve the
-$x$-update using the PDIP method. 
-% As it 
-% was explained in \autoref{sec:l1tf_var}, to use PDIP we need to
-% compute the dual problem, which in turn needs the computation of the
-% Fenchel conjugate of the loss function. In addition, referring to
-% \autoref{alg:pdip}, we need to compute the gradient and Jacobian of
-% the conjugate functions. It is important to note that compare to the
-% optimization problem \autoref{eq:l1tf_var}, the loss function in the
-% x-update step in \autoref{eq:ADMM_steps} includes the quadratic term
-% $(\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2$. This makes the
-% computation more involved than \autoref{sec:l1tf_var}. 
-\autoref{alg:conADMM} gives the general version of this algorithm. A
-more detailed discussion of the is in the Supplement.
+penalties corresponding to $x_i$ only in one sub-cube.
+Thus, we now need to use PDIP to solve $N$ problems each of size $n_i$, which is feasible for small enough $n_i$.
+\autoref{alg:conADMM} gives the general version of this algorithm.
+A more detailed discussion of this is in Appendix B of the Supplement where we show how to compute the dual and the derivatives of the augmented Lagrangian.
 
@@
 \end{align}
-where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu < \rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
-$
- \prox_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \norm{ x-u}_2^2.
-$
+where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu < \rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is defined as $\prox_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \norm{ x-u}_2^2$.
+Proximal algorithms are feasible when these proximal operators can be evaluated efficiently which, as we show next, is the case for our problem.
 
@@
 % $\prox_{\mu f}$ and $\prox_{\rho g}$.
-Proximal
-algorithms are feasible when these proximal operators can be
-evaluated efficiently which, as we show next, is the case for our
-problem.
 
@@
 larger than optimal value of $\lambda_t$ leads to estimated values
-which are ``too smooth".
+which are ``too smooth". The middle panel of \autoref{fig:true_fitted_var} shows the
+convergence of both methods. Each iteration of the linearized
+algorithm takes 0.01 seconds on average while each iteration of the
+consensus ADMM takes about 20 seconds.
 
@@
 
-The middle panel of \autoref{fig:true_fitted_var} shows the
-convergence of both methods. Each iteration of the linearized
-algorithm takes 0.01 seconds on average while each iteration of the
-consensus ADMM takes about 20 seconds.
-
 To further examine the performance of the proposed model, we next
@@
 spatial smoothness (equivalent to fitting the model in
-\autoref{sec:l1tf_var} to each time-series separately), a model which
-does not consider imposes only spatial smoothness, and a GARCH(1,1) model. We
+\autoref{sec:l1tf_var} to each time-series separately), a model which only imposes spatial smoothness, and a GARCH(1,1) model. We
 simulated 100 datasets using the method explained above with $\sigma_s
@@
 the 2 meter temperature measured daily at 12 p.m from January 1, 1960
-to December 24, 2010.
-
-The Supplement explains some preprocessing and investigates some
+to December 24, 2010. The Supplement explains some preprocessing and investigates some
 properties of the time-series of different locations on
-earth. \autoref{fig:bloom_estimatedSD} a processed time-series for a
+earth. \autoref{fig:bloom_estimatedSD} shows a processed time-series for a
 single location.
@@
 parameters and then choose the values which minimize a model selection
-criterion. However, such analyses needs the computation of the degrees
+criterion. However, such analysis needs the computation of the degrees
 of freedom. Several previous work have investigated the df in
@@
 generalized lasso with general objective function has not been
-considered yet.
-
-In this paper, we use a heuristic method for choosing $\lambda_t$ and
+considered yet. In this paper, we use a heuristic method for choosing $\lambda_t$ and
 $\lambda_s$: we compute the optimal solution for a range of values of
@@
 \lVert D_{total}h \lVert$ will be smaller but with the cost of larger
-$-l(y|h)$.
-
-We computed the optimal solution for all the combinations of the
+$-l(y|h)$. We computed the optimal solution for all the combinations of the
 following sets of values: $\lambda_t \in \{0,2,4,8,10,15,200,1000\} \, \, ,
@@
   \centering
-  \includegraphics[width=.9 \columnwidth]{Figures/ts_estimatedVar}\\
+  \includegraphics[width=.8 \columnwidth]{Figures/ts_estimatedVar}\\
   \caption{Detrended data and the estimated SD for
@@
 changed in each location, we subtract the average of the estimated
-variance in \attn{1992} from the average in the following years and compute
+variance in 1961 from the average in the following years and compute
 their sum. The value of this change in the variance in each location
@@
 
+\newpage
 \bibliographystyle{abbrvnat}

### 2018-06-12 Risk estimation/risk-estim.tex
Update ignore, begin df doc

\documentclass[11pt]{article}

% Include statements
\usepackage{graphicx}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage[numbers,square,sort]{natbib}
\usepackage[left=1in,top=1in,right=1in,bottom=1in,nohead]{geometry}
\usepackage{pdfsync}
\usepackage{multirow}
\usepackage{enumitem}
\setlist{nosep}
\usepackage{hyperref}
\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}
\usepackage{algorithm,algorithmic}
% \SetKw{KwSet}{Set}
% \SetKw{KwSolve}{Solve}
\def\algorithmautorefname{Algorithm}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{6pt plus
  4pt minus 2pt}
\titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{6pt plus
  4pt minus 2pt}
\titlespacing\paragraph{0pt}{6pt plus 4pt minus 2pt}{2pt plus
  2pt minus 0pt}

\usepackage[showonlyrefs]{mathtools}
%\usepackage{autonum}

% Bibliography

\renewcommand*{\figureautorefname}{Figure}%
\renewcommand*{\tableautorefname}{Table}%
\renewcommand*{\partautorefname}{Part}%
\renewcommand*{\chapterautorefname}{Chapter}%
\renewcommand*{\sectionautorefname}{Section}%
\renewcommand*{\subsectionautorefname}{Section}%
\renewcommand*{\subsubsectionautorefname}{Section}%

% Macros
\DeclareMathOperator*{\argmin}{argmin}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\one}{\mathbf{1}}
%\renewcommand{\S}{\mathcal{S}}
\newcommand{\Expect}[1]{\E\left[#1\right]}
\newcommand{\given}{\mbox{ }\vert\mbox{ }}
\renewcommand{\hat}{\widehat}
\DeclareMathOperator*{\trace}{tr}
\DeclareMathOperator*{\diag}{diag}
\renewcommand{\tilde}{\widetilde}
\newcommand{\Expectwrt}[2]{\mathbb{E}_{ #1 }\left[ #2 \right]}
\usepackage{xspace}
\makeatletter
\newcommand*{\iid}{%
    \@ifnextchar{.}%
        {i.i.d.}%
        {i.i.d.\@\xspace}%
}
\makeatother

\newcommand{\Pnd}{P_{\mathcal{N}(D)}}

\newtheorem{result}{Result}
\newtheorem{theorem}{Theorem}[result]
\renewcommand*{\sectionautorefname}{Section}%

\renewcommand{\P}{\mathbb{P}}
\DeclareMathOperator*{\argmax}{argmax}
\def\indep{\perp\!\!\!\perp}

\newcommand{\makeHeader}{\begin{center} 
DJM \hfill Risk estimation for Gen-Gen Lasso \hfill \today

\rule{\textwidth}{1pt}
\end{center}
}

\begin{document}
\makeHeader

This document combines a number of threads for model selection in our
problem (and other related optimization problems). First, we consider
estimating the degrees of freedom. Then we give two results which can
be used to select models.

\section{Degrees of freedom for Gen-Gen Lasso}
\label{sec:degrees-freedom-gen}

Suppose we are minimizing the negative log likelihood from a natural
exponential family subject to the generalized lasso penalty. That is,
given $Y_i \sim p(y \given \eta_i(\tau))$ for $\tau \in \R^p$, $i=1,\ldots,n$, our goal is to solve
\begin{equation}
  \label{eq:1}
  \hat\tau = \argmin_\tau \sum_{i=1}^n \phi(\eta_i(\tau)) - \langle y_i,\
  \eta_i(\tau)\rangle + \lambda\norm{D\tau}_1,
\end{equation}
where $D\in \R^{q\times p}$.

This generalizes the various standard models. For example,
$\ell_1$-trend filtering has $\phi(x) = x^2/2$, $\eta_i(x)=x_i$, and
$D$ the first-order discrete difference operator. Variance estimation
in our context has $\phi(x) = x$, $y_i = z^2_i$ (where $z_i\sim
N(0,\exp(-h)$) and $\eta_i(x) = -e^{-\tau_i}$. Logistic loss also
falls into this category.

We first consider the case that $\eta_i(\tau) = \eta_i(\tau_i)$, i.e.,
$\tau \in \R^n$ and $p(y_i \given \tau) = p(y_i \given \tau_i)$.

\begin{theorem}
  The divergence of $\hat\tau(y)$ is given by
  \begin{equation}
    \label{eq:2}
    \trace\left(D \hat\tau(y)\right) = -\Pnd \left(\Pnd
      \diag\left(\frac{d^2}{d\tau_i^2}\ell
        \vert_{y_i,\hat{\eta}_i}\right) \Pnd\right)^\dagger \Pnd
      \diag\left(\frac{d^2}{d\tau_i d y_i} \ell \vert_{y_i,\hat{\eta}_i}\right),
    \end{equation}
    where $\ell = \ell(y,\eta(\tau_i)) = \sum_{i=1}^n \phi(\eta_i(\tau_i)) - \langle y_i,\
    \eta_i(\tau_i)\rangle$, $\hat{\eta_i} = \eta_i(\hat\tau_i)$, and 
    \begin{equation}
      \label{eq:3}
      \Pnd = I_n - D_S^\top (D_S^\top D_S)^\dagger D_S
    \end{equation}
    is the projection onto the null-space of $D_S$, where $S = \{j \in
    [q] : D\hat\tau = 0\}$, and the notation $D_S$ means the rows of
    $D$ whose indices are in $S$.
\end{theorem}

\paragraph{Special cases:}
\begin{enumerate}
\item If we are interested in the natural exponential family (that is
  $\eta_i(\tau_i)=\tau_i$), then $\frac{d^2}{d\tau_i d y_i} \ell = -1$
  and $\frac{d^2}{d\tau_i^2}\phi_i(\eta_i(\tau_i)) = \textrm{Var}[y_i]$.
\item For Gaussian likelihood, $\frac{d^2}{d\tau_i^2}\ell = 1$
  and $\frac{d^2}{d\tau_i d y_i} \ell = -1$, so the divergence is the
  dimension of $\mathcal{N}(D_S)$ as shown in
  \citet{TibshiraniTaylor2012}. 
\item For logistic loss, $\frac{d^2}{d\tau_i^2}\ell =
  e^{\tau_i}/(1+e^{\tau_i})^2$ and 
  $\frac{d^2}{d\tau_i d y_i} \ell = -1$.
\item For the case of variance estimation, we have $\frac{d^2}{d\tau_i^2}\ell =
  y^2_ie^{-h}$ and 
  $\frac{d^2}{d\tau_i d y_i} \ell = -e^{-h}$.
\end{enumerate}

Now we generalize to the regression setting. Define
$\mu_i = x_i^\top \beta$. Furthermore, write
\begin{equation}
  \label{eq:4}
  \hat\beta = \argmin_\tau \sum_{i=1}^n \phi(\mu_i(\beta)) - \langle y_i,\
  \mu_i(\beta)\rangle + \lambda\norm{D\beta}_1.
\end{equation}

\begin{theorem}
  The divergence of $\hat\mu(y)$ is given by
  \begin{equation}
    \label{eq:5}
    \trace\left(D \hat\mu(y)\right) = -X_P \left(X^\top_P
      \diag\left(\frac{d^2}{d\mu_i^2}\ell
        \vert_{y_i,\hat{\mu}_i}\right) X_P\right)^\dagger X^\top_P
      \diag\left(\frac{d^2}{d\mu_i d y_i} \ell \vert_{y_i,\hat{\mu}_i}\right),
    \end{equation}
    where $\ell = \ell(y,\eta(\mu_i)) = \sum_{i=1}^n \phi(\eta_i(\mu_i)) - \langle y_i,\
    \mu_i\rangle$, $\hat{\eta_i} = \eta_i(x_i^\top\hat\beta)$, and 
    \begin{equation}
      \label{eq:6}
      X_P =( I_n - D_S^\top (D_S^\top D_S)^\dagger D_S) X
    \end{equation}
    is the projection of $X$ onto the null-space of $D_S$.
\end{theorem}

\begin{proof}
  This follows mainly from Theorem 2 in~\citep{VaiterDeledalle2017},
  though the conditions are non-trivial. To be expanded.
\end{proof}

\bibliographystyle{mybibsty}
\bibliography{AllReferences}
\end{document}

### 2018-06-29 Risk estimation/risk-estim.tex
Drafty but complete risk estimation

@@
 
-\usepackage[showonlyrefs]{mathtools}
 %\usepackage{autonum}
@@
 \DeclareMathOperator*{\trace}{tr}
+\DeclareMathOperator*{\Cov}{Cov}
 \DeclareMathOperator*{\diag}{diag}
@@
 
-\newcommand{\Pnd}{P_{\mathcal{N}(D)}}
+\newcommand{\Pnd}{P_{\mathcal{N}(G)}}
 
-\newtheorem{result}{Result}
-\newtheorem{theorem}{Theorem}[result]
-\renewcommand*{\sectionautorefname}{Section}%
+\usepackage{aliascnt}
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newaliascnt{lemma}{theorem}
+\newtheorem{lemma}[lemma]{Lemma}
+\aliascntresetthe{lemma}
+\providecommand*{\lemmaautorefname}{Lemma}
+\newaliascnt{cor}{theorem}
+\newtheorem{corollary}[cor]{Corollary}
+\aliascntresetthe{cor}
+\providecommand*{\corautorefname}{Corollary}
+\newaliascnt{def}{theorem}
+\newtheorem{definition}[def]{Definition}
+\aliascntresetthe{def}
+\providecommand*{\defautorefname}{Definition}
+\newtheorem{example}{Example}
 
@@
 \def\indep{\perp\!\!\!\perp}
+
+\usepackage{mathtools}
+\mathtoolsset{showonlyrefs}
 
@@
 
-Suppose we are minimizing the negative log likelihood from a natural
-exponential family subject to the generalized lasso penalty. That is,
-given $Y_i \sim p(y \given \eta_i(\tau))$ for $\tau \in \R^p$, $i=1,\ldots,n$, our goal is to solve
+Consider a random vector $Y \in \R^n$ with distribution in the
+exponential family with independent
+components. Specifically, write the density of $Y$ as
+\begin{equation}
+  \label{eq:4}
+  p(Y \given \theta) = \left(\prod_{i=1}^n h(y_i)\right)\exp\left\{
+    \sum_{i=1}^n w(y_i) \mu_i(\theta) -\psi_i(\theta) \right\}.
+\end{equation}
+We will consider estimation of $\theta$ subject to the generalized lasso penalty. That is,
+our goal is to solve 
 \begin{equation}
   \label{eq:1}
-  \hat\tau = \argmin_\tau \sum_{i=1}^n \phi(\eta_i(\tau)) - \langle y_i,\
-  \eta_i(\tau)\rangle + \lambda\norm{D\tau}_1,
+  \hat\theta(y) = \argmin_\theta \ell(y | \theta) +
+  \lambda\norm{G\theta}_1 = \argmin_\theta\sum_{i=1}^n \psi_i(\theta) - w(y_i)\mu_i(\theta) + \lambda\norm{G\theta}_1,
 \end{equation}
-where $D\in \R^{q\times p}$.
+where $G\in \R^{q\times p}$ and $\theta\in\R^p$.
 
-This generalizes the various standard models. For example,
-$\ell_1$-trend filtering has $\phi(x) = x^2/2$, $\eta_i(x)=x_i$, and
-$D$ the first-order discrete difference operator. Variance estimation
-in our context has $\phi(x) = x$, $y_i = z^2_i$ (where $z_i\sim
-N(0,\exp(-h)$) and $\eta_i(x) = -e^{-\tau_i}$. Logistic loss also
-falls into this category.
+This generalizes many standard models. For example,
+$\ell_1$-trend filtering has $p=n$, $\psi_i(x) = x_i^2/2$,
+$\mu_i(x)=x_i$, $w(x)=2$, and
+$G$ the first-order discrete difference operator. Variance estimation
+in our context has $\psi_i(x) = x_i$, $w(x) = y^2_i$ and $\mu_i(x) = e^{-x_i}$. Logistic loss also
+falls into this category. We use $G$ rather than the more common
+choice of $D$ to avoid confusion with the differential operator that
+we will require below.
 
-We first consider the case that $\eta_i(\tau) = \eta_i(\tau_i)$, i.e.,
-$\tau \in \R^n$ and $p(y_i \given \tau) = p(y_i \given \tau_i)$.
+\paragraph{Notation} For a vector-valued function $f:\R^p\rightarrow
+\R^q$, we use the notation $Df$ to denote the Jacobian matrix
+and $\nabla f$ to be the gradient.
+
+We first consider the case that $\theta \in\R^n$ and that $\psi_i$ and
+$\mu_i$ operate component-wise on
+$\theta$. That is $p(y \given \theta) = \prod_{i=1}^n p(y_i \given \theta_i)$.
 
 \begin{theorem}
-  The divergence of $\hat\tau(y)$ is given by
+\label{thm:simple-divergence}
+  Suppose $p(y \given \theta) = \prod_{i=1}^n p(y_i \given \theta_i)$
+  and that $\hat{\theta}$ is a solution of \eqref{eq:1}. Then, the
+  divergence of $\hat\theta(y)$ is given by 
   \begin{equation}
     \label{eq:2}
-    \trace\left(D \hat\tau(y)\right) = -\Pnd \left(\Pnd
-      \diag\left(\frac{d^2}{d\tau_i^2}\ell
-        \vert_{y_i,\hat{\eta}_i}\right) \Pnd\right)^\dagger \Pnd
-      \diag\left(\frac{d^2}{d\tau_i d y_i} \ell \vert_{y_i,\hat{\eta}_i}\right),
+    \trace\left(D \hat\theta(y)\right) = -\Pnd \left(\Pnd
+      \diag\left(\frac{d^2}{d\theta_i^2}\ell
+        \bigg\vert_{y_i,\hat{\theta}_i}\right) \Pnd\right)^\dagger \Pnd
+      \diag\left(\frac{d^2}{d\theta_i d y_i} \ell \bigg\vert_{y_i,\hat{\theta}_i}\right),
     \end{equation}
-    where $\ell = \ell(y,\eta(\tau_i)) = \sum_{i=1}^n \phi(\eta_i(\tau_i)) - \langle y_i,\
-    \eta_i(\tau_i)\rangle$, $\hat{\eta_i} = \eta_i(\hat\tau_i)$, and 
+    where 
     \begin{equation}
       \label{eq:3}
-      \Pnd = I_n - D_S^\top (D_S^\top D_S)^\dagger D_S
+      \Pnd = I_n - G_S^\top (G_S^\top G_S)^\dagger G_S
     \end{equation}
-    is the projection onto the null-space of $D_S$, where $S = \{j \in
-    [q] : D\hat\tau = 0\}$, and the notation $D_S$ means the rows of
-    $D$ whose indices are in $S$.
+    is the projection onto the null-space of $G_S$, where $S = \{j \in
+    [q] : G\hat\theta = 0\}$, and the notation $G_S$ means the rows of
+    $G$ whose indices are in $S$.
 \end{theorem}
@@
 \item If we are interested in the natural exponential family (that is
-  $\eta_i(\tau_i)=\tau_i$), then $\frac{d^2}{d\tau_i d y_i} \ell = -1$
-  and $\frac{d^2}{d\tau_i^2}\phi_i(\eta_i(\tau_i)) = \textrm{Var}[y_i]$.
-\item For Gaussian likelihood, $\frac{d^2}{d\tau_i^2}\ell = 1$
-  and $\frac{d^2}{d\tau_i d y_i} \ell = -1$, so the divergence is the
-  dimension of $\mathcal{N}(D_S)$ as shown in
+  $\mu_i(\theta)=\theta_i$ and $w(y_i)=y_i$), then $\frac{d^2}{d\theta_i d y_i} \ell = -1$
+  and $\frac{d^2}{d\theta_i^2}\psi_i(\theta_i) = \textrm{Var}[Y_i]$.
+\item In particular, if $p(y_i \given \theta_i) = \mbox{N}(\theta_i, \sigma^2)$, then $\frac{d^2}{d\theta_i^2}\ell = \sigma^2$
+  and $\frac{d^2}{d\tau_i d y_i} \ell = -1$, so the divergence is
+  $\sigma^2$ times the
+  dimension of $\mathcal{N}(G_S)$ as shown in
   \citet{TibshiraniTaylor2012}. 
-\item For logistic loss, $\frac{d^2}{d\tau_i^2}\ell =
-  e^{\tau_i}/(1+e^{\tau_i})^2$ and 
+\item For logistic loss, $\frac{d^2}{d\theta_i^2}\ell =
+  e^{\theta_i}/(1+e^{\theta_i})^2$ and 
   $\frac{d^2}{d\tau_i d y_i} \ell = -1$.
-\item For the case of variance estimation, we have $\frac{d^2}{d\tau_i^2}\ell =
-  y^2_ie^{-h}$ and 
-  $\frac{d^2}{d\tau_i d y_i} \ell = -e^{-h}$.
+\item For the case of variance estimation, we have $\frac{d^2}{d\theta_i^2}\ell =
+  y^2_ie^{-\theta}$ and 
+  $\frac{d^2}{d\theta_i d y_i} \ell = -e^{-\theta}$.
 \end{enumerate}
@@
 Now we generalize to the regression setting. Define
-$\mu_i = x_i^\top \beta$. Furthermore, write
+$\theta_i = x_i^\top \beta$, $\beta \in \R^p$. And let
 \begin{equation}
-  \label{eq:4}
-  \hat\beta = \argmin_\tau \sum_{i=1}^n \phi(\mu_i(\beta)) - \langle y_i,\
-  \mu_i(\beta)\rangle + \lambda\norm{D\beta}_1.
+  \label{eq:7}
+   \hat\beta(y) = \argmin_\beta \sum_{i=1}^n \psi_i(x_i^\top \beta) -
+   w(y_i)\mu_i(x_i^\top \beta) + \lambda\norm{G\beta}_1,
 \end{equation}
@@
 \begin{theorem}
-  The divergence of $\hat\mu(y)$ is given by
+\label{thm:regression-divergence}
+  The divergence of $\hat\theta(y) := X\hat\beta(y)$ is given by
   \begin{equation}
     \label{eq:5}
-    \trace\left(D \hat\mu(y)\right) = -X_P \left(X^\top_P
-      \diag\left(\frac{d^2}{d\mu_i^2}\ell
-        \vert_{y_i,\hat{\mu}_i}\right) X_P\right)^\dagger X^\top_P
-      \diag\left(\frac{d^2}{d\mu_i d y_i} \ell \vert_{y_i,\hat{\mu}_i}\right),
+    \trace\left(D \hat\theta(y)\right) = -X_P \left(X^\top_P
+      \diag\left(\frac{d^2}{d\theta_i^2}\ell
+        \bigg\vert_{y_i,\hat{\theta}_i}\right) X_P\right)^\dagger X^\top_P
+      \diag\left(\frac{d^2}{d\theta_i d y_i} \ell \bigg\vert_{y_i,\hat{\theta}_i}\right),
     \end{equation}
-    where $\ell = \ell(y,\eta(\mu_i)) = \sum_{i=1}^n \phi(\eta_i(\mu_i)) - \langle y_i,\
-    \mu_i\rangle$, $\hat{\eta_i} = \eta_i(x_i^\top\hat\beta)$, and 
+    where 
     \begin{equation}
       \label{eq:6}
-      X_P =( I_n - D_S^\top (D_S^\top D_S)^\dagger D_S) X
+      X_P =( I_n - G_S^\top (G_S^\top G_S)^\dagger G_S) X
     \end{equation}
-    is the projection of $X$ onto the null-space of $D_S$.
+    is the projection of $X$ onto the null-space of $G_S$.
 \end{theorem}
@@
 
+\section{Risk estimation}
+\label{sec:risk-estimation}
+
+If $Y \sim \mbox{N}(\theta, \sigma^2)$, a now common method of risk
+estimation makes use of Stein's Lemma.
+\begin{lemma}[Stein's Lemma]
+  Assume $f(Y)$ is weakly differentiable with essentially
+  bounded weak partial derivatives on $\R^n$, then
+  \begin{equation}
+    \label{eq:8}
+    \trace \Cov(Y,f(Y)) = \Expect{\left\langle Y,\ f(Y)\right\rangle} = \sigma^2\Expect{\trace Df(Y) \bigg\vert_y }.
+  \end{equation}
+\end{lemma}
+The utility of this result comes from examining the decomposition of
+the mean squared error of $f(Y)$ as an estimator of $\theta$.
+\begin{align}
+  \label{eq:9}
+  \Expect{\norm{\theta-f(Y)}_2^2} 
+  &= \Expect{\norm{Y-f(Y)}} -n\sigma^2 + 2
+    \trace\Cov(Y,f(Y))\\
+  &= \Expect{\norm{Y-f(Y)}} -n\sigma^2 + 2\sigma^2
+    \Expect{\trace Df(y) \bigg\vert_Y} .
+\end{align}
+This characterization motivates the definition of degrees-of-freedom
+for linear predictors ($\textrm{df} :=\frac{1}{\sigma^2} \trace
+Df(y)\big\vert_Y$)~\citep{efron1986biased}, where $f(y)=Hy$. Using 
+Stein's Lemma, assuming $\sigma^2$ is known, we have Stein's Unbiased
+Risk Estimator
+\begin{equation}
+  \label{eq:10}
+  SURE_\theta = \norm{Y-f(Y)} -n\sigma^2 + 2\sigma^2\trace Df(y) \bigg\vert_y.
+\end{equation}
+Note that this is a risk for estimating the $n$-dimensional parameter $\theta$.
+
+The following result generalizes this idea to certain continuous
+exponential families.
+\begin{lemma}[Generalized Stein Lemma~\citep{Eldar2009}]
+  \label{thm:g-stein-lemma}
+  Assume $f(Y)$ is weakly differentiable with essentially
+  bounded weak partial derivatives on $\R^n$. Let $Y$ be distributed
+  according to a natural exponential family
+  \begin{equation}
+    \label{eq:11}
+    p(Y \given \beta) = \left(\prod_{i=1}^n h(y_i)\right)\exp\left\{
+    \sum_{i=1}^n y_i \theta_i(\beta_i) -\psi(\beta_i) \right\},
+  \end{equation}
+  and assume that $h$ is weakly differentiable and that $\theta_i$ is
+  one-to-one. Then,
+  \begin{equation}
+    \label{eq:12}
+    \Expect{\left\langle \theta(\beta),\ f(Y)\right\rangle} = - \Expect{
+      \left\langle\frac{\nabla h(Y)}{h(Y)},\ f(Y)\right\rangle + \trace Df(y) \bigg\vert_Y}.
+  \end{equation}
+Note that $\nabla h(Y)$ here means the vector $[d/dy h(y)
+\vert_{y_i}]$ and $h(Y)$ means the vector $[h(y_i)]$.
+\end{lemma}
+Therefore we define the Generalized SURE~\citep{Eldar2009} along the lines of the multivariate Gaussian case.
+\begin{lemma}
+  \label{lem:gsure}
+  Assume $h$ and $\nabla h$ are weakly differentiable, $\theta_i$ is one-to-one,
+  $f(Y)$ is weakly differentiable with essentially bounded partial
+  derivatives and that $p(Y\given\beta)$ is as in \eqref{eq:11}. Then
+  \begin{equation}
+    \label{eq:13}
+    GSURE_{\theta} = \norm{f(Y)}^2_2 + 2 \left\langle\frac{\nabla
+        h(Y)}{h(Y)},\ f(Y)\right\rangle + 2\trace Df(y) \bigg\vert_Y +
+    \frac{1}{h(Y)} \trace \frac{\partial^2 h(y)}{\partial y^2}\bigg\vert_Y
+  \end{equation}
+  is an unbiased estimator for the MSE of an estimator $f(Y)$ of
+  $\theta$: $\Expect{\norm{f(Y) - \theta(\beta)}_2^2}$.
+\end{lemma}
+
+\begin{proof}
+  We have
+  \begin{align}
+    \label{eq:14}
+    \Expect{\norm{f(Y) - \theta(\beta)}_2^2} 
+    &= \Expect{\norm{f(Y)}_2^2} + \Expect{\norm{\theta}_2^2} -
+      2\Expect{\left\langle \theta(\beta),\ f(Y)\right\rangle}.
+  \end{align}
+  Now, the first term is a function of the data only, and to the last term, we simply apply
+  \autoref{thm:g-stein-lemma}. For the second term,
+  \begin{align}
+    \Expect{\norm{\theta}_2^2} 
+    &=\Expect{\langle \theta,\ \theta \rangle}
+     = -\Expect{\left\langle\frac{\nabla h(Y)}{h(Y)},\ \theta\right\rangle}\\ 
+    &= \Expect{\left\langle\frac{\nabla h(Y)}{h(Y)},\
+      \frac{\nabla h(Y)}{h(Y)}\right\rangle} + \Expect{\trace
+      \frac{\partial}{\partial y} \frac{\nabla h(y)}{h(y)}
+      \bigg\vert_Y}\\
+    &= \Expect{\frac{\norm{\nabla h(Y)}_2^2}{h(Y)^2}} + \Expect{\trace
+      \frac{\norm{\nabla h(Y)}_2^2 +h(Y) \partial^2/\partial y^2 h(y)\big\vert_Y} 
+      {h(Y)^2}}\\ 
+    &=\Expect{ \frac{1}{h(Y)} \trace \frac{\partial^2 h(y)}{\partial y^2}\bigg\vert_Y},
+  \end{align}
+  by applying \autoref{thm:g-stein-lemma} twice along with the
+  quotient rule.
+\end{proof}
+
+\begin{theorem}
+  Let $Y_i \sim N(0, e^{x_i})$. Consider estimating $x$ by solving
+  \begin{equation}
+    \label{eq:15}
+    \min_h \sum_{i=1}^n x_i + y_i^2e^{-x_i} + \lambda \norm{Gx}_1.
+  \end{equation}
+  Then an unbiased estimator of $\Expect{\norm{\frac{1}{2e^x} -
+      \frac{1}{2e^{\hat x}}}}_2^2$ is given by
+  \begin{equation}
+    \label{eq:16}
+    \norm{\frac{1}{2e^{\hat x}}}_2^2 - \frac{1}{2}\left\langle
+      \frac{1}{y^2},\ \frac{1}{e^{\hat x}}\right\rangle + 2 \Pnd \left(\Pnd
+      \diag\left(y^2e^{-\hat x}\right) \Pnd\right)^\dagger \Pnd
+      \diag\left( e^{-\hat x}\right) + \frac{3}{4y^4},
+  \end{equation}
+where $\Pnd$ is as in \eqref{eq:3}.
+\end{theorem}
+
+\begin{proof}
+  Define $z_i := y_i^2$ and $\theta_i = -\frac{1}{2e^{x_i}}$. Then
+  \begin{align}
+    \label{eq:17}
+    p(z_i \given x_i) 
+    &= \frac{1}{\sqrt{2\pi z_i e^{x_i}}} \exp\left\{-\frac{1}{2}
+      \frac{z_i}{e^{x_i}}\right\}\one_{(0,\infty)}(z_i)\\
+    &= \frac{1}{\sqrt{\pi z}}\one_{(0,\infty)}(z_i) \exp\left\{z_i \cdot
+      \left(-\frac{1}{2e^{x_i}}\right) -
+      \left(-\frac{1}{2}\log\left(\frac{1}{2 e^{x_i}}\right)
+      \right)\right\}\\
+    &= \frac{1}{\sqrt{\pi z}}\one_{(0,\infty)}(z_i)
+      \exp\left\{z_i\cdot \theta_i - \psi(\theta_i)\right\},
+  \end{align}
+  which is a natural exponential family with $\psi(t) =
+  -\frac{1}{2}\log(-t)$ and $h(t) = \frac{1}{\sqrt{\pi t}}
+  \one_{(0,\infty)}(t)$. Therefore, $d/dt \log h(t) = -1/(2t)$ and
+  $(d^2/dt^2 h(t))/h(t) = 3 / (4t^2)$. Finally, we have $\frac{d^2}{d\theta_i^2}\ell =
+  y^2_ie^{-\theta}$ and 
+  $\frac{d^2}{d\theta_i d y_i} \ell = -e^{-\theta}$. The result
+  follows from \autoref{thm:simple-divergence} and \autoref{lem:gsure}.
+
+\end{proof}
+
+An alternate model selection technique is to examine the
+Kullback-Leibler Divergence between the density under
+$\theta=\hat\theta(Y)$ and that under $\theta=\theta$:
+$\Expect{KL(p(Y\given \hat\theta(Y)) \Vert p(Y\given \theta))}$. For
+exponential families, we have
+\begin{equation}
+\Expect{KL(\hat\theta(Y) \Vert 
+  \theta)} = \Expect{KL(p(Y\given \hat\theta(Y)) \Vert p(Y\given
+  \theta))} = \left\langle \hat\theta(Y)-\theta,\ \nabla
+  \psi(\hat\theta(Y))\right\rangle + \psi(\theta) - \psi(\hat\theta(Y)).
+\end{equation}
+
+An application of \autoref{thm:g-stein-lemma} thus provides an
+unbiased estimator of this quantity~\citep{Deledalle2017}.
+\begin{lemma} 
+  \label{lem:sukls}
+  Assume $h$ and $\nabla h$ are weakly differentiable, $\theta_i$ is one-to-one,
+  $f(Y)$ is weakly differentiable with essentially bounded partial
+  derivatives and that $p(Y\given\beta)$ is as in \eqref{eq:11}. Then
+  \begin{equation}
+    SUKLS = \left\langle \hat\theta + \frac{\nabla h(Y)}{h(Y)},\
+      \nabla \psi(\hat\theta(Y))\right\rangle + \trace Df(y) \bigg\vert_Y
+    - \psi(\hat\theta(Y))
+  \end{equation}
+  is unbiased for $\Expect{KL(\hat\theta(Y) \Vert
+  \theta)} - \psi(\theta)$.
+\end{lemma}
+
+\begin{corollary}
+   Let $Y_i \sim N(0, e^{x_i})$. Consider estimating $x$ by solving
+  \begin{equation}
+    \label{eq:15}
+    \min_h \sum_{i=1}^n x_i + y_i^2e^{-x_i} + \lambda \norm{Gx}_1.
+  \end{equation}
+  Then an unbiased estimator of $\Expect{KL(\hat x \Vert
+    x} - \psi(\theta)$ is given by
+  \begin{equation}
+    \label{eq:18}
+    -\frac{n}{2} - \left\langle \frac{1}{2y^2},\ e^{\hat
+        x}\right\rangle + 
+    -\frac{1}{2}\log\left(2e^{\hat
+        x_i}\right) + \Pnd \left(\Pnd
+      \diag\left(y^2e^{-\hat x}\right) \Pnd\right)^\dagger \Pnd
+      \diag\left( e^{-\hat x}\right). 
+  \end{equation}
+\end{corollary}
+\begin{proof}
+  As $\hat \theta_i = -\frac{1}{2e^{\hat x_i}}$, 
+  \[
+    \psi(\hat\theta_i) = -\frac{1}{2}\log\left(\frac{1}{2e^{\hat
+          x_i}}\right)
+  \]
+  and $d/d\theta \psi(\theta) = -\frac{1}{2\theta}$. Thus
+  \begin{align}
+    SUKLS 
+    &= \left\langle -\frac{1}{2e^{\hat x}} - \frac{1}{2y^2},\ e^{\hat
+      x}\right\rangle + \trace Df(y) \bigg\vert_Y
+      +\frac{1}{2}\log\left(\frac{1}{2e^{\hat
+      x_i}}\right)\\
+    &= -\frac{n}{2} - \left\langle \frac{1}{2y^2},\ e^{\hat
+      x}\right\rangle + \trace Df(y) \bigg\vert_Y
+      -\frac{1}{2}\log\left(2e^{\hat
+      x_i}\right).
+  \end{align}
+NOTE: I think some signs are wrong here.
+\end{proof}
+
 \bibliographystyle{mybibsty}

### 2018-08-30 aaai-2018/khodadadi-mcdonald-aaai19.tex
initial aaai version

--- 2018-05-18 nips-2018/paper.tex
@@
-\documentclass{article}
-\PassOptionsToPackage{sort,numbers,square}{natbib}
-
-\usepackage{nips_2018}
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
+%File: formatting-instruction.tex
+\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
+\usepackage{aaai19}  %Required
+\usepackage{times}  %Required
+\usepackage{helvet}  %Required
+\usepackage{courier}  %Required
+\usepackage{url}  %Required
+\usepackage{graphicx}  %Required
+\frenchspacing  %Required
+\setlength{\pdfpagewidth}{8.5in}  %Required
+\setlength{\pdfpageheight}{11in}  %Required
+%PDF Info Is Required:
+  \pdfinfo{
+/Title (Algorithms for Estimating Trends in Global Temperature Volatility)
+%/Author (Khodadadi and McDonald)
+/Author (John Doe, Jane Doe)}
+\setcounter{secnumdepth}{2}
+
+% \usepackage[bookmarks=false]{hyperref}       
+% \AtBeginDocument{\NoHyper} % turn off all links, colors
 \usepackage{booktabs}       % professional-quality tables
@@
 % For citations
-\usepackage{natbib}
+%\usepackage[sort,numbers,square]{natbib}
+
+\newcommand{\citet}[1]
+{\citeauthor{#1} ̃\shortcite{#1}}
+\newcommand{\citep}{\cite}
+\newcommand{\citealp}[1]
+{\citeauthor{#1} ̃\citeyear{#1}}
 
@@
 \usepackage{algorithmic}
-\usepackage{hyperref}
-\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}
+%\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}
 
 %\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-\usepackage{graphicx}
+\usepackage{cleveref}
+
 \usepackage{subfigure} 
-\graphicspath{{../nips-2018/}}
+\graphicspath{{../Figures/}}
 
@@
 
-\renewcommand*{\figureautorefname}{Figure}%
-\renewcommand*{\tableautorefname}{Table}%
-\renewcommand*{\sectionautorefname}{Section}%
-\renewcommand*{\subsectionautorefname}{Section}%
-\renewcommand*{\subsubsectionautorefname}{Section}%
-
 \newtheorem{theorem}{Theorem}%[section]
@@
 \newtheorem{lemma}[theorem]{Lemma}
+
+\newcommand{\autoref}{\Cref}
 
@@
 \usepackage{xr}
-\externaldocument{paper-supplement}
-
-\title{Modeling trend in temperature volatility using generalized LASSO}
-
-\author{Arash Khodadadi\\
- Department of Statistics\\
- Indiana University\\
- Bloomington, IN 47408 \\
- \texttt{arakhoda@indiana.edu}
-\And
-  Daniel J.\ McDonald \\
- Department of Statistics\\
- Indiana University\\
- Bloomington, IN 47408 \\
- \texttt{dajmcdon@indiana.edu}}
+\externaldocument{aaai-paper-supplement}
 
@@
 
+\title{Algorithms for Estimating Trends in Global Temperature Volatility}
+
+\author{Author 1 \ and Author 2\\
+Address line\\
+Address line
+}
+% \author{Arash Khodadadi \and Daniel J. McDonald\\
+%  Department of Statistics\\
+%  Indiana University\\
+%  Bloomington, IN 47408 \\
+%  \{arakhoda,dajmcdon\}@indiana.edu}
+
 \maketitle
@@
 \begin{abstract}
-In this paper, we present methodology for estimating trends in
-spatio-temporal volatility. We give two algorithms for computing our
-estimator which are tailored for dense, gridded observations over both
-space and time, though these can be easily extended to other
-structures (time-varying network flows, neuroimaging). We motivate our
-methodology by applying it to a massive climate dataset and
-discuss the implications for climate analysis.
-\end{abstract}
-
-\section{Introduction}
-
-Estimating smooth trends over time for large collections of time
-series is a common problem in economics, finance, meteorology,
-neuroscience and more. Most of this work focuses on analyzing trends
-(or removing trends) in the temporal average, but trends in variance
-can be more relevant, especially for financial data but also for
-climate science.
-
 Trends in terrestrial temperature variability are perhaps more relavant for species
 viability than trends in mean
-temperature~\citep{huntingford_no_2013},
-because, an increase in the
+temperature. In this paper, we develop methodology for estimating such
+trends using multi-resolution climate data from polar orbiting weather
+satellites. We derive two novel algorithms for computation that are
+tailored for dense, gridded observations over both 
+space and time. We evaluate our methods with a simulation constructed
+to mimic these data's 
+features and on a large, publicly available, global
+temperature dataset with the goal of tracking trends in
+cloud reflectance temperature variability.
+\end{abstract}
+
+\section{Introduction}
+
+% Estimating smooth trends over time for large collections of time
+% series is a common problem in economics, finance, meteorology,
+% neuroscience and more. Most of this work focuses on analyzing trends
+% (or removing trends) in the temporal average, but trends in variance
+% can be more relevant, especially for financial data but also for
+% climate science.
+
+The amount of sunlight reflected from clouds is among the largest
+sources of uncertainty in climate
+prediction~\citep{BoucherRandall2013}. But climate models fail to
+reproduce global cloud statistics and understanding the reasons for
+this failure is a grand challenge of the World Climate Research
+Programme~\citep{BonyStevens2015}.
+By understanding how cloudliness changes over time, examining changes
+in brightness temperature within various cloud types, and inferring
+changes in cloud radiative effects, atmospheric scientists can better model climate
+change. Modeling inter-satellite biases, satellite drift, and
+seasonal and long-term climate phenomena like El Ni\~{n}o will lead to
+a better understanding of climate
+change~\citep{SchreierKahn2014,BaumWielicki1994,BaumWielicki1992,FreyAckerman1996}.
+
+Numerous
+studies have examined the overall impacts of clouds on climate
+variability~\citep{MyersMechoso2018,GrisePolvani2013,BenderRamanathan2012},
+but such investigations have been hampered by the lack of a suitable
+dataset. Ideal data would have global coverage at high spatial
+resolution, a long enough record to recover temporal trends, and be
+multispectral~\citep{WielickiYoung2013}. The International Satellite
+Cloud Climatology Project (ISCCP)
+has been producing cloud property information from
+limited spectral channels for over three
+decades~\citep{RossowSchiffer1991}. These data have been used in 
+various climate variability
+studies~\citep[e.g.]{BenderRamanathan2012}, however, the utility of
+these data for long-term study of climate variability has been
+questioned~\citep{EvanHeidinger2007}.
+
+ISCCP provides a relatively long record, but it only incorporates
+radiances from limited visible and infrared channels from Advanced
+Very High Resolution Radiometer (AVHRR) instruments. Ongoing work seeks to create a
+more spectrally-detailed dataset which can avoid the above issues by
+combining radiance data from AVHRR imagers with readings from High-resolution
+Infrared Radiation Sounders on board legacy
+satellites~\citep{StatenKahn2016,SchreierKahn2010,KahnFishbein2007}. In
+anticipation of this new dataset, our work develops novel methodology
+for examining the trends in variability of climate data across space
+and time.
+
+\subsection{Variability Rather Than Average}
+\label{sec:vari-rath-than}
+
+Trends in terrestrial temperature variability are perhaps more relavant for species
+viability than trends in mean
+temperature~\cite{huntingford_no_2013},
+because an increase in
 temperature variability will increase the probability of extreme hot
-outliers~\citep{VasseurDeLong2014}. Recent climate literature
+outliers~\cite{VasseurDeLong2014}. Recent climate literature
 suggests that it is more difficult for society to adapt to these
 extremes than to the gradual increase in the mean temperature
-\citep{hansen_perception_2012,huntingford_no_2013}. Furthermore, the willingness of popular media to
+\cite{hansen_perception_2012,huntingford_no_2013}. Furthermore, the willingness of popular media to
 emphasize the prevalence extreme cold events coupled with a
@@
 short-term, localized behavior) leads to public misunderstanding of climate
-change. In fact, increased frequency of extreme cold events in the
-northern hemisphere can be partially attributed to increases in
-mean climate but is also due to increases in temperature
-variance~\citep{Screen2014,FischerBeyerle2013,TrenberthZhang2014}.
+change. In fact, a point of active debate is the extent to which the observed
+increased frequency of extreme cold events in the 
+northern hemisphere can be attributed to increases in temperature
+variance  rather than to increases in
+mean climate~\cite{Screen2014,FischerBeyerle2013,TrenberthZhang2014}.
 
 Nevertheless, research examining trends in the volatility of
-spatio-temporal climate data is scarce. \citep{hansen_perception_2012} studied the change in
+spatio-temporal climate data is scarce. \citet{hansen_perception_2012} studied the change in
 the standard deviation (SD) of the surface temperature in the NASA
@@
 that these estimates are increasing.
-\citep{huntingford_no_2013} took a similar
+\citet{huntingford_no_2013} took a similar
 approach in analyzing the ERA-40 data set. Their results showed that there
@@
 is much less than what is obtained from the method used in
-\citep{hansen_perception_2012}. The authors also computed the
+\cite{hansen_perception_2012}. The authors also computed the
 time-evolving global SD from the de-trended time-series at each
@@
 These and other related research, e.g.,
-\citep{rhines_frequent_2013}, have several shortcomings. First, no
+\citet{rhines_frequent_2013}, have several shortcomings. First, no
 statistical analysis has been performed to examine if the changes in
@@
 and most importantly, temporal and spatial correlations between the
-observations are ignored.
-
-\subsection{Related work}
-
-Variance estimation for financial time series has a lengthy history,
-focused especially on parametric models like the generalized
-autoregressive conditional heteroskedasticity (GARCH) process~\citep{engle2002dynamic} and
-stochastic volatility models~\citep{HarveyRuiz1994}. These models (and
-related AR processes) are specifically for parametric modelling of
-short ``bursts'' of high volatility, behavior typical of financial
-instruments. Parametric models for spatial data go back at least
-to~\citep{besag1974spatial} who proposed a conditional probability
-model on the lattice for examining plant ecology.
-
-More recently, nonparametric models for both spatial and temporal data
-have focused on using $\ell_1$-regularization for trend
-estimation. \citep{KimKoh2009} proposed $\ell_1$-trend filtering for
-univariate time series, which forms the basis of our methods. These
-methods have been generalized to higher order temporal smoothness
-~\citep{Tibshirani2014}, graph dependencies~\citep{WangSharpnack2016},
-and, most recently, small, time-varying
-graphs~\citep{HallacPark2017}. Our methodology is similar in flavor
-to~\citep{HallacPark2017}, though it uses a different likelihood
-function to emphasize variance estimation rather than trends in mean
-signal. Furthermore, the focus is in high-dimensional, regular data
-rather than a small number of changing graphs.
-
-\subsection{Main contributions}
+observations are completely ignored.
+
+In the present work, we examine variance (rather than the mean) for a number of reasons.
+First, instrument bias in the satellites increases over time so
+examining the mean over time conflates that bias with any actual
+change in mean (though the variance is unaffected). 
+Second, extreme weather events (hurricanes, droughts, wildfires in
+California, heatwaves in Europe) may be driven more strongly by increases
+in variance than by increases in mean. 
+Finally, even if the global mean temperature is constant, there may still
+be climate change. In fact, atmospheric physics suggests that,
+across space, average temperatures should not change (extreme cold
+in one location is offset by heat in another). But if swings across
+space are becoming more rapid, then, even with no change in mean global
+temperature over time, increasing variance can lead to
+increases in the prevalence of extreme events.
+
+\subsection{Main Contributions}
 
@@
 are then estimated by maximizing the likelihood of the observed
-data. We show that this formulation is not appropriate for
-detecting the trend, so, following ~\citep{Tibshirani2014}, we penalize the
+data. Following~\cite{Tibshirani2014}, we penalize the
 differences between the estimated variances which
 are temporally and spatially ``close'', resulting
-in a generalized LASSO problem. % However, the
-% dimension of this optimization problem is very high and so the
-% standard methods solvers are inadequate. We develop two algorithms
-% which are computationally feasible. In the first method, we adopt an optimization technique
-% called alternative direction method of multipliers (ADMM)
-% \citep{boyd_distributed_2011}, to divide the total problem into
-% several sub-problems of much lower dimension and show how the total
-% problem can be solved by iteratively solving these sub-problems. The
-% second method, called the \textit{linearized ADMM algorithm}
-% \citep{parikh_proximal_2014} solves the main problem by iteratively
-% solving a linearized version of it. We will compare the benefits of
-% each method.
+in a generalized LASSO problem. However, in our application, the
+dimension of this optimization problem is massive, and so the
+standard solvers are inadequate. We develop two algorithms
+which are computationally feasible. In the first method, we adopt an optimization technique
+called alternating direction method of multipliers
+(ADMM, \citealp{boyd_distributed_2011}), to divide the total problem into
+several sub-problems of much lower dimension and show how the total
+problem can be solved by iteratively solving these sub-problems. The
+second method, called the \textit{linearized ADMM algorithm}
+\cite{parikh_proximal_2014} solves the main problem by iteratively
+solving a linearized version of it. We will compare the benefits of
+each method.
 
@@
 \item We propose a model for nonparametric variance estimation for a
-  spatio-temporal process (\autoref{sec:ell_1-trend-filt}).
+  spatio-temporal process and discuss the relationship between our
+  methods and those existing in the machine learning literature (\autoref{sec:ell_1-trend-filt}).
 \item We derive two alternating direction method of multiplier
@@
   very large data (\autoref{sec:prop-optim-meth}). We give situations
-  under which each algorithm is most likely to be useful.
-\item We illustrate our methods on a large global temperature dataset
+  under which each algorithm is most likely to be useful. Open-source
+  Python code is available on {\tt github}.
+\item We illustrate our methods on a large, publicly available, global temperature dataset
   with the goal of tracking world-wide trends in variance as well as a
@@
 
-While we motivate and illustrate our methods on large, gridded climate
-data, we note that our algorithms are also applicable to neuroimaging
-data or large collections of financial instruments.
-
-\section{Estimating the variance of spatio-temporal data}
+While the motivation for our methodology is its application to large, gridded climate
+data, we note that our algorithms are also applicable to neuroimaging,
+image denoising, or examining large collections of financial instruments.
+
+\section{Smooth Spatio-temporal Variance Estimation}
 \label{sec:ell_1-trend-filt}
 
-$\ell_1$-trend filtering was proposed by \citep{KimKoh2009} as a
+\citet{KimKoh2009} proposed $\ell_1$-trend filtering as a
 method for estimating a smooth, time-varying trend. It is formulated
@@
 \end{equation}
- where $y_t$ is an observed time-series, $\beta$ is the smooth trend,
+ where $y=\{y_t\}_{t=1}^T$ is an observed time-series, $\beta\in \mathbb{R}^T$ is the smooth trend,
  $D$ is a $(T-2)\times T$ matrix, and $\lambda$ is a tuning parameter
@@
 With the penalty matrix $D$, the estimated $\beta$ will be piecewise
-linear. \citep{KimKoh2009} proposed a specialized primal-dual
+linear. \citet{KimKoh2009} proposed a specialized primal-dual
 interior point (PDIP) algorithm for solving \eqref{eq:l1tf}. From a
-statistical perspective, \eqref{eq:l1tf} is a constrained maximum
+statistical perspective, \eqref{eq:l1tf} can be viewed as a constrained maximum
 likelihood problem with independent observations from a normal
 distribution with common variance, $y_t \sim \mbox{N}(\beta_t,
-\sigma^2)$, subject to a piecewise linear constraint on $\beta$.
+\sigma^2)$, subject to a piecewise linear constraint on
+$\beta$. Alternatively, solutions to \eqref{eq:l1tf} are equivalent to
+maximum a posteriori Bayesian estimators based on Gaussian likelihood
+with a special Laplace prior distribution on $\beta$. Note that the
+structure of the estimator is determined by the penalty function
+$\lambda\norm{D\beta}_1$ rather than any parametric trend
+assumptions---autoregressive, moving average, sinusoidal seasonal
+component, etc. The resulting trend is therefore essentially
+nonparametric in the same way that splines are. In fact,
+using squared $\ell_2$ norm as the penalty instead of $\ell_1$ results exactly in
+regression splines.
 
@@
 
-\subsection{Estimating the variance}
+\subsection{Modifications for Variance}
 \label{sec:l1tf_var}
@@
 non-parametric model for estimating the variance of a time-series. To
-this end, we assume that at each time step $t$, there is a hidden
-variable $h_t$ such that conditioned on $h_t$ the observations $y_t$
+this end, we assume that at each time step $t$, there is a parameter
+$h_t$ such that the observations $y_t$
 are independent normal variables with zero mean and variance
@@
 model is $l(y\given h) \propto -\sum_{t=1}^T h_t - y_t^2e^{-h_t}$. Crucially,
-we assume that the hidden variables $h_t$ vary smoothly. To impose
+we assume that the parameters $h_t$ vary smoothly. To impose
 this assumption, we estimate $h_t$ by solving the penalized, negative
@@
 PDIP algorithm (as in, e.g.,
-\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}). In each iteration of
+\texttt{cvxopt}, \citealp{andersen_cvxopt:_2013}). In each iteration of
 PDIP we need to compute a search direction by taking a Newton
 step on a system of nonlinear equations. Due to space limitations, we
-defer details to Appendix A in the Supplement, where we show how to derive the dual of this optimization problem and compute the first and second derivatives of the dual objective function.
+defer details to Appendix A in the Supplement, where we show how to
+derive the dual of this optimization problem and compute the first and
+second derivatives of the dual objective function.
 
@@
 
-\subsection{Adding spatial constraints}
+\subsection{Adding Spatial Constraints}
 \label{sec:exten}
@@
 
-At a specific time $t$, the data is measured on a grid of points with
+At a specific time $t$, the data are measured on a grid of points with
 $n_r$ rows and $n_c$ columns for a total of $S=n_r\times n_c$ spatial
@@
 by solving the following optimization problem: 
-\begin{align}
-\min_h &\sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}}
-+\lambda_1 \sum_{i,j} \sum_{t=1}^{T-2} \left|h_{ijt}-2h_{ij(t+1)}+h_{ij(t+2)}\right|\\
-&\quad+\lambda_2 \sum_{t,j} \sum_{i=1}^{n_r-1} \left|h_{ijt}-h_{(i+1)jt}\right|
+{\scriptsize
+\begin{equation}
+\label{eq:l1tf_var_st}.
+\begin{aligned}
+&\min_h \sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}} +\lambda_1
+\sum_{i,j} \sum_{t=1}^{T-2}
+\left|h_{ijt}-2h_{ij(t+1)}+h_{ij(t+2)}\right|\\ 
+&+\lambda_2 \sum_{t,j} \sum_{i=1}^{n_r-1} \left|h_{ijt}-h_{(i+1)jt}\right|
 +\lambda_2 \sum_{t,i} \sum_{j=1}^{n_c-1} \left|h_{ijt}-h_{i(j+1)t}\right|
-\label{eq:l1tf_var_st}.
-\end{align}
+\end{aligned}
+\end{equation}
+}%
 
@@
 component of this
-penalty is a special case of trend filtering on graphs~\citep{WangSharpnack2016} which penalizes the difference between the estimated values of the signal on the connected nodes.
-As before, we can write \eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times S$ vector and $D$ is replaced by $D_{ST} \in \mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and spatial constraints, respectively\footnote{$N_s$ is obtained by counting the number of unique constraints at each location and at all times.}.
-The exact form of this matrix in clarified in Appendix A in the Supplement.
-Then, as we have two different tuning parameters for the temporal and spatial components, we write $\Lambda =\left[\lambda_1\one_{N_t}^\top,\;  \lambda_2\one_{N_s}^\top\right]^\top$ leading to:\footnote{Throughout the paper, we use $|x|$ for both scalars and vectors.
-For vectors we use this to denote a vector obtained by taking the absolute value of each entry of $x$.}
+penalty is a special case of trend filtering on
+graphs~\cite{WangSharpnack2016} which penalizes the difference between
+the estimated values of the signal on the connected nodes (though the
+likelihood is different). As before,
+we can write \eqref{eq:l1tf_var_st} in matrix form where $h$ is an
+$T\times S$ vector and $D$ is replaced by $D_{ST} \in
+\mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot
+(T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and
+spatial constraints, respectively\footnote{$N_s$ is obtained by
+  counting the number of unique constraints at each location and at
+  all times.}. The exact form of this matrix in clarified in Appendix
+A in the Supplement. Then, as we have two different tuning parameters
+for the temporal and spatial components, we write $\Lambda
+=\left[\lambda_1\one_{N_t}^\top,\;
+  \lambda_2\one_{N_s}^\top\right]^\top$ leading
+to:\footnote{Throughout the paper, we use $|x|$ for both scalars and
+  vectors. For vectors we use this to denote a vector obtained by
+  taking the absolute value of each entry of $x$.}  
 % let $h$ be a
@@
 
-\section{Proposed optimization methods}
+\subsection{Related Work}
+
+\attn{I've moved this here rather than above to focus the intro more
+  on the data.}
+
+Variance estimation for financial time series has a lengthy history,
+focused especially on parametric models like the generalized
+autoregressive conditional heteroskedasticity (GARCH) process~\cite{engle2002dynamic} and
+stochastic volatility models~\cite{HarveyRuiz1994}. These models (and
+related AR processes) are specifically for parametric modelling of
+short ``bursts'' of high volatility, behavior typical of financial
+instruments. Parametric models for spatial data go back at least
+to~\cite{besag1974spatial} who proposed a conditional probability
+model on the lattice for examining plant ecology.
+
+More recently, nonparametric models for both spatial and temporal data
+have focused on using $\ell_1$-regularization for trend
+estimation. \citet{KimKoh2009} proposed $\ell_1$-trend filtering for
+univariate time series, which forms the basis of our methods. These
+methods have been generalized to higher order temporal smoothness
+~\cite{Tibshirani2014}, graph dependencies~\cite{WangSharpnack2016},
+and, most recently, small, time-varying
+graphs~\cite{HallacPark2017}. Our methodology is similar in flavor
+to~\cite{HallacPark2017}, though it uses a different likelihood
+function to emphasize variance estimation rather than trends in mean
+signal. Furthermore, the focus is in high-dimensional, regular data
+rather than a small number of changing graphs.
+
+\attn{Paragraph discussing relationship with estimating time-varying covariance}
+
+\section{Optimization Methods}
 \label{sec:prop-optim-meth}
@@
 1^\circ$ grid over the entire northern hemisphere and daily data over
-10 years, we have $n_r=90$, $n_c=360$, $T=3650$ and so $D_{ST}$ has
+10 years, we have $S=90\times 360\approx 32,000$ spatial locations and
+$T=3650$ time points, and so $D_{ST}$ has
 approximately $10^8$ columns and $10^8$ rows. In each step of the PDIP
@@
 applying the PDIP directly is infeasible for our data.\footnote{We
-  note that this is a highly structured and sparse matrix, but, unlike
+  note that $D_{ST}$ is a highly structured, sparse matrix, but, unlike
   trend filtering alone, it is not banded. We are unaware of general
@@
 problem efficiently. The first casts the problem as a
-so-called consensus optimization problem~\citep{boyd_distributed_2011}
+so-called consensus optimization problem~\cite{boyd_distributed_2011}
 which solves smaller sub-problems using PDIP and then recombines the results. The
@@
 
-\subsection{Consensus optimization}
+\subsection{Consensus Optimization}
 \label{sec:consOpt}
@@
 constraint. Following the notation of 
-\citep{boyd_distributed_2011}, let $k=\mathscr{G}(i,j)$ which
+\cite{boyd_distributed_2011}, let $k=\mathscr{G}(i,j)$ which
 means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or
@@
 into several smaller sub-problems that can be solved independently in
-each iteration of optimization.
+each iteration.
 
@@
 \begin{aligned}
-\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)\\
- s.t. \,\,\, \tilde{h}_i=x_i. \,\,\,\,\,\,
+&\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i) &
+s.t. &\quad \tilde{h}_i=x_i.
 \end{aligned}
@@
   \centering
-  \includegraphics[height=.13\textheight]{Figures/data_cube}
+  \includegraphics[width=.6\columnwidth]{Figures/data_cube}
   \caption{The cube represents the global variable $h$ in space and
@@
 as a cube. We decompose $h$
-into sub-cubes as shown by white lines.
-Each local variable inside the sub-cubes corresponds to only one global variable.
-The local variables on the border (white lines), however, correspond to more than one global variable.
-With
+into sub-cubes as shown by white lines. Each global variable inside
+the sub-cubes corresponds to only one local variable. The global
+variables on the border (white lines), however, correspond to more
+than one local variable. With 
 this definition of $x_i$, the objective
@@
 $\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
-penalties corresponding to $x_i$ only in one sub-cube.
-Thus, we now need to use PDIP to solve $N$ problems each of size $n_i$, which is feasible for small enough $n_i$.
-\autoref{alg:conADMM} gives the general version of this algorithm.
-A more detailed discussion of this is in Appendix B of the Supplement where we show how to compute the dual and the derivatives of the augmented Lagrangian.
+penalties corresponding to $x_i$ only in one sub-cube along with its
+boundary. Thus, we now need to use PDIP to solve $N$ problems each of
+size $n_i$, which is feasible for small enough
+$n_i$. \autoref{alg:conADMM} gives the general version of this
+algorithm. A more detailed discussion of this is in Appendix B of the
+Supplement where we show how to compute the dual and the derivatives
+of the augmented Lagrangian.
 
@@
   \label{alg:conADMM}
-  \begin{algorithmic}
+  \begin{algorithmic}[1]
     \STATE {\bfseries Input:} data $y$, penalty matrix $D$, 
@@
     \REPEAT
-    \STATE $x_i\leftarrow\argmin_{x_i} -l(y_i\given
-    x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|$
-    \STATE $\quad\quad\quad+ (u_i)^\top x_i +
-    (\rho/2)  \lVert x_i-\tilde{h}_i \rVert_2^2$. \COMMENT{Update local
+    \STATE $\begin{aligned}x_i&\leftarrow\argmin_{x_i} -l(y_i\given
+    x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|\\&\quad+ (u_i)^\top x_i +
+    (\rho/2)  \lVert x_i-\tilde{h}_i \rVert_2^2\end{aligned}$. \\\COMMENT{Update local
       vars using PDIP}
@@
     \STATE $ u_i\leftarrow u_i + \rho (x_i-\tilde{h}_i)$. \COMMENT{Dual update}
-    \UNTIL {$\norm{h^{m+1}-h^m}_2^2\ < \epsilon$}
+    \UNTIL {$\max\left\{\norm{h^{m+1}-h^m},\ \norm{h^m-x^m}\right\} < \epsilon$}
     \STATE {\bf Return:} $h$.
@@
 % time is not feasible. 
-However, this algorithm is only useful if we can parallelize the
-computation over several machines. In the next section, we describe
+This algorithm is most useful if we can parallelize the
+computation over several machines with low communication cost between
+machines. In the next section, we describe 
 another algorithm which makes the computation feasible on a single
@@
 
-% In this section, we describe \textit{Linearized ADMM algorithm} \citep{parikh_proximal_2014} which, as we will see, makes the computation on a single machine feasible.
+% In this section, we describe \textit{Linearized ADMM algorithm} \cite{parikh_proximal_2014} which, as we will see, makes the computation on a single machine feasible.
 
@@
 iteration of the linearized ADMM
-algorithm~\citep{parikh_proximal_2014} for solving this problem 
+algorithm~\cite{parikh_proximal_2014} for solving this problem 
 has the form
-\begin{align}
+\begin{equation}
+  \label{eq:linADMM_steps}
+  \begin{aligned}
     x & \leftarrow \prox_{\mu f} \left(x - (\mu/\rho)D^\top (D x - z + u )\right)\\
-z & \leftarrow \prox_{\rho g} \left(z + u\right) \label{eq:linADMM_steps}\\
+    z & \leftarrow \prox_{\rho g} \left(z + u\right)\\
     u & \leftarrow u + D x - z
-\end{align}
-where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu < \rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is defined as $\prox_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \norm{ x-u}_2^2$.
-Proximal algorithms are feasible when these proximal operators can be evaluated efficiently which, as we show next, is the case for our problem.
+  \end{aligned}
+\end{equation}
+where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu <
+\rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is
+defined as $\prox_{\alpha \varphi}(u) = \min_x \,\, \alpha \cdot
+\varphi(x)+\frac{1}{2} \norm{ x-u}_2^2$. Proximal algorithms are feasible
+when these proximal operators can be evaluated efficiently which, as
+we show next, is the case for our problem.
 
@@
 \begin{lemma}
-  Let $f(h) = \sum_{k} h_{k} + y_{k}^2e^{-h_{k}}$ and $g(x) =
+  Let $f(x) = \sum_{k} x_{k} + y_{k}^2e^{-x_{k}}$ and $g(x) =
   \norm{x}_1$. Then, 
@@
   \end{align}
-where $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996},  $[S_{\alpha}(u)]_k = \sign(u_k)(|u_k| -\alpha_k)_+$ and
-$(v)_+=v\vee 0$.
+  where $\mathscr{W}(\cdot)$ is the \textit{Lambert W function}
+  \cite{corless_lambertw_1996},  $[S_{\alpha}(u)]_k = \sign(u_k)(|u_k|
+  -\alpha_k)_+$ and $(v)_+=v\vee 0$.
 \end{lemma}
-\begin{proof}
-  If $f(x)=\sum_k f_k(x_k)$ then $[\prox_{\mu f}(x)]_k =
-  \prox_{\mu f_k}(u_k)$. So 
-  $[\prox_{\mu f}(u)]_k=\min_{x_k} \,\,
-  \mu\big(x_k+y_{k}^2e^{-x_{k}}\big)+\frac{1}{2}  (x_k-u_k)^2.$
-  Setting the derivative to 0 and solving for $u_k$ gives the
-  result. Similarly, $[\prox_{\rho g}(u)]_\ell=\rho
-  \lambda_\ell |z_\ell|+1/2(z_\ell-u_\ell)^2$. This is not differentiable,
-  but the solution must satisfy $\rho \cdot \lambda_\ell \cdot \partial
-  \big(|z_\ell| \big)=u_\ell-z_\ell$ where $\partial \big(|z_\ell| \big)$ is the
-  sub-differential of $|z_\ell|$. The solution is the soft-thresholding
-  operator $S_{\rho\lambda_\ell}(u_\ell)$.
-\end{proof}
+% \begin{proof}
+%   If $f(x)=\sum_k f_k(x_k)$ then $[\prox_{\mu f}(x)]_k =
+%   \prox_{\mu f_k}(u_k)$. So 
+%   $[\prox_{\mu f}(u)]_k=\min_{x_k} \,\,
+%   \mu\big(x_k+y_{k}^2e^{-x_{k}}\big)+\frac{1}{2}  (x_k-u_k)^2.$
+%   Setting the derivative to 0 and solving for $u_k$ gives the
+%   result. Similarly, $[\prox_{\rho g}(u)]_\ell=\rho
+%   \lambda_\ell |z_\ell|+1/2(z_\ell-u_\ell)^2$. This is not differentiable,
+%   but the solution must satisfy $\rho \cdot \lambda_\ell \cdot \partial
+%   \big(|z_\ell| \big)=u_\ell-z_\ell$ where $\partial \big(|z_\ell| \big)$ is the
+%   sub-differential of $|z_\ell|$. The solution is the soft-thresholding
+%   operator $S_{\rho\lambda_\ell}(u_\ell)$.
+% \end{proof}
 
 Therefore, \autoref{alg:linADMM} gives a different method for solving
-the same problem.
+the same problem. For this algorithm, both the primal update and the
+soft thresholding step are performed elementwise at each point of the
+spatio-temporal grid. It can therefore be extremely fast to perform
+these steps. However, because there are now many more dual variables,
+this algorithm will require more outer iterations to achieve
+consensus. It therefore is highly problem and architecture dependent
+whether \autoref{alg:conADMM} or \autoref{alg:linADMM} will be more
+useful in any particular context.
 
@@
   \label{alg:linADMM}
-  \begin{algorithmic}
+  \begin{algorithmic}[1]
     \STATE {\bfseries Input:} data $y$, penalty matrix $D$, 
@@
     \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu
-      u_k}{\mu}\quad k=1,\ldots TS$. \COMMENT{Primal update}
+      u_k}{\mu}$ for all $k=1,\ldots TS$. \COMMENT{Primal update}
     \STATE $z\leftarrow S_{\rho\lambda}(u)$. \COMMENT{Elementwise soft thresholding}
     \STATE $u\leftarrow u + Dh-z$. \COMMENT{Dual update}
-    \UNTIL {$\max\{\norm{h-z}_2^2,\; \norm{z^{m+1}-z^m}_2^2\} < \epsilon$}
+    \UNTIL {$\max\{\norm{Dh-z},\; \norm{z^{m+1}-z^m}\} < \epsilon$}
     \STATE {\bf Return:} $z$.
@@
 
-\section{Empirical evaluation}
+\section{Empirical Evaluation}
 \label{sec:empirical-evaluation}
@@
 
-We generate observations at all time steps and all locations from
+Before examining real data, we apply our model to some synthetic
+data. This example was constructed to mimic the types of spatial and
+temporal phenomena observable in typical climate data. We generate a
+complete spatio temporal field wherein 
+observations at all time steps and all locations are
 independent Gaussian random variables with zero mean. However, the
 variance of these random variables follows a smoothly varying function
-in time and space
-
+in time and space given by the following parametric model:
 \begin{align}
-\sigma^2(t,r,c) & =\sum_{s=1}^{S} W_s(t) \cdot \exp\bigg( \frac{(r-r_s)^2+(c-c_s)^2}{2\sigma_s^2} \bigg); &
-W_s(t) & =\alpha_s \cdot t + \exp(\sin(2\pi\omega_s t+\phi_s)) .
+  \sigma^2(t,r,c) & =\sum_{k=1}^{K} W_k(t) \cdot \exp\bigg(
+                    \frac{(r-r_k)^2+(c-c_k)^2} {2\sigma_k^2} \bigg)\\
+  W_k(t) & =\alpha_k \cdot t + \exp(\sin(2\pi\omega_k t+\phi_k)) .
 \label{eq:sourceVar}
 \end{align}
-
 In words, the variance at each time and location is computed as the
-weighted sum of $S$ bell-shaped functions where the weights are
-time-varying, consist of a linear trend $\alpha_s \cdot t$ and a
-periodic term $\beta_s \cdot \sin(2\pi\omega_s t+\phi_s)$. The
-bell-shaped functions impose the spatial smoothness, and the linear
+weighted sum of $K$ bell-shaped functions where the weights are
+time-varying, consist of a linear trend $\alpha_k \cdot t$ and a
+periodic term $\beta_k \cdot \sin(2\pi\omega_k t+\phi_k)$. The
+bell-shaped functions impose spatial smoothness while the linear
 trend and the periodic terms enforce the temporal smoothness similar
 to the seasonal component in the real climate data. We simulated the
-data on a 5 by 7 grid and for 780 time steps with $S=4$. 
+data on a 5 by 7 grid and for 780 time steps with $K=4$. This yields a
+small enough problem so that it can be evaluated many times while
+still mimicking important properties of climate data.
 Specific parameter choices of the variance function are shown in
-\autoref{tab:sim_params} in the Supplement. For illustration, we also plot the
+\autoref{tab:sim_params}.
+For illustration, we also plot the
 variance function for all locations at $t=25$ and $t=45$ in as well as
-the variance across time at $(0,0)$ in \autoref{fig:true_var_spatial}
-in the Supplement.
+the variance across time at $(0,0)$ in \autoref{fig:true_var_spatial}.
+\begin{table}[tb]
+  \caption{Parameters used to simulate data.}
+  \label{tab:sim_params}
+  \begin{center}
+    \begin{tabular}{ccccccc}
+      \hline
+      $s$ & $r_s$ & $c_s$ & $\sigma_s$ &$\alpha_s$ & $\omega_s$ & $\phi_s$\\
+      \hline
+      1 & 0 & 0 & 5 & 0.5 & 0.121 & 0 \\
+      2 & 0 & 5 & 5 & 0.1 & 0.121 & 0 \\
+      3 & 3 & 0 & 5 & -0.5 & 0.121 & $\pi/2$ \\
+      4 & 3 & 5 & 5 & -0.1 & 0.121 & $\pi/2$ \\
+      \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\begin{figure}[tb]
+  \centering	
+  \includegraphics[height=.15\textheight]{Figures/true_var_spatial}\\
+  \includegraphics[height=.15\textheight]{Figures/true_fitted_var}
+  \caption{Top: Variance function at $t=25$ (left) and $t=45$
+    (right). Bottom: The true (orange) and estimated standard deviation 
+    function at the location (0,0). The estimated values are
+    obtained using linearized ADMM with $\lambda_s=0.1$ and two
+    values of $\lambda_t$: $\lambda_t=5$ (blue) and
+    $\lambda_t=100$ (green).} \label{fig:true_var_spatial}
+\end{figure}
 
@@
 algorithm takes 0.01 seconds on average while each iteration of the
-consensus ADMM takes about 20 seconds.
+consensus ADMM takes about 20 seconds. \attn{fix this discription}
 
@@
     values of $\lambda_t$: $\lambda_t=5$ (blue) and
-    $\lambda_t=100$ (green).
-Middle: Convergence speed of linearized and consensus ADMM.
-Right: MAE for four models: admm\_opt: the proposed model with optimal values of $\lambda_t$ and $\lambda_s$, admm\_temp: no spatial penalty, admm\_sp: no temporal penalty.} \label{fig:true_fitted_var}
+    $\lambda_t=100$ (green). Middle: Convergence speed of linearized
+    and consensus ADMM. Right: MAE for four models: admm\_opt: the
+    proposed model with optimal values of $\lambda_t$ and $\lambda_s$,
+    admm\_temp: no spatial penalty, admm\_sp: no temporal
+    penalty.\attn{Fix these}} \label{fig:true_fitted_var}
 \end{figure}
@@
 \autoref{fig:true_fitted_var} shows the boxplot of the MAE for these
-models. Interestingly, the proposed model with optimal parameters
+models. \attn{What is this next sentence trying to say?}
+Interestingly, the proposed model with optimal parameters 
 outperforms GARCH(1,1) in estimating the true value of the variance.
 
-\subsection{Data analysis}
+\subsection{Data Analysis}
 \label{sec:data}
@@
 applied this algorithm to the northern hemisphere of the ERA-20C
-dataset available from the \href {https://www.ecmwf.int}{European Center for Medium-Range
-  Weather Forecasts}. The data are
+dataset available from the European Center for Medium-Range
+Weather Forecasts\footnote{\url{https://www.ecmwf.int}}. 
 the 2 meter temperature measured daily at 12 p.m from January 1, 1960
-to December 24, 2010. The Supplement explains some preprocessing and investigates some
-properties of the time-series of different locations on
-earth. \autoref{fig:bloom_estimatedSD} shows a processed time-series for a
-single location.
-The variance of this time-series has an irregular cyclic
-behavior. Additionally, the time-series of other locations show 
-different patterns. These observations motivated the need to develop a
-non-parametric framework for this
-problem. \autoref{fig:bloom_estimatedSD} also shows the
-estimated SD obtained using the method of
-\autoref{sec:l1tf_var}. 
+to December 24, 2010.
+
+\paragraph{Preprocessing}
+
+\attn{Can we redo these but with, say, Indianapolis instead of
+  Bloomington?} Examination of the time-series alone demonstrates
+strong differences 
+between trend and cyclic behavior across spatial locations. One might
+try to model the cycles by the summation of 
+sinusoidal terms with different frequencies. However, for some
+time-series this may need many terms to be included in the summation
+to achieve a reasonable level of accuracy. In addition, such a model
+cannot capture the non-stationarity in the cycles.
+
+\autoref{fig:cities_ts} shows the time-series of the 
+temperature of three cities: Bloomington (USA), San Diego (USA) and
+Manaus (Brazil). The time-series of Bloomington and San Diego show
+clear cyclic behavior, though the amplitude of the cycles
+changes. The time-series of Manaus does not
+show any regular cyclic behavior. For this reason, we
+first apply trend filtering to remove seasonal terms and de-trend
+every time series. For each time-series, we found the optimal value of
+the penalty parameter using \textit{k-fold cross-validation} with
+$k=5$. We used the R package \textbf{genlasso} to perform these
+computations \cite{arnold_efficient_2016}.  
+\begin{figure}[tb]
+	\centering
+	\includegraphics[width=.2\textheight]{Figures/cities_ts}
+ 	\caption{Time-series of the temperature (in Kelvin) of three cities.}
+ 	\label{fig:cities_ts}
+\end{figure} 
+The blue curve in the left panel of Figure
+\autoref{fig:bloom_estimatedSD} shows the time-series of the
+temperature of Bloomington after detrending using this method.
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}
+  \caption{Left: The residuals of the time-series of Bloomington
+    (averaged weekly) and the estimated SD obtained from the method of
+    \autoref{sec:l1tf_var} (red). Middle: the estimated SDs (red) and
+    their annual average (black) without imposing the long horizon
+    penalty. Right: the same as middle panel but here the long horizon
+    penalty is imposed. See the text for more details.} 
+  \label{fig:bloom_estimatedSD}
+\end{figure}
+
+The red curve in the left panel of Figure \autoref{fig:bloom_estimatedSD}
+shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
+the time-series of Bloomington obtained from out proposed model. For
+ease of analysis, we compute the average
+of the estimated SD for each year. Both are shown in the middle panel of
+\autoref{fig:bloom_estimatedSD}. To smooth the annual trend, we add
+a long horizon penalty to \eqref{eq:l1tf_var}.
+The estimated, smoothed SDs are shown
+the right panel of \autoref{fig:bloom_estimatedSD}. The annual average
+of the estimated SDs shows a linear trend with a positive slope.
+
+% The Supplement explains some preprocessing and investigates some
+% properties of the time-series of different locations on
+% earth. \autoref{fig:bloom_estimatedSD} shows a processed time-series for a
+% single location.
+% The variance of this time-series has an irregular cyclic
+% behavior. Additionally, the time-series of other locations show 
+% different patterns. These observations motivated the need to develop a
+% non-parametric framework for this
+% problem. \autoref{fig:bloom_estimatedSD} also shows the
+% estimated SD obtained using the method of
+% \autoref{sec:l1tf_var}.
+
 % Since the estimated SD captures the periodic
@@
 
-\paragraph{Model selection}
+\paragraph{Model Selection}
 One common method for choosing the penalty parameters in the Lasso
@@
 generalized lasso problems
-\citep{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}. However,
+\cite{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}. However,
 all these studies have considered the linear regression problem and,
@@
 
-\paragraph{Analysis of trend in temperature volatility}
+\paragraph{Analysis of Trends in Temperature Volatility}
 
@@
   \centering
-  \includegraphics[width=.8 \columnwidth]{Figures/ts_estimatedVar}\\
+  \includegraphics[width=.8\columnwidth]{Figures/ts_estimatedVar}
   \caption{Detrended data and the estimated SD for
-    a small midwestern city (left) and San Diego (right).} 
+    a small midwestern city (left) and San Diego (right).\attn{Fix these}} 
   \label{fig:avg_change_estimatedSD}
@@
   \centering
-  \includegraphics[width=.45\linewidth]{Figures/avg_logVar.pdf}
-  \includegraphics[width=.45\linewidth]{Figures/avg_change_logVar.pdf}
+  \includegraphics[width=.9\linewidth]{Figures/avg_logVar.pdf}\\
+  \includegraphics[width=.9\linewidth]{Figures/avg_change_logVar.pdf}
   \caption{The
@@
 
-It is interesting to note that the trend in volatility is almost zero
-over the oceans. The most positive trend can be observed in Asia and
-particularly in south-east Asia.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=.3\linewidth]{Figures/hist_avg_change.pdf}
-  \caption{The histogram of changes in estimated SD.}
-  \label{fig:hist}
-\end{figure}
-
-\autoref{fig:hist} shows the
-histogram of change in the estimated SD across the northern
-hemisphere. The SD in most locations on the northern
+The SD in most locations on the northern
 hemisphere had a negative trend in this time period, though spatially,
@@
 regions display fewer days of extreme cold.
+% It is interesting to note that the trend in volatility is almost zero
+% over the oceans. 
+The most positive trend can be observed in Asia and
+particularly in south-east Asia.
+
+% \begin{figure}[tb]
+%   \centering
+%   \includegraphics[width=.9\linewidth]{Figures/hist_avg_change.pdf}
+%   \caption{The histogram of changes in estimated SD.}
+%   \label{fig:hist}
+% \end{figure}
+
+% \autoref{fig:hist} shows the
+% histogram of change in the estimated SD across the northern
+% hemisphere. The SD in most locations on the northern
+% hemisphere had a negative trend in this time period, though spatially,
+% this decreasing pattern is localized mainly toward the extreme
+% northern latitudes and over oceans. In many ways, this is consistent
+% with climate change predictions: oceans tend to operate as a local
+% thermostat, regulating deviations in local temperature, while warming polar
+% regions display fewer days of extreme cold.
 
 \section{Discussion}
+\label{sec:discussion}
+
 In this paper, we proposed a new method for estimating the variance of
@@
 
-\newpage
-\bibliographystyle{abbrvnat}
-\small
-\bibliography{nips-references}
+\fontsize{9.0pt}{10.0pt} \selectfont
+\bibliographystyle{aaai}
+\bibliography{aaai-references}
+\clearpage
+
+\appendix
+\section{PDIP for $\ell_1$ Trend Filtering}
+\label{sec:app_l1tf_var}
+
+In this appendix we provide more details on how to solve the optimization problem \ref{eq:l1tf_var} using PDIP.
+The objective function is convex but not differentiable.
+Therefore, to be able to use PDIP we first need to derive the dual of this problem.
+We note that this is a generalized LASSO problem \citep{TibshiraniTaylor2011}.
+The dual of a generalized LASSO with the objective $f(x)+\lambda \norm{ Dx }_1$ is:
+
+\begin{align}
+\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
+\end{align}
+
+where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that
+
+\begin{equation}
+f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
+\label{eq:conj}
+\end{equation}
+
+Each iteration of PDIP involves computing a search direction by taking a Newton step for the system of nonlinear equations $r_w(v,\mu_1,\mu_2)=0$, where $w>0$ is a parameter and
+
+\begin{equation}
+  r_w(v,\mu_1,\mu_2):=
+	\begin{bmatrix}
+	r_{dual}\\
+	r_{cent}	
+	\end{bmatrix}=
+  \begin{bmatrix}
+    \nabla f^*(-D^\top v) + \mu_1 - \mu_2\\
+    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
+  \end{bmatrix}
+\label{eq:resid}
+\end{equation}
+
+for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form
+
+\begin{equation}
+r_w(v,\mu_1,\mu_2)+A
+\begin{bmatrix}
+	\nabla v\\
+	\nabla \mu_1\\
+	\nabla \mu_2	
+	\end{bmatrix}= 0
+\label{eq:newton_step}
+\end{equation}
+
+We have:
+
+\begin{equation}
+A=
+\begin{bmatrix}
+	\nabla^2 f^*(-D^\top v) & I & -I\\
+	-\mathbf{diag(\mu_1)}\one & -v+\lambda\one & \mathbf{0}\\
+	\mathbf{diag(\mu_2)}\one & v+\lambda\one & \mathbf{0}
+	\end{bmatrix}
+\label{eq:delta_r}
+\end{equation}
+
+Therefore, to perform the Newton step we need to compute $\nabla f^*(-D^\top v)$ and $\nabla^2 f^*(-D^\top v)$. It is straightforward to show that
+
+\begin{align}
+\nabla f^*(-D^\top v) = -\nabla_u f^*(u) D^\top ,\quad u=-D^\top v, \quad (\nabla_u f^*(u))_j=\log\bigg(\frac{y_j^2}{1-u_j}\bigg) \\
+\nabla^2 f^*(-D^\top v)=D\nabla_u^2 f^*(u)D^\top, \quad (\nabla_u^2 f^*(u))_j=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg)
+\end{align}
+
+The matrix $D$ in these equations have the following form:  $D_{total}^t=[D_{t}^t|D_{s}^t]$. The matrix $D_t$ is the following block-diagonal matrix and corresponds to the temporal penalty: 
+ \begin{equation}
+ D_t=\begin{bmatrix}
+ D &  & \\ 
+  & \ddots & \\
+  &  & D
+ \end{bmatrix}
+ \label{eq:d_t_matrix}
+ \end{equation}
+
+where $D$ was defined in Section 2 of the main text.
+The number of the diagonal blocks is equal to the grid size $n_r \times n_c$.
+Each row of the matrix $D_s$ corresponds to one spatial constraint in Equation \eqref{eq:l1tf_var_st} in the text.
+For example, the first $T$ rows correspond to $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to $|h_{11t}-h_{12t}|$, and so on.
+
+\section{PDIP Update in~\autoref{alg:conADMM}}
+\label{sec:app_consADMM}
+
+In this section we give more details on performing the $x$-update step in Algorithm 1. We need to solve the following optimization problem:
+
+\begin{equation}
+\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
+\label{eq:x_update_opt}
+\end{equation}
+
+\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$.
+Let $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$.
+As it was explained in \autoref{sec:l1tf_var}, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$.
+So to use PDIP we first need to compute the conjugate function $f^*(\cdot)$.
+We have:
+
+\begin{align}
+f^*(\xi) & = \max_x \quad \xi^\top x - f(x)\\
+& =  \max_x \quad \sum_{j=1}^{n_b} (\xi_jx_j - x_j - y_j^2e^{-x_j} - (\rho/2)(x_j-\tilde{z}_j+u_j))
+\label{eq:conjugate}
+\end{align}
+
+Setting the derivative of the terms inside the summation to 0, we obtain:
+
+\begin{equation}
+\xi_j-y_j^2e^{-x_j^*}-\rho x_j^* + \rho (\tilde{z}_j-u_j)=0
+\label{eq:x_start}
+\end{equation}
+
+\noindent where $x^*$ is the maximizer in \ref{eq:conjugate}. Then, it can be shown that $x_j^*$ which satisfies \eqref{eq:x_start} can be obtained as follows:
+
+\begin{align}
+x^*_j & = \mathscr{W}\bigg(\frac{y_j^2}{\rho} e^{\phi_j} \bigg) - \phi_j \\
+\phi_j & =\frac{1-\xi_j-\rho(\tilde{z}_j-u_j)}{\rho}
+\end{align}
+
+In this equation, $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996}.
+Finally, the conjugate function is: $f^*(\xi) = \sum_{j=1}^{n_b} (\xi_jx^*_j - x^*_j - y_j^2e^{-x^*_j} - (\rho/2)(x^*_j-\tilde{z}_j+u_j))$.
+
+To use PDIP, we also need to evaluate $\nabla f^*$ and $\nabla^2 f^*$.
+First note that $\frac{\partial \mathscr{W}(q)}{\partial q} = \frac{\mathscr{W}(q)}{q(1+\mathscr{W}(q))}$ and $\frac{\partial^2 \mathscr{W}(q)}{\partial q^2} = - \frac{\mathscr{W}^2(q)(\mathscr{W}(q)+q)}{q^2(1+\mathscr{W}(q))^3}$.
+Using the chain rule we get:
+
+%\x^*_j + \xi_j + \frac{\partial \x^*_j}{\partial \xi_j}
+
+\begin{equation}
+\frac{\partial f^*(\xi)}{\partial \xi_j}  =  x^*_j  + \frac{\partial x^*_j}{\partial \xi_j} \bigg[ \xi_j -1 + y_j^2 e^{-x_j^*} + \rho (\tilde{z}_j - u_j - x_j^*) \bigg]
+\label{eq:d_f*_start}
+\end{equation}
+where we have:
+
+\begin{equation}
+\frac{\partial x_j^*}{\partial \xi_j}  = \frac{1}{\rho(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))}
+\label{eq:d_x*_start}
+\end{equation}
+
+By some tedious but straightforward computation we can obtain the second derivatives:
+
+\begin{align}
+\frac{\partial^2 f^*(\xi)}{\partial \xi_j^2} & =  \frac{\partial x_j^*}{\partial \xi_j} - \rho \frac{\partial^2 x_j^*}{\partial \xi_j^2} \bigg[ \phi_j +x_j^* - \tilde{z}_j + u_j \bigg]\\
+& \quad + \frac{\partial x_j^*}{\partial \xi_j} \bigg[ 1-y_j^2 \frac{\partial x_j^*}{\partial \xi_j} e^{-x_j^*} -\rho \frac{\partial x_j^*}{\partial \xi_j} \bigg]
+\label{eq:d2_f*_start}
+\end{align}
+
+\begin{equation}
+\frac{\partial^2 x_j^*}{\partial \xi_j^2}  = \frac{\mathscr{W}((y_j^2/\rho) e^{-\phi_j} )}{\rho^2(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))^3}
+\label{eq:d2_x*_start}
+\end{equation}
+
+Having computed the conjugate function and its gradient and Jacobian, now we can use a number of convex optimization software packages which have an implementation of PDIP to perform the x-update step inside the ADMM loop.
+We chose the python API of the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.
+
+% \section{Appendix C}
+% \autoref{tab:sim_params} lists the parameters used for simulating data
+% in \autoref{sec:simulations}. \autoref{fig:true_var_spatial} shows the
+% variance function obtained from there parameters at $t=25$ (left) and
+% $t=45$ (center).
+
+% \begin{table}[tb]
+%   \caption{Parameters used to simulate data.}
+%   \label{tab:sim_params}
+%   \begin{center}
+%     \begin{tabular}{ccccccc}
+%       \hline
+%       $s$ & $r_s$ & $c_s$ & $\sigma_s$ &$\alpha_s$ & $\omega_s$ & $\phi_s$\\
+%       \hline
+%       1 & 0 & 0 & 5 & 0.5 & 0.121 & 0 \\
+%       2 & 0 & 5 & 5 & 0.1 & 0.121 & 0 \\
+%       3 & 3 & 0 & 5 & -0.5 & 0.121 & $\pi/2$ \\
+%       4 & 3 & 5 & 5 & -0.1 & 0.121 & $\pi/2$ \\
+%       \hline
+%     \end{tabular}
+%   \end{center}
+% \end{table}
+
+% \begin{figure}[tb]
+%   \centering	
+%   \includegraphics[height=.15\textheight]{Figures/true_var_spatial}
+%   \includegraphics[height=.15\textheight]{Figures/true_fitted_var}
+%   \caption{Variance function at $t=25$ (left) and $t=45$
+%     (center). Right: the true (orange) and estimated standard deviation 
+%     function at the location (0,0). The estimated values are
+%     obtained using linearized ADMM with $\lambda_s=0.1$ and two
+%     values of $\lambda_t$: $\lambda_t=5$ (blue) and
+%     $\lambda_t=100$ (green).} \label{fig:true_var_spatial}
+% \end{figure}
+
+\section{Data Exploration}
+
+In this appendix we examine some of the properties of the time-series
+of the temperature in ERA-20C dataset. The goal here is to demonstrate
+some of the difficulties in modeling the trend in the temperature
+volatility and motivate our methodology.
+
+Figure \autoref{fig:cities_ts} shows the time-series of the
+temperature of three cities: Bloomington (USA), San Diego (USA) and
+Manaus (Brazil). The time-series of Bloomington and San Diego show
+clear cyclic behavior. However, while it seems (qualitatively) that
+these cycles can be modeled by a sinusoidal function for Bloomington,
+the same is not true for San Diego. Also, the amplitude of the cycles
+changes from some years to others. The time-series of Manaus does not
+show any regular cyclic behavior. This demonstrates the first
+difficulty in analyzing the variance of this data: to analyze the
+variance, we first need to remove the cyclic terms from all
+time-series. However, there is a lot of variations in the cyclic
+behavior of the time-series of different locations. In addition, some
+of these cycles cannot be easily modeled by a parametric function
+\footnote{One might try to model the cycles by the summation of
+  sinusoidal terms with different frequencies. However, for some
+  time-series this may need many terms to be included in the summation
+  to achieve a reasonable level of accuracy. In addition, this model
+  cannot capture the non-stationarity in the cycles.}. To overcome
+these issues, we use a non-parametric approach to remove the cyclic
+terms from the time-series and de-trend them. This approach, called
+\textit{$\ell_1$-trend filtering} is explained in Section 2 of the
+text. We detrended each time-series separately using this method. For
+each time-series, we found the optimal value of the penalty parameter
+using \textit{k-fold cross-validation} with $k=5$. We used the R
+package \textbf{genlasso} to perform these computations
+\cite{arnold_efficient_2016}.
+
+\begin{figure}[tb]
+	\centering
+	\includegraphics[width=.2\textheight]{Figures/cities_ts}
+ 	\caption{Time-series of the temperature (in Kelvin) of three cities.}
+ 	\label{fig:cities_ts}
+\end{figure}
+
+The blue curve in the left panel of Figure \autoref{fig:bloom_estimatedSD} shows the time-series of the temperature of Bloomington after detrending using this method.
+This figure, reveals another difficulty in estimating the trend of volatility in this data: the variance of this signal, shows cyclic behavior.
+Also, the cycles are not regular and their amplitude and frequency change.
+Even if one can describe the behavior of the variance of the time-series at all locations using a single parametric model (for example a variant of the GARCH models \citep{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework.
+These observations motivate the need to develop a non-parametric framework for the problem at hand.
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}
+  \caption{Left: The residuals of the time-series of Bloomington
+    (averaged weekly) and the estimated SD obtained from the method of
+    \autoref{sec:l1tf_var} (red). Middle: the estimated SDs (red) and
+    their annual average (black) without imposing the long horizon
+    penalty. Right: the same as middle panel but here the long horizon
+    penalty is imposed. See the text for more details.} 
+  \label{fig:bloom_estimatedSD}
+\end{figure}
+
+The red curve in the left panel of Figure \autoref{fig:bloom_estimatedSD}
+shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
+the time-series of Bloomington obtained from out proposed model.
+To reduce the number of time-steps we work on the weekly averaged of the data.
+The curve of the estimated SD captures the periodic variations in the
+SD of the signal. Just by looking at this curve, it is hard to say if
+the SD is decreasing or increasing. Therefore, we compute the average
+of the estimated SD for each year. The estimated SD together with this
+annual average is shown in the middle panel of
+\autoref{fig:bloom_estimatedSD}. As it can be seen, the annual trend
+is not smooth. This is because in the optimization problem
+\eqref{eq:l1tf_var}, the smoothness of the annual trend is not
+encouraged. To remedy this, we add the following long horizon penalty to \eqref{eq:l1tf_var}:
+
+\begin{equation}
+\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
+\label{eq:lh_penalty}
+\end{equation}
+
+where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$. Also,
+$N_{year}$ is the number of years over which we are performing our
+analysis (here $N_{year}=10$). Since we are working on the weekly
+averaged data, each year corresponds to 52 observations. In the matrix
+form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the
+matrix $D$. The estimated SDs using this penalty matrix is shown in
+the right panel of \autoref{fig:bloom_estimatedSD}. The annual average
+of the estimated SDs shows a linear trend with a positive slope.
 

### 2018-08-30 nips-2018/nips-response.tex
initial aaai version

\documentclass[11pt]{article}
% Include statements
\usepackage{graphicx}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage[numbers,square]{natbib}
\usepackage[left=1in,top=1in,right=1in,bottom=1in,nohead]{geometry}
\usepackage{pdfsync}
\usepackage{hyperref}
\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}
\renewcommand{\qedsymbol}{$\blacksquare$}

% Bibliography
\bibliographystyle{mybibsty}

% Theorem environments
\newcommand{\email}[1]{\href{mailto:#1}{#1}}

\title{Author response: NIPS 2018, Paper 3656}
\author{Modeling trend in temperature volatility using generalized
  LASSO}
\date{\today}

\begin{document}
\noindent {\bf Author response: NIPS 2018, Paper 3656\\}
Modeling trend in temperature volatility using generalized
  LASSO\\
\today

\hspace{10pt}

We wish to thank the reviewers for their careful and well-informed
comments. We are especially grateful for the literature suggestions
given by reviewer 1. The Gibberd et al.\ (2017) and Monti et al.\ (2014) papers especially deserve
mention in our paper, and neither was familier to us. Both take a similar approach to ours, and while
different in a number of key ways, deserve a more careful discussion
than can be provided here. In particular, we should, at minimum, try
their algorithms or give reasons why we cannot. However, we feel these papers provide
a good launching point to discuss the main
issue, mentioned by all three reviewers.

\paragraph{Main issue} All three reviewers felt that this paper lacked sufficient novelty
relative to existing literature. The novelty of our paper is best understood as an attempt to
apply a slightly modified $\ell_1$-trend filter to a very large and
very important dataset. It is certainly true that $\ell_1$
regularization is everywhere, as is trend filtering, and variance
estimation under such a penalty. In that sense, our paper is not
particularly new. However, we found that the size of our problem
(coupled with a non-quadratic likelihood function) meant that standard
methods were computationally infeasible. The time-varying networks
used in Monti, Hallac et al.\ (2017), and Gibberd, have on the order of hundreds of
time points and dozens to hundreds of nodes. Our data have (see lines
100--105) 3500 time points and ~25,000 nodes. It is this massive
increase in throughput which makes our methods both relevant and
novel. Even building the penalty matrix (or loading the data) is
relatively infeasible. This is also the reason that our simulation
size is relatively paltry: to enable many replications performed
quickly. Note however that Gibberd uses 10 nodes and 50 timepoints in
their simulations, so our example with 35 nodes and 780 timepoints is
not that out of the ordinary.

\paragraph{Why variance} Finally, there was some confusion as to why the variance would be
important here (rather than the mean). This is for a number of reasons
which we tried to articulate carefully in the introduction.
\begin{enumerate}
\item Instrument bias in the satellite increases over time so
  examining the mean over time conflates that bias with any actual
  change in mean (though the variance is unaffected).
\item Extreme weather events (hurricanes, droughts, wildfires in
  California, heatwaves in Europe) are driven more strongly by
  increases in variance than by increases in mean.
\item Even if the global mean temperature is constant, there may still
  be climate change. In fact, atmospheric physics suggests that,
  across space, average temperatures should not change (extreme cold
  in one location is offset by heat in another). But if swings across
  space are becoming more rapid, then even with no change in mean global
  temperature over time, the variance is increasing, again leading to
  increases in extreme events.
\end{enumerate}

\paragraph{Minor issues} We will of course be careful to address typos
(thank you for pointing them out!) as well as attempt to clarify
points of the paper which were less precise than they should have
been. And, really, thank you for your reviews.

\bibliography{AllReferences.bib}
\end{document}

### 2018-09-03 aaai-2018/khodadadi-mcdonald-aaai19.tex
update paper

@@
 %PDF Info Is Required:
-  \pdfinfo{
-/Title (Algorithms for Estimating Trends in Global Temperature Volatility)
+
+%\pdfinfo{
+%	/Title (Algorithms for Estimating Trends in Global Temperature Volatility)
 	%/Author (Khodadadi and McDonald)
-/Author (John Doe, Jane Doe)}
+%	/Author (John Doe, Jane Doe)}
+
 \setcounter{secnumdepth}{2}
@@
 % climate science.
+
+\textcolor{red}{How is this first part related to the rest of the paper?}
 
@@
 and, most recently, small, time-varying
-graphs~\cite{HallacPark2017}. Our methodology is similar in flavor
-to~\cite{HallacPark2017}, though it uses a different likelihood
+graphs~\cite{HallacPark2017}.
+
+\textcolor{red}{Our methodology is similar in flavor to~\cite{HallacPark2017}, with several fundamental differences.
+The main goal of Hallac et al.
+is to discover the time-varying structure of a network.
+To achieve this goal, they introduced two penalty terms into the objective function.
+One term penalizes the sum of the absolute values of the off-diagonal elements of the inverse covariance matrix.
+This term enforces sparsity on the estimated inverse covariance matrix and so helps in discovering which nodes in the network are connected.
+The other penalty term, penalizes differences in the estimated inverse covariance matrix from one time-step to the next.
+Our goal in the present work is to detect the temporal trend in the variance of each point in the network.
+The network in our problem is the grid of points over the earth and so the network's structure remains the same through time.
+The variance of each point, however changes.
+To capture these properties in our data using the objective function used in Hallac et al.
+(see Equation 2 in their paper), we need to have different temporal penalty terms for the diagonal and off-diagonal elements in the inverse covariance matrix.
+This results in unnecessary complication in the optimization problem.
+To overcome these difficulties, we introduced the new objective function \ref{eq:l1tf_var_st_mat}.
+This leads to a new optimization problem which we consider its solution next.}
+
+though it uses a different likelihood
 function to emphasize variance estimation rather than trends in mean
@@
 
+\attn{Arash: Can you be more specific?}
+
 \begin{figure}[tb]
@@
 models. \attn{What is this next sentence trying to say?}
+
+\attn{that our model is better than GARCH in estimating the variance.}
+
 Interestingly, the proposed model with optimal parameters 
@@
 \autoref{fig:cities_ts} shows the time-series of the 
-temperature of three cities: Bloomington (USA), San Diego (USA) and
-Manaus (Brazil). The time-series of Bloomington and San Diego show
+temperature of three cities: Indianapolis (USA), San Diego (USA) and
+Manaus (Brazil). The time-series of Indianapolis and San Diego show
 clear cyclic behavior, though the amplitude of the cycles
@@
 \autoref{fig:bloom_estimatedSD} shows the time-series of the
-temperature of Bloomington after detrending using this method.
+temperature of Indianapolis after detrending using this method.
 
@@
   \includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}
-  \caption{Left: The residuals of the time-series of Bloomington
+  \caption{Left: The residuals of the time-series of Indianapolis
     (averaged weekly) and the estimated SD obtained from the method of
@@
 shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
-the time-series of Bloomington obtained from out proposed model. For
+the time-series of Indianapolis obtained from out proposed model. For
 ease of analysis, we compute the average
@@
   \centering
-  \includegraphics[width=.8\columnwidth]{Figures/ts_estimatedVar}
+  \includegraphics[width=1\columnwidth]{Figures/ts_estimatedVar}
   \caption{Detrended data and the estimated SD for
@@
 \begin{align}
-\nabla f^*(-D^\top v) = -\nabla_u f^*(u) D^\top ,\quad u=-D^\top v, \quad (\nabla_u f^*(u))_j=\log\bigg(\frac{y_j^2}{1-u_j}\bigg) \\
-\nabla^2 f^*(-D^\top v)=D\nabla_u^2 f^*(u)D^\top, \quad (\nabla_u^2 f^*(u))_j=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg)
+& \nabla f^*(-D^\top v) =  -\nabla_u f^*(u) D^\top ,\\
+& \quad u=-D^\top v, \quad (\nabla_u f^*(u))_j=\log\bigg(\frac{y_j^2}{1-u_j}\bigg) \\
+& \nabla^2 f^*(-D^\top v)=D\nabla_u^2 f^*(u)D^\top, \quad (\nabla_u^2 f^*(u))_j=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg)
 \end{align}
@@
 Figure \autoref{fig:cities_ts} shows the time-series of the
-temperature of three cities: Bloomington (USA), San Diego (USA) and
-Manaus (Brazil). The time-series of Bloomington and San Diego show
+temperature of three cities: Indianapolis (USA), San Diego (USA) and
+Manaus (Brazil). The time-series of Indianapolis and San Diego show
 clear cyclic behavior. However, while it seems (qualitatively) that
-these cycles can be modeled by a sinusoidal function for Bloomington,
+these cycles can be modeled by a sinusoidal function for Indianapolis,
 the same is not true for San Diego. Also, the amplitude of the cycles
@@
 
-The blue curve in the left panel of Figure \autoref{fig:bloom_estimatedSD} shows the time-series of the temperature of Bloomington after detrending using this method.
+The blue curve in the left panel of Figure \autoref{fig:bloom_estimatedSD} shows the time-series of the temperature of Indianapolis after detrending using this method.
 This figure, reveals another difficulty in estimating the trend of volatility in this data: the variance of this signal, shows cyclic behavior.
@@
 shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
-the time-series of Bloomington obtained from out proposed model.
+the time-series of Indianapolis obtained from out proposed model.
 To reduce the number of time-steps we work on the weekly averaged of the data.

### 2018-09-03 code/AnalyseSyntheticData_conADMM.py
update paper

@@ -16,7 +16,7 @@ saveFigDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Reports/051218/Figures'
 
 lam_t_vec=[5.];
 lam_s_vec=[0.1];
-rho=.013
+rho=.01
 
 ca = ConsensusADMM()#construct linearizedADMM object
 data_fn=join(dstDir,'simulated_data')

### 2018-09-03 code/AnalyseSyntheticData_linADMM.py

@@ -6,8 +6,8 @@ import matplotlib.pyplot as plt
 
 bl=[False,True]
 ifSimulateData=bl[0]
-ifFitModel=bl[0]
-ifVisualize=bl[1]
+ifFitModel=bl[1]
+ifVisualize=bl[0]
 ifModelSelection=bl[0]
 
 dstDir='/home/arash/MEGA/MEGAsync/Projects/Cloud/Data/Simulated data'
@@ -67,10 +67,10 @@ if ifSimulateData:
 ###################
 ###Fit model###
 ###################
-lam_t_vec=[5.];
+lam_t_vec=[100.];
 lam_s_vec=[0.0,.05,.1,.2,.3];
 lam_s_vec=[0.1];
-mu=.01
+mu=.0001
 #lam_t_vec=[10];lam_s_vec=[.05];mu=.01
 
 la = LinearizedADMM()#construct linearizedADMM object
@@ -110,7 +110,7 @@ if ifVisualize:
     fig.add_subplot(122)
     plt.pcolor(covMat[:,45].reshape((n_rows,n_cols),order='F'))    
     plt.xticks([]);plt.yticks([]);    
-    fig.savefig(join(saveFigDir,'true_var_spatial.pdf'))
+    #fig.savefig(join(saveFigDir,'true_var_spatial.pdf'))
     #---true spatial variance ---
 
     #---true vs estimated temporal variance --- 
@@ -122,12 +122,12 @@ if ifVisualize:
     i=0;plt.plot(np.exp(la.fittedVar[i,]/2));plt.plot(covMat[i,:])
     plt.xlabel('time');plt.ylabel('standard deviation')
     
-#    lam_t=100.;lam_s=.1;mu=.0001
-#    fn='mu_{}_lam_t_{}_lam_s_{}'.format(mu,lam_t,lam_s)
-#    filepath=join(savedResultsDir,'X_'+fn)
-#    la.analyseFittedValues(filepath)
-#    plt.plot(np.exp(la.fittedVar[i,]/2))
-#    plt.legend(['estimated, $\lambda_t=5$','true','estimated, $\lambda_t=100$'])
+    lam_t=100.;lam_s=.1;mu=.01
+    fn='mu_{}_lam_t_{}_lam_s_{}'.format(mu,lam_t,lam_s)
+    filepath=join(savedResultsDir,'X_'+fn)
+    la.analyseFittedValues(filepath)
+    plt.plot(np.exp(la.fittedVar[i,]/2))
[…]

### 2018-09-03 code/preProcessing.py

@@ -48,8 +48,9 @@ spatialSubSample(dataDir,data_fn=fn+'_data_avg',
 
 #===save data into csv file===
 #prepare data for genlasso R package. genlasso is used to detrend the data.
-metadata=load(open(join(dataDir,
-                        '1960-01-01_to_2010-12-31_metadata_avg_sub')))
+with open(join(dataDir,'1960-01-01_to_2010-12-31_metadata_avg_sub')) as f:
+    metadata=load(f)
+
 dataMat=np.fromfile(join(dataDir,
                          '1960-01-01_to_2010-12-31_data_avg_sub'),
                     'float32')

### 2018-09-03 code/volatilitytrend/data_utils/ecmwf_dataUtils.py

@@ -1,7 +1,11 @@
 '''
 This module contains functions for loading and analyzing
 datasets from ecmwf. See:
-https://software.ecmwf.int/wiki/display/WEBAPI/Access+ECMWF+Public+Datasets        
+https://software.ecmwf.int/wiki/display/WEBAPI/Access+ECMWF+Public+Datasets 
+
+and 
+
+http://apps.ecmwf.int/datasets/       
 '''
 
 from .. import config
@@ -210,7 +214,7 @@ def computeAveragedData(dataDir,data_fn,metadata_fn,win_size=7):
     #===sub-sample from dates===
     idx=np.arange(0,(n_win*win_size),win_size)
     new_dates=[dates[i] for i in idx]
-    metadata.update({'dates':new_dates})
+    metadata.update({'dates':new_dates,'T':n_win})
     with open(join(dataDir,metadata_fn+'_avg'),'wb') as f:
         dump(metadata,f)
     #===sub-sample from dates=== 

### 2018-09-04 aaai-2018/khodadadi-mcdonald-aaai19.tex
update paper

@@
 This leads to a new optimization problem which we consider its solution next.}
-
-though it uses a different likelihood
-function to emphasize variance estimation rather than trends in mean
-signal. Furthermore, the focus is in high-dimensional, regular data
-rather than a small number of changing graphs.
 
@@
   \includegraphics[height=.15\textheight]{Figures/true_var_spatial}\\
-  \includegraphics[height=.15\textheight]{Figures/true_fitted_var}
+  \includegraphics[height=.2\textheight]{Figures/true_fitted_var}
   \caption{Top: Variance function at $t=25$ (left) and $t=45$
@@
   \centering	
-  \includegraphics[height=.12\textheight]{Figures/true_fitted_var}
-  \includegraphics[height=.12\textheight]{Figures/convergence}  
-  \includegraphics[height=.12\textheight]{Figures/modelComp_MAE}  
-  \caption{Left: The true (orange) and estimated standard deviation 
-    function at the location (0,0). The estimated values are
-    obtained using linearized ADMM with $\lambda_s=0.1$ and two
-    values of $\lambda_t$: $\lambda_t=5$ (blue) and
-    $\lambda_t=100$ (green). Middle: Convergence speed of linearized
-    and consensus ADMM. Right: MAE for four models: admm\_opt: the
+  \includegraphics[height=.16\textheight]{Figures/convergence}  
+  \includegraphics[height=.16\textheight]{Figures/modelComp_MAE}  
+  \caption{Top: Convergence speed of linearized
+    and consensus ADMM. Bottom: MAE for four models: admm\_opt: the
     proposed model with optimal values of $\lambda_t$ and $\lambda_s$,
@@
 	\centering
-	\includegraphics[width=.2\textheight]{Figures/cities_ts}
+	\includegraphics[width=.3\textheight]{Figures/cities_ts}
  	\caption{Time-series of the temperature (in Kelvin) of three cities.}
@@
   \centering
-  \includegraphics[width=1\columnwidth]{Figures/ts_estimatedVar}
+  \includegraphics[width=1.1\columnwidth]{Figures/ts_estimatedVar}
   \caption{Detrended data and the estimated SD for

### 2018-09-04 aaai-2018/khodadadi-mcdonald-aaai19.tex
update figures
update appendix

@@
 \begin{equation}
-\min_{\beta} \frac{1}{2} \norm{ y-\beta }_2^2+\lambda \norm{ D\beta}_1
+\min_{\beta} \frac{1}{2} \norm{ y-\beta }_2^2+\lambda \norm{ D_t \beta}_1
 \label{eq:l1tf}
@@
  where $y=\{y_t\}_{t=1}^T$ is an observed time-series, $\beta\in \mathbb{R}^T$ is the smooth trend,
- $D$ is a $(T-2)\times T$ matrix, and $\lambda$ is a tuning parameter
+ $D_t$ is a $(T-2)\times T$ matrix, and $\lambda$ is a tuning parameter
  which balances fidelity to the data (small errors in the first term)
@@
 %The first term in \eqref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
-With the penalty matrix $D$, the estimated $\beta$ will be piecewise
-linear. \citet{KimKoh2009} proposed a specialized primal-dual
+In Appendix A, we specify the form of the matrix $D_t$ which results in piecewise
+linear estimated $\beta$. \citet{KimKoh2009} proposed a specialized primal-dual
 interior point (PDIP) algorithm for solving \eqref{eq:l1tf}. From a
@@
 structure of the estimator is determined by the penalty function
-$\lambda\norm{D\beta}_1$ rather than any parametric trend
+$\lambda\norm{D_t\beta}_1$ rather than any parametric trend
 assumptions---autoregressive, moving average, sinusoidal seasonal
@@
 \begin{equation}
-\min_h -l(y\given h)+\lambda \norm{ Dh }_1
+\min_h -l(y\given h)+\lambda \norm{ D_th }_1
 \label{eq:l1tf_var}
 \end{equation}
-where $D$ has the same structure as above.
+where $D_t$ has the same structure as above.
 
@@
 we can write \eqref{eq:l1tf_var_st} in matrix form where $h$ is an
-$T\times S$ vector and $D$ is replaced by $D_{ST} \in
+$T\times S$ vector and $D_t$ is replaced by $D \in
 \mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot
@@
 \begin{equation}
-\min_h -l(y\given h)+ \Lambda^\top | D_{ST}h |.
+\min_h -l(y\given h)+ \Lambda^\top | Dh |.
 \label{eq:l1tf_var_st_mat}
@@
 
-For a spatial grid of size $S$ and $T$ time steps, $D_{ST}$ will have
-$3Tn_rn_c-2n_rn_c-Tn_r$ rows and $ST$ columns. For a $1^\circ\times
+For a spatial grid of size $S$ and $T$ time steps, $D$ in equation \ref{eq:l1tf_var_st_mat} will have
+$3Tn_rn_c-2n_rn_c-Tn_r$ rows and $S \cdot T$ columns. For a $1^\circ\times
 1^\circ$ grid over the entire northern hemisphere and daily data over
 10 years, we have $S=90\times 360\approx 32,000$ spatial locations and
-$T=3650$ time points, and so $D_{ST}$ has
+$T=3650$ time points, and so $D$ has
 approximately $10^8$ columns and $10^8$ rows. In each step of the PDIP
 algorithm, we need to solve a linear system of equations which
-depends on $D_{ST}^\top D_{ST}$ (see appendix A and B). Therefore,
+depends on $D^\top D$ (see appendix A and B). Therefore,
 applying the PDIP directly is infeasible for our data.\footnote{We
-  note that $D_{ST}$ is a highly structured, sparse matrix, but, unlike
+  note that $D$ is a highly structured, sparse matrix, but, unlike
   trend filtering alone, it is not banded. We are unaware of general
@@
   \caption{Top: Convergence speed of linearized
-    and consensus ADMM. Bottom: MAE for four models: admm\_opt: the
+    and consensus ADMM. Bottom: MAE for four models: our method: the
     proposed model with optimal values of $\lambda_t$ and $\lambda_s$,
-    admm\_temp: no spatial penalty, admm\_sp: no temporal
+    temporal: no spatial penalty, spatial: no temporal
     penalty.\attn{Fix these}} \label{fig:true_fitted_var}
@@
   \centering
-  \includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}
+  \includegraphics[width=.8 \columnwidth]{Figures/indy_estimatedSD_1}
+  \includegraphics[width=.45 \columnwidth]{Figures/indy_estimatedSD_2}
+  \includegraphics[width=.45 \columnwidth]{Figures/indy_estimatedSD_3}  
   \caption{Left: The residuals of the time-series of Indianapolis
@@
     penalty. Right: the same as middle panel but here the long horizon
-    penalty is imposed. See the text for more details.} 
+    penalty is imposed. See the text for more details. \attn{Fix the caption (top-bottom or left right)}} 
   \label{fig:bloom_estimatedSD}
@@
   \caption{The
-    average of the estimated variance over the northern hemisphere (left) and the
-    change in the variance from 1961 to 2011 (right).}
+    average of the detrended estimated variance over the northern hemisphere (top) and the
+    change in the variance from 1961 to 2011 (bottom).}
 \label{fig:avg_chg}
@@
 \appendix
-\section{PDIP for $\ell_1$ Trend Filtering}
+\section{PDIP for $\ell_1$ Trend Filtering of variance}
 \label{sec:app_l1tf_var}
 
-In this appendix we provide more details on how to solve the optimization problem \ref{eq:l1tf_var} using PDIP.
+In this appendix we provide more details on how to solve the optimization problem with the objective specified in equation \ref{eq:l1tf_var} using PDIP.
 The objective function is convex but not differentiable.
@@
 
-where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that
+\noindent where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that for the objective function of Equation \ref{eq:l1tf_var}
 
@@
 
-for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form
+\noindent for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form
 
@@
 
-The matrix $D$ in these equations have the following form:  $D_{total}^t=[D_{t}^t|D_{s}^t]$. The matrix $D_t$ is the following block-diagonal matrix and corresponds to the temporal penalty: 
- \begin{equation}
- D_t=\begin{bmatrix}
- D &  & \\ 
+Having computed the conjugate function and its gradient and Jacobian, now we can use a number of convex optimization software packages which have an implementation of PDIP to solve the optimization problem with the objective function \ref{eq:l1tf_var}.
+We chose the python API of the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.
+
+\section{PDIP Update in~\autoref{alg:conADMM}}
+\label{sec:app_consADMM}
+
+In this section we give more details on performing the $x$-update step in Algorithm 1. We need to solve the following optimization problem:
+
+\begin{equation}
+\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
+\label{eq:x_update_opt}
+\end{equation}
+
+\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$.
+
+The matrix $D$ has the following form:  $D^t=[D_{temp}^t|D_{spat}^t]$. The matrix $D_{temp}$ is the following block-diagonal matrix and corresponds to the temporal penalty: 
+\begin{equation}
+D_{temp}=\begin{bmatrix}
+D_t &  & \\ 
 & \ddots & \\
-  &  & D
+&  & D_t
 \end{bmatrix}
@@
 
-where $D$ was defined in Section 2 of the main text.
-The number of the diagonal blocks is equal to the grid size $n_r \times n_c$.
-Each row of the matrix $D_s$ corresponds to one spatial constraint in Equation \eqref{eq:l1tf_var_st} in the text.
+\noindent where $D_t$ was first introduced in Section 2 of the main text and has the following form:
+
+\begin{equation}
+ D_t=\begin{bmatrix}
+ 1 & -2 & 1 &  &  &  &\\ 
+ & 1 & -2 & 1 &  &  &\\ 
+ &  & \ddots & \ddots & \ddots  &  &\\ 
+ &  & & 1 & -2 & 1 &  \\ 
+ &  &  &   & 1 & -2 & 1 
+ \end{bmatrix}
+\label{eq:d_matrix}
+\end{equation}
+
+The number of the diagonal blocks in $D_{temp}$ is equal to the grid size $n_r \times n_c$.
+Each row of the matrix $D_{spat}$ corresponds to one spatial constraint in Equation \eqref{eq:l1tf_var_st} in the text.
 For example, the first $T$ rows correspond to $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to $|h_{11t}-h_{12t}|$, and so on.
 
-\section{PDIP Update in~\autoref{alg:conADMM}}
-\label{sec:app_consADMM}
-
-In this section we give more details on performing the $x$-update step in Algorithm 1. We need to solve the following optimization problem:
-
-\begin{equation}
-\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
-\label{eq:x_update_opt}
-\end{equation}
-
-\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$.
-Let $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$.
-As it was explained in \autoref{sec:l1tf_var}, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$.
-So to use PDIP we first need to compute the conjugate function $f^*(\cdot)$.
+This optimization problem, is again a generalized LASSO problem with $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$.
+
+As it was explained in Appendix A, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$.
+To use PDIP we first need to compute the conjugate function $f^*(\cdot)$.
 We have:
@@
 \end{equation}
-
-Having computed the conjugate function and its gradient and Jacobian, now we can use a number of convex optimization software packages which have an implementation of PDIP to perform the x-update step inside the ADMM loop.
-We chose the python API of the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{Figures/bloom_estimatedSD}
+  \includegraphics[width=\columnwidth]{Figures/indy_estimatedSD_1}
   \caption{Left: The residuals of the time-series of Indianapolis

### 2018-09-04 aaai-2018/khodadadi-mcdonald-aaai19.tex
"shrink figs, cutting down manuscript"

@@
 \begin{aligned}
-&\min_h \sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}} +\lambda_1
+&\min_h \sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}} +\lambda_t
 \sum_{i,j} \sum_{t=1}^{T-2}
 \left|h_{ijt}-2h_{ij(t+1)}+h_{ij(t+2)}\right|\\ 
-&+\lambda_2 \sum_{t,j} \sum_{i=1}^{n_r-1} \left|h_{ijt}-h_{(i+1)jt}\right|
-+\lambda_2 \sum_{t,i} \sum_{j=1}^{n_c-1} \left|h_{ijt}-h_{i(j+1)t}\right|
+&+\lambda_s \sum_{t,j} \sum_{i=1}^{n_r-1} \left|h_{ijt}-h_{(i+1)jt}\right|
++\lambda_s \sum_{t,i} \sum_{j=1}^{n_c-1} \left|h_{ijt}-h_{i(j+1)t}\right|
 \end{aligned}
@@
 
-\textcolor{red}{Our methodology is similar in flavor to~\cite{HallacPark2017}, with several fundamental differences.
-The main goal of Hallac et al.
-is to discover the time-varying structure of a network.
-To achieve this goal, they introduced two penalty terms into the objective function.
-One term penalizes the sum of the absolute values of the off-diagonal elements of the inverse covariance matrix.
-This term enforces sparsity on the estimated inverse covariance matrix and so helps in discovering which nodes in the network are connected.
-The other penalty term, penalizes differences in the estimated inverse covariance matrix from one time-step to the next.
-Our goal in the present work is to detect the temporal trend in the variance of each point in the network.
-The network in our problem is the grid of points over the earth and so the network's structure remains the same through time.
-The variance of each point, however changes.
-To capture these properties in our data using the objective function used in Hallac et al.
-(see Equation 2 in their paper), we need to have different temporal penalty terms for the diagonal and off-diagonal elements in the inverse covariance matrix.
-This results in unnecessary complication in the optimization problem.
-To overcome these difficulties, we introduced the new objective function \ref{eq:l1tf_var_st_mat}.
-This leads to a new optimization problem which we consider its solution next.}
-
-\attn{Paragraph discussing relationship with estimating time-varying covariance}
+Our methodology is similar in flavor to~\cite{HallacPark2017} or
+related work in~\citep{GibberdNelson2017,MontiHellyer2014}, but with
+several fundamental differences. These papers aim to
+discover the time-varying structure of a network. To achieve this
+goal, they use Gaussian likelihood with unknown precision matrix and
+introduce penalty terms which (1) encourage sparsity among the
+off-diagonal elements and (2) discourage changes in the estimated inverse
+covariance matrix from one time-step to the next.
+Our goal in the present work is to detect the temporal trend in the variance of each
+point in the network, but the network is known (corresponding to the grid
+over the earth). The variance of each point, however changes. To modify the
+objective function in~\cite[Eq.\ 2]{HallacPark2017}, we would
+enforce complete sparsity on the off-diagonal elements (since
+they are not estimated) and add a new penalty to enforce spatial
+behavior across the diagonal elements. Thus, 
+\eqref{eq:l1tf_var_st_mat} is not simply a special case of these
+existing methods.
 
@@
   \norm{x}_1$. Then, 
-  \begin{align}
+  \begin{equation}
+    \begin{aligned}
       [\prox_{\mu f}(u)]_k &= \mathscr{W}\bigg(\frac{y_k^2}{\mu}
@@
       \prox_{\rho g}(u) &= S_{\rho\lambda}(u)
-  \end{align}
+    \end{aligned}
+  \end{equation}
   where $\mathscr{W}(\cdot)$ is the \textit{Lambert W function}
@@
 \end{align}
-In words, the variance at each time and location is computed as the
+The variance at each time and location is computed as the
 weighted sum of $K$ bell-shaped functions where the weights are
@@
 For illustration, we also plot the
-variance function for all locations at $t=25$ and $t=45$ in as well as
-the variance across time at $(0,0)$ in \autoref{fig:true_var_spatial}.
+variance function for all locations at $t=25$ and $t=45$
+(\autoref{fig:true_var_spatial}, top panel) in as well as
+the variance across time at $(0,0)$ (bottom panel, orange).
 \begin{table}[tb]
@@
 \end{table} 
-
 \begin{figure}[tb]
   \centering	
-  \includegraphics[height=.15\textheight]{Figures/true_var_spatial}\\
-  \includegraphics[height=.2\textheight]{Figures/true_fitted_var}
+  \includegraphics[height=.1\textheight]{Figures/true_var_spatial}\\
+  \includegraphics[height=.15\textheight]{Figures/true_fitted_var}
   \caption{Top: Variance function at $t=25$ (left) and $t=45$
@@
 estimated variance and the true variance at all locations and all time
-steps. For $\lambda_t=5$ and $\lambda_s=0.1$, the MAE was minimized. The
-left panel of \autoref{fig:true_fitted_var} shows the true and the
-estimated standard deviation at location (0,0) using $\lambda_s=0.1$
-and $\lambda_t=5$ (blue) and $\lambda_t=100$ (green). As we can see,
-larger than optimal value of $\lambda_t$ leads to estimated values
-which are ``too smooth". The middle panel of \autoref{fig:true_fitted_var} shows the
-convergence of both methods. Each iteration of the linearized
-algorithm takes 0.01 seconds on average while each iteration of the
-consensus ADMM takes about 20 seconds. \attn{fix this discription}
-
-\attn{Arash: Can you be more specific?}
-
+steps. For $\lambda_t=5$ and $\lambda_s=0.1$, the MAE was minimized. 
 \begin{figure}[tb]
   \centering	
-  \includegraphics[height=.16\textheight]{Figures/convergence}  
-  \includegraphics[height=.16\textheight]{Figures/modelComp_MAE}  
-  \caption{Top: Convergence speed of linearized
-    and consensus ADMM. Bottom: MAE for four models: our method: the
-    proposed model with optimal values of $\lambda_t$ and $\lambda_s$,
-    temporal: no spatial penalty, spatial: no temporal
-    penalty.\attn{Fix these}} \label{fig:true_fitted_var}
+  \includegraphics[width=.46\columnwidth]{Figures/convergence}  
+  \includegraphics[width=.46\columnwidth]{Figures/modelComp_MAE}  
+  \caption{Left: Value of the objective function for linearized (orange)
+    and consensus (blue) ADMM against iteration. Right: MAE for (1) our
+    method with optimal values of $\lambda_t$ and $\lambda_s$ (2)
+    temporal penalty only (3) spatial
+    penalty only and (4) a GARCH(1,1).} 
+  \label{fig:true_fitted_var}
 \end{figure}
+The
+bottom panel of \autoref{fig:true_var_spatial} shows the true and the
+estimated standard deviation at location (0,0) 
+and $\lambda_t=5$ (blue) and $\lambda_t=100$ (green) ($\lambda_s=0.1$).
+Larger values of $\lambda_t$ lead to estimated values
+which are ``too smooth". The left panel of \autoref{fig:true_fitted_var} shows the
+convergence of both methods as a function of iteration. It is
+important to note that each iteration of the linearized
+algorithm takes 0.01 seconds on average while each iteration of the
+consensus ADMM takes about 20 seconds. Thus, where the lines meet at
+400 iterations requires about 4 seconds for the linearized method and
+2 hours for the consensus method.
 
@@
 \autoref{fig:true_fitted_var} shows the boxplot of the MAE for these
-models. \attn{What is this next sentence trying to say?}
-
-\attn{that our model is better than GARCH in estimating the variance.}
-
-Interestingly, the proposed model with optimal parameters 
-outperforms GARCH(1,1) in estimating the true value of the variance.
+models. We note that, using an algorithm akin
+to~\citep{HallacPark2017} ignores the spatial component and
 
@@
 
-Consensus ADMM in~\autoref{sec:consOpt} is appropriate only if
-we parallelize it over multiple machines, and it is significantly
-slower on our simulated data, so we do not pursue it
-further here. All the results reported in this section are obtained
+Consensus ADMM in~\autoref{sec:consOpt} is appropriate when we can
+easily parallelize it over multiple machines.  Otherwise, it is significantly
+slower, so all the results reported in this section are obtained
 using \autoref{alg:linADMM}. We
-applied this algorithm to the northern hemisphere of the ERA-20C
+applied this algorithm to the Northern hemisphere of the ERA-20C
 dataset available from the European Center for Medium-Range
-Weather Forecasts\footnote{\url{https://www.ecmwf.int}}. 
+Weather Forecasts\footnote{\url{https://www.ecmwf.int}}. We use 
 the 2 meter temperature measured daily at 12 p.m from January 1, 1960
@@
 
-\attn{Can we redo these but with, say, Indianapolis instead of
-  Bloomington?} Examination of the time-series alone demonstrates
+Examination of the time-series alone demonstrates
 strong differences 

### 2018-09-04 aaai-2018/khodadadi-mcdonald-aaai19.tex

@@
 
-%\pdfinfo{
-%	/Title (Algorithms for Estimating Trends in Global Temperature Volatility)
-	%/Author (Khodadadi and McDonald)
-%	/Author (John Doe, Jane Doe)}
+\pdfinfo{
+  /Title (Algorithms for Estimating Trends in Global Temperature Volatility)
+  % /Author (Khodadadi and McDonald)
+  /Author (John Doe, Jane Doe)}
 
@@
 features and on a large, publicly available, global
-temperature dataset with the goal of tracking trends in
+temperature dataset with the eventual goal of tracking trends in
 cloud reflectance temperature variability.
@@
 
-\textcolor{red}{How is this first part related to the rest of the paper?}
-
 The amount of sunlight reflected from clouds is among the largest
@@
 prediction~\citep{BoucherRandall2013}. But climate models fail to
-reproduce global cloud statistics and understanding the reasons for
+reproduce global cloud statistics, and understanding the reasons for
 this failure is a grand challenge of the World Climate Research
 Programme~\citep{BonyStevens2015}.
-By understanding how cloudliness changes over time, examining changes
-in brightness temperature within various cloud types, and inferring
-changes in cloud radiative effects, atmospheric scientists can better model climate
-change. Modeling inter-satellite biases, satellite drift, and
-seasonal and long-term climate phenomena like El Ni\~{n}o will lead to
-a better understanding of climate
-change~\citep{SchreierKahn2014,BaumWielicki1994,BaumWielicki1992,FreyAckerman1996}.
-
-Numerous
+% By understanding how cloudliness changes over time, examining changes
+% in brightness temperature within various cloud types, and inferring
+% changes in cloud radiative effects, atmospheric scientists can better model climate
+% change. Modeling inter-satellite biases, satellite drift, and
+% seasonal and long-term climate phenomena like El Ni\~{n}o will lead to
+% a better understanding of climate
+% change~\citep{SchreierKahn2014,BaumWielicki1994,BaumWielicki1992,FreyAckerman1996}.
+While numerous
 studies have examined the overall impacts of clouds on climate
 variability~\citep{MyersMechoso2018,GrisePolvani2013,BenderRamanathan2012},
-but such investigations have been hampered by the lack of a suitable
+such investigations have been hampered by the lack of a suitable
 dataset. Ideal data would have global coverage at high spatial
 resolution, a long enough record to recover temporal trends, and be
-multispectral~\citep{WielickiYoung2013}. The International Satellite
-Cloud Climatology Project (ISCCP)
-has been producing cloud property information from
-limited spectral channels for over three
-decades~\citep{RossowSchiffer1991}. These data have been used in 
-various climate variability
-studies~\citep[e.g.]{BenderRamanathan2012}, however, the utility of
-these data for long-term study of climate variability has been
-questioned~\citep{EvanHeidinger2007}.
-
-ISCCP provides a relatively long record, but it only incorporates
-radiances from limited visible and infrared channels from Advanced
-Very High Resolution Radiometer (AVHRR) instruments. Ongoing work seeks to create a
-more spectrally-detailed dataset which can avoid the above issues by
-combining radiance data from AVHRR imagers with readings from High-resolution
+multispectral~\citep{WielickiYoung2013}. 
+% The International Satellite
+% Cloud Climatology Project (ISCCP)
+% has been producing cloud property information from
+% limited spectral channels for over three
+% decades~\citep{RossowSchiffer1991}. These data have been used in 
+% various climate variability
+% studies~\citep[e.g.]{BenderRamanathan2012}, however, the utility of
+% these data for long-term study of climate variability has been
+% questioned~\citep{EvanHeidinger2007}. 
+% ISCCP provides a relatively long record, but it only incorporates
+% radiances from limited visible and infrared channels from Advanced
+% Very High Resolution Radiometer (AVHRR) instruments. 
+Current work~\cite{StatenKahn2016,SchreierKahn2010,KahnFishbein2007} seeks to create a
+spectrally-detailed dataset by %which can avoid the above issues by
+combining radiance data from Advanced Very High Resolution Radiometer
+imagers with readings from High-resolution 
 Infrared Radiation Sounders on board legacy
-satellites~\citep{StatenKahn2016,SchreierKahn2010,KahnFishbein2007}. In
-anticipation of this new dataset, our work develops novel methodology
+satellites. 
+In anticipation of this new dataset, our work develops novel methodology
 for examining the trends in variability of climate data across space
@@
 \citet{huntingford_no_2013} took a similar
-approach in analyzing the ERA-40 data set. Their results showed that there
-still is an increase in the SDs from 1958-1970 to 1991-2001, but this
-is much less than what is obtained from the method used in
-\cite{hansen_perception_2012}. The authors also computed the
+approach in analyzing the ERA-40 data set. They argued that, while there
+is an increase in the SDs from 1958-1970 to 1991-2001, it
+is much smaller than found in
+\cite{hansen_perception_2012}. \citet{huntingford_no_2013} also computed the
 time-evolving global SD from the de-trended time-series at each
-position, which suggests that the global SD has been
+position and argued that the global SD has been
 stable.
 
-These and other related research, e.g.,
-\citet{rhines_frequent_2013}, have several shortcomings. First, no
+These and other related work  
+\cite[e.g.,]{rhines_frequent_2013} have several shortcomings. First, no
 statistical analysis has been performed to examine if the changes in
@@
 computing the SDs are highly sensitive to the choice of base period. Third,
-and most importantly, temporal and spatial correlations between the
+and most importantly, temporal and spatial correlations between 
 observations are completely ignored.
@@
 detecting the trend in the volatility of spatio-temporal data. In this
-methodology, the variance at each position and time is considered as
-a hidden variable. The values of these hidden variables
-are then estimated by maximizing the likelihood of the observed
-data. Following~\cite{Tibshirani2014}, we penalize the
+methodology, the variance at each position and time are estimated by
+maximizing penalized likelihood
+maximization. Following~\cite{Tibshirani2014}, we penalize the 
 differences between the estimated variances which
@@
 in a generalized LASSO problem. However, in our application, the
-dimension of this optimization problem is massive, and so the
-standard solvers are inadequate. We develop two algorithms
-which are computationally feasible. In the first method, we adopt an optimization technique
+dimension of this optimization problem is massive, so the
+standard solvers are inadequate.
+
+We develop two algorithms
+which are computationally feasible on extremely large data. In the
+first method, we adopt an optimization technique 
 called alternating direction method of multipliers
@@
 problem can be solved by iteratively solving these sub-problems. The
-second method, called the \textit{linearized ADMM algorithm}
-\cite{parikh_proximal_2014} solves the main problem by iteratively
+second method, called linearized ADMM
+\cite{parikh_proximal_2014}, solves the main problem by iteratively
 solving a linearized version of it. We will compare the benefits of
@@
   Python code is available on {\tt github}.
-\item We illustrate our methods on a large, publicly available, global temperature dataset
-  with the goal of tracking world-wide trends in variance as well as a
+\item Because the construction of satellite-based datasets is ongoing
+  and currently proprietary, we illustrate our methods on a large,
+  publicly available, global temperature dataset.
+  The goal is to demonstrate the feasibility of these methods for tracking
+  world-wide trends in variance in standard atmospheric data and a
   simulation constructed to mimic these data's features
@@
 % following convex optimization problem: 
-$
+$$
 \min_{\beta} \frac{1}{2} \sum_{t=1}^{T} (y_t-\beta_t)^2+\lambda
 \sum_{t=1}^{T-2} \left|\beta_t-2\beta_{t+1}+\beta_{t+2}\right| 
-$
+$$
 or equivalently:
@@
 %The first term in \eqref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
-In Appendix A, we specify the form of the matrix $D_t$ which results in piecewise
-linear estimated $\beta$. \citet{KimKoh2009} proposed a specialized primal-dual
+% In Appendix A, we specify the form of the matrix $D$ which results in piecewise
+% linear estimated $\beta$. 
+\citet{KimKoh2009} proposed a specialized primal-dual
 interior point (PDIP) algorithm for solving \eqref{eq:l1tf}. From a
@@
 \sigma^2)$, subject to a piecewise linear constraint on
-$\beta$. Alternatively, solutions to \eqref{eq:l1tf} are equivalent to
+$\beta$. Alternatively, solutions to \eqref{eq:l1tf} are %equivalent to
 maximum a posteriori Bayesian estimators based on Gaussian likelihood
@@
 component, etc. The resulting trend is therefore essentially
-nonparametric in the same way that splines are. In fact,
+nonparametric in the same way that splines are nonparametric. In fact,
 using squared $\ell_2$ norm as the penalty instead of $\ell_1$ results exactly in
@@
 model is $l(y\given h) \propto -\sum_{t=1}^T h_t - y_t^2e^{-h_t}$. Crucially,
-we assume that the parameters $h_t$ vary smoothly. To impose
-this assumption, we estimate $h_t$ by solving the penalized, negative
+we assume that the parameters $h_t$ vary smoothly and estimate them by
+minimizing the penalized, negative
 log-likelihood:  
@@
 PDIP we need to compute a search direction by taking a Newton
-step on a system of nonlinear equations. Due to space limitations, we
-defer details to Appendix A in the Supplement, where we show how to
+step on a system of nonlinear equations. For completeness, we
+provide the details in Appendix A of the Supplement, where we show how to
 derive the dual of this optimization problem and compute the first and
@@
 The method in the previous section can be used to estimate the
-variance of a single time-series. In this section, we extend this
-method to the estimation of the variance of spatio-temporal data.
+variance of a single time-series. Here we extend this
+method to the case of spatio-temporal data.
 
@@
 observation at time $t$ on the $i^\text{th}$ row and $j^\text{th}$
-column of the grid, and $h_{ijt}$ denote the corresponding hidden
-variable. We seek to impose both temporal and spatial smoothness
+column of the grid, and $h_{ijt}$ denote the corresponding
+parameter. We seek to impose both temporal and spatial smoothness 
 constraints on the hidden variables. Specifically, we seek a solution
@@
 likelihood is different). As before,
-we can write \eqref{eq:l1tf_var_st} in matrix form where $h$ is an
-$T\times S$ vector and $D_t$ is replaced by $D \in
+we can write \eqref{eq:l1tf_var_st} in matrix form where $h$ is a
+vector of length
+$TS$ and $D_t$ is replaced by $D \in
 \mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot
 (T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and
-spatial constraints, respectively\footnote{$N_s$ is obtained by
-  counting the number of unique constraints at each location and at
-  all times.}. The exact form of this matrix in clarified in Appendix
-A in the Supplement. Then, as we have two different tuning parameters
+spatial constraints, respectively. 
+% The exact form of this matrix in clarified in Appendix
+% B in the Supplement. 
+Then, as we have two different tuning parameters
 for the temporal and spatial components, we write $\Lambda
-=\left[\lambda_1\one_{N_t}^\top,\;
-  \lambda_2\one_{N_s}^\top\right]^\top$ leading
+=\left[\lambda_t\one_{N_t}^\top,\;
+  \lambda_s\one_{N_s}^\top\right]^\top$ leading
 to:\footnote{Throughout the paper, we use $|x|$ for both scalars and
@@
 \subsection{Related Work}
-
-\attn{I've moved this here rather than above to focus the intro more
-  on the data.}
 
@@
 point in the network, but the network is known (corresponding to the grid
-over the earth). The variance of each point, however changes. To modify the
+over the earth). To modify the
 objective function in~\cite[Eq.\ 2]{HallacPark2017}, we would
@@
 \eqref{eq:l1tf_var_st_mat} is not simply a special case of these
-existing methods.
+existing methods. Finally, these papers examine networks with hundreds
+of nodes and dozens to hundreds of time points. As discussed next, our data are
+significantly larger than these networks and attempting to estimate a
+full covariance would be prohibitive, were it required.
 
@@
 10 years, we have $S=90\times 360\approx 32,000$ spatial locations and
-$T=3650$ time points, and so $D$ has
-approximately $10^8$ columns and $10^8$ rows. In each step of the PDIP
-algorithm, we need to solve a linear system of equations which
-depends on $D^\top D$ (see appendix A and B). Therefore,
+$T=3650$ time points, so $D$ has
+approximately $10^8$ columns and $10^8$ rows. In principal, we could
+solve~\eqref{eq:l1tf_var_st_mat} using PDIP as before, however, each iteration
+requires solving a linear system of equations which
+depends on $D^\top D$. Therefore,
 applying the PDIP directly is infeasible for our data.\footnote{We
@@
   linear algebra techniques for inverting such matrix, despite our
-  best efforts.}
+  best efforts to find them.}
 
@@
 \begin{aligned}
-&\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i) &
+&\min_{\{x_1,...,x_N  \}} \sum_{i=1}^N f_i(x_i) &
 s.t. &\quad \tilde{h}_i=x_i.
@@
 variables corresponding to the same global variable.  
-The \textit{augmented Lagrangian} corresponding to 
+The augmented Lagrangian corresponding to 
 \eqref{eq:consADMM} is $L_\rho(x,h,y)=\sum_i
 \big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert
-x_i-\tilde{h}_i \lVert_2^2 \big)$.
-Now, we can apply ADMM to $L_\rho$.
-This results in solving $N$ independent optimization problems followed by a step to achieve consensus among the solutions in each iteration.
+x_i-\tilde{h}_i \lVert_2^2 \big)$. Now, we can apply ADMM to
+$L_\rho$. This results in solving $N$ independent optimization
+problems followed by a step to achieve consensus among the solutions
+in each iteration. 
 % which results in the following ADMM updates:
@@
 will be small. In addition, our experiments with several values of
-$\lambda_t$ and $\lambda_s$ showed that the algorithm converges in few
-hundreds iterations. 
+$\lambda_t$ and $\lambda_s$ showed that the algorithm converges in a few
+hundred iterations. 
 % \attn{Need to redo this:}Solving each sub-problem on a machine with four
@@
 when these proximal operators can be evaluated efficiently which, as
-we show next, is the case for our problem.
+we show next, is the case.
 
@@
 % \end{proof}
-
+The proof is fairly straightforward and given in Appendix B in the Supplement. 
 Therefore, \autoref{alg:linADMM} gives a different method for solving
@@
 whether \autoref{alg:conADMM} or \autoref{alg:linADMM} will be more
-useful in any particular context.
+useful in any particular context. In our experience,
+\autoref{alg:conADMM} converges in many fewer iterations, but
+each iteration is much slower unless carefully parallelized.
 
@@
 temporal phenomena observable in typical climate data. We generate a
-complete spatio temporal field wherein 
+complete spatio-temporal field wherein 
 observations at all time steps and all locations are
@@
 variance function for all locations at $t=25$ and $t=45$
-(\autoref{fig:true_var_spatial}, top panel) in as well as
+(\autoref{fig:true_var_spatial}, top panel) as well as
 the variance across time at $(0,0)$ (bottom panel, orange).
@@
   \centering	
-  \includegraphics[width=.46\columnwidth]{Figures/convergence}  
-  \includegraphics[width=.46\columnwidth]{Figures/modelComp_MAE}  
+  \includegraphics[width=.48\columnwidth]{Figures/convergence}  
+  \includegraphics[width=.48\columnwidth]{Figures/modelComp_MAE}  
   \caption{Left: Value of the objective function for linearized (orange)
@@
     method with optimal values of $\lambda_t$ and $\lambda_s$ (2)
-    temporal penalty only (3) spatial
+    spatial penalty only (3) temporal
     penalty only and (4) a GARCH(1,1).} 
@@
 \autoref{fig:true_fitted_var} shows the boxplot of the MAE for these
-models. We note that, using an algorithm akin
-to~\citep{HallacPark2017} ignores the spatial component and
+models. As discussed above, using an algorithm akin
+to~\citep{HallacPark2017} ignores the spatial component and thus gives
+results which are similar to the second column when the covariances are shrunk to
+zero and massively worse otherwise.
 
@@
 Consensus ADMM in~\autoref{sec:consOpt} is appropriate when we can
-easily parallelize it over multiple machines.  Otherwise, it is significantly
+easily parallelize over multiple machines.  Otherwise, it is significantly
 slower, so all the results reported in this section are obtained
 using \autoref{alg:linADMM}. We
-applied this algorithm to the Northern hemisphere of the ERA-20C
+applied this algorithm to the Northern Hemisphere of the ERA-20C
 dataset available from the European Center for Medium-Range
@@
 
-\paragraph{Preprocessing}
+\paragraph{Preprocessing and other considerations}
 
@@
 strong differences 
-between trend and cyclic behavior across spatial locations. One might
+between trend and cyclic behavior across spatial locations (the data
+are not mean-zero). One might
 try to model the cycles by the summation of 
@@
 every time series. For each time-series, we found the optimal value of
-the penalty parameter using \textit{k-fold cross-validation} with
-$k=5$. We used the R package \textbf{genlasso} to perform these
-computations \cite{arnold_efficient_2016}.  
+the penalty parameter using 5-fold cross-validation with
+$k=5$. 
+% We used the R package \textbf{genlasso} to perform these
+% computations \cite{arnold_efficient_2016}.  
 \begin{figure}[tb]
 	\centering
-	\includegraphics[width=.3\textheight]{Figures/cities_ts}
+	\includegraphics[width=.66\columnwidth]{Figures/cities_ts}
  	\caption{Time-series of the temperature (in Kelvin) of three cities.}
@@
 \end{figure} 
-The blue curve in the left panel of Figure
+The blue curve in the left panel of 
 \autoref{fig:bloom_estimatedSD} shows the time-series of the
@@
   \includegraphics[width=.45 \columnwidth]{Figures/indy_estimatedSD_3}  
-  \caption{Left: The residuals of the time-series of Indianapolis
-    (averaged weekly) and the estimated SD obtained from the method of
-    \autoref{sec:l1tf_var} (red). Middle: the estimated SDs (red) and
+  \caption{Top: The variability of the time-series of Indianapolis
+    (weekly) and the estimated SD obtained from the method of
+    \autoref{sec:l1tf_var} (red). Lower left: the estimated SDs (red) and
     their annual average (black) without imposing the long horizon
-    penalty. Right: the same as middle panel but here the long horizon
-    penalty is imposed. See the text for more details. \attn{Fix the caption (top-bottom or left right)}} 
+    penalty. Lower right: the same but with the long horizon
+    penalty is imposed.} 
   \label{fig:bloom_estimatedSD}
 \end{figure} 
-
-The red curve in the left panel of Figure \autoref{fig:bloom_estimatedSD}
-shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
-the time-series of Indianapolis obtained from out proposed model. For
-ease of analysis, we compute the average
-of the estimated SD for each year. Both are shown in the middle panel of
-\autoref{fig:bloom_estimatedSD}. To smooth the annual trend, we add
-a long horizon penalty to \eqref{eq:l1tf_var}.
-The estimated, smoothed SDs are shown
+The red curve 
+shows the estimated SD (which is $\exp(h_t/2)$) obtained from out
+proposed model. For ease of analysis, we compute the average
+of the estimated SD for each year. Both are shown in the lower left panel of
+\autoref{fig:bloom_estimatedSD}.
+
+In addition to the constraints discussed in \autoref{sec:exten}, we
+, we add
+a long horizon penalty to smooth the annual trend.
+$\sum_{i=1}^{N_{year}-2}\left| \sum_t h_{t_1}-2h_{t_2}+h_{t_3}\right|$
+where $N_{year}$ is the number of years, $t_1=52(i-1)+t$, $t_2=52i+t$
+and $t_3=52(i+1)+t$. Finally, because the observations are on the surface
+of a hemisphere rather than a grid, we add extra spatial constraints with the
+obvious form to handle the boundary between
+$180^{\circ}$W and $180^{\circ}$E as well as the region near the North Pole.
+The estimated SDs for Indianapolis are shown
 the right panel of \autoref{fig:bloom_estimatedSD}. The annual average
@@
 
-\paragraph{Convergence}
+%\paragraph{Convergence and warm starts}
 
@@
 $\epsilon=0.001\%$ of the MSE of the data.
-Our simulation experiments showed that the
+Our simulation experiments indicated that the
 convergence speed depends on the value of $\lambda_t$ and
-$\lambda_s$. Furthermore, using the solution obtained for smaller values
-of these parameters as a warm start for the larger values, the
-converges speed improves.
+$\lambda_s$. For the temperature data, we used the solutions obtained for smaller values
+of these parameters as a warm start for the larger values resulting in
+faster convergence.
 
@@
 criterion. However, such analysis needs the computation of the degrees
-of freedom. Several previous work have investigated the df in
-generalized lasso problems
+of freedom. Several previous work have investigated the degrees of freedom in
+generalized lasso problems with Gaussian likelihood
 \cite{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}. However,
-all these studies have considered the linear regression problem and,
-to the best of our knowledge, the problem of computing the df for
-generalized lasso with general objective function has not been
-considered yet. In this paper, we use a heuristic method for choosing $\lambda_t$ and
-$\lambda_s$: we compute the optimal solution for a range of values of
+similar computations for non-Gaussian likelihood remains an open problem.
+In this paper, therefore, we use a heuristic method for choosing $\lambda_t$ and
+$\lambda_s$: we compute the solutions for a range of values of
 these parameters and choose the values which minimize
@@
 \lambda_s \in \{0,.1,.5,2,5,10\}$. The best combination was
-$\lambda_t=4$ and $\lambda_s=2$. All the analyses in the next section
-are performed on the solution for these values.
+$\lambda_t=4$ and $\lambda_s=2$.
 
@@
   \centering
-  \includegraphics[width=1.1\columnwidth]{Figures/ts_estimatedVar}
-  \caption{Detrended data and the estimated SD for
-    a small midwestern city (left) and San Diego (right).\attn{Fix these}} 
+  \includegraphics[width=.66\columnwidth]{Figures/ts_estimatedVarLeft}
+  \includegraphics[width=.66\columnwidth]{Figures/ts_estimatedVarRight}
+  \caption{Difference in observed data from the estimated trend
+    (blue), the estimated SDs (orange), and annual average SD (green) for
+   Indianapolis (top) and San Diego (bottom). The $y$-axis shows $\textrm{K}^{\circ}$.} 
   \label{fig:avg_change_estimatedSD}
@@
 
-The top row of \autoref{fig:avg_change_estimatedSD} shows the
+\autoref{fig:avg_change_estimatedSD} shows the
 detrended data, the estimated standard deviation and the yearly
-average of these estimates for two cities in the US: a small
-midwestern city (left) and San Diego (right). The estimated SD
+average of these estimates for two cities in the US: Indianapolis
+(top) and San Diego (bottom). The estimated SD 
 captures the periodic behavior in the variance of the time-series. In
@@
 temperature volatility. For example, we can see that the variance in
-the midwestern city displays a small positive trend. To determine how the volatility has
+Indianapolis displays a small positive trend. To determine how the volatility has
 changed in each location, we subtract the average of the estimated
 variance in 1961 from the average in the following years and compute
-their sum. The value of this change in the variance in each location
-is depicted in the right panel of
-\autoref{fig:avg_chg}. The left panel of this
-shows the average estimated variance in each location. Since the
+their sum. The average estimated variance at each location is shown in
+the top panel of~\autoref{fig:avg_chg} while the change from 1961
+is depicted in bottom panel. Since the
 optimal value of the spatial penalty is rather large ($\lambda_s=2$)
@@
 In this paper, we proposed a new method for estimating the variance of
-spatio-temporal data. The main idea is to cast this problem as a
+spatio-temporal data with the goal of analyzing global temperature
+data. The main idea is to cast this problem as a
 constrained optimization problem where the constraints enforce smooth
@@
 
-The consensus ADMM algorithm converges in a few hundreds of iterations
+The consensus ADMM algorithm converges in a few hundred iterations
 but each iteration takes much longer than the linearized ADMM
@@
 \end{equation}
-
-\noindent for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form
-
+where $\mu_1$ and $\mu_2$ are dual variables for the $\ell_\infty$ constraint. Let $A=[\nabla r_{dual}^\top , \nabla r_{cent}^\top]^\top$. The newton step takes the following form
 \begin{equation}
@@
 \end{equation}
-
-We have:
-
+with
 \begin{equation}
@@
 	\mathbf{diag(\mu_2)}\one & v+\lambda\one & \mathbf{0}
-	\end{bmatrix}
+	\end{bmatrix}.
 \label{eq:delta_r}
 \end{equation}
-
-Therefore, to perform the Newton step we need to compute $\nabla f^*(-D^\top v)$ and $\nabla^2 f^*(-D^\top v)$. It is straightforward to show that
-
+Therefore, to perform the Newton step we need to compute $\nabla
+f^*(-D^\top v)$ and $\nabla^2 f^*(-D^\top v)$.
+
+It is straightforward to show that
 \begin{align}
-& \nabla f^*(-D^\top v) =  -\nabla_u f^*(u) D^\top ,\\
-& \quad u=-D^\top v, \quad (\nabla_u f^*(u))_j=\log\bigg(\frac{y_j^2}{1-u_j}\bigg) \\
-& \nabla^2 f^*(-D^\top v)=D\nabla_u^2 f^*(u)D^\top, \quad (\nabla_u^2 f^*(u))_j=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg)
+\nabla f^*(-D^\top v) &=  -\nabla_u f^*(u) D^\top ,\\
+u&=-D^\top v, \\
+ (\nabla_u f^*(u))_j&=\log\bigg(\frac{y_j^2}{1-u_j}\bigg), \\
+\nabla^2 f^*(-D^\top v)&=D\nabla_u^2 f^*(u)D^\top,\\  
+(\nabla_u^2 f^*(u))_j&=\mathbf{diag}\bigg(\frac{1}{1-u_j}\bigg).
 \end{align}
 
-Having computed the conjugate function and its gradient and Jacobian, now we can use a number of convex optimization software packages which have an implementation of PDIP to solve the optimization problem with the objective function \ref{eq:l1tf_var}.
-We chose the python API of the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.
+Having computed the conjugate function and its gradient and Jacobian,
+now we can use a number of convex optimization software packages which
+have an implementation of PDIP to solve the optimization problem with
+the objective function \eqref{eq:l1tf_var}. We chose the python API of
+the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.
+
+\section{Proof of Lemma 1}
+
+\begin{proof}
+  If $f(x)=\sum_k f_k(x_k)$ then $[\prox_{\mu f}(x)]_k =
+  \prox_{\mu f_k}(u_k)$. So 
+  $[\prox_{\mu f}(u)]_k=\min_{x_k} \,\,
+  \mu\big(x_k+y_{k}^2e^{-x_{k}}\big)+\frac{1}{2}  (x_k-u_k)^2.$
+  Setting the derivative to 0 and solving for $u_k$ gives the
+  result. Similarly, $[\prox_{\rho g}(u)]_\ell=\rho
+  \lambda_\ell |z_\ell|+1/2(z_\ell-u_\ell)^2$. This is not differentiable,
+  but the solution must satisfy $\rho \cdot \lambda_\ell \cdot \partial
+  \big(|z_\ell| \big)=u_\ell-z_\ell$ where $\partial \big(|z_\ell| \big)$ is the
+  sub-differential of $|z_\ell|$. The solution is the soft-thresholding
+  operator $S_{\rho\lambda_\ell}(u_\ell)$.
+\end{proof}
 
@@
 &  & D_t
-\end{bmatrix}
+\end{bmatrix},
 \label{eq:d_t_matrix}
 \end{equation}
-
-\noindent where $D_t$ was first introduced in Section 2 of the main text and has the following form:
-
+where $D_t$ was first introduced in Section 2 of the main text and has the following form:
 \begin{equation}
@@
  &  &  &   & 1 & -2 & 1 
- \end{bmatrix}
+ \end{bmatrix}.
 \label{eq:d_matrix}
@@
 
-The number of the diagonal blocks in $D_{temp}$ is equal to the grid size $n_r \times n_c$.
-Each row of the matrix $D_{spat}$ corresponds to one spatial constraint in Equation \eqref{eq:l1tf_var_st} in the text.
-For example, the first $T$ rows correspond to $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to $|h_{11t}-h_{12t}|$, and so on.
-
-This optimization problem, is again a generalized LASSO problem with $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$.
-
-As it was explained in Appendix A, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$.
-To use PDIP we first need to compute the conjugate function $f^*(\cdot)$.
-We have:
+The number of the diagonal blocks in $D_{temp}$ is equal to the grid
+size $n_r \times n_c$. Each row of the matrix $D_{spat}$ corresponds
+to one spatial constraint in Equation \eqref{eq:l1tf_var_st} in the
+text. For example, the first $T$ rows correspond to
+$|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to
+$|h_{11t}-h_{12t}|$, and so on.
+
+This optimization problem, is again a generalized LASSO problem with
+$f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert
+x-\tilde{z} + u \lVert_2^2$.
+
+As it was explained in Appendix A, the dual of this optimization
+problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k|
+\le \Lambda_k$. To use PDIP we first need to compute the conjugate
+function $f^*(\cdot)$. We have:
 
@@
 % \end{figure}
+
+\end{document}
 

### 2018-09-05 aaai-2018/khodadadi-mcdonald-aaai19.tex
proof

@@
   /Title (Algorithms for Estimating Trends in Global Temperature Volatility)
-  % /Author (Khodadadi and McDonald)
+   /Author (Khodadadi and McDonald)
   /Author (John Doe, Jane Doe)}
@@
 \begin{abstract}
-Trends in terrestrial temperature variability are perhaps more relavant for species
+Trends in terrestrial temperature variability are perhaps more relevant for species
 viability than trends in mean
@@
 
-Trends in terrestrial temperature variability are perhaps more relavant for species
+Trends in terrestrial temperature variability are perhaps more relevant for species
 viability than trends in mean
@@
  where $y=\{y_t\}_{t=1}^T$ is an observed time-series, $\beta\in \mathbb{R}^T$ is the smooth trend,
- $D_t$ is a $(T-2)\times T$ matrix, and $\lambda$ is a tuning parameter
+ $D_t$ is a $(T-2)\times T$ matrix (see Appendix C), and $\lambda$ is a tuning parameter
  which balances fidelity to the data (small errors in the first term)
@@
 $TS$ and $D_t$ is replaced by $D \in
-\mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot
+\mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$ (see Appendix C), where $N_t=S \cdot
 (T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and
@@
 \end{figure} 
-The blue curve in the left panel of 
+The blue curve in the top panel of 
 \autoref{fig:bloom_estimatedSD} shows the time-series of the
@@
 The red curve 
-shows the estimated SD (which is $\exp(h_t/2)$) obtained from out
+shows the estimated SD (which is $\exp(h_t/2)$) obtained from our
 proposed model. For ease of analysis, we compute the average
@@
 $180^{\circ}$W and $180^{\circ}$E as well as the region near the North Pole.
-The estimated SDs for Indianapolis are shown
-the right panel of \autoref{fig:bloom_estimatedSD}. The annual average
+The estimated SDs for Indianapolis are shown in the lower right panel of \autoref{fig:bloom_estimatedSD}. The annual average
 of the estimated SDs shows a linear trend with a positive slope.

### 2018-09-06 aaai-2018/aaai-paper-supplement.tex
aaai submitted version

--- 2018-09-04 aaai-2018/khodadadi-mcdonald-aaai19.tex
@@
-%File: formatting-instruction.tex
 \documentclass[letterpaper]{article} %DO NOT CHANGE THIS
@@
 \pdfinfo{
-  /Title (Algorithms for Estimating Trends in Global Temperature Volatility)
-  % /Author (Khodadadi and McDonald)
+  /Title (Supplement for Algorithms for Estimating Trends in Global Temperature
+  Volatility )
+   %/Author (Khodadadi and McDonald)
   /Author (John Doe, Jane Doe)}
@@
 \newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
+\newcommand{\y}{\textrm{year}}
 \DeclareMathOperator*{\prox}{\bf prox}
@@
 \usepackage{xr}
-\externaldocument{aaai-paper-supplement}
+\externaldocument{khodadadi-mcdonald-aaai19}
 
@@
 
-\title{Algorithms for Estimating Trends in Global Temperature Volatility}
+\title{Supplement for Algorithms for Estimating Trends in Global Temperature Volatility}
 
@@
 \maketitle
-
-\begin{abstract}
-Trends in terrestrial temperature variability are perhaps more relavant for species
-viability than trends in mean
-temperature. In this paper, we develop methodology for estimating such
-trends using multi-resolution climate data from polar orbiting weather
-satellites. We derive two novel algorithms for computation that are
-tailored for dense, gridded observations over both 
-space and time. We evaluate our methods with a simulation constructed
-to mimic these data's 
-features and on a large, publicly available, global
-temperature dataset with the eventual goal of tracking trends in
-cloud reflectance temperature variability.
-\end{abstract}
-
-\section{Introduction}
-
-% Estimating smooth trends over time for large collections of time
-% series is a common problem in economics, finance, meteorology,
-% neuroscience and more. Most of this work focuses on analyzing trends
-% (or removing trends) in the temporal average, but trends in variance
-% can be more relevant, especially for financial data but also for
-% climate science.
-
-The amount of sunlight reflected from clouds is among the largest
-sources of uncertainty in climate
-prediction~\citep{BoucherRandall2013}. But climate models fail to
-reproduce global cloud statistics, and understanding the reasons for
-this failure is a grand challenge of the World Climate Research
-Programme~\citep{BonyStevens2015}.
-% By understanding how cloudliness changes over time, examining changes
-% in brightness temperature within various cloud types, and inferring
-% changes in cloud radiative effects, atmospheric scientists can better model climate
-% change. Modeling inter-satellite biases, satellite drift, and
-% seasonal and long-term climate phenomena like El Ni\~{n}o will lead to
-% a better understanding of climate
-% change~\citep{SchreierKahn2014,BaumWielicki1994,BaumWielicki1992,FreyAckerman1996}.
-While numerous
-studies have examined the overall impacts of clouds on climate
-variability~\citep{MyersMechoso2018,GrisePolvani2013,BenderRamanathan2012},
-such investigations have been hampered by the lack of a suitable
-dataset. Ideal data would have global coverage at high spatial
-resolution, a long enough record to recover temporal trends, and be
-multispectral~\citep{WielickiYoung2013}. 
-% The International Satellite
-% Cloud Climatology Project (ISCCP)
-% has been producing cloud property information from
-% limited spectral channels for over three
-% decades~\citep{RossowSchiffer1991}. These data have been used in 
-% various climate variability
-% studies~\citep[e.g.]{BenderRamanathan2012}, however, the utility of
-% these data for long-term study of climate variability has been
-% questioned~\citep{EvanHeidinger2007}. 
-% ISCCP provides a relatively long record, but it only incorporates
-% radiances from limited visible and infrared channels from Advanced
-% Very High Resolution Radiometer (AVHRR) instruments. 
-Current work~\cite{StatenKahn2016,SchreierKahn2010,KahnFishbein2007} seeks to create a
-spectrally-detailed dataset by %which can avoid the above issues by
-combining radiance data from Advanced Very High Resolution Radiometer
-imagers with readings from High-resolution 
-Infrared Radiation Sounders on board legacy
-satellites. 
-In anticipation of this new dataset, our work develops novel methodology
-for examining the trends in variability of climate data across space
-and time.
-
-\subsection{Variability Rather Than Average}
-\label{sec:vari-rath-than}
-
-Trends in terrestrial temperature variability are perhaps more relavant for species
-viability than trends in mean
-temperature~\cite{huntingford_no_2013},
-because an increase in
-temperature variability will increase the probability of extreme hot
-outliers~\cite{VasseurDeLong2014}. Recent climate literature
-suggests that it is more difficult for society to adapt to these
-extremes than to the gradual increase in the mean temperature
-\cite{hansen_perception_2012,huntingford_no_2013}. Furthermore, the willingness of popular media to
-emphasize the prevalence extreme cold events coupled with a
-fundamental misunderstanding of the relationship between climate (the global
-distribution of weather over the long run) and weather (observed
-short-term, localized behavior) leads to public misunderstanding of climate
-change. In fact, a point of active debate is the extent to which the observed
-increased frequency of extreme cold events in the 
-northern hemisphere can be attributed to increases in temperature
-variance  rather than to increases in
-mean climate~\cite{Screen2014,FischerBeyerle2013,TrenberthZhang2014}.
-
-Nevertheless, research examining trends in the volatility of
-spatio-temporal climate data is scarce. \citet{hansen_perception_2012} studied the change in
-the standard deviation (SD) of the surface temperature in the NASA
-Goddard Institute for Space Studies gridded temperature data set by
-examining the empirical SD at each spatial location relative to that
-location's SD over a base period, and showed
-that these estimates are increasing.
-\citet{huntingford_no_2013} took a similar
-approach in analyzing the ERA-40 data set. They argued that, while there
-is an increase in the SDs from 1958-1970 to 1991-2001, it
-is much smaller than found in
-\cite{hansen_perception_2012}. \citet{huntingford_no_2013} also computed the
-time-evolving global SD from the de-trended time-series at each
-position and argued that the global SD has been
-stable.
-
-These and other related work  
-\cite[e.g.,]{rhines_frequent_2013} have several shortcomings. First, no
-statistical analysis has been performed to examine if the changes in
-the SD are statistically significant. Second, the methodologies for
-computing the SDs are highly sensitive to the choice of base period. Third,
-and most importantly, temporal and spatial correlations between 
-observations are completely ignored.
-
-In the present work, we examine variance (rather than the mean) for a number of reasons.
-First, instrument bias in the satellites increases over time so
-examining the mean over time conflates that bias with any actual
-change in mean (though the variance is unaffected). 
-Second, extreme weather events (hurricanes, droughts, wildfires in
-California, heatwaves in Europe) may be driven more strongly by increases
-in variance than by increases in mean. 
-Finally, even if the global mean temperature is constant, there may still
-be climate change. In fact, atmospheric physics suggests that,
-across space, average temperatures should not change (extreme cold
-in one location is offset by heat in another). But if swings across
-space are becoming more rapid, then, even with no change in mean global
-temperature over time, increasing variance can lead to
-increases in the prevalence of extreme events.
-
-\subsection{Main Contributions}
-
-The main contribution of this work is to develop a new methodology for
-detecting the trend in the volatility of spatio-temporal data. In this
-methodology, the variance at each position and time are estimated by
-maximizing penalized likelihood
-maximization. Following~\cite{Tibshirani2014}, we penalize the 
-differences between the estimated variances which
-are temporally and spatially ``close'', resulting
-in a generalized LASSO problem. However, in our application, the
-dimension of this optimization problem is massive, so the
-standard solvers are inadequate.
-
-We develop two algorithms
-which are computationally feasible on extremely large data. In the
-first method, we adopt an optimization technique 
-called alternating direction method of multipliers
-(ADMM, \citealp{boyd_distributed_2011}), to divide the total problem into
-several sub-problems of much lower dimension and show how the total
-problem can be solved by iteratively solving these sub-problems. The
-second method, called linearized ADMM
-\cite{parikh_proximal_2014}, solves the main problem by iteratively
-solving a linearized version of it. We will compare the benefits of
-each method.
-
-Our main contributions are as follows:
-\begin{enumerate}
-\item We propose a model for nonparametric variance estimation for a
-  spatio-temporal process and discuss the relationship between our
-  methods and those existing in the machine learning literature (\autoref{sec:ell_1-trend-filt}).
-\item We derive two alternating direction method of multiplier
-  algorithms (ADMM) to fit our estimator when applied to 
-  very large data (\autoref{sec:prop-optim-meth}). We give situations
-  under which each algorithm is most likely to be useful. Open-source
-  Python code is available on {\tt github}.
-\item Because the construction of satellite-based datasets is ongoing
-  and currently proprietary, we illustrate our methods on a large,
-  publicly available, global temperature dataset.
-  The goal is to demonstrate the feasibility of these methods for tracking
-  world-wide trends in variance in standard atmospheric data and a
-  simulation constructed to mimic these data's features
-  (\autoref{sec:empirical-evaluation}). 
-\end{enumerate}
-
-While the motivation for our methodology is its application to large, gridded climate
-data, we note that our algorithms are also applicable to neuroimaging,
-image denoising, or examining large collections of financial instruments.
-
-\section{Smooth Spatio-temporal Variance Estimation}
-\label{sec:ell_1-trend-filt}
-
-\citet{KimKoh2009} proposed $\ell_1$-trend filtering as a
-method for estimating a smooth, time-varying trend. It is formulated
-as the optimization problem 
-% : a time-series $y_t$, $t=1,...,T$ is observed and we believe that
-% it consists of a smooth trend and a stochastic component and the
-% goal is to estimate the trend. The estimated trend is desired to be
-% smooth, but at the same time, we seek small estimated residuals
-% (stochastic component). These two objectives cannot be achieved
-% simultaneously and so a trade-off between them should be made. In
-% $\ell_1$-trend filtering, this trade-off is formulated as the
-% following convex optimization problem: 
-$$
-\min_{\beta} \frac{1}{2} \sum_{t=1}^{T} (y_t-\beta_t)^2+\lambda
-\sum_{t=1}^{T-2} \left|\beta_t-2\beta_{t+1}+\beta_{t+2}\right| 
-$$
-or equivalently:
-\begin{equation}
-\min_{\beta} \frac{1}{2} \norm{ y-\beta }_2^2+\lambda \norm{ D_t \beta}_1
-\label{eq:l1tf}
-\end{equation}
- where $y=\{y_t\}_{t=1}^T$ is an observed time-series, $\beta\in \mathbb{R}^T$ is the smooth trend,
- $D_t$ is a $(T-2)\times T$ matrix, and $\lambda$ is a tuning parameter
- which balances fidelity to the data (small errors in the first term)
- with a desire for smoothness.  
-% The \textit{penalty matrix} is:
-% \begin{equation}
-% D=\begin{bmatrix}
-% 1 & -2 & 1 &  &  &  &\\ 
-% & 1 & -2 & 1 &  &  &\\ 
-% &  & \ddots & \ddots & \ddots  &  &\\ 
-% &  & & 1 & -2 & 1 &  \\ 
-% &  &  &   & 1 & -2 & 1 
-% \end{bmatrix}
-% \label{eq:d_matrix}
-% \end{equation}
-%The first term in \eqref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
-% In Appendix A, we specify the form of the matrix $D$ which results in piecewise
-% linear estimated $\beta$. 
-\citet{KimKoh2009} proposed a specialized primal-dual
-interior point (PDIP) algorithm for solving \eqref{eq:l1tf}. From a
-statistical perspective, \eqref{eq:l1tf} can be viewed as a constrained maximum
-likelihood problem with independent observations from a normal
-distribution with common variance, $y_t \sim \mbox{N}(\beta_t,
-\sigma^2)$, subject to a piecewise linear constraint on
-$\beta$. Alternatively, solutions to \eqref{eq:l1tf} are %equivalent to
-maximum a posteriori Bayesian estimators based on Gaussian likelihood
-with a special Laplace prior distribution on $\beta$. Note that the
-structure of the estimator is determined by the penalty function
-$\lambda\norm{D_t\beta}_1$ rather than any parametric trend
-assumptions---autoregressive, moving average, sinusoidal seasonal
-component, etc. The resulting trend is therefore essentially
-nonparametric in the same way that splines are nonparametric. In fact,
-using squared $\ell_2$ norm as the penalty instead of $\ell_1$ results exactly in
-regression splines.
-
-% The results of applying $\ell_1$-trend filtering on the Bloomington
-% time-series are shown in \autoref{fig:bloom_detrended}. The left
-% panel shows the original time-series (blue) together with the
-% estimated $\beta$ (red) and the right panel shows the residuals
-% $y_t-\beta_t$ (blue). As it can be seen, the estimated $\beta$s are
-% piecewise linear. The number of linear segments is determined by the
-% value of $\lambda$. Here, we chose $\lambda=500$. The choice of
-% $\lambda$ is a model selection problem and we will return to it in
-% the next section ???.
-
-\subsection{Modifications for Variance}
-\label{sec:l1tf_var}
-
-Inspired by the $\ell_1$-trend filtering algorithm, we propose a
-non-parametric model for estimating the variance of a time-series. To
-this end, we assume that at each time step $t$, there is a parameter
-$h_t$ such that the observations $y_t$
-are independent normal variables with zero mean and variance
-$\exp(h_t)$. The negative log-likelihood of the observed data in this
-model is $l(y\given h) \propto -\sum_{t=1}^T h_t - y_t^2e^{-h_t}$. Crucially,
-we assume that the parameters $h_t$ vary smoothly and estimate them by
-minimizing the penalized, negative
-log-likelihood:  
-\begin{equation}
-\min_h -l(y\given h)+\lambda \norm{ D_th }_1
-\label{eq:l1tf_var}
-\end{equation}
-where $D_t$ has the same structure as above.
-
-As with~\eqref{eq:l1tf}, one can solve~\eqref{eq:l1tf_var} using the
-PDIP algorithm (as in, e.g.,
-\texttt{cvxopt}, \citealp{andersen_cvxopt:_2013}). In each iteration of
-PDIP we need to compute a search direction by taking a Newton
-step on a system of nonlinear equations. For completeness, we
-provide the details in Appendix A of the Supplement, where we show how to
-derive the dual of this optimization problem and compute the first and
-second derivatives of the dual objective function.
-
-\begin{comment}
-\begin{algorithm}[tb]
-  \caption{PDIP for $\ell_1$ variance estimation}
-  \label{alg:pdip}
-  \begin{algorithmic}
-    \REQUIRE $\lambda>0$, $w>0$, $\nu\leftarrow 0$, $\mu_1\leftarrow
-    0$, $\mu_2\leftarrow 0$, $J\in\mathbb{Z}^+$, $\{w_k\}$
-    \COMMENT{Initialization} 
-    \FOR[Central path]{$k=1,2,\ldots$}
-    \FOR[Newton updates]{$j=1,\ldots,J$}
-    \STATE Solve $A [\Delta\nu\; \Delta\mu_1\;
-    \Delta\mu_2]^\top=r_{w_k}$ to find the search direction. 
-    \STATE $\quad\quad A$ is given in \ref{eq:delta_r}
-    \STATE Update $[\nu^{j+1}\; \mu^{j+1}_1\; \mu^{j+1}_2] \leftarrow
-    [\nu^{j}\; \mu^{j}_1\; \mu^{j}_2] + [\Delta\nu\; \Delta\mu_1\;
-    \Delta\mu_2]$. 
-    \ENDFOR
-   \ENDFOR
-   \RETURN $h=\log\frac{y^2}{1+D^\top\nu}$ 
-   \\
-   \attn{Is this right? Explicit form of $A$}\\
-   \attn{Arash: see appendix A}
-  \end{algorithmic}
-\end{algorithm}
-\end{comment}
-
-\subsection{Adding Spatial Constraints}
-\label{sec:exten}
-
-The method in the previous section can be used to estimate the
-variance of a single time-series. Here we extend this
-method to the case of spatio-temporal data.
-
-At a specific time $t$, the data are measured on a grid of points with
-$n_r$ rows and $n_c$ columns for a total of $S=n_r\times n_c$ spatial
-locations. Let $y_{ijt}$ denote the value of the 
-observation at time $t$ on the $i^\text{th}$ row and $j^\text{th}$
-column of the grid, and $h_{ijt}$ denote the corresponding
-parameter. We seek to impose both temporal and spatial smoothness 
-constraints on the hidden variables. Specifically, we seek a solution
-for $h$ which is piecewise linear in time and piecewise constant in
-space (although higher-order smoothness can be imposed with minimal
-alterations to the methodology). We achieve this goal
-by solving the following optimization problem: 
-{\scriptsize
-\begin{equation}
-\label{eq:l1tf_var_st}.
-\begin{aligned}
-&\min_h \sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}} +\lambda_t
-\sum_{i,j} \sum_{t=1}^{T-2}
-\left|h_{ijt}-2h_{ij(t+1)}+h_{ij(t+2)}\right|\\ 
-&+\lambda_s \sum_{t,j} \sum_{i=1}^{n_r-1} \left|h_{ijt}-h_{(i+1)jt}\right|
-+\lambda_s \sum_{t,i} \sum_{j=1}^{n_c-1} \left|h_{ijt}-h_{i(j+1)t}\right|
-\end{aligned}
-\end{equation}
-}%
-
-The first term in the objective is proportional to the negative
-log-likelihood, the second is the temporal penalty for the
-time-series at each location $(i,j)$, while the third and fourth,
-penalize the difference between the estimated variance of two
-vertically and horizontally adjacent points, respectively. The spatial
-component of this
-penalty is a special case of trend filtering on
-graphs~\cite{WangSharpnack2016} which penalizes the difference between
-the estimated values of the signal on the connected nodes (though the
-likelihood is different). As before,
-we can write \eqref{eq:l1tf_var_st} in matrix form where $h$ is a
-vector of length
-$TS$ and $D_t$ is replaced by $D \in
-\mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot
-(T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and
-spatial constraints, respectively. 
-% The exact form of this matrix in clarified in Appendix
-% B in the Supplement. 
-Then, as we have two different tuning parameters
-for the temporal and spatial components, we write $\Lambda
-=\left[\lambda_t\one_{N_t}^\top,\;
-  \lambda_s\one_{N_s}^\top\right]^\top$ leading
-to:\footnote{Throughout the paper, we use $|x|$ for both scalars and
-  vectors. For vectors we use this to denote a vector obtained by
-  taking the absolute value of each entry of $x$.}  
-% let $h$ be a
-% vector whose first $T$ entries are $h_{11t}$ for $t=1,...,T$, its next
-% $T$ entries are $h_{21t}$ and so on, then the optimization problem in
-% the matrix form is as
-\begin{equation}
-\min_h -l(y\given h)+ \Lambda^\top | Dh |.
-\label{eq:l1tf_var_st_mat}
-\end{equation}
-
-\subsection{Related Work}
-
-Variance estimation for financial time series has a lengthy history,
-focused especially on parametric models like the generalized
-autoregressive conditional heteroskedasticity (GARCH) process~\cite{engle2002dynamic} and
-stochastic volatility models~\cite{HarveyRuiz1994}. These models (and
-related AR processes) are specifically for parametric modelling of
-short ``bursts'' of high volatility, behavior typical of financial
-instruments. Parametric models for spatial data go back at least
-to~\cite{besag1974spatial} who proposed a conditional probability
-model on the lattice for examining plant ecology.
-
-More recently, nonparametric models for both spatial and temporal data
-have focused on using $\ell_1$-regularization for trend
-estimation. \citet{KimKoh2009} proposed $\ell_1$-trend filtering for
-univariate time series, which forms the basis of our methods. These
-methods have been generalized to higher order temporal smoothness
-~\cite{Tibshirani2014}, graph dependencies~\cite{WangSharpnack2016},
-and, most recently, small, time-varying
-graphs~\cite{HallacPark2017}.
-
-Our methodology is similar in flavor to~\cite{HallacPark2017} or
-related work in~\citep{GibberdNelson2017,MontiHellyer2014}, but with
-several fundamental differences. These papers aim to
-discover the time-varying structure of a network. To achieve this
-goal, they use Gaussian likelihood with unknown precision matrix and
-introduce penalty terms which (1) encourage sparsity among the
-off-diagonal elements and (2) discourage changes in the estimated inverse
-covariance matrix from one time-step to the next.
-Our goal in the present work is to detect the temporal trend in the variance of each
-point in the network, but the network is known (corresponding to the grid
-over the earth). To modify the
-objective function in~\cite[Eq.\ 2]{HallacPark2017}, we would
-enforce complete sparsity on the off-diagonal elements (since
-they are not estimated) and add a new penalty to enforce spatial
-behavior across the diagonal elements. Thus, 
-\eqref{eq:l1tf_var_st_mat} is not simply a special case of these
-existing methods. Finally, these papers examine networks with hundreds
-of nodes and dozens to hundreds of time points. As discussed next, our data are
-significantly larger than these networks and attempting to estimate a
-full covariance would be prohibitive, were it required.
-
-\section{Optimization Methods}
-\label{sec:prop-optim-meth}
-
-For a spatial grid of size $S$ and $T$ time steps, $D$ in equation \ref{eq:l1tf_var_st_mat} will have
-$3Tn_rn_c-2n_rn_c-Tn_r$ rows and $S \cdot T$ columns. For a $1^\circ\times
-1^\circ$ grid over the entire northern hemisphere and daily data over
-10 years, we have $S=90\times 360\approx 32,000$ spatial locations and
-$T=3650$ time points, so $D$ has
-approximately $10^8$ columns and $10^8$ rows. In principal, we could
-solve~\eqref{eq:l1tf_var_st_mat} using PDIP as before, however, each iteration
-requires solving a linear system of equations which
-depends on $D^\top D$. Therefore,
-applying the PDIP directly is infeasible for our data.\footnote{We
-  note that $D$ is a highly structured, sparse matrix, but, unlike
-  trend filtering alone, it is not banded. We are unaware of general
-  linear algebra techniques for inverting such matrix, despite our
-  best efforts to find them.}
-
-In the next section, we develop two ADMM algorithms for solving this
-problem efficiently. The first casts the problem as a
-so-called consensus optimization problem~\cite{boyd_distributed_2011}
-which solves smaller sub-problems using PDIP and then recombines the results. The
-second uses proximal methods to avoid matrix inversions.
-
-\subsection{Consensus Optimization}
-\label{sec:consOpt}
-
-\begin{comment}
-General (unconstrained) optimization problems $\min_z
-\sum_i f_i(z)$, $z\in\mathbb{R}^n$ can be rewritten equivalently using a
-collection of local variables, $x_i\in \mathbb{R}^{n_i}$, and a
-constraint. Following the notation of 
-\cite{boyd_distributed_2011}, let $k=\mathscr{G}(i,j)$ which
-means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or
-$(x_i)_j=z_k$) and define $\tilde{z}_i \in \mathbb{R}^{n_i}$ by
-$(\tilde{z}_i)_j=(x_i)_j$. Then the original unconstrained
-optimization problem is equivalent to the following constrained
-problem
-$\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)$ subject to $\tilde{z}_i=x_i$.
-Now, we can apply ADMM to the augmented Lagrangian of this
-problem: 
-\begin{align}
-x_i&\leftarrow \argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
-     (\rho/2) \norm{ x_i-\tilde{z}_i }_2^2\\
-%\label{eq:ADMM_steps}
-z_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j\\
-u_i&\leftarrow u_i + \rho (x_i-\tilde{z}_i)
-\end{align}
-\end{comment}
-
-Consider an optimization problem of the form $\min_h f(h)$, where
-$h\in\mathbb{R}^n$ is the \textit{global variable} and
-$f(h):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is
-convex. Consensus optimization breaks this problem
-into several smaller sub-problems that can be solved independently in
-each iteration.
-
-Assume it is possible to define a set of local variables
-$x_i \in \mathbb{R}^{n_i}$ such that $f(h)=\sum_i f_i(x_i)$, where
-each $x_i$ is a subset of the global variable $h$. More specifically,
-each entry of the local variables corresponds to an entry of the
-global variable. Therefore we can define a mapping $\mathscr{G}(i,j)$
-from the local variable indices into the global variable indices:
-$k=\mathscr{G}(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is
-$h_k$ (or $(x_i)_j=h_k$). For ease of notation, define $\tilde{h}_i
-\in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{\mathscr{G}(i,j)}$. Then,
-the original optimization problem is equivalent to the following
-problem:  
-\begin{equation}
-\begin{aligned}
-&\min_{\{x_1,...,x_N  \}} \sum_{i=1}^N f_i(x_i) &
-s.t. &\quad \tilde{h}_i=x_i.
-\end{aligned}
-\label{eq:consADMM}
-\end{equation}
-It is important to note that each entry of the global variable may
-correspond to several entries of the local variables and so the
-constraints $\tilde{h}_i=x_i$ enforce the consensus between the local
-variables corresponding to the same global variable.  
-The augmented Lagrangian corresponding to 
-\eqref{eq:consADMM} is $L_\rho(x,h,y)=\sum_i
-\big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert
-x_i-\tilde{h}_i \lVert_2^2 \big)$. Now, we can apply ADMM to
-$L_\rho$. This results in solving $N$ independent optimization
-problems followed by a step to achieve consensus among the solutions
-in each iteration. 
-% which results in the following ADMM updates:
-% \begin{align}
-% x_i&\leftarrow\argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
-%      (\rho/2)  \norm{ x_i-\tilde{h}_i }_2^2\\
-% h_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j \label{eq:ADMM_steps}\\
-% u_i&\leftarrow u_i + \rho (x_i-\tilde{h}_i).
-% \end{align}
-% \attn{If you hate the left arrow, that's fine. But change it
-%   everywhere, and use $=$. The $:=$ notation means that ``thing on
-%   left is defined to be thing on right'' and isn't really appropriate here.}
-% Here, $S_k$ is the number of local variable entries that correspond to
-% $h_k$, and $u_i$ are the Lagrange multipliers.  
-To solve the optimization problem \eqref{eq:l1tf_var_st_mat} using
-this method, we need to address two 
-questions: first, how to choose the 
-local variables $x_i$, and second, how to the update them.
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=.6\columnwidth]{Figures/data_cube}
-  \caption{The cube represents the global variable $h$ in space and
-    time. The sub-cubes specified by the white lines are
-    $x_i$.}
-  \label{fig:data_cube}
-\end{figure}
-
-In \autoref{fig:data_cube}, the global variable $h$ is represented
-as a cube. We decompose $h$
-into sub-cubes as shown by white lines. Each global variable inside
-the sub-cubes corresponds to only one local variable. The global
-variables on the border (white lines), however, correspond to more
-than one local variable. With 
-this definition of $x_i$, the objective
-\eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where
-$f_i(x_i)=-l(y_i\given x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|$, and
-$\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
-penalties corresponding to $x_i$ only in one sub-cube along with its
-boundary. Thus, we now need to use PDIP to solve $N$ problems each of
-size $n_i$, which is feasible for small enough
-$n_i$. \autoref{alg:conADMM} gives the general version of this
-algorithm. A more detailed discussion of this is in Appendix B of the
-Supplement where we show how to compute the dual and the derivatives
-of the augmented Lagrangian.
-
-\begin{algorithm}[tb]
-  \caption{Consensus ADMM }
-  \label{alg:conADMM}
-  \begin{algorithmic}[1]
-    \STATE {\bfseries Input:} data $y$, penalty matrix $D$, 
-    $\epsilon, \rho,\lambda_t,\lambda_s >0$.
-    \STATE {\bf Set:} $h\leftarrow 0$, $z\leftarrow 0$, $u\leftarrow
-    0$. \COMMENT{Initialization} 
-    \REPEAT
-    \STATE $\begin{aligned}x_i&\leftarrow\argmin_{x_i} -l(y_i\given
-    x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|\\&\quad+ (u_i)^\top x_i +
-    (\rho/2)  \lVert x_i-\tilde{h}_i \rVert_2^2\end{aligned}$. \\\COMMENT{Update local
-      vars using PDIP}
-    \STATE $h_k\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j
-    $. \COMMENT{Global update.}
-    \STATE $ u_i\leftarrow u_i + \rho (x_i-\tilde{h}_i)$. \COMMENT{Dual update}
-    \UNTIL {$\max\left\{\norm{h^{m+1}-h^m},\ \norm{h^m-x^m}\right\} < \epsilon$}
-    \STATE {\bf Return:} $h$.
-  \end{algorithmic}
-\end{algorithm}
-
-Because consensus ADMM breaks the large optimization into
-sub-problems that can be solved independently, it is amenable to a
-split-gather parallelization strategy via, e.g., the map reduce framework.
-In each iteration, the
-computation time will be equal to the time to solve each sub-problem
-plus the time to communicate the solutions on the master processor
-and perform the consensus step. Since each sub-problem is
-small, with parallelization, the computation time in each iteration
-will be small. In addition, our experiments with several values of
-$\lambda_t$ and $\lambda_s$ showed that the algorithm converges in a few
-hundred iterations. 
-% \attn{Need to redo this:}Solving each sub-problem on a machine with four
-% 3.20GHz Intel i5-3470 cores takes less than 3 seconds on average, and
-% so for example if we assume that communication time is 10 seconds and
-% the algorithm converges in 300 iterations, with parallelization on
-% $N_{sub-cubes}$ machines, the algorithm will converge in about 1
-% hour. Assuming that we use $N_{sub-cubes}$ machines and that the
-% convergence rate of the algorithm is independent of the grid size,
-% this time will be independent of the grid size. 
-% If we perform these computations on a single machine, the computation
-% time grows linearly with $N_{sub-cubes}$. For example, for the data in
-% a grid over the united states and using $3\times3\times521$ sub-cubes
-% each iteration of the algorithm will take about 20 minutes on a single
-% machine and so with 300 iterations it will take several days to
-% converge. Given that we need to compute the solution for several
-% values of the parameters $\lambda_t$ and $\lambda_s$, this computation
-% time is not feasible. 
-This algorithm is most useful if we can parallelize the
-computation over several machines with low communication cost between
-machines. In the next section, we describe 
-another algorithm which makes the computation feasible on a single
-machine.
-
-\subsection{Linearized ADMM}
-\label{sec:linADMM}
-
-% In this section, we describe \textit{Linearized ADMM algorithm} \cite{parikh_proximal_2014} which, as we will see, makes the computation on a single machine feasible.
-
-Consider the generic optimization problem
-$\min_x f(x)+g(Dx)$
-where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each
-iteration of the linearized ADMM
-algorithm~\cite{parikh_proximal_2014} for solving this problem 
-has the form
-\begin{equation}
-  \label{eq:linADMM_steps}
-  \begin{aligned}
-    x & \leftarrow \prox_{\mu f} \left(x - (\mu/\rho)D^\top (D x - z + u )\right)\\
-    z & \leftarrow \prox_{\rho g} \left(z + u\right)\\
-    u & \leftarrow u + D x - z
-  \end{aligned}
-\end{equation}
-where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu <
-\rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is
-defined as $\prox_{\alpha \varphi}(u) = \min_x \,\, \alpha \cdot
-\varphi(x)+\frac{1}{2} \norm{ x-u}_2^2$. Proximal algorithms are feasible
-when these proximal operators can be evaluated efficiently which, as
-we show next, is the case.
-
-%Clearly, \eqref{eq:l1tf_var_st} has this form necessary for using this algorithm.
-% The optimization problem 
-% form \eqref{eq:linADMM_opt} by defining
-% \begin{align}
-% & f(x):= \sum_{k} f_k(x_k) := \sum_{k}x_k+y_{k}^2e^{-x_{k}}\\
-% & g(z):= \sum_{l} g_l(z_l) := \sum_{l} \lambda_l |z_l|\label{eq:linADMM_fg}\\
-% & z = Dx
-% \end{align}
-% where $y_k$ is the $k^{th}$ entry of the vector whose entries are
-% $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector
-% $\Lambda^\top=(\lambda_t\textbf{e}_{n_t}^\top|\lambda_s\textbf{e}_{n_s}^\top)$
-% (see \autoref{sec:exten}). 
-% To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate
-% $\prox_{\mu f}$ and $\prox_{\rho g}$.
-
-\begin{lemma}
-  Let $f(x) = \sum_{k} x_{k} + y_{k}^2e^{-x_{k}}$ and $g(x) =
-  \norm{x}_1$. Then, 
-  \begin{equation}
-    \begin{aligned}
-      [\prox_{\mu f}(u)]_k &= \mathscr{W}\bigg(\frac{y_k^2}{\mu}
-      \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu u_k}{\mu},\\
-      \prox_{\rho g}(u) &= S_{\rho\lambda}(u)
-    \end{aligned}
-  \end{equation}
-  where $\mathscr{W}(\cdot)$ is the \textit{Lambert W function}
-  \cite{corless_lambertw_1996},  $[S_{\alpha}(u)]_k = \sign(u_k)(|u_k|
-  -\alpha_k)_+$ and $(v)_+=v\vee 0$.
-\end{lemma}
-% \begin{proof}
-%   If $f(x)=\sum_k f_k(x_k)$ then $[\prox_{\mu f}(x)]_k =
-%   \prox_{\mu f_k}(u_k)$. So 
-%   $[\prox_{\mu f}(u)]_k=\min_{x_k} \,\,
-%   \mu\big(x_k+y_{k}^2e^{-x_{k}}\big)+\frac{1}{2}  (x_k-u_k)^2.$
-%   Setting the derivative to 0 and solving for $u_k$ gives the
-%   result. Similarly, $[\prox_{\rho g}(u)]_\ell=\rho
-%   \lambda_\ell |z_\ell|+1/2(z_\ell-u_\ell)^2$. This is not differentiable,
-%   but the solution must satisfy $\rho \cdot \lambda_\ell \cdot \partial
-%   \big(|z_\ell| \big)=u_\ell-z_\ell$ where $\partial \big(|z_\ell| \big)$ is the
-%   sub-differential of $|z_\ell|$. The solution is the soft-thresholding
-%   operator $S_{\rho\lambda_\ell}(u_\ell)$.
-% \end{proof}
-The proof is fairly straightforward and given in Appendix B in the Supplement. 
-Therefore, \autoref{alg:linADMM} gives a different method for solving
-the same problem. For this algorithm, both the primal update and the
-soft thresholding step are performed elementwise at each point of the
-spatio-temporal grid. It can therefore be extremely fast to perform
-these steps. However, because there are now many more dual variables,
-this algorithm will require more outer iterations to achieve
-consensus. It therefore is highly problem and architecture dependent
-whether \autoref{alg:conADMM} or \autoref{alg:linADMM} will be more
-useful in any particular context. In our experience,
-\autoref{alg:conADMM} converges in many fewer iterations, but
-each iteration is much slower unless carefully parallelized.
-
-\begin{algorithm}[tb]
-  \caption{Linearized ADMM }
-  \label{alg:linADMM}
-  \begin{algorithmic}[1]
-    \STATE {\bfseries Input:} data $y$, penalty matrix $D$, 
-    $\epsilon, \rho,\lambda_t,\lambda_s >0$.
-    \STATE {\bf Set:} $h\leftarrow 0$, $z\leftarrow 0$, $u\leftarrow
-    0$. \COMMENT{Initialization} 
-    \REPEAT
-    \STATE $h_k\leftarrow \mathscr{W}\bigg(\frac{y_k^2}{\mu} 
-    \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu
-      u_k}{\mu}$ for all $k=1,\ldots TS$. \COMMENT{Primal update}
-    \STATE $z\leftarrow S_{\rho\lambda}(u)$. \COMMENT{Elementwise soft thresholding}
-    \STATE $u\leftarrow u + Dh-z$. \COMMENT{Dual update}
-    \UNTIL {$\max\{\norm{Dh-z},\; \norm{z^{m+1}-z^m}\} < \epsilon$}
-    \STATE {\bf Return:} $z$.
-  \end{algorithmic}
-\end{algorithm}
-
-\section{Empirical Evaluation}
-\label{sec:empirical-evaluation}
-
-In this section, we examine both simulated and real spatio-temporal
-climate data. All the computations were performed on a Linux machine
-with four 3.20GHz Intel i5-3470 cores.
-
-\subsection{Simulations}
-\label{sec:simulations}
-
-Before examining real data, we apply our model to some synthetic
-data. This example was constructed to mimic the types of spatial and
-temporal phenomena observable in typical climate data. We generate a
-complete spatio-temporal field wherein 
-observations at all time steps and all locations are
-independent Gaussian random variables with zero mean. However, the
-variance of these random variables follows a smoothly varying function
-in time and space given by the following parametric model:
-\begin{align}
-  \sigma^2(t,r,c) & =\sum_{k=1}^{K} W_k(t) \cdot \exp\bigg(
-                    \frac{(r-r_k)^2+(c-c_k)^2} {2\sigma_k^2} \bigg)\\
-  W_k(t) & =\alpha_k \cdot t + \exp(\sin(2\pi\omega_k t+\phi_k)) .
-\label{eq:sourceVar}
-\end{align}
-The variance at each time and location is computed as the
-weighted sum of $K$ bell-shaped functions where the weights are
-time-varying, consist of a linear trend $\alpha_k \cdot t$ and a
-periodic term $\beta_k \cdot \sin(2\pi\omega_k t+\phi_k)$. The
-bell-shaped functions impose spatial smoothness while the linear
-trend and the periodic terms enforce the temporal smoothness similar
-to the seasonal component in the real climate data. We simulated the
-data on a 5 by 7 grid and for 780 time steps with $K=4$. This yields a
-small enough problem so that it can be evaluated many times while
-still mimicking important properties of climate data.
-Specific parameter choices of the variance function are shown in
-\autoref{tab:sim_params}.
-For illustration, we also plot the
-variance function for all locations at $t=25$ and $t=45$
-(\autoref{fig:true_var_spatial}, top panel) as well as
-the variance across time at $(0,0)$ (bottom panel, orange).
-\begin{table}[tb]
-  \caption{Parameters used to simulate data.}
-  \label{tab:sim_params}
-  \begin{center}
-    \begin{tabular}{ccccccc}
-      \hline
-      $s$ & $r_s$ & $c_s$ & $\sigma_s$ &$\alpha_s$ & $\omega_s$ & $\phi_s$\\
-      \hline
-      1 & 0 & 0 & 5 & 0.5 & 0.121 & 0 \\
-      2 & 0 & 5 & 5 & 0.1 & 0.121 & 0 \\
-      3 & 3 & 0 & 5 & -0.5 & 0.121 & $\pi/2$ \\
-      4 & 3 & 5 & 5 & -0.1 & 0.121 & $\pi/2$ \\
-      \hline
-    \end{tabular}
-  \end{center}
-\end{table} 
-\begin{figure}[tb]
-  \centering	
-  \includegraphics[height=.1\textheight]{Figures/true_var_spatial}\\
-  \includegraphics[height=.15\textheight]{Figures/true_fitted_var}
-  \caption{Top: Variance function at $t=25$ (left) and $t=45$
-    (right). Bottom: The true (orange) and estimated standard deviation 
-    function at the location (0,0). The estimated values are
-    obtained using linearized ADMM with $\lambda_s=0.1$ and two
-    values of $\lambda_t$: $\lambda_t=5$ (blue) and
-    $\lambda_t=100$ (green).} \label{fig:true_var_spatial}
-\end{figure}
-
-We estimated the linearized ADMM for all combinations of values of
-$\lambda_t$ and $\lambda_s$ from the sets $\lambda_t \in
-\{0,1,5,10,50,100\}$ and $\lambda_s \in \{0,0.05,0.1,0.2,0.3\}$. For
-each pair, we then compute the mean absolute error (MAE) between the
-estimated variance and the true variance at all locations and all time
-steps. For $\lambda_t=5$ and $\lambda_s=0.1$, the MAE was minimized. 
-\begin{figure}[tb]
-  \centering	
-  \includegraphics[width=.48\columnwidth]{Figures/convergence}  
-  \includegraphics[width=.48\columnwidth]{Figures/modelComp_MAE}  
-  \caption{Left: Value of the objective function for linearized (orange)
-    and consensus (blue) ADMM against iteration. Right: MAE for (1) our
-    method with optimal values of $\lambda_t$ and $\lambda_s$ (2)
-    spatial penalty only (3) temporal
-    penalty only and (4) a GARCH(1,1).} 
-  \label{fig:true_fitted_var}
-\end{figure}
-The
-bottom panel of \autoref{fig:true_var_spatial} shows the true and the
-estimated standard deviation at location (0,0) 
-and $\lambda_t=5$ (blue) and $\lambda_t=100$ (green) ($\lambda_s=0.1$).
-Larger values of $\lambda_t$ lead to estimated values
-which are ``too smooth". The left panel of \autoref{fig:true_fitted_var} shows the
-convergence of both methods as a function of iteration. It is
-important to note that each iteration of the linearized
-algorithm takes 0.01 seconds on average while each iteration of the
-consensus ADMM takes about 20 seconds. Thus, where the lines meet at
-400 iterations requires about 4 seconds for the linearized method and
-2 hours for the consensus method.
-
-To further examine the performance of the proposed model, we next
-compare it to three alternatives: a model which does not consider the
-spatial smoothness (equivalent to fitting the model in
-\autoref{sec:l1tf_var} to each time-series separately), a model which only imposes spatial smoothness, and a GARCH(1,1) model. We
-simulated 100 datasets using the method explained above with $\sigma_s
-\sim \mathsf{uniform}(4,7)$. The right panel of
-\autoref{fig:true_fitted_var} shows the boxplot of the MAE for these
-models. As discussed above, using an algorithm akin
-to~\citep{HallacPark2017} ignores the spatial component and thus gives
-results which are similar to the second column when the covariances are shrunk to
-zero and massively worse otherwise.
-
-\subsection{Data Analysis}
-\label{sec:data}
-
-Consensus ADMM in~\autoref{sec:consOpt} is appropriate when we can
-easily parallelize over multiple machines.  Otherwise, it is significantly
-slower, so all the results reported in this section are obtained
-using \autoref{alg:linADMM}. We
-applied this algorithm to the Northern Hemisphere of the ERA-20C
-dataset available from the European Center for Medium-Range
-Weather Forecasts\footnote{\url{https://www.ecmwf.int}}. We use 
-the 2 meter temperature measured daily at 12 p.m from January 1, 1960
-to December 24, 2010.
-
-\paragraph{Preprocessing and other considerations}
-
-Examination of the time-series alone demonstrates
-strong differences 
-between trend and cyclic behavior across spatial locations (the data
-are not mean-zero). One might
-try to model the cycles by the summation of 
-sinusoidal terms with different frequencies. However, for some
-time-series this may need many terms to be included in the summation
-to achieve a reasonable level of accuracy. In addition, such a model
-cannot capture the non-stationarity in the cycles.
-
-\autoref{fig:cities_ts} shows the time-series of the 
-temperature of three cities: Indianapolis (USA), San Diego (USA) and
-Manaus (Brazil). The time-series of Indianapolis and San Diego show
-clear cyclic behavior, though the amplitude of the cycles
-changes. The time-series of Manaus does not
-show any regular cyclic behavior. For this reason, we
-first apply trend filtering to remove seasonal terms and de-trend
-every time series. For each time-series, we found the optimal value of
-the penalty parameter using 5-fold cross-validation with
-$k=5$. 
-% We used the R package \textbf{genlasso} to perform these
-% computations \cite{arnold_efficient_2016}.  
-\begin{figure}[tb]
-	\centering
-	\includegraphics[width=.66\columnwidth]{Figures/cities_ts}
- 	\caption{Time-series of the temperature (in Kelvin) of three cities.}
- 	\label{fig:cities_ts}
-\end{figure} 
-The blue curve in the left panel of 
-\autoref{fig:bloom_estimatedSD} shows the time-series of the
-temperature of Indianapolis after detrending using this method.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=.8 \columnwidth]{Figures/indy_estimatedSD_1}
-  \includegraphics[width=.45 \columnwidth]{Figures/indy_estimatedSD_2}
-  \includegraphics[width=.45 \columnwidth]{Figures/indy_estimatedSD_3}  
-  \caption{Top: The variability of the time-series of Indianapolis
-    (weekly) and the estimated SD obtained from the method of
-    \autoref{sec:l1tf_var} (red). Lower left: the estimated SDs (red) and
-    their annual average (black) without imposing the long horizon
-    penalty. Lower right: the same but with the long horizon
-    penalty is imposed.} 
-  \label{fig:bloom_estimatedSD}
-\end{figure} 
-The red curve 
-shows the estimated SD (which is $\exp(h_t/2)$) obtained from out
-proposed model. For ease of analysis, we compute the average
-of the estimated SD for each year. Both are shown in the lower left panel of
-\autoref{fig:bloom_estimatedSD}.
-
-In addition to the constraints discussed in \autoref{sec:exten}, we
-, we add
-a long horizon penalty to smooth the annual trend.
-$\sum_{i=1}^{N_{year}-2}\left| \sum_t h_{t_1}-2h_{t_2}+h_{t_3}\right|$
-where $N_{year}$ is the number of years, $t_1=52(i-1)+t$, $t_2=52i+t$
-and $t_3=52(i+1)+t$. Finally, because the observations are on the surface
-of a hemisphere rather than a grid, we add extra spatial constraints with the
-obvious form to handle the boundary between
-$180^{\circ}$W and $180^{\circ}$E as well as the region near the North Pole.
-The estimated SDs for Indianapolis are shown
-the right panel of \autoref{fig:bloom_estimatedSD}. The annual average
-of the estimated SDs shows a linear trend with a positive slope.
-
-% The Supplement explains some preprocessing and investigates some
-% properties of the time-series of different locations on
-% earth. \autoref{fig:bloom_estimatedSD} shows a processed time-series for a
-% single location.
-% The variance of this time-series has an irregular cyclic
-% behavior. Additionally, the time-series of other locations show 
-% different patterns. These observations motivated the need to develop a
-% non-parametric framework for this
-% problem. \autoref{fig:bloom_estimatedSD} also shows the
-% estimated SD obtained using the method of
-% \autoref{sec:l1tf_var}.
-
-% Since the estimated SD captures the periodic
-% behavior of 
-% volatility, it is hard to distinguish any long-term trend. Therefore,
-% we compute the annual average of these 
-% estimates. However, as this figure shows, the annual trend is not
-% smooth. This is because in the optimization problem
-% \eqref{eq:l1tf_var}, the smoothness of the annual trend is not
-% encouraged. Therefore, in Appendix D we propose a simple method to
-% encourage the smoothness of the annual average trend. The estimated
-% SDs using this method is shown in the right panel of
-% \autoref{fig:bloom_estimatedSD}. The annual average of the estimated
-% SDs shows a linear trend with a positive slope.
-
-%\paragraph{Convergence and warm starts}
-
-As shown in \autoref{alg:linADMM}, we evaluated convergence using
-$\epsilon=0.001\%$ of the MSE of the data.
-Our simulation experiments indicated that the
-convergence speed depends on the value of $\lambda_t$ and
-$\lambda_s$. For the temperature data, we used the solutions obtained for smaller values
-of these parameters as a warm start for the larger values resulting in
-faster convergence.
-
-\paragraph{Model Selection}
-One common method for choosing the penalty parameters in the Lasso
-problems is to find the solution for a range of the values of these
-parameters and then choose the values which minimize a model selection
-criterion. However, such analysis needs the computation of the degrees
-of freedom. Several previous work have investigated the degrees of freedom in
-generalized lasso problems with Gaussian likelihood
-\cite{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}. However,
-similar computations for non-Gaussian likelihood remains an open problem.
-In this paper, therefore, we use a heuristic method for choosing $\lambda_t$ and
-$\lambda_s$: we compute the solutions for a range of values of
-these parameters and choose the values which minimize
-$\mathscr{L}(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D_{total}h
-\lVert$. This objective is a compromise between the negative log
-likelihood ($-l(y|h)$) and the complexity of the solution ($\sum
-\lVert D_{total}h \lVert$). For smoother solutions the value of $\sum
-\lVert D_{total}h \lVert$ will be smaller but with the cost of larger
-$-l(y|h)$. We computed the optimal solution for all the combinations of the
-following sets of values: $\lambda_t \in \{0,2,4,8,10,15,200,1000\} \, \, ,
-\lambda_s \in \{0,.1,.5,2,5,10\}$. The best combination was
-$\lambda_t=4$ and $\lambda_s=2$.
-
-\paragraph{Analysis of Trends in Temperature Volatility}
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=.66\columnwidth]{Figures/ts_estimatedVarLeft}
-  \includegraphics[width=.66\columnwidth]{Figures/ts_estimatedVarRight}
-  \caption{Difference in observed data from the estimated trend
-    (blue), the estimated SDs (orange), and annual average SD (green) for
-   Indianapolis (top) and San Diego (bottom). The $y$-axis shows $\textrm{K}^{\circ}$.} 
-  \label{fig:avg_change_estimatedSD}
-\end{figure}
-
-\autoref{fig:avg_change_estimatedSD} shows the
-detrended data, the estimated standard deviation and the yearly
-average of these estimates for two cities in the US: Indianapolis
-(top) and San Diego (bottom). The estimated SD 
-captures the periodic behavior in the variance of the time-series. In
-addition, the number of linear segments changes adaptively in each
-time window depending on how fast the variance is changing.  
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=.9\linewidth]{Figures/avg_logVar.pdf}\\
-  \includegraphics[width=.9\linewidth]{Figures/avg_change_logVar.pdf}
-  \caption{The
-    average of the detrended estimated variance over the northern hemisphere (top) and the
-    change in the variance from 1961 to 2011 (bottom).}
-\label{fig:avg_chg}
-\end{figure}
-
-The yearly average of the estimated SD captures the trend in the
-temperature volatility. For example, we can see that the variance in
-Indianapolis displays a small positive trend. To determine how the volatility has
-changed in each location, we subtract the average of the estimated
-variance in 1961 from the average in the following years and compute
-their sum. The average estimated variance at each location is shown in
-the top panel of~\autoref{fig:avg_chg} while the change from 1961
-is depicted in bottom panel. Since the
-optimal value of the spatial penalty is rather large ($\lambda_s=2$)
-the estimated variance is spatially very smooth.
-
-The SD in most locations on the northern
-hemisphere had a negative trend in this time period, though spatially,
-this decreasing pattern is localized mainly toward the extreme
-northern latitudes and over oceans. In many ways, this is consistent
-with climate change predictions: oceans tend to operate as a local
-thermostat, regulating deviations in local temperature, while warming polar
-regions display fewer days of extreme cold.
-% It is interesting to note that the trend in volatility is almost zero
-% over the oceans. 
-The most positive trend can be observed in Asia and
-particularly in south-east Asia.
-
-% \begin{figure}[tb]
-%   \centering
-%   \includegraphics[width=.9\linewidth]{Figures/hist_avg_change.pdf}
-%   \caption{The histogram of changes in estimated SD.}
-%   \label{fig:hist}
-% \end{figure}
-
-% \autoref{fig:hist} shows the
-% histogram of change in the estimated SD across the northern
-% hemisphere. The SD in most locations on the northern
-% hemisphere had a negative trend in this time period, though spatially,
-% this decreasing pattern is localized mainly toward the extreme
-% northern latitudes and over oceans. In many ways, this is consistent
-% with climate change predictions: oceans tend to operate as a local
-% thermostat, regulating deviations in local temperature, while warming polar
-% regions display fewer days of extreme cold.
-
-\section{Discussion}
-\label{sec:discussion}
-
-In this paper, we proposed a new method for estimating the variance of
-spatio-temporal data with the goal of analyzing global temperature
-data. The main idea is to cast this problem as a
-constrained optimization problem where the constraints enforce smooth
-changes in the variance for neighboring points in time and space. In
-particular, the solution is piecewise linear in time and piecewise
-constant in space. The resulting optimization is in the form of a
-generalized LASSO problem with high-dimension, and so applying the
-PDIP method directly is infeasible. We therefore developed two
-ADMM-based algorithms to solve this problem: the consensus ADMM and
-linearized ADMM.
-
-The consensus ADMM algorithm converges in a few hundred iterations
-but each iteration takes much longer than the linearized ADMM
-algorithm. The appealing feature of the consensus ADMM algorithm is
-that if it is parallelized on enough machines the
-computation time per iteration remains constant as the problem size
-increases. The linearized ADMM algorithm on the other hand converges
-in a few thousand iterations but each iteration is performed in a
-split second. However, since the algorithm converges in many
-iterations it is not very appropriate for parallelization. The reason
-is that after each iteration the solution computed in each machine
-should be broadcast to the master machine and this operation takes
-some time which depends on the speed of the network connecting the
-slave machines to the master. A direction for future research would be
-to combine these two algorithms in the following way: the problem
-should be split into the sub-problems (as in the consensus ADMM) but
-each sub-problem can be solved using linearized ADMM.
-
-% \attn{We did not do this. Change the below.}
-% We applied the linearized ADMM algorithm to the surface temperature
-% data on a grid over the united states, for years 1992-2002. The
-% results showed that in many locations the variance of the temperature
-% has increased about 1 unit in 10 years.
-
-% The goal of this paper, however, is not to make any conclusions about
-% the trend in the variance because we solved the problem only for a
-% grid over the united states and for 10 years of the data. A thorough
-% analysis, needs the full solution over the globe and for a longer time
-% period. The goal of the paper, was to propose the idea of estimating
-% the trend in variance of spatio-temporal signals using generalized
-% lasso and to investigate the algorithms for solving the resulting
-% optimization problem.
-
-\fontsize{9.0pt}{10.0pt} \selectfont
-\bibliographystyle{aaai}
-\bibliography{aaai-references}
-\clearpage
 
@@
 \begin{equation}
-  r_w(v,\mu_1,\mu_2):=
+\begin{aligned}
+  &r_w(v,\mu_1,\mu_2)\\
+  &:=
 	\begin{bmatrix}
@@
 	r_{cent}	
-	\end{bmatrix}=
+	\end{bmatrix}\\ &=
   \begin{bmatrix}
@@
   \end{bmatrix}
+\end{aligned}
 \label{eq:resid}
@@
 the objective function \eqref{eq:l1tf_var}. We chose the python API of
-the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.
-
-\section{Proof of Lemma 1}
-
-\begin{proof}
-  If $f(x)=\sum_k f_k(x_k)$ then $[\prox_{\mu f}(x)]_k =
-  \prox_{\mu f_k}(u_k)$. So 
-  $[\prox_{\mu f}(u)]_k=\min_{x_k} \,\,
-  \mu\big(x_k+y_{k}^2e^{-x_{k}}\big)+\frac{1}{2}  (x_k-u_k)^2.$
-  Setting the derivative to 0 and solving for $u_k$ gives the
-  result. Similarly, $[\prox_{\rho g}(u)]_\ell=\rho
-  \lambda_\ell |z_\ell|+1/2(z_\ell-u_\ell)^2$. This is not differentiable,
-  but the solution must satisfy $\rho \cdot \lambda_\ell \cdot \partial
-  \big(|z_\ell| \big)=u_\ell-z_\ell$ where $\partial \big(|z_\ell| \big)$ is the
-  sub-differential of $|z_\ell|$. The solution is the soft-thresholding
-  operator $S_{\rho\lambda_\ell}(u_\ell)$.
-\end{proof}
+the \texttt{cvxopt} package~\citep{andersen_cvxopt:_2013}.
 
@@
 
+\begin{small}
 \begin{equation}
@@
 \end{equation}
+\end{small}
 
@@
 
-The matrix $D$ has the following form:  $D^t=[D_{temp}^t|D_{spat}^t]$. The matrix $D_{temp}$ is the following block-diagonal matrix and corresponds to the temporal penalty: 
+The matrix $D$ has the following form:  $D=[D_{temp}|D_{spat}]$. The matrix $D_{temp}$ is the following block-diagonal matrix and corresponds to the temporal penalty: 
 \begin{equation}
@@
 function $f^*(\cdot)$. We have: 
+\begin{equation}
+\begin{aligned}
+&f^*(\xi)\\ & = \max_x  \xi^\top x - f(x)\\
+& =  \max_x \sum_{j=1}^{n_b} (\xi_jx_j - x_j - y_j^2e^{-x_j} -
+(\rho/2)(x_j-\tilde{z}_j+u_j)). 
+\end{aligned}
+\label{eq:conjugate}
+\end{equation}
+Setting the derivative of the terms inside the summation to 0, we obtain:
+\begin{equation}
+\xi_j-y_j^2e^{-x_j^*}-\rho x_j^* + \rho (\tilde{z}_j-u_j)=0,
+\label{eq:x_start}
+\end{equation}
+where $x^*$ is the maximizer in \ref{eq:conjugate}. Then, it
+can be shown that $x_j^*$ which satisfies \eqref{eq:x_start} can be
+obtained as follows:
 
 \begin{align}
-f^*(\xi) & = \max_x \quad \xi^\top x - f(x)\\
-& =  \max_x \quad \sum_{j=1}^{n_b} (\xi_jx_j - x_j - y_j^2e^{-x_j} - (\rho/2)(x_j-\tilde{z}_j+u_j))
-\label{eq:conjugate}
+x^*_j & = \mathscr{W}\bigg(\frac{y_j^2}{\rho} e^{\phi_j} \bigg) - \phi_j, \\
+\phi_j & =\frac{1-\xi_j-\rho(\tilde{z}_j-u_j)}{\rho}.
 \end{align}
 
-Setting the derivative of the terms inside the summation to 0, we obtain:
-
-\begin{equation}
-\xi_j-y_j^2e^{-x_j^*}-\rho x_j^* + \rho (\tilde{z}_j-u_j)=0
-\label{eq:x_start}
-\end{equation}
-
-\noindent where $x^*$ is the maximizer in \ref{eq:conjugate}. Then, it can be shown that $x_j^*$ which satisfies \eqref{eq:x_start} can be obtained as follows:
-
-\begin{align}
-x^*_j & = \mathscr{W}\bigg(\frac{y_j^2}{\rho} e^{\phi_j} \bigg) - \phi_j \\
-\phi_j & =\frac{1-\xi_j-\rho(\tilde{z}_j-u_j)}{\rho}
-\end{align}
-
-In this equation, $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996}.
+In this equation, $\mathscr{W}(\cdot)$ is the Lambert W function \cite{corless_lambertw_1996}.
 Finally, the conjugate function is: $f^*(\xi) = \sum_{j=1}^{n_b} (\xi_jx^*_j - x^*_j - y_j^2e^{-x^*_j} - (\rho/2)(x^*_j-\tilde{z}_j+u_j))$.
@@
 Using the chain rule we get:
-
 %\x^*_j + \xi_j + \frac{\partial \x^*_j}{\partial \xi_j}
-
 \begin{equation}
-\frac{\partial f^*(\xi)}{\partial \xi_j}  =  x^*_j  + \frac{\partial x^*_j}{\partial \xi_j} \bigg[ \xi_j -1 + y_j^2 e^{-x_j^*} + \rho (\tilde{z}_j - u_j - x_j^*) \bigg]
+\frac{\partial f^*(\xi)}{\partial \xi_j}  =  x^*_j  + \frac{\partial
+  x^*_j}{\partial \xi_j} \left[ \xi_j -1 + y_j^2 e^{-x_j^*} + \rho
+  (\tilde{z}_j - u_j - x_j^*) \right], 
 \label{eq:d_f*_start}
 \end{equation}
-where we have:
-
+where we have
 \begin{equation}
-\frac{\partial x_j^*}{\partial \xi_j}  = \frac{1}{\rho(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))}
+\frac{\partial x_j^*}{\partial \xi_j}  = \frac{1}{\rho(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))}.
 \label{eq:d_x*_start}
@@
 By some tedious but straightforward computation we can obtain the second derivatives:
-
 \begin{align}
-\frac{\partial^2 f^*(\xi)}{\partial \xi_j^2} & =  \frac{\partial x_j^*}{\partial \xi_j} - \rho \frac{\partial^2 x_j^*}{\partial \xi_j^2} \bigg[ \phi_j +x_j^* - \tilde{z}_j + u_j \bigg]\\
-& \quad + \frac{\partial x_j^*}{\partial \xi_j} \bigg[ 1-y_j^2 \frac{\partial x_j^*}{\partial \xi_j} e^{-x_j^*} -\rho \frac{\partial x_j^*}{\partial \xi_j} \bigg]
-\label{eq:d2_f*_start}
+\frac{\partial^2 f^*(\xi)}{\partial \xi_j^2} 
+  & =  \frac{\partial
+    x_j^*}{\partial \xi_j}
+    - \rho \frac{\partial^2
+    x_j^*}{\partial
+    \xi_j^2} \bigg[ \phi_j
+    +x_j^* - \tilde{z}_j +
+    u_j \bigg],\\ 
+  & \quad + \frac{\partial x_j^*}{\partial \xi_j} \bigg[ 1-y_j^2
+    \frac{\partial x_j^*}{\partial \xi_j} e^{-x_j^*} -\rho
+    \frac{\partial x_j^*}{\partial \xi_j} \bigg], \label{eq:d2_f*_start}\\
+  \frac{\partial^2 x_j^*}{\partial \xi_j^2}  &=
+                                               \frac{\mathscr{W}((y_j^2/\rho)
+                                               e^{-\phi_j}
+                                               )}{\rho^2(1+\mathscr{W}((y_j^2/\rho)
+                                               e^{-\phi_j} ))^3} .
+\label{eq:d2_x*_start}
 \end{align}
 
-\begin{equation}
-\frac{\partial^2 x_j^*}{\partial \xi_j^2}  = \frac{\mathscr{W}((y_j^2/\rho) e^{-\phi_j} )}{\rho^2(1+\mathscr{W}((y_j^2/\rho) e^{-\phi_j} ))^3}
-\label{eq:d2_x*_start}
-\end{equation}
+\section{Proof of Lemma 1}
+\label{sec:proof-lemma-1}
 
-% \section{Appendix C}
-% \autoref{tab:sim_params} lists the parameters used for simulating data
-% in \autoref{sec:simulations}. \autoref{fig:true_var_spatial} shows the
-% variance function obtained from there parameters at $t=25$ (left) and
-% $t=45$ (center).
+\begin{proof}
+  If $f(x)=\sum_k f_k(x_k)$ then $[\prox_{\mu f}(x)]_k =
+  \prox_{\mu f_k}(u_k)$. So 
+  $[\prox_{\mu f}(u)]_k=\min_{x_k} \,\,
+  \mu\big(x_k+y_{k}^2e^{-x_{k}}\big)+\frac{1}{2}  (x_k-u_k)^2.$
+  Setting the derivative to 0 and solving for $u_k$ gives the
+  result. Similarly, $[\prox_{\rho g}(u)]_\ell=\rho
+  \lambda_\ell |z_\ell|+1/2(z_\ell-u_\ell)^2$. This is not differentiable,
+  but the solution must satisfy $\rho \cdot \lambda_\ell \cdot \partial
+  \big(|z_\ell| \big)=u_\ell-z_\ell$ where $\partial \big(|z_\ell| \big)$ is the
+  sub-differential of $|z_\ell|$. The solution is the soft-thresholding
+  operator $S_{\rho\lambda_\ell}(u_\ell)$.
+\end{proof}
 
-% \begin{table}[tb]
-%   \caption{Parameters used to simulate data.}
-%   \label{tab:sim_params}
-%   \begin{center}
-%     \begin{tabular}{ccccccc}
-%       \hline
-%       $s$ & $r_s$ & $c_s$ & $\sigma_s$ &$\alpha_s$ & $\omega_s$ & $\phi_s$\\
-%       \hline
-%       1 & 0 & 0 & 5 & 0.5 & 0.121 & 0 \\
-%       2 & 0 & 5 & 5 & 0.1 & 0.121 & 0 \\
-%       3 & 3 & 0 & 5 & -0.5 & 0.121 & $\pi/2$ \\
-%       4 & 3 & 5 & 5 & -0.1 & 0.121 & $\pi/2$ \\
-%       \hline
-%     \end{tabular}
-%   \end{center}
-% \end{table}
-
-% \begin{figure}[tb]
-%   \centering	
-%   \includegraphics[height=.15\textheight]{Figures/true_var_spatial}
-%   \includegraphics[height=.15\textheight]{Figures/true_fitted_var}
-%   \caption{Variance function at $t=25$ (left) and $t=45$
-%     (center). Right: the true (orange) and estimated standard deviation 
-%     function at the location (0,0). The estimated values are
-%     obtained using linearized ADMM with $\lambda_s=0.1$ and two
-%     values of $\lambda_t$: $\lambda_t=5$ (blue) and
-%     $\lambda_t=100$ (green).} \label{fig:true_var_spatial}
-% \end{figure}
+\bibliographystyle{aaai}
+\bibliography{aaai-references}
 
 \end{document}
-
-\section{Data Exploration}
-
-In this appendix we examine some of the properties of the time-series
-of the temperature in ERA-20C dataset. The goal here is to demonstrate
-some of the difficulties in modeling the trend in the temperature
-volatility and motivate our methodology.
-
-Figure \autoref{fig:cities_ts} shows the time-series of the
-temperature of three cities: Indianapolis (USA), San Diego (USA) and
-Manaus (Brazil). The time-series of Indianapolis and San Diego show
-clear cyclic behavior. However, while it seems (qualitatively) that
-these cycles can be modeled by a sinusoidal function for Indianapolis,
-the same is not true for San Diego. Also, the amplitude of the cycles
-changes from some years to others. The time-series of Manaus does not
-show any regular cyclic behavior. This demonstrates the first
-difficulty in analyzing the variance of this data: to analyze the
-variance, we first need to remove the cyclic terms from all
-time-series. However, there is a lot of variations in the cyclic
-behavior of the time-series of different locations. In addition, some
-of these cycles cannot be easily modeled by a parametric function
-\footnote{One might try to model the cycles by the summation of
-  sinusoidal terms with different frequencies. However, for some
-  time-series this may need many terms to be included in the summation
-  to achieve a reasonable level of accuracy. In addition, this model
-  cannot capture the non-stationarity in the cycles.}. To overcome
-these issues, we use a non-parametric approach to remove the cyclic
-terms from the time-series and de-trend them. This approach, called
-\textit{$\ell_1$-trend filtering} is explained in Section 2 of the
-text. We detrended each time-series separately using this method. For
-each time-series, we found the optimal value of the penalty parameter
-using \textit{k-fold cross-validation} with $k=5$. We used the R
-package \textbf{genlasso} to perform these computations
-\cite{arnold_efficient_2016}.
-
-\begin{figure}[tb]
-	\centering
-	\includegraphics[width=.2\textheight]{Figures/cities_ts}
- 	\caption{Time-series of the temperature (in Kelvin) of three cities.}
- 	\label{fig:cities_ts}
-\end{figure}
-
-The blue curve in the left panel of Figure \autoref{fig:bloom_estimatedSD} shows the time-series of the temperature of Indianapolis after detrending using this method.
-This figure, reveals another difficulty in estimating the trend of volatility in this data: the variance of this signal, shows cyclic behavior.
-Also, the cycles are not regular and their amplitude and frequency change.
-Even if one can describe the behavior of the variance of the time-series at all locations using a single parametric model (for example a variant of the GARCH models \citep{bollerslev_generalized_1986}), it is not clear how the trend in the variance should be investigated in this framework.
-These observations motivate the need to develop a non-parametric framework for the problem at hand.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=\columnwidth]{Figures/indy_estimatedSD_1}
-  \caption{Left: The residuals of the time-series of Indianapolis
-    (averaged weekly) and the estimated SD obtained from the method of
-    \autoref{sec:l1tf_var} (red). Middle: the estimated SDs (red) and
-    their annual average (black) without imposing the long horizon
-    penalty. Right: the same as middle panel but here the long horizon
-    penalty is imposed. See the text for more details.} 
-  \label{fig:bloom_estimatedSD}
-\end{figure}
-
-The red curve in the left panel of Figure \autoref{fig:bloom_estimatedSD}
-shows the estimated SD (which is $\exp(h_t/2)$) of the residuals of
-the time-series of Indianapolis obtained from out proposed model.
-To reduce the number of time-steps we work on the weekly averaged of the data.
-The curve of the estimated SD captures the periodic variations in the
-SD of the signal. Just by looking at this curve, it is hard to say if
-the SD is decreasing or increasing. Therefore, we compute the average
-of the estimated SD for each year. The estimated SD together with this
-annual average is shown in the middle panel of
-\autoref{fig:bloom_estimatedSD}. As it can be seen, the annual trend
-is not smooth. This is because in the optimization problem
-\eqref{eq:l1tf_var}, the smoothness of the annual trend is not
-encouraged. To remedy this, we add the following long horizon penalty to \eqref{eq:l1tf_var}:
-
-\begin{equation}
-\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
-\label{eq:lh_penalty}
-\end{equation}
-
-where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$. Also,
-$N_{year}$ is the number of years over which we are performing our
-analysis (here $N_{year}=10$). Since we are working on the weekly
-averaged data, each year corresponds to 52 observations. In the matrix
-form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the
-matrix $D$. The estimated SDs using this penalty matrix is shown in
-the right panel of \autoref{fig:bloom_estimatedSD}. The annual average
-of the estimated SDs shows a linear trend with a positive slope.
-
-\end{document}

### 2018-09-06 aaai-2018/khodadadi-mcdonald-aaai19.tex
aaai submitted version

@@
   /Title (Algorithms for Estimating Trends in Global Temperature Volatility)
-   /Author (Khodadadi and McDonald)
+   %/Author (Khodadadi and McDonald)
   /Author (John Doe, Jane Doe)}
@@
 tailored for dense, gridded observations over both 
-space and time. We evaluate our methods with a simulation constructed
-to mimic these data's 
+space and time. We evaluate our methods with a simulation that mimics these data's 
 features and on a large, publicly available, global
@@
 % Very High Resolution Radiometer (AVHRR) instruments. 
-Current work~\cite{StatenKahn2016,SchreierKahn2010,KahnFishbein2007} seeks to create a
+To address this gap, current
+work~\cite{StatenKahn2016,SchreierKahn2010,KahnFishbein2007} seeks to
+create a 
 spectrally-detailed dataset by %which can avoid the above issues by
@@
 imagers with readings from High-resolution 
-Infrared Radiation Sounders on board legacy
+Infrared Radiation Sounders, instruments onboard legacy weather
 satellites. 
@@
 northern hemisphere can be attributed to increases in temperature
-variance  rather than to increases in
+variance  rather than to changes in
 mean climate~\cite{Screen2014,FischerBeyerle2013,TrenberthZhang2014}.
@@
 the standard deviation (SD) of the surface temperature in the NASA
-Goddard Institute for Space Studies gridded temperature data set by
+Goddard Institute for Space Studies gridded temperature dataset by
 examining the empirical SD at each spatial location relative to that
-location's SD over a base period, and showed
+location's SD over a base period and showed
 that these estimates are increasing.
@@
 \cite{hansen_perception_2012}. \citet{huntingford_no_2013} also computed the
-time-evolving global SD from the de-trended time-series at each
+time-evolving global SD from the detrended time-series at each
 position and argued that the global SD has been
@@
 
-These and other related work  
-\cite[e.g.,]{rhines_frequent_2013} have several shortcomings. First, no
+These and other related work (e.g., \citealp{rhines_frequent_2013}) have
+several shortcomings which our work seeks to remedy. First, no 
 statistical analysis has been performed to examine if the changes in
@@
 
-In the present work, we examine variance (rather than the mean) for a number of reasons.
+Importantly, existing literature and our present work examines variance (rather
+than the mean) for a number of reasons. 
 First, instrument bias in the satellites increases over time so
@@
 methodology, the variance at each position and time are estimated by
-maximizing penalized likelihood
-maximization. Following~\cite{Tibshirani2014}, we penalize the 
+minimizing the penalized negative
+loglikelihood. Following methods for mean estimation~\cite{Tibshirani2014}, we penalize the  
 differences between the estimated variances which
@@
 \cite{parikh_proximal_2014}, solves the main problem by iteratively
-solving a linearized version of it. We will compare the benefits of
+solving a linearized version. We will compare the benefits of
 each method.
@@
 \begin{enumerate}
-\item We propose a model for nonparametric variance estimation for a
+\item We propose a method for nonparametric variance estimation for a
   spatio-temporal process and discuss the relationship between our
@@
 \item We derive two alternating direction method of multiplier
-  algorithms (ADMM) to fit our estimator when applied to 
+  algorithms to fit our estimator when applied to 
   very large data (\autoref{sec:prop-optim-meth}). We give situations
   under which each algorithm is most likely to be useful. Open-source
-  Python code is available on {\tt github}.
+  Python code is available on {\tt github}.%\footnote{Link removed for anonymity.}
 \item Because the construction of satellite-based datasets is ongoing
@@
 \min_{\beta} \frac{1}{2} \sum_{t=1}^{T} (y_t-\beta_t)^2+\lambda
-\sum_{t=1}^{T-2} \left|\beta_t-2\beta_{t+1}+\beta_{t+2}\right| 
+\sum_{t=2}^{T-1} \left|\beta_{t-1}-2\beta_{t}+\beta_{t+1}\right| 
 $$
@@
  where $y=\{y_t\}_{t=1}^T$ is an observed time-series, $\beta\in \mathbb{R}^T$ is the smooth trend,
- $D_t$ is a $(T-2)\times T$ matrix (see Appendix C), and $\lambda$ is a tuning parameter
+ $D_t$ is a $(T-2)\times T$ matrix, and $\lambda$ is a tuning parameter
  which balances fidelity to the data (small errors in the first term)
@@
 nonparametric in the same way that splines are nonparametric. In fact,
-using squared $\ell_2$ norm as the penalty instead of $\ell_1$ results exactly in
+using squared $\ell_2$-norm as the penalty instead of $\ell_1$ results exactly in
 regression splines.
@@
 $\exp(h_t)$. The negative log-likelihood of the observed data in this
-model is $l(y\given h) \propto -\sum_{t=1}^T h_t - y_t^2e^{-h_t}$. Crucially,
+model is $l(y\given h) \propto -\sum_t h_t - y_t^2e^{-h_t}$. Crucially,
 we assume that the parameters $h_t$ vary smoothly and estimate them by
@@
 parameter. We seek to impose both temporal and spatial smoothness 
-constraints on the hidden variables. Specifically, we seek a solution
+constraints on the parameters. Specifically, we seek a solution
 for $h$ which is piecewise linear in time and piecewise constant in
@@
 by solving the following optimization problem: 
-{\scriptsize
+{\small
 \begin{equation}
@@
 \begin{aligned}
-&\min_h \sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}} +\lambda_t
-\sum_{i,j} \sum_{t=1}^{T-2}
-\left|h_{ijt}-2h_{ij(t+1)}+h_{ij(t+2)}\right|\\ 
+&\min_h \sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}}\\& +\lambda_t
+\sum_{i,j} \sum_{t=2}^{T-1}
+\left|h_{ij(t-1)}-2h_{ijt}+h_{ij(t+1)}\right|\\ 
 &+\lambda_s \sum_{t,j} \sum_{i=1}^{n_r-1} \left|h_{ijt}-h_{(i+1)jt}\right|
@@
 point in the network, but the network is known (corresponding to the grid
-over the earth). To modify the
-objective function in~\cite[Eq.\ 2]{HallacPark2017}, we would
+over the earth) and fixed in time. To apply these methods in our context
+(e.g., \citealp{HallacPark2017}, Eq.\ 2), we would
 enforce complete sparsity on the off-diagonal elements (since
@@
 significantly larger than these networks and attempting to estimate a
-full covariance would be prohibitive, were it required.
+full covariance would be prohibitive, were it necessary.
 
@@
 
-For a spatial grid of size $S$ and $T$ time steps, $D$ in equation \ref{eq:l1tf_var_st_mat} will have
-$3Tn_rn_c-2n_rn_c-Tn_r$ rows and $S \cdot T$ columns. For a $1^\circ\times
+For a spatial grid of size $S$ and $T$ time steps, $D$ in Equation
+\eqref{eq:l1tf_var_st_mat} will have 
+$3TS-2S-Tn_r$ rows and $TS$ columns. For a $1^\circ\times
 1^\circ$ grid over the entire northern hemisphere and daily data over
@@
 depends on $D^\top D$. Therefore,
-applying the PDIP directly is infeasible for our data.\footnote{We
+applying the PDIP directly is infeasible.\footnote{We
   note that $D$ is a highly structured, sparse matrix, but, unlike
   trend filtering alone, it is not banded. We are unaware of general
-  linear algebra techniques for inverting such matrix, despite our
+  linear algebra techniques for inverting such a matrix, despite our
   best efforts to find them.}
 
-In the next section, we develop two ADMM algorithms for solving this
+In the next section, we develop two algorithms for solving this
 problem efficiently. The first casts the problem as a
@@
 which solves smaller sub-problems using PDIP and then recombines the results. The
-second uses proximal methods to avoid matrix inversions.
+second uses proximal methods to avoid matrix inversions. Either may be more
+appropriate depending on the particular computing infrastructure.
 
@@
 Consider an optimization problem of the form $\min_h f(h)$, where
-$h\in\mathbb{R}^n$ is the \textit{global variable} and
+$h\in\mathbb{R}^n$ is the global variable and
 $f(h):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is
@@
 \in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{\mathscr{G}(i,j)}$. Then,
-the original optimization problem is equivalent to the following
-problem:  
+the original optimization problem is equivalent to:  
 \begin{equation}
@@
 correspond to several entries of the local variables and so the
-constraints $\tilde{h}_i=x_i$ enforce the consensus between the local
+constraints $\tilde{h}_i=x_i$ enforce consensus between the local
 variables corresponding to the same global variable.  
@@
 $n_i$. \autoref{alg:conADMM} gives the general version of this
-algorithm. A more detailed discussion of this is in Appendix B of the
+procedure. A more detailed discussion of this is in~\autoref{sec:app_consADMM} of the
 Supplement where we show how to compute the dual and the derivatives
@@
 sub-problems that can be solved independently, it is amenable to a
-split-gather parallelization strategy via, e.g., the map reduce framework.
+split-gather parallelization strategy via, e.g., the MapReduce framework.
 In each iteration, the
 computation time will be equal to the time to solve each sub-problem
-plus the time to communicate the solutions on the master processor
+plus the time to communicate the solutions to the master processor
 and perform the consensus step. Since each sub-problem is
@@
   \end{equation}
-  where $\mathscr{W}(\cdot)$ is the \textit{Lambert W function}
+  where $\mathscr{W}(\cdot)$ is the Lambert W function
   \cite{corless_lambertw_1996},  $[S_{\alpha}(u)]_k = \sign(u_k)(|u_k|
@@
 % \end{proof}
-The proof is fairly straightforward and given in Appendix B in the Supplement. 
+The proof is fairly straightforward and given
+in~\autoref{sec:proof-lemma-1} in the Supplement.  
 Therefore, \autoref{alg:linADMM} gives a different method for solving
-the same problem. For this algorithm, both the primal update and the
+the same problem. In this case, both the primal update and the
 soft thresholding step are performed elementwise at each point of the
@@
 useful in any particular context. In our experience,
-\autoref{alg:conADMM} converges in many fewer iterations, but
+\autoref{alg:conADMM} requires an order of magnitude fewer iterations, but
 each iteration is much slower unless carefully parallelized.
@@
 weighted sum of $K$ bell-shaped functions where the weights are
-time-varying, consist of a linear trend $\alpha_k \cdot t$ and a
-periodic term $\beta_k \cdot \sin(2\pi\omega_k t+\phi_k)$. The
+time-varying, consist of a linear trend %$\alpha_k \cdot t$ 
+and a periodic term. 
+%$\beta_k \cdot \sin(2\pi\omega_k t+\phi_k)$. 
+The
 bell-shaped functions impose spatial smoothness while the linear
 trend and the periodic terms enforce the temporal smoothness similar
-to the seasonal component in the real climate data. We simulated the
-data on a 5 by 7 grid and for 780 time steps with $K=4$. This yields a
-small enough problem so that it can be evaluated many times while
+to the seasonal component in real climate data. We simulated the
+data on a 5$\times$7 grid for 780 time steps with $K=4$. This yields a
+small enough problem to be evaluated many times while
 still mimicking important properties of climate data.
@@
 (\autoref{fig:true_var_spatial}, top panel) as well as
-the variance across time at $(0,0)$ (bottom panel, orange).
+the variance across time at location $(0,0)$ (\autoref{fig:true_var_spatial}, bottom panel, orange).
 \begin{table}[tb]
@@
 which are ``too smooth". The left panel of \autoref{fig:true_fitted_var} shows the
-convergence of both methods as a function of iteration. It is
+convergence of both Algorithms as a function of iteration. It is
 important to note that each iteration of the linearized
@@
 to~\citep{HallacPark2017} ignores the spatial component and thus gives
-results which are similar to the second column when the covariances are shrunk to
-zero and massively worse otherwise.
+results which are similar to the second column if the covariances are shrunk to
+zero (massively worse if they are estimated).
 
@@
 Weather Forecasts\footnote{\url{https://www.ecmwf.int}}. We use 
-the 2 meter temperature measured daily at 12 p.m from January 1, 1960
+the 2 meter temperature measured daily at noon local time from January 1, 1960
 to December 24, 2010.
@@
 show any regular cyclic behavior. For this reason, we
-first apply trend filtering to remove seasonal terms and de-trend
-every time series. For each time-series, we found the optimal value of
-the penalty parameter using 5-fold cross-validation with
-$k=5$. 
+first apply trend filtering to remove seasonal terms and detrend
+every time-series. For each time-series, we found the optimal value of
+the penalty parameter using 5-fold cross-validation with.
 % We used the R package \textbf{genlasso} to perform these
@@
 \end{figure}
+
 The blue curve in the top panel of 
@@
 temperature of Indianapolis after detrending using this method. 
-
 \begin{figure}[tb]
@@
     \autoref{sec:l1tf_var} (red). Lower left: the estimated SDs (red) and
-    their annual average (black) without imposing the long horizon
+    their annual average (black) without the long horizon
     penalty. Lower right: the same but with the long horizon
-    penalty is imposed.} 
+    penalty.} 
   \label{fig:bloom_estimatedSD}
@@
 The red curve 
-shows the estimated SD (which is $\exp(h_t/2)$) obtained from our
+shows the estimated SD, $\exp(h_t/2)$, obtained from our
 proposed model. For ease of analysis, we compute the average
@@
 
-In addition to the constraints discussed in \autoref{sec:exten}, we
-, we add
-a long horizon penalty to smooth the annual trend.
-$\sum_{i=1}^{N_{year}-2}\left| \sum_t h_{t_1}-2h_{t_2}+h_{t_3}\right|$
-where $N_{year}$ is the number of years, $t_1=52(i-1)+t$, $t_2=52i+t$
-and $t_3=52(i+1)+t$. Finally, because the observations are on the surface
+In addition to the constraints discussed in \autoref{sec:exten}, we add
+a long horizon penalty to smooth the annual trend:
+$\sum_{i=2}^{N_{\y}-1}\left| \sum_t h_{\mathcal{A}(-1)}-2h_{\mathcal{A}(0)}+h_{\mathcal{A}(1)}\right|$
+where $N_{\y}$ is the number of years and $\mathcal{A}(b)=\{t : t\in
+(\y_i+b)\}$. Finally, because the observations are on the
+surface 
 of a hemisphere rather than a grid, we add extra spatial constraints with the
@@
 
-As shown in \autoref{alg:linADMM}, we evaluated convergence using
+As shown in \autoref{alg:linADMM}, we checked convergence using
 $\epsilon=0.001\%$ of the MSE of the data.
-Our simulation experiments indicated that the
+Our simulations indicated that the
 convergence speed depends on the value of $\lambda_t$ and
@@
 \paragraph{Model Selection}
-One common method for choosing the penalty parameters in the Lasso
+One common method for choosing the penalty parameters in lasso
 problems is to find the solution for a range of the values of these
 parameters and then choose the values which minimize a model selection
-criterion. However, such analysis needs the computation of the degrees
-of freedom. Several previous work have investigated the degrees of freedom in
+criterion. However, such analysis needs either the computation of the degrees
+of freedom or requires cross validation. Previous work has
+investigated the degrees of freedom in 
 generalized lasso problems with Gaussian likelihood
-\cite{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}. However,
-similar computations for non-Gaussian likelihood remains an open problem.
+\cite{tibshirani_degrees_2012,hu_dual_2015,zeng_geometry_2017}, but,
+results for non-Gaussian likelihood remains an open problem, and cross
+validation is too expensive. 
 In this paper, therefore, we use a heuristic method for choosing $\lambda_t$ and
 $\lambda_s$: we compute the solutions for a range of values of
-these parameters and choose the values which minimize
-$\mathscr{L}(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D_{total}h
+and choose those which minimize
+$\mathscr{L}(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D h
 \lVert$. This objective is a compromise between the negative log
-likelihood ($-l(y|h)$) and the complexity of the solution ($\sum
-\lVert D_{total}h \lVert$). For smoother solutions the value of $\sum
-\lVert D_{total}h \lVert$ will be smaller but with the cost of larger
-$-l(y|h)$. We computed the optimal solution for all the combinations of the
+likelihood and the complexity of the solution. For smoother solutions the value of $\sum
+\lVert Dh \lVert$ will be smaller but with the cost of larger
+$-l(y|h)$. We computed the solution for all the combinations of the
 following sets of values: $\lambda_t \in \{0,2,4,8,10,15,200,1000\} \, \, ,
@@
   \includegraphics[width=.66\columnwidth]{Figures/ts_estimatedVarRight}
-  \caption{Difference in observed data from the estimated trend
+  \caption{Residuals from the estimated trend
     (blue), the estimated SDs (orange), and annual average SD (green) for
-   Indianapolis (top) and San Diego (bottom). The $y$-axis shows $\textrm{K}^{\circ}$.} 
+   Indianapolis (top) and San Diego (bottom). Units are $\textrm{K}^{\circ}$.} 
   \label{fig:avg_change_estimatedSD}
@@
     average of the detrended estimated variance over the northern hemisphere (top) and the
-    change in the variance from 1961 to 2011 (bottom).}
+    change in the variance from 1961 to 2011 (bottom). Units are K$^\circ$.}
 \label{fig:avg_chg}
@@
 temperature volatility. For example, we can see that the variance in
-Indianapolis displays a small positive trend. To determine how the volatility has
+Indianapolis displays a small positive trend (easiest to see in
+\autoref{fig:bloom_estimatedSD}). To determine how the volatility has 
 changed in each location, we subtract the average of the estimated
@@
 % over the oceans. 
-The most positive trend can be observed in Asia and
-particularly in south-east Asia.
+The most positive trend can be observed in Asia,
+particularly South-East Asia.
 
@@
 In this paper, we proposed a new method for estimating the variance of
-spatio-temporal data with the goal of analyzing global temperature
-data. The main idea is to cast this problem as a
+spatio-temporal data with the goal of analyzing global
+temperatures. The main idea is to cast this problem as a 
 constrained optimization problem where the constraints enforce smooth
@@
 iterations it is not very appropriate for parallelization. The reason
-is that after each iteration the solution computed in each machine
-should be broadcast to the master machine and this operation takes
-some time which depends on the speed of the network connecting the
+is that after each iteration the solution computed on each local machine
+must be collected by the master machine, and this operation takes
+depends on the speed of the network connecting the
 slave machines to the master. A direction for future research would be
@@
 \bibliography{aaai-references}
-\clearpage
+\end{document}
 
@@
 the objective function \eqref{eq:l1tf_var}. We chose the python API of
-the \texttt{cvxopt} package ~\citep{andersen_cvxopt:_2013}.
+the \texttt{cvxopt} package~\citep{andersen_cvxopt:_2013}.
 
@@
 
-In this equation, $\mathscr{W}(\cdot)$ is the \textit{Lambert function} \cite{corless_lambertw_1996}.
+In this equation, $\mathscr{W}(\cdot)$ is the Lambert W function \cite{corless_lambertw_1996}.
 Finally, the conjugate function is: $f^*(\xi) = \sum_{j=1}^{n_b} (\xi_jx^*_j - x^*_j - y_j^2e^{-x^*_j} - (\rho/2)(x^*_j-\tilde{z}_j+u_j))$.

### 2018-11-07 aaai-2018/khodadadi-mcdonald-aaai19-camera.tex
update ignore

--- 2018-09-06 aaai-2018/khodadadi-mcdonald-aaai19.tex
@@
   /Title (Algorithms for Estimating Trends in Global Temperature Volatility)
-   %/Author (Khodadadi and McDonald)
-  /Author (John Doe, Jane Doe)}
+   /Author (Khodadadi and McDonald)
+   % /Author (John Doe, Jane Doe)
+ }
 
@@
 \newcommand{\citet}[1]
-{\citeauthor{#1} ̃\shortcite{#1}}
+{\citeauthor{#1} ~\shortcite{#1}}
 \newcommand{\citep}{\cite}
 \newcommand{\citealp}[1]
-{\citeauthor{#1} ̃\citeyear{#1}}
+{\citeauthor{#1} ~\citeyear{#1}}
 
@@
 
-%\newcommand{\theHalgorithm}{\arabic{algorithm}}
-\usepackage{cleveref}
+%\newcommand{\theHalgorithm}{\arabic{algorithm}
 
@@
 \mathtoolsset{showonlyrefs}
+
+\usepackage{cleveref}
 
@@
 
-\author{Author 1 \ and Author 2\\
-Address line\\
-Address line
-}
-% \author{Arash Khodadadi \and Daniel J. McDonald\\
-%  Department of Statistics\\
-%  Indiana University\\
-%  Bloomington, IN 47408 \\
-%  \{arakhoda,dajmcdon\}@indiana.edu}
+% \author{Author 1 \ and Author 2\\
+% Address line\\
+% Address line
+% }
+\author{Arash Khodadadi \and Daniel J. McDonald\\
+ Department of Statistics\\
+ Indiana University\\
+ Bloomington, IN 47408 \\
+ \{arakhoda,dajmcdon\}@indiana.edu}
 
@@
   under which each algorithm is most likely to be useful. Open-source
-  Python code is available on {\tt github}.%\footnote{Link removed for anonymity.}
+  Python code is available.\footnote{\tt github.com/dajmcdon/VolatilityTrend}
 \item Because the construction of satellite-based datasets is ongoing
@@
 While the motivation for our methodology is its application to large, gridded climate
-data, we note that our algorithms are also applicable to neuroimaging,
-image denoising, or examining large collections of financial instruments.
+data, we note that our algorithms are easily generalizable to
+spatio-temporal data under convex loss, e.g. exponential family
+likelihood. Furthermore the spatial structure can be broadly construed
+to include general graph dependencies. Our current application uses
+Gamma likelihood which lends itself well to modeling trends in
+pollutant emissions or in astronomical phenomena like microwave
+background radiation. Volatility estimation in natural gas/oil markets
+or with financial data is another possibility. These methods can also
+be applied to resting-state fMRI data (though the penalty structure changes).
 
@@
 400 iterations requires about 4 seconds for the linearized method and
-2 hours for the consensus method.
+2 hours for the consensus method. For consensus ADMM, computation per
+iteration per core is requires $\sim$10 seconds with $\sim$4 seconds
+for communication. In general, the literature suggests linear
+convergence for ADMM~\citep{NishiharaLessard2015} and~\autoref{fig:true_fitted_var} seems to fit in the linear framework for both algorithms, though with different constants.
 
@@
 of these parameters as a warm start for the larger values resulting in
-faster convergence.
+faster convergence. The timing of the estimation depends on $\lambda_t$ and $\lambda_s$ but takes between 1 and 4 hours for convergence.
 
@@
   \centering
-  \includegraphics[width=.66\columnwidth]{Figures/ts_estimatedVarLeft}
-  \includegraphics[width=.66\columnwidth]{Figures/ts_estimatedVarRight}
+  \includegraphics[width=.45\columnwidth]{Figures/ts_estimatedVarLeft}
+  \includegraphics[width=.45\columnwidth]{Figures/ts_estimatedVarRight}
   \caption{Residuals from the estimated trend
     (blue), the estimated SDs (orange), and annual average SD (green) for
-   Indianapolis (top) and San Diego (bottom). Units are $\textrm{K}^{\circ}$.} 
+   Indianapolis (left) and San Diego (right). Units are $\textrm{K}^{\circ}$.} 
   \label{fig:avg_change_estimatedSD}
@@
 average of these estimates for two cities in the US: Indianapolis
-(top) and San Diego (bottom). The estimated SD 
+(left) and San Diego (right). The estimated SD 
 captures the periodic behavior in the variance of the time-series. In
@@
 % optimization problem.
+
+\section*{Acknowledgements}
+
+This material is based upon work supported by the National Science
+Foundation under Grant Nos.\ DMS--1407439 and DMS--1753171.
 

### 2018-11-07 aaai-2018/referee/response.txt
update ignore

We would like to thank the reviewers for their comments, and we appreciate the positive reception. We will address the reviewers' main concern first before responding to specific details.

The major issue for all three reviewers was lack of novelty/suitability for AAAI. We agree that the algorithmic contributions here are marginal (though we would argue nontrivial) from the point of view of standard AI/ML methodological papers. Our major goal was to create tailored algorithms which worked for a specific applied problem (though they do generalize, see below). For this reason, the paper was submitted to the "AI for Social Impact" track (see https://aaai.org/Conferences/AAAI-19/aaai19emergingcall/) rather than the technical track. We chose AAAI specifically because of this particular emphasis, feeling that this type of contribution fits well with that call. We hope the reviewers will (or already have) taken this into account.

Reviewer #1:

"it is hard to quantitatively evaluate the results...the real data experiments is still not very convincing"

This is an issue in our minds as well. We found it very difficult to find meaningful competing methods. While our simulation is enough to convince us that the estimator "works" for data of the appropriate structure, there aren't really any appropriate techniques to compare with in the literature. The state-of-the-art for this data is essentially the unpenalized estimator. We considered comparing to some image denoising algorithms (reformulated to estimate the size of the noise), but the performance was very poor since they ignore the temporal dependence.

Reviewer #2:

"the concern that the paper is too narrow by giving a few examples of other extant problems across the sciences to which their methodology could be applied"

Our methods are easily generalizable to spatio-temporal data under convex loss, e.g. exponential family likelihood. And "spatio" can be broadly construed to include more complicated graph dependencies. Our particular application uses Gamma likelihood (rather than the standard Gaussian). This lends itself well to modeling trends in pollutant emissions, say, or in astronomical phenomena like microwave background radiation. Volatility estimation in natural gas/oil markets (see e.g. Pindyck, R. S. (2004)) or with financial data is another possibility. We've also considered applying it to resting-state fMRI data (though the penalty structure changes). In ML, as mentioned above, it could conceivably be used for video denoising, though we doubt it would be competitive with the state-of-the-art.

Reviewer #3:

"For consensus ADMM, can you provide results on how much time is spent on...communication...computation"?

Computation per iteration per core is ~10s. Communication is ~4. Though this is platform specific.

"The model selection criterion"

We agree completely that this is heuristic. Cross validation has no theoretical support in  the spatio-temporal setting, and is too expensive computationally. Our current (theoretical) work derives unbiased risk estimators (like AIC) for this setup. Previous work requires Gaussian likelihood, and our results extend it to exponential families with this as a motivating application. It was not yet ready at the submission deadline.

"If you change the model selection criterion,... will you still get this trend"?

The short answer is: probably here, though not elsewhere. We chose Indianapolis for that figure because it has variable climate but not noisy weather, so that location isn't particularly sensitive to the tuning parameters. But other locations are more sensitive. Essentially, specific locations in the map in Fig 7 will likely change, but the broad pattern won't change too much (based on visual comparisons at other choices). However, this is strong motivation for the need for a theoretically sound risk estimator.

"Can you provide the convergence rate for the ADMM algorithms?"

The literature suggests linear convergence for ADMM (see Hong and Luo, 2016 or Nishihara, Lessard, Recht, et al. 2015). Fig 3 seems to fit in the linear framework for both algorithms, though with different constants.

"Can you provide the computing time for the algorithms?"

For the simulation, the times are in the first paragraph of p6 and the left panel of Fig 3. For the temperature data we will provide these times in the revised version of the paper. The timing depends on the lambda_t and lambda_s, but it takes between 1 and 4 hours for convergence.

### 2018-11-07 aaai-2018/aaai-paper-supplement.tex
update ignore

@@
 \pdfinfo{
-  /Title (Supplement for Algorithms for Estimating Trends in Global Temperature
-  Volatility )
-   %/Author (Khodadadi and McDonald)
-  /Author (John Doe, Jane Doe)}
+  /Title (Supplement for ``Algorithms for Estimating Trends in Global Temperature
+  Volatility'' )
+   /Author (Khodadadi and McDonald)}
 
@@
 %\newcommand{\theHalgorithm}{\arabic{algorithm}}
-\usepackage{cleveref}
 
@@
 \usepackage{xr}
-\externaldocument{khodadadi-mcdonald-aaai19}
+\externaldocument{khodadadi-mcdonald-aaai19-camera}
 
@@
 
-\title{Supplement for Algorithms for Estimating Trends in Global Temperature Volatility}
+\title{Supplement for ``Algorithms for Estimating Trends in Global Temperature Volatility''}
 
-\author{Author 1 \ and Author 2\\
-Address line\\
-Address line
-}
+\author{Arash Khodadadi \and Daniel J. McDonald\\
+ Department of Statistics\\
+ Indiana University\\
+ Bloomington, IN 47408 \\
+ \{arakhoda,dajmcdon\}@indiana.edu}
+
 % \author{Arash Khodadadi \and Daniel J. McDonald\\
@@
 The dual of a generalized LASSO with the objective $f(x)+\lambda \norm{ Dx }_1$ is:
-
 \begin{align}
@@
 \end{align}
-
-\noindent where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that for the objective function of Equation \ref{eq:l1tf_var}
-
+where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x u^\top x-f(x)$. It is simple to show that for the objective function of Equation \ref{eq:l1tf_var}
 \begin{equation}
@@
 Each iteration of PDIP involves computing a search direction by taking a Newton step for the system of nonlinear equations $r_w(v,\mu_1,\mu_2)=0$, where $w>0$ is a parameter and
-
 \begin{equation}
 \begin{aligned}
-  &r_w(v,\mu_1,\mu_2)\\
-  &:=
+  &r_w(v,\mu_1,\mu_2)
+  :=
 	\begin{bmatrix}
@@
 f^*(-D^\top v)$ and $\nabla^2 f^*(-D^\top v)$. 
-
 It is straightforward to show that
@@
 In this section we give more details on performing the $x$-update step in Algorithm 1. We need to solve the following optimization problem:
-
-\begin{small}
-\begin{equation}
-\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
+\begin{align}
+\hat{x} &:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j +
+                y_j^2e^{-x_j})\\
+  &\quad+ (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
 \label{eq:x_update_opt}
-\end{equation}
-\end{small}
-
-\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$.
+\end{align}
+where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$.
 
@@
  1 & -2 & 1 &  &  &  &\\ 
- & 1 & -2 & 1 &  &  &\\ 
- &  & \ddots & \ddots & \ddots  &  &\\ 
- &  & & 1 & -2 & 1 &  \\ 
+ & 1 & -2 & 1 &  &0  &\\ 
+ &  &  & \ddots &  &  &\\ 
+ & 0 & & 1 & -2 & 1 &  \\ 
  &  &  &   & 1 & -2 & 1 
@@
 \begin{aligned}
-&f^*(\xi)\\ & = \max_x  \xi^\top x - f(x)\\
+&f^*(\xi)  = \max_x  \xi^\top x - f(x)\\
 & =  \max_x \sum_{j=1}^{n_b} (\xi_jx_j - x_j - y_j^2e^{-x_j} -
@@
 obtained as follows: 
-
 \begin{align}
@@
 \end{align}
-
 In this equation, $\mathscr{W}(\cdot)$ is the Lambert W function \cite{corless_lambertw_1996}.

### 2018-11-14 aaai-2018/aaai-paper-supplement.tex
final proofing of camera-ready

@@
 
-In this appendix we provide more details on how to solve the optimization problem with the objective specified in equation \ref{eq:l1tf_var} using PDIP.
+In this appendix we provide more details on how to solve the optimization problem with the objective specified in \autoref{eq:l1tf_var} using PDIP.
 The objective function is convex but not differentiable.
@@
 have an implementation of PDIP to solve the optimization problem with
-the objective function \eqref{eq:l1tf_var}. We chose the python API of
+the objective function \autoref{eq:l1tf_var}. We chose the python API of
 the \texttt{cvxopt} package~\citep{andersen_cvxopt:_2013}.

### 2018-11-14 aaai-2018/khodadadi-mcdonald-aaai19-camera.tex
final proofing of camera-ready

@@
 \DeclareMathOperator*{\argmin}{argmin}
-\usepackage{siunitx}
-\usepackage{comment}
-\usepackage{mathtools}
-\mathtoolsset{showonlyrefs}
+%\usepackage{siunitx}
+%\usepackage{comment}
 
@@
 \newcommand{\autoref}{\Cref}
+
+%\usepackage{autonum}
 
@@
 temperature variability will increase the probability of extreme hot
+or cold
 outliers~\cite{VasseurDeLong2014}. Recent climate literature
@@
 is an increase in the SDs from 1958-1970 to 1991-2001, it
-is much smaller than found in
-\cite{hansen_perception_2012}. \citet{huntingford_no_2013} also computed the
+is much smaller than found by
+\citet{hansen_perception_2012}. \citet{huntingford_no_2013} also computed the
 time-evolving global SD from the detrended time-series at each
@@
 pollutant emissions or in astronomical phenomena like microwave
-background radiation. Volatility estimation in natural gas/oil markets
-or with financial data is another possibility. These methods can also
+background radiation. Volatility estimation in oil and natural gas markets
+or with financial data is another possibility. Our methods can also
 be applied to resting-state fMRI data (though the penalty structure changes).
@@
 \begin{equation}
+  \label{eq:l1tf_var}
   \min_h -l(y\given h)+\lambda \norm{ D_th }_1
-\label{eq:l1tf_var}
 \end{equation}
@@
 second derivatives of the dual objective function.
-
-\begin{comment}
-\begin{algorithm}[tb]
-  \caption{PDIP for $\ell_1$ variance estimation}
-  \label{alg:pdip}
-  \begin{algorithmic}
-    \REQUIRE $\lambda>0$, $w>0$, $\nu\leftarrow 0$, $\mu_1\leftarrow
-    0$, $\mu_2\leftarrow 0$, $J\in\mathbb{Z}^+$, $\{w_k\}$
-    \COMMENT{Initialization} 
-    \FOR[Central path]{$k=1,2,\ldots$}
-    \FOR[Newton updates]{$j=1,\ldots,J$}
-    \STATE Solve $A [\Delta\nu\; \Delta\mu_1\;
-    \Delta\mu_2]^\top=r_{w_k}$ to find the search direction. 
-    \STATE $\quad\quad A$ is given in \ref{eq:delta_r}
-    \STATE Update $[\nu^{j+1}\; \mu^{j+1}_1\; \mu^{j+1}_2] \leftarrow
-    [\nu^{j}\; \mu^{j}_1\; \mu^{j}_2] + [\Delta\nu\; \Delta\mu_1\;
-    \Delta\mu_2]$. 
-    \ENDFOR
-   \ENDFOR
-   \RETURN $h=\log\frac{y^2}{1+D^\top\nu}$ 
-   \\
-   \attn{Is this right? Explicit form of $A$}\\
-   \attn{Arash: see appendix A}
-  \end{algorithmic}
-\end{algorithm}
-\end{comment}
 
@@
 {\small
-\begin{equation}
-\label{eq:l1tf_var_st}.
-\begin{aligned}
-&\min_h \sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}}\\& +\lambda_t
-\sum_{i,j} \sum_{t=2}^{T-1}
-\left|h_{ij(t-1)}-2h_{ijt}+h_{ij(t+1)}\right|\\ 
+\begin{align}
+  &\min_h \sum_{i,j,t}h_{ijt}+y_{ijt}^2e^{-h_{ijt}}\notag\\
+  & +\lambda_t \sum_{i,j} \sum_{t=2}^{T-1}
+    \left|h_{ij(t-1)}-2h_{ijt}+h_{ij(t+1)}\right|\label{eq:l1tf_var_st}\\ 
   &+\lambda_s \sum_{t,j} \sum_{i=1}^{n_r-1} \left|h_{ijt}-h_{(i+1)jt}\right|
-+\lambda_s \sum_{t,i} \sum_{j=1}^{n_c-1} \left|h_{ijt}-h_{i(j+1)t}\right|
-\end{aligned}
-\end{equation}
+    +\lambda_s \sum_{t,i} \sum_{j=1}^{n_c-1}
+    \left|h_{ijt}-h_{i(j+1)t}\right|\notag
+\end{align}
 }%
@@
 
-\begin{comment}
-General (unconstrained) optimization problems $\min_z
-\sum_i f_i(z)$, $z\in\mathbb{R}^n$ can be rewritten equivalently using a
-collection of local variables, $x_i\in \mathbb{R}^{n_i}$, and a
-constraint. Following the notation of 
-\cite{boyd_distributed_2011}, let $k=\mathscr{G}(i,j)$ which
-means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or
-$(x_i)_j=z_k$) and define $\tilde{z}_i \in \mathbb{R}^{n_i}$ by
-$(\tilde{z}_i)_j=(x_i)_j$. Then the original unconstrained
-optimization problem is equivalent to the following constrained
-problem
-$\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)$ subject to $\tilde{z}_i=x_i$.
-Now, we can apply ADMM to the augmented Lagrangian of this
-problem: 
-\begin{align}
-x_i&\leftarrow \argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
-     (\rho/2) \norm{ x_i-\tilde{z}_i }_2^2\\
-%\label{eq:ADMM_steps}
-z_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j\\
-u_i&\leftarrow u_i + \rho (x_i-\tilde{z}_i)
-\end{align}
-\end{comment}
-
 Consider an optimization problem of the form $\min_h f(h)$, where
@@
 each entry of the local variables corresponds to an entry of the
-global variable. Therefore we can define a mapping $\mathscr{G}(i,j)$
+global variable. Therefore we can define a mapping $G(i,j)$
 from the local variable indices into the global variable indices:
-$k=\mathscr{G}(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is
+$k=G(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is
 $h_k$ (or $(x_i)_j=h_k$). For ease of notation, define $\tilde{h}_i
-\in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{\mathscr{G}(i,j)}$. Then,
+\in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{G(i,j)}$. Then,
 the original optimization problem is equivalent to:  
@@
       vars using PDIP}
-    \STATE $h_k\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j
+    \STATE $h_k\leftarrow (1/S_k)\sum_{G(i,j)=k} (x_i)_j
     $. \COMMENT{Global update.}
@@
 has the form
-\begin{equation}
+\begin{equation*}
   \label{eq:linADMM_steps}
@@
   \end{aligned}
-\end{equation}
+\end{equation*}
 where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu <
@@
   \norm{x}_1$. Then, 
-  \begin{equation}
+  \begin{equation*}
     \begin{aligned}
-      [\prox_{\mu f}(u)]_k &= \mathscr{W}\bigg(\frac{y_k^2}{\mu}
+      \bigl[\prox_{\mu f}(u) \bigl]_k &= \mathscr{W}\bigg(\frac{y_k^2}{\mu}
       \exp\bigg(\frac{1-\mu u_k}{\mu}\bigg) \bigg) + \frac{1-\mu u_k}{\mu},\\
@@
     \end{aligned}
-  \end{equation}
+  \end{equation*}
   where $\mathscr{W}(\cdot)$ is the Lambert W function
@@
 in time and space given by the following parametric model:
-\begin{align}
+\begin{align*}
   \sigma^2(t,r,c) & =\sum_{k=1}^{K} W_k(t) \cdot \exp\bigg(
@@
 \label{eq:sourceVar}
-\end{align}
+\end{align*}
 The variance at each time and location is computed as the
@@
 2 hours for the consensus method. For consensus ADMM, computation per
-iteration per core is requires $\sim$10 seconds with $\sim$4 seconds
+iteration per core requires $\sim$10 seconds with $\sim$4 seconds
 for communication. In general, the literature suggests linear
@@
 sinusoidal terms with different frequencies. However, for some
-time-series this may need many terms to be included in the summation
-to achieve a reasonable level of accuracy. In addition, such a model
+locations, this would require many frequencies
+to achieve a reasonable level of accuracy, while other locations would
+require relatively few. In addition, such a model
 cannot capture the non-stationarity in the cycles.
@@
 Manaus (Brazil). The time-series of Indianapolis and San Diego show
-clear cyclic behavior, though the amplitude of the cycles
-changes. The time-series of Manaus does not
+clear cyclic behavior, though the amplitude is different. The
+time-series of Manaus does not 
 show any regular cyclic behavior. For this reason, we
@@
 every time-series. For each time-series, we found the optimal value of
-the penalty parameter using 5-fold cross-validation with.
+the penalty parameter using 5-fold cross-validation.
 % We used the R package \textbf{genlasso} to perform these
@@
 The blue curve in the top panel of 
-\autoref{fig:bloom_estimatedSD} shows the time-series of the
-temperature of Indianapolis after detrending using this method. 
+\autoref{fig:bloom_estimatedSD} shows the daily
+temperature for Indianapolis after detrending. 
 \begin{figure}[tb]
@@
 $\lambda_s$. For the temperature data, we used the solutions obtained for smaller values
-of these parameters as a warm start for the larger values resulting in
-faster convergence. The timing of the estimation depends on $\lambda_t$ and $\lambda_s$ but takes between 1 and 4 hours for convergence.
+of these parameters as warm starts for larger values. Estimation takes
+between 1 and 4 hours for convergence for each pair of tuning parameters.
 
@@
 and choose those which minimize
-$\mathscr{L}(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D h
+$L(\lambda_t,\lambda_s)=-l(y|h)+ \sum \lVert D h
 \lVert$. This objective is a compromise between the negative log
